### 2013-06-04 text/Chapter01.tex
Initial commit (working on chapter 2)

\chapter{Introduction} 
\label{ch:introduction}

%----------------------------------------------------------------------------------------

Some intro

\section{How lambda works}
The key difference - how lambda works

Effect systems, introduced by Gifford and Lucassen
\cite{effects-gifford}, track \emph{effects} of computations, such as
memory access or message-based
communication~\cite{effects-messagepassing}. Their approach augments
typing judgments with effect information: $\Gamma \vdash e :
\tau, F$.  Wadler and Thiemann explain how this shapes
effect analysis of lambda abstraction~\cite{monads-effects-marriage}:
%
\begin{quote}
\emph{In the rule for abstraction, the effect is empty because evaluation immediately
returns the function, with no side effects. The effect on the function arrow
is the same as the effect for the function body, because applying the function will
have the same side effects as evaluating the body.}
\end{quote}
%
In contrast to the static analysis of \emph{effects}, the analysis of \emph{context-dependence} 
does not match this pattern. In the systems we consider, lambda abstraction places 
requirements on both the \emph{call-site} (latent requirements) and the \emph{declaration-site} 
(immediate requirements), resulting in different syntactic properties. 
We informally discuss three examples first that demonstrate how contextual
requirements propagate. Section ? then unifies these 
in a single calculus.

Also \cite{effects-producer-semantics}

\begin{quote}
\emph{We will define an effect as a producer effect if all computations with that effect 
can be thunked as "pure" computations for a domain-specific notion of purity.}
\end{quote}

Demonstrate using distributed computations

Demonstrate using dataflow/liveness

Perhaps mention LINQ as a motivation for cross-compilation..
\cite{app-linq}

\section{Associating with context}

-> Type providers make this important

%---------------------------------------------------------------------------------------------------

\section{Flat coeffect system}
~

resources?

%---------------------------------------------------------------------------------------------------

\section{Structural coeffect system}
\label{sec:structural}

The \emph{flat coeffect system} presented in the previous sections has a number of uses, but often
we need to track context-dependence in a more fine-grained way. To track neededness or security, 
we need to associate information with individual \emph{variables} of the context.

At the same time, we want to avoid developing multiple variants of coeffect systems -- indeed, 
our motivation is to develop a \emph{single unified mechanism} for tracking context-dependence.
In this section, we present a more powerful \emph{structural coeffect calculus}, which is a 
generalization of the flat calculus.

%We show that \clflt~can be obtained as a special case of \clstr. The
%structural version of our system is also where our work differs more significantly from well-known
%effect systems.

% --------------------------------------------------------------------------------------------------

\subsection{Motivation: Tracking array accesses}
\label{sec:structural-motivation}
\newcommand{\pastval}[2]{\ident{#1}_{[#2]}}
\newcommand{\dnat}{\ident{D}_\ident{nat}}
\newcommand{\cprd}{\times}
\newcommand{\cvop}{~\ident{max}~}

Similarly to the flat version, the \emph{structural coeffect calculus} works with contexts and
functions annotated with a coeffectt tags, written $\ctyp{r}{\Gamma}$ and $\ctyp{r}{\tau_1} \rightarrow \tau_2$, 
respectively, but we use richer tag structure.

As an example, consider a language that allows us to get a value of a variable (representing
some changing data-source) \ident{x} versions back using the syntax $\pastval{a}{x}$. 
To track information about individual variables, we use a product-like operation $\cprd$ on tags 
to mirrors the product structure of variables. For example:
%
\begin{equation*}
\begin{array}{l}
\ctyp{5 \cprd 10}{(\ident{a}:\dnat, \ident{b}:\dnat)}
  \vdash
    \pastval{a}{5}+\pastval{b}{10}: \ident{nat}
\end{array}
\end{equation*}
%
The coeffect tag $5 \cprd 10$ corresponds to the free-variable context $\ident{a}, \ident{b}$, denoting
that we need at most 5 and 10 past values of \ident{a} and \ident{b}. If we substitute \ident{c}
for both \ident{a} and \ident{b}, we need another operation to combine multiple tags associated with 
a single variable:
%
\begin{equation*}
\begin{array}{l}
\ctyp{5 \cvop 10}{(\ident{c}:\dnat)}
  \vdash
    \pastval{c}{5}+\pastval{c}{10}: \ident{nat}
\end{array}
\end{equation*}
%
In this example, the operation $\cvop$ would be the \emph{max} function and so $5 \cvop 10 = 10$.
Before looking at the formal definition, consider the typing of let bindings:
%
\begin{equation*}
\begin{array}{l}
\kvd{let}~\ident{c} = \kvd{if}~\ident{test}()~\kvd{then}~\ident{a}~\kvd{else}~\ident{b}\\
\pastval{a}{15}+\pastval{c}{10}
\end{array}
\end{equation*}
%
The expression has free variables \ident{a} and \ident{b} (we ignore \ident{test}, which
is not a data source). It defines \ident{c}, which may be assigned either \ident{a} or \ident{b}.
The variable \ident{a} may be used directly (second line) or indirectly via \ident{c}.

The expression assigned to \ident{c} uses variables \ident{a} and \ident{b}, so its typing
context is $\ctyp{0 \cprd 0}{(\ident{a}, \ident{b})}$. The value $0$ is the unit
of $\cvop$ and it denotes empty coeffect. The typing context of the
body is $\ctyp{15 \cprd 10}{(\ident{a}, \ident{c})}$.

To combine the tags, we take the coeffect associated with \ident{c} and apply it to the tags
of the context in which \ident{c} was defined using the $\cvop$ operation. This is then combined 
with the remaining tags from the body yielding the overall context:
$\ctyp{15 \cprd (10 \cvop (0 \cprd 0))}{(\ident{a}, (\ident{a}, \ident{b}))}$. Using a simple
normalization mechanism (described later), this can be further reduced to
$\ctyp{(15 \cvop 10) \cprd 10}{(\ident{a}, \ident{b})}$. This gives us the required information --
we need at most $\emph{max}(15,10)$ past values of \ident{a} and at most $10$ past values of 
\ident{b}.

%===================================================================================================

\section{Contributions}