### 2014-03-27 text/Chapter01.tex
Finish introduction chapter and restructure the rest

% ==================================================================================================

\chapter{Why context-aware programming matters} 
\label{ch:intro}

Many advances in programming language design are driven by some practical motivations. Sometimes, 
the practical motivations are easy to see -- for example, when they come from an external change 
such as the rise of multi-core processors. Sometimes, discovering the practical motivations is a 
difficult task -- perhaps because we are so used to a certain way of doing things that we do not 
even \emph{see} the flaws of our approach.

Before exploring the motivations leading to this thesis, we briefly consider two recent practical 
concerns that led to the development of new programming languages. This helps to explain why 
context-aware programming is of importance. The examples are by no means representative, but they 
illustrate different kinds of motivations well.

\paragraph{Parallel programming.}

The rise of multi-core CPUs is a clear example of an external development influencing
programming language research. As multi-core and multi-processor systems became de-facto standard, 
languages had to provide better abstractions for parallel programming. This led to the industrial 
popularity of \emph{immutable} data structures (and functional programming in general), software 
transactional memory \cite{app-haskell-stm}, data-parallelism and also asynchronous 
computing \cite{app-fsharp-async}.

In this case, the motivation is easy to see -- writing multi-core programs using earlier 
abstractions, such as threads and locks, is difficult and error-prone. At the same time, 
multi-core CPUs become the standard very quickly and so the lack of good language abstractions
was apparent.

\paragraph{Data access.}

Accessing data is an example of a more subtle challenge. Initiatives like open government data\footnote{
In the UK, the open government data portal is available at: \url{http://data.gov.uk/}}
certainly make more data available. However, to access the data, one has to parse CSV and Excel files, 
issue SQL or SPARQL queries (to query database and the semantic web, respectively).

Technologies like LINQ \cite{app-linq} make querying data significantly easier. But perhaps 
because accessing data became important more gradually, it was not easy to see that inline
SQL is a poor solution \emph{before} better approaches were developed.

This is even more the case for \emph{type providers} -- a recent feature in F\# that integrates
external data sources directly into the type system of the language and thus makes data explorable
directly from the source code editor (through features such as auto-completion on object members).
It is not easy to see the limitations of standard techniques (using HTTP requests to query REST 
services or parsing CSV files and using string-based lookup) until one sees how type providers 
change the data-scientist's workflow\footnote{This is difficult to explain in writing and so the 
reader is encouraged to watch a video showing type providers for the WorldBank and CSV data 
sources \cite{app-fsharp-world}.}.

\paragraph{Context-aware programming.}

In this chapter, we argue that the next important practical challenge for programming language 
designers is designing languages that are better at working with (and understanding) the 
\emph{context in which programs are executed}.

This challenge is of the kind that is not easy to see -- perhaps because we are so used to 
doing things in certain ways that we cannot see their flaws. In this chapter, we aim to
uncover such flaws -- we look at a number of basic programs that rely on contextual 
information, we explain why they are inappropriate and then we briefly outline how this
thesis remedies the situation.

Putting deeper philosophical questions about the nature of scientific prog\-ress aside,
the goal of programming language research is generally to design languages that provide 
more \emph{appropriate abstractions} for capturing common problems, are \emph{simple} and 
more \emph{unified}. 
These are exactly the aims that we follow in this thesis. In this chapter, we explain what
the common problems are. In Chapter~\ref{ch:flat} and Chapter~\ref{ch:structural},
we develop two simple calculi to understand and capture the structure of the problems and,
finally, Chapter~\ref{ch:unified} unifies the two abstractions.

% ==================================================================================================

\section{Why context-aware programming matters}

The phrase \emph{context in which programs are executed} sounds rather abstract and generic.
What notions of \emph{context} can be identified in modern software systems?
Different environments provide different resources (\eg~database
or GPS sensor), environments are increasingly diverse (\eg~multiple versions of different
mobile platforms). Web applications are split between client, server and mobile components; mobile
applications must be aware of the physical environment while the ``internet
of things`` makes the environments even more heterogeneous. At the same time, applications access
rich data sources and need to be aware of security policies and provenance information from the
environment.

Writing such context-aware (or environment-aware) applications is a fundamental problem of modern 
software engineering. The state of the art relies on ad-hoc approaches -- using hand-written conditions 
or pre-processors for conditional compilation. Common problems that developers face include:

\begin{itemize}
\item \textbf{System capabilities.} When writing code that is cross-compiled to multiple 
  targets (e.g. SQL \cite{app-linq}, OpenCL or JavaScript \cite{app-hop-lang}) a part of the compilation 
  (generating the SQL query) often occurs at runtime and developers have no guarantee that it will succeed 
  until the program is executed.

\item \textbf{Platform versions.} When developing cross-platform applications, different platforms
  (and different versions of the same platform) provide different API functions. Writing a 
  cross-platform code usually relies on (fragile) conditional compilation or (equally fragile) 
  dynamic loading.

\item \textbf{Security and provenance.} When working with data (be it sensitive database or social 
   network data), we have permissions to access only some of the data and we may want to track 
   \emph{provenance} information. However, this is not checked -- if a program attempts to access 
   unavailable data, the access will be refused at run-time.

\item \textbf{Resources \& data availability.} When creating a mobile application, the program may
  (or may not) be granted access to device capabilities such as GPS sensor, social updates or battery
  status. We would like to know which of the capabilities are required and which are optional
  (\ie~enhance the user experience, but there is a fallback strategy). Equally, on the server-side,
  we might have access to different database tables and other information sources.
\end{itemize}

Most developers do not perceive the above as programming language flaws -- they are simply common
programming problems (at most somewhat annoying and tedious) that had to be solved. However, this 
is because we do not realize that a suitable language extension could make the above problems 
significantly easier to solve. As the number of distinct contexts and their diversity increases,
these problems will become even more commonplace.

The following sub-sections explore four examples in more details. The examples are chosen to 
demonstrate two distinct forms of contexts that are studied in this thesis.

% --------------------------------------------------------------------------------------------------

\begin{figure}
\begin{equation*}
\begin{array}{l}
\kvd{for}~\ident{header},\ident{value}~\kvd{in}~\ident{header}~\kvd{do}\\[-0.25em]
\hspace{2em}\kvd{match}~\ident{header}~\kvd{with}\\[-0.25em]
\hspace{2em}|~\str{accept} \rightarrow~ \ident{req.Accept} \leftarrow \ident{value}\\[-0.25em]
%
\prepk{\#if}~\prepi{FX\_NO\_WEBREQUEST\_USERAGENT}\\[-0.25em]
\hspace{2em}|~\str{user-agent} \rightarrow~ \ident{req.UserAgent} \leftarrow \ident{value}\\[-0.25em]
\prepk{\#else}\\[-0.25em]
\hspace{2em}|~\str{user-agent} \rightarrow~
  \ident{req.Headers}.[\,\ident{HttpHeader.UserAgent}\,] \leftarrow \ident{value}\\[-0.25em]
\prepk{\#endif}\\[-0.25em]
%
\prepk{\#if}~\prepi{FX\_NO\_WEBREQUEST\_REFERER}\\[-0.25em]
\hspace{2em}|~\str{referer} \rightarrow~ \ident{req.Referer} \leftarrow \ident{value}\\[-0.25em]
\prepk{\#else}\\[-0.25em]
\hspace{2em}|~\str{user-agent} \rightarrow~
  \ident{req.Headers}.[\,\ident{HttpHeader.Referer}\,] \leftarrow \ident{value}\\[-0.25em]
\prepk{\#endif}\\[-0.25em]
%
\hspace{2em}|~\ident{other} \rightarrow~ \ident{req.Headers}.[\,\ident{other}\,] \leftarrow \ident{value}\\[-0.25em]
\end{array}
\end{equation*}

\vspace{1em}
\noindent\makebox[\linewidth]{\rule{\textwidth}{0.5pt}} 
\vspace{-1.5em}
\caption{Conditional compilation in the HTTP module of the F\# Data library}
\label{fig:introduction-context-http}
\end{figure}

% --------------------------------------------------------------------------------------------------

\subsection{Context awareness \#1: Platform versioning}

The diversity across devices means that developers need to target an increasing number of platforms
and possibly also multiple versions of each platform. For Android, there is a number called 
API level \cite{app-android-apilevel} which ``uniquely identifies the framework API revision offered 
by a version of the Android platform''. Most changes in the libraries (but not all) are additive.

Equally, in the .NET ecosystem, there are multiple versions of the .NET runtime, mobile and 
portable versions of the framework etc. The differences may be subtle -- for example, some members
are omitted to make the mobile version of the library smaller, some functionality is not available
at all, but naming can also vary between versions.

For example, the Figure~\ref{fig:introduction-context-http} shows an excerpt from the \ident{Http} module 
in the F\# Data library\footnote{The file version shown here is available at:
\url{https://github.com/fsharp/FSharp.Data/blob/b4c58f4015a63bb9f8bb4449ab93853b90f93790/src/Net/Http.fs}}.
The example uses conditional compilation to target multiple versions of the .NET framework. Such code
is difficult to write -- to see whether a change is correct, it had to be recompiled for all combinations
of pre-processor flags -- and maintaining the code is equally hard. The above example could be refactored
and the .NET API could be cleaner, but the fundamental issue remains. If the language does not understand
the context (here, the different platforms and platform versions), it cannot provide any static
guarantees about the code.

As an alternative to conditional compilation, developers can use dynamic loading. For example, 
on Android, programs can access API from higher level platform dynamically using techniques like 
reflection and writing wrappers. This is even more error prone. As noted in an article introducing
the technique\footnote{Retrieved from: \url{http://android-developers.blogspot.com/2009/04/backward-}\\\url{compatibility-for-android.html}}:
\emph{``Remember the mantra: if you haven't tried it, it doesn't work''}. Again, it would be 
reasonable to expect that statically-typed languages could provide a better solution.

% --------------------------------------------------------------------------------------------------

\subsection{Context awareness \#2: System capabilities}

Another example related to the previous one is when libraries use meta-programming techniques
(such as LINQ \cite{app-linq} or F\# quotations \cite{app-fsharp-metaprog}) to translate code
written in a subset of a host language to some other target language, such as SQL, OpenCL or 
JavaScript. This is an important technique, because it lets developers targets multiple 
heterogeneous runtimes that have limited execution capabilities.

For example, consider the following LINQ query written in C\# that queries a database and
selects product names where the first upper case letter is "C":
%
\begin{equation*}
\begin{array}{l}
\kvd{var}~\ident{db} = \kvd{new}~\ident{NorthwindDataContext}();\\[0.25em]
\kvd{from}~\ident{p}~\kvd{in}~\ident{db.Products}\\[-0.25em]
\kvd{where}~\ident{p.ProductName.First}(\ident{c} \Rightarrow \ident{Char.IsUpper}(\ident{c})) == \str{C}\\[-0.25em]
\kvd{select}~\ident{p.ProductName};
\end{array}
\end{equation*}
%
This appears as a perfectly valid code ant the C\# compiler accepts it. However, when the program
is executed, it fails with the following error:
%
\begin{quote}
Unhandled Exception: \texttt{System.NotSupportedException}: Sequence operators 
not supported for type \texttt{System.String}.
\end{quote}
%
The problem is that LINQ can only translate a \emph{subset} of normal C\# code. The above snippet
uses \ident{First} method to iterate over characters of a string, which is not supported. This is
not a technical limitation of LINQ, but a fundamental problem of the approach.

When cross-compiling to a limited environment, we cannot always support the full source language.
The example with LINQ and SQL demonstrates the importance of this problem. As of March 2014,
Google search returns 11800 results for the message above and even more (44100 results) for a
LINQ error message \emph{``Method X has no supported translation to SQL``} caused by a similar 
limitation.

% --------------------------------------------------------------------------------------------------

\subsection{Context awareness \#3: Confidentiality and provenance}

The previous two examples were related to the non-existence of some library functions in another
environment. Another common factor was that they were related to the execution context of the whole 
program or a scope. However, contextual properties can be also related to specific variables.

For example, consider the following code sample that accesses database by building a SQL query
using string concatenation:
%
\begin{equation*}
\begin{array}{l}
\kvd{let}~\ident{query} = \ident{sprintf}~\str{SELECT * FROM Products WHERE Name='\%s'}~\ident{name}\\[-0.25em]
\kvd{let}~\ident{cmd} = \kvd{new}~\ident{SqlCommand}(\ident{query})\\[-0.25em]
\kvd{let}~\ident{reader} = \ident{cmd.ExecuteReader}()
\end{array}
\end{equation*}
%
The code compiles without error, but it contains a major security flaw called \emph{SQL injection}
(an attacker could enter \str{'; DROP TABLE Products -{}-} as their \ident{name} and delete the 
database table with products). For this reason, most libraries discourage building SQL commands by
string concatenation, but there are still many systems that do so.

The example demonstrates a more general property. Sometimes, it is desirable to track additional 
meta-data about variables that are in some ways special. Such meta-data can determine how the variables
can be used. Here, \ident{name} comes from the user input. This \emph{provenance} information should 
be propagated to \ident{query}. The \ident{SqlCommand} object should then require arguments that can
not directly contain user input (in an unchecked form). Such marking of values (but at run-time)
is also called tainting \cite{app-tainting-sql}.

Similarly, if we had \ident{password} or \ident{creditCard} variables in a client/server web application,
these should be annotated as sensitive and it should not be possible to send their values over an
unsecured network connection.

In another context, when working with data (\eg~in data journalism), it would be desirable to 
track meta-data about the quality and the source of the data. For example, is the source trustworthy?
Is the data up-to-date? Such meta-data could propagate to the result and tell us important 
information about the calculated results.

% --------------------------------------------------------------------------------------------------

\subsection{Context-awareness \#4: Checking array access patterns}

The last example leaves the topic of cross-platform and distributed computing. We focus on 
checking how arrays are accessed. This is a simpler version of the data-flow programming
examples used later in the thesis.

Consider a simple programming language with arrays where $n^{\textnormal{th}}$ element of an
array \ident{arr} is accessed using $\ident{arr}[n]$. Furthermore, we focus on performing local
transformations and we assume that the keyword $\kvd{cursor}$ returns the \emph{current} location
in the array.

The following example implements a simple one-dimensional cellular automata, reading from the
\ident{input} array and writing to \ident{output}:
%
\begin{equation*}
\begin{array}{l}
\kvd{let}~\ident{sum}=\ident{input}[\kvd{cursor} - 1] ~+~ \ident{input}[\kvd{cursor}] ~+~ \ident{input}[\kvd{cursor} + 1]\\[-0.25em]
\kvd{if}~\ident{sum}=2 ~||~ (\ident{sum}=1 ~\&\&~ \ident{input}[\kvd{cursor} - 1] = 0)\\[-0.25em] 
\kvd{then}~\ident{output}[\kvd{cursor}] \leftarrow 1 ~\kvd{else}~\ident{output}[\kvd{cursor}] \leftarrow 0
\end{array}
\end{equation*}
%
In this example, we use the term \emph{context} to refer to the values in the array around the 
current location provided by \kvd{cursor}. The interesting question is, how much of the context
(\ie~how far in the array) does the program access.

This is contextual information attached to individual (array) variables. 
In the above example, we want to track that \ident{input} is accessed in the range $\langle-1, 1 \rangle$
while \ident{output} is accessed in the range $\langle 0, 0 \rangle$. When calculating the ranges,
we need to be able to compose ranges $\langle-1,-1$, $\langle 0,0 \rangle$ and $\langle 1,1 \rangle$
(based on the accesses on the first line).

The information about access patterns can be used to efficiently compile the computation
(as we know which sub-range of the array might be accessed) and it also allows better handling
of boundaries. For example, wrap-around behaviour we could pad the \ident{input} with a known
number of elements from the other side of the array.

% ==================================================================================================

\section{Towards context-aware languages}

The four examples presented in the previous section cover different notions of \emph{context}.
The context can be viewed as execution environment, capabilities provided by the environment 
or input and meta-data about the input.

The different notions of context can be broadly classified into two categories -- those that
speak about the environment and those that speak about individual inputs (variables). In this
thesis, we refer to them as \emph{flat coeffects} and \emph{structural coeffects}, respectively:

\begin{itemize}
\item \textbf{Flat coeffects} represent additional data, resources and meta-data that are 
   available in the execution environment (regardless of how they are accessed in a program).
   Examples include resources such as GPS sensors and battery status (on a phone), databases 
   (on the server), or framework version.

\item \textbf{Structural coeffects} capture additional meta-data related to inputs. This can include
   provenance (source of the input value), usage information (how often is the value
   accessed and in what ways) or security information (whether it contain sensitive data or not). 
\end{itemize}
%
This thesis follows the tradition of statically typed programming languages. As such, we
attempt to capture such contextual information in the type system of context-aware programming
languages. The type system should provide both safety guarantees (as in the first three examples)
and also static analysis useful for optimization (as in the last example).

Although the main focus of this thesis is on the underlying theory of \emph{coeffects} and on
their structure, the following section briefly demonstrates the features that a practical
context-aware language, based on the theory of coeffects, can provide.

% --------------------------------------------------------------------------------------------------

\subsection{Context-aware language in action}

As an example, consider a news reader consisting of server-side (which stores the news in a 
database) and a number of clients applications for popular platforms (Android, Windows Phone, etc.). 
A simplified code excerpt that might appear somewhere in the implementation is shown in
Figure~\ref{fig:introduction-sample}.

We assume that the language supports cross-compilation and splits the single program into three
components: one for the server-side and two for the client-side, for iPhone and Windows platforms, 
respectively. The cross-compilation could be done in a way similar to Links \cite{app-distributed-links},
but we do not require explicit annotations specifying the target platform.

If we were writing the code using current main-stream technologies, we would have to create three 
completely separate components. The server-side would include the \ident{fetchNews} function, which
queries the database. The iPhone version would include \ident{fetchLocalNews}, which gets the current
GPS location and performs a call to the remote server and \ident{iPhoneMain}, which constructs the
user-interface. For Windows, we would also need \ident{fetchLocalNews}, but this time with 
\ident{windowsMain}. When using a language that can be compiled for all of the platforms, we would need
a number of \prepk{\#if} blocks to delimit the platform-specific parts.

To support cross-compilation, the language needs to be context-aware. Each of the function has a 
number of context requirements. The \ident{fetchNews} function needs to have access to a database;
\ident{fetchLocalNews} needs access to a GPS sensor and to a network (to perform the remote call).
However, it does not need a specific platform -- it can work on both iPhone and Windows. The last
two platform-specific functions inherit the requirements of \ident{fetchLocalNews} and additionally
also require a specific platform.

% --------------------------------------------------------------------------------------------------

\begin{figure}
\begin{equation*}
\begin{array}{l}
\kvd{let}~\ident{fetchNews}(\ident{loc}) =\\[-0.25em]
\hspace{1em} \kvd{let}~\ident{cmd} = \ident{sprintf}~\str{SELECT * FROM News WHERE Location='\%s'}~\ident{loc}\\[-0.25em]
\hspace{1em} \ident{query}(\ident{cmd}, \ident{password})\\[-0.25em]
\\[-1.0em]
\kvd{let}~\ident{fetchLocalNews}() =\\[-0.25em]
\hspace{1em} \kvd{let}~\ident{loc} = \ident{gpsLocation}()\\[-0.25em]
\hspace{1em} \kvd{remote}~\ident{fetchNews}(\ident{loc})\\[-0.25em]
\\[-1.0em]
\kvd{let}~\ident{iPhoneMain}() = \\[-0.25em]
\hspace{1em} \ident{createCocoaListing}(\ident{fetchLocalNews})\\[-0.25em]
\\[-1.0em]
\kvd{let}~\ident{windowsMain}() = \\[-0.25em]
\hspace{1em} \ident{createMetroListing}(\ident{fetchLocalNews})
\end{array}
\end{equation*}

\vspace{1em}
\noindent\makebox[\linewidth]{\rule{\textwidth}{0.5pt}} 
\vspace{-1.5em}
\caption{News reader implemented in a context-aware language}
\label{fig:introduction-sample}
\end{figure}

% --------------------------------------------------------------------------------------------------

\subsection{Understanding context with types}

The approach advocated in this thesis is to track information about context requirements using
the type system. To make this practical, the system needs to provide at least partial support for 
automatic type inference, as the information about context requirements makes the types more complex.
An inspiring example might be the F\# support for units of measure \cite{app-units} -- the user has
to explicitly annotate constants, but the rest of the information is inferred automatically.

Furthermore, integrating contextual information into the type system can provide information for
modern developer tools. For example, many editors for F\# display inferred types when placing mouse
pointer over an identifier. For \ident{fetchLocalNews}, the tip could appear as follows:
%
\definecolor{tooltipy}{rgb}{0.98,0.97,0.76}
\definecolor{tooltipb}{rgb}{0.7,0.7,0.7}
\newcommand{\introfunc}[3]{ \ident{#1 \hspace{-1em}\footnotesize{@} \{ #2 \}} $\rightarrow$ \ident{#3}  }
\newcommand{\introvar}[2]{ \ident{#1 \hspace{-1em}\footnotesize{@} #2 } }
\vspace{0.5em}
\begin{equation*}
\hspace{5em}
\begin{array}{l}
\ident{\textbf{fetchLocalNews}}\\
\hspace{1em} \fcolorbox{tooltipb}{tooltipy}{
	\introfunc{unit}{gps, rpc}{(news~list)~async}
}
\end{array}
\end{equation*}
\vspace{0.5em}

Here, we use the notation $\tau_1 \,\ident{\hspace{-1em}\footnotesize{@}}\, c \rightarrow \tau_2$ to denote 
a function that takes an input of type $\tau_1$, produces a result of type $\tau_2$ and has additional 
context requirements specified by $c$. In the above example, the annotation $c$ is simply a set of 
required resources or capabilities. However, a more complex structure could be used as well, for example,
including the Android API level.

The following summary shows the types of the functions from the code sample in 
Figure~\ref{fig:introduction-sample}. These guide the code generation by specifying which 
function should be compiled for which of the platforms, but they also provide documentation
for the developers:

\begin{equation*}
\begin{array}{lcl}
\ident{password} &:& $\introvar{string}{sensitive}$\\[-0.25em]
\ident{fetchNews} &:& $\introfunc{location}{database}{news~list}$\\[0.5em]
\ident{gpsLocation} &:& $\introfunc{unit}{gps}{location}$\\[-0.25em]
\ident{fetchLocalNews} &:& $\introfunc{location}{gps, rpc}{news~list}$\\[0.5em]
\ident{iPhoneMain} &:& $\introfunc{unit}{ cocoa, gps, rpc}{unit}$\\[-0.25em]
\ident{windowsMain} &:& $\introfunc{unit}{ windows, gps, rpc}{unit}$
\end{array}
\end{equation*}

As mentioned earlier, the concrete syntax used here is just for illustration. Furthermore,
some information could even be mapped to other visual representations -- for example, differently
coloured backgrounds for platform-specific functions. The key point is that the type provides a
number of useful information:

\begin{itemize}
\item The \ident{password} variable is available in the context (we assume it has been declared 
  earlier), but is marked as sensitive, which restricts how it can be used. In particular, we 
  cannot return it as a result of a function that is called via a remote call (\eg~\ident{fetchNews})
  as that would leak sensitive data over an unsecured connection.

\item The \ident{fetchNews} function requires database access and so it can only run on the server-side
  (or on a thick client with local copy of the database, such as a desktop computer with an offline mode).

\item The \ident{gpsLocation} function accesses the GPS sensor and since we call it in 
  from \texttt{fetchLocalNews}, this function also requires GPS (the requirement is propagated 
  automatically).

\item We can compile the program for two client-side platforms - the entry points are \ident{iPhoneMain}
  and \ident{windowsMain} and require Cocoa and Windows user-interface libraries, together with
  GPS and the ability to perform remote calls over the network.
\end{itemize}

The details of how the cross-compilation would work are out of the scope of this thesis. However,
one can imagine that the compiler would take multiple sets of references (representing the different 
platforms), expose the \emph{union} of the functions, but annotate each with the required platform.
Then, it would produce multiple different binaries -- here, one for the server-side (containing
\ident{fetchNews}), one for iPhone and one for Windows.

In this scenario, the main benefit of using an integrated context-aware language would be the 
ability to design appropriate abstractions using standard mechanisms of the language. For 
cross-compilation, we can structure code using functions, rather than relying on \prepk{\#if}
directives. Similarly, the splitting between client-side, server-side and shared code can be
done using ordinary functions and modules -- rather than having to split the application into 
separate independent libraries or projects.

The purpose of this section was to show that many modern programs rely on the context in which 
they execute in non-trivial ways. Thus designing context-aware languages is an important 
practical problem for language designers. The sample serves more as a motivation than as a 
technical background for this thesis. We explore more concrete examples of properties that 
can be tracked using the systems developed in this thesis in Chapter~\ref{ch:applications}.

% --------------------------------------------------------------------------------------------------

\section{Theory of context dependence}

The previous section introduced the idea of context-aware languages from the practical perspective.
As already discussed, we approach the problem from the perspective of statically typed programming
languages. This section outlines how can contextual information be integrated into the 
standard framework of static typing. This section is intended only as an informal overview and the
related work is discussed in Chapter~\ref{ch:pathways}.

\paragraph{Type systems.}
A type system is a form of static analysis that is usually specified by \emph{typing judgements}
such as $\Gamma \vdash e : \tau$. The judgement specifies that, given some variables described by
the context $\Gamma$, the expression $e$ has a type $\tau$. 
The variable context $\Gamma$ is necessary to determine the type of expressions. Consider an 
expression $x + y$. In many languages, including Java, C\# and F\#, the type could be \ident{int}
or \ident{float}, depending on the types of the variables. For example, the following is a valid
typing judgement in F\#:
%
\begin{equation*}
x\-:\-\ident{int},~y\-:\-\ident{int} \vdash x+y : \ident{int}
\end{equation*}
%
This judgement assumes that the type of both $x$ and $y$ is \ident{int} and so the result must also
be \ident{int}. The expression would also be typeable in a context $x\-:\-\ident{int},~y\-:\-\ident{int}$,
but not, for example, in a context where $x$ has a type \ident{unit}.

\paragraph{Tracking evaluation effects.}
Type systems can be extended in numerous ways. The types can be more precise, for example, by specifying
the range of an integer. However, it is also possible to track what program \emph{does} when executed.
In ML-like languages, the following is a valid judgement:
%
\begin{equation*}
x\-:\-\ident{int} \vdash \ident{print}~x : \ident{unit}
\end{equation*}
%
The judgement states that the expression $\ident{print}~x$ has a type \ident{unit}. This is correct, 
but it ignores the important fact that the expression has a \emph{side-effect} and prints a number
to the console. In purely functional languages, this would not be possible. For example, in Haskell,
the type would be $\ident{IO unit}$ meaning that the result is a \emph{computation} that performs I/O
effects and then returns \ident{unit} value.

Another option for tracking effects is to extend the judgement with additional information 
about the effects. The judgement in a language with effect system would look as follows:
%
\begin{equation*}
x\-:\-\ident{int} \vdash \ident{print}~x : \ident{unit~\&~\{ console \}}
\end{equation*}
%
Effect systems add \emph{effect annotation} as another component of the typing judgement. In the above
example, the return type is \ident{unit}, but the effect annotation informs us that the expression
also accesses \ident{console} as part of the evaluation. To track such information, the compiler needs
to understand the effects of primitive built-in functions -- such as \ident{print}.

The crucial part of type systems is dealing with different forms of composition. For example, assume
we have a function \ident{read} that reads from the console and a function \ident{send} that sends
data over the network. In that case, the type system should correctly infer that the effects of an 
expression $\ident{send}(\ident{read}())$ are $\{ \ident{console}, \ident{network} \}$.

Effect systems are an established idea, but they are suitable only for tracking properties of a
certain kind. They can be used for properties that describe how programs \emph{affect} the environment.
For context-aware languages, we instead need to track what programs \emph{require} from the environment.

\paragraph{Tracking context requirements.}

The systems for tracking of context requirements developed in this thesis are inspired by the idea
of effect systems. To demonstrate our approach, consider the following call from the sample program
shown earlier -- first using standard ML-like type system:
%
\begin{equation*}
\ident{password}\-:\-\ident{string},~\ident{cmd}\-:\-\ident{string} 
  \vdash \ident{query}(\ident{cmd}, \ident{password}) : \ident{news list}
\end{equation*}
%
The expression queries a database and gets back a list of news values as the result. Recall from the
earlier discussion that there are two contextual information that are desirable to track for this 
expression. First, the call to the \ident{query} primitive requires \emph{database access}. Second,
the \ident{password} argument needs to be marked as \emph{sensitive value} to avoid sending it over
an unsecure network connection.
The \emph{coeffect systems} developed in this thesis capture this information in the following way: 
%
\begin{equation*}
\begin{array}{l}
 ( \ident{password}\-:\-\ident{string \hspace{-1em}{\footnotesize @ sensitive}},
   \ident{cmd}\-:\-\ident{string})
 \hspace{-0.5em}
 \ident{\footnotesize @ \{ database \}} \vdash \\[-0.1em]
 \hspace{3em} \ident{query}(\ident{cmd}, \ident{password}) : \ident{news list}
\end{array}
\end{equation*}
%
Rather than attaching the annotation to the \emph{resulting type}, we attach them to the variable
context $\Gamma$. In other words, coeffect systems track more detailed information about the context,
not just the available variables. In the above example, it tracks meta-data about the variables
and annotates \ident{password} as sensitive. Furthermore, it tracks requirements about the execution
environment -- for example, that the execution requires an access to database.

The example demonstrates the two kinds of coeffect systems outlined earlier. The tracking of 
\emph{whole-context} information (such as environment requirements) is captured by the \emph{flat 
coeffect calculus} developed in Chapter~\ref{ch:flat}, while the tracking of \emph{per-variable} 
information is captured by the \emph{structural coeffect calculus} developed in Chapter~\ref{ch:structural}.

As mentioned earlier, it is well-known fact that \emph{effects} correspond to \emph{monads} and
languages such sa Haskell use monads to provide a limited form of effect system. An interesting 
observation made in this thesis is that \emph{coeffects}, or systems for tracking contextual
information, correspond to the category theoretical dual of monads called \emph{comonads}. The
details are explained when discussing the semantics of coeffects throughout the thesis.

% ==================================================================================================

\section{Thesis outline}

The key claim of this thesis is that programming languages need to provide better ways of capturing
how programs rely on the context in which they execute. This chapter shows why this is an important
problem. We looked at a number of properties related to context that are currently handled in 
ad-hoc and error-prone ways. Next, we considered the properties in a simplified, but realistic 
example of a client/server application for displaying local news.

Tracking of contextual properties may not be initially perceived as a major problem -- perhaps 
because we are so used to write code in certain ways that prevent us from seeing the flaws. The 
purpose of this chapter was to uncover the flaws and convince the reader that there should be a 
better solution. Finding the foundations of such better solution is the goal of this thesis:

\begin{itemize}
\item In Chapter~\ref{ch:pathways} we give an overview of related work. Most importantly, we show
  that the idea of context-aware computations can be naturally approached from a number of 
  directions developed recently in theories of programming languages. Chapter~\ref{ch:applications}
  follows by showing practical motivation for coeffects -- we look at a number of systems that 
  can be captured using the systems developed later.

\item In Chapter~\ref{ch:pathways} and Chapter~\ref{ch:applications}, we present the key novel
  contributions of this thesis. We develop the \emph{flat} and \emph{structural} calculi, show 
  how they capture important contextual properties and develop their categorical semantics using
  a notion based on comonads. Chapter~\ref{ch:unified} links the two systems into a single 
  formalism that is capable of capturing both flat and structural properties.

\item Related work is presented in Chapter~\ref{ch:pathways} and throughout the thesis, but 
  one important direction deserves further exploration. In Chapter~\ref{ch:meta}, we look at 
  a different approach to tracking contextual information that arises from modal logics. 
  Finally, Chapter~\ref{ch:conclusions} discusses approaches for implementing the presented theory
  in main-stream programming languages and concludes.
\end{itemize}