### 2014-06-02 text/AppendixA.tex
Syntactic properties (Chapter 4)

\chapter{Appendix A} 
\label{ch:appendix}

%---------------------------------------------------------------------------------------------------

\section{Internalized substitution}

\subsection{First transformation}

\begin{equation*}
(\kvd{glet}~x=e_1~\kvd{in}~e_2)~e_3 \leadsto \kvd{glet}~x=e_1~\kvd{in}~(e_2~e_3)
\end{equation*}

\begin{equation*}
\tyrule{app}
  { \tyrule{glet}
      { \coctx{\Gamma}{\cclrd{s}} \vdash e_1 : \tau_1 &
        \coctx{\Gamma, x\!:\!\tau_1}{\cclrd{r}} \vdash e_2 : \tau_3 \xrightarrow{\cclrd{t}} \tau_2 }
      { \coctx{\Gamma}{\cclrd{r}\;\cpar\;({\cclrd{s}\,\cseq\,\cclrd{r})}} \vdash \kvd{glet}~x=e_1~\kvd{in}~e_2 : \tau_3 \xrightarrow{\cclrd{t}} \tau_2 } &
    \coctx{\Gamma}{\cclrd{u}} \vdash e_3 : \tau_3  }
  { \coctx{\Gamma}{(\cclrd{r}\;\cpar\;(\cclrd{s}\,\cseq\,\cclrd{r})) \;\cpar\; (\cclrd{u}\,\cseq\,\cclrd{t}) } \vdash (\kvd{glet}~x=e_1~\kvd{in}~e_2)~e_3 : \tau_2 }
\end{equation*}

to

\begin{equation*}
\tyrule{glet}
  { \coctx{\Gamma}{\cclrd{s}} \vdash e_1 : \tau_1 &
    \tyrule{app}
      { \coctx{\Gamma, x\!:\!\tau_1}{\cclrd{r}} \vdash e_2 : \tau_3 \xrightarrow{\cclrd{t}} \tau_2 & 
        \coctx{\Gamma}{\cclrd{u}} \vdash e_3 : \tau_3 }
      { \coctx{\Gamma}{\cclrd{r}\;\cpar\;(\cclrd{u}\,\cseq\,\cclrd{t})} \vdash e_2~e_3 : \tau_2 } }
  { \coctx{\Gamma}{ (\cclrd{r}\;\cpar\;(\cclrd{u}\,\cseq\,\cclrd{t})) \;\cpar\; (\cclrd{s}\,\cseq\,(\cclrd{r}\;\cpar\;(\cclrd{u}\,\cseq\,\cclrd{t})))  }
      \vdash \kvd{glet}~x=e_1~\kvd{in}~(e_2~e_3) : \tau_2 }
\end{equation*}

meaning
\begin{equation*}
\begin{array}{l}
(\cclrd{r}\;\cpar\;(\cclrd{s}\,\cseq\,\cclrd{r})) \;\cpar\; (\cclrd{u}\,\cseq\,\cclrd{t}) =
\end{array}
\end{equation*}

\subsection{Second transformation}

Second transformation

\begin{equation*}
\tyrule{glet}
  { \coctx{\Gamma}{\cclrd{s}} \vdash e_s : \tau_s &
    \coctx{\Gamma, x\!:\!\tau_1}{\cclrd{r}} \vdash e_r : \tau_r &
    \coctx{\Gamma, x\!:\!\tau_1}{\cclrd{t}} \vdash e_t : \tau_t }
  { \coctx{\Gamma}{
      \cclrd{t}
      \;\cpar\;
      (  (\cclrd{r}\;\cpar\;(\cclrd{s}\,\cseq\,\cclrd{r}))
         \,\cseq\, 
         \cclrd{t} )
    } 
    \vdash \kvd{glet}~x_r=(\kvd{glet}~x_s=e_s~\kvd{in}~e_r)~\kvd{in}~e_t : \tau_t }
\end{equation*}

or

\begin{equation*}
\tyrule{glet}
  { \coctx{\Gamma}{\cclrd{s}} \vdash e_s : \tau_s &
    \coctx{\Gamma, x\!:\!\tau_1}{\cclrd{r}} \vdash e_r : \tau_r &
    \coctx{\Gamma, x\!:\!\tau_1}{\cclrd{t}} \vdash e_t : \tau_t }
  { \coctx{\Gamma}{
      (\cclrd{t}\;\cpar\;(\cclrd{r}\,\cseq\,\cclrd{t})) 
      \;\cpar\;
      (  \cclrd{s}
         \,\cseq\, 
         (\cclrd{t}\;\cpar\;(\cclrd{r}\,\cseq\,\cclrd{t})) )
    } 
    \vdash \kvd{glet}~x_s=e_s~\kvd{in}~(\kvd{glet}~x_r=e_r~\kvd{in}~e_t) : \tau_t }
\end{equation*}

\begin{equation*}
\begin{array}{l}
  \cclrd{t}
  \;\cpar\;
  (  (\cclrd{r}\;\cpar\;(\cclrd{s}\,\cseq\,\cclrd{r}))
     \,\cseq\, 
     \cclrd{t} ) = 
\\
  \cclrd{t}
  \;\cpar\;
  (\cclrd{r}\,\cseq\, \cclrd{t})
  \;\cpar\;
  (\cclrd{s}\,\cseq\,\cclrd{r}\,\cseq\, \cclrd{t}) = 
\\
  \cclrd{s}\,\cseq\,\cclrd{r}\,\cseq\, \cclrd{t} 
\end{array}
\end{equation*}

\begin{equation*}
\begin{array}{l}
  (\cclrd{t}\;\cpar\;(\cclrd{r}\,\cseq\,\cclrd{t})) 
  \;\cpar\;
  (  \cclrd{s}
     \,\cseq\, 
     (\cclrd{t}\;\cpar\;(\cclrd{r}\,\cseq\,\cclrd{t})) ) = 
\\     
  \cclrd{t}\;\cpar\;(\cclrd{r}\,\cseq\,\cclrd{t})
  \;\cpar\;
  (\cclrd{s} \,\cseq\, \cclrd{t})
  \;\cpar\;
  (\cclrd{s} \,\cseq\, \cclrd{r}\,\cseq\,\cclrd{t})  = 
\\     
  \cclrd{s} \,\cseq\, \cclrd{r}\,\cseq\,\cclrd{t}
\end{array}
\end{equation*}

require

\begin{equation*}
\cclrd{r}\;\cpar\;(\cclrd{r} \,\cseq\, \cclrd{s}) = \cclrd{r} \,\cseq\, \cclrd{s}
\end{equation*}