### 2014-09-05 text/Abstract.tex
Equational theory (Chapter 5) and Abstract

%!TEX root = ../main.tex
\pdfbookmark[1]{Abstract}{Abstract} % Bookmark name visible in a PDF viewer

\begingroup
\let\clearpage\relax
\let\cleardoublepage\relax
\let\cleardoublepage\relax

\chapter*{Abstract} % Abstract name
The development of programming languages needs to reflect important changes in the way
programs execute. In recent years, this included \eg~the development of parallel programming 
models (in reaction to the multi-core revolution) or improvements in data access technologies. 
This thesis is a response to another such revolution -- the diversification of devices and 
systems where programs run.

The key point made by this thesis is the realization that execution environment or
\emph{context} is fundamental for writing modern applications and that programming 
languages should provide abstractions for programming with context and verifying how 
it is accessed.

We identify a number of program properties that were not connected before, but model some notion 
of context. Our examples include tracking different execution platforms (and their versions) 
in cross-platform development, resources available in different execution environments (\eg~GPS
sensor on a phone and database on the server), but also more traditional notions such as 
variable usage (\eg~in liveness analaysis) or past values in stream-based data-flow programming.

Our first contribution is the discovery of the connection between the above examples and
their novel presentation -- in the form of calculi with type systems and formal semantics --
that highlights this relationship. Our second and third contribution are two \emph{coeffect}
calculi that capture commonalities in the presented examples. In particular, our 
\emph{flat} coeffect calculus models languages with contextual properties of the execution 
environment and our \emph{structural} coeffect calculus models languages where the contextual
properties are attached to the variable usage.

Although the focus of this thesis is on the syntactical properties of the presented 
systems, we also discuss their category-theoretical motivation. We introduce the notion of
an \emph{indexed} comonad (the dual of monads) and show how they provide semantics of the
presented two calculi.

\endgroup

\vfill