### 2014-11-17 text/Preface.tex

%!TEX root = ../main.tex
\pdfbookmark[1]{Preface}{Preface} % Bookmark name visible in a PDF viewer

\begingroup
\let\clearpage\relax
\let\cleardoublepage\relax
\let\cleardoublepage\relax

\chapter*{Declaration} % Abstract name

This dissertation is my own work and includes nothing which is the outcome of work done in 
collaboration with others except where specifically indicated.
The work presented in Chapter~\ref{ch:flat} and Chapter~\ref{ch:structural} has been previously published
in peer-reviewed papers \cite{coeffects-icalp13} and \cite{coeffects-icfp14}.

The published papers present the key ideas of the dissertation from Chapters~\ref{ch:flat} and 
Chapter~\ref{ch:structural} as well as the motivation discussed in Chapter~\ref{ch:intro}.
They present the flat and structural coeffect calculi and their equational theory as 
developed in this dissertation.

The categorical semantics presented in the published papers has been developed in collaboration and
differs from the one presented in this dissertation. The semantics 
presented here (Section~\ref{sec:flat-semantics} and Section~\ref{sec:struct-semantics}) 
has been developed by the author.

\endgroup

\vfill