### 2014-12-12 text/AppendixB.tex
Proofreading and English conventions (Chapter 4)

%!TEX root = ../main.tex

\chapter{Appendix B} 
\label{ch:appendix}

This appendinx provides additional details for some of the proofs for equational theory
of flat coeffect calculus from Chapter~\ref{ch:flat} and structural coeffect calculus
from Chapter~\ref{ch:structural}.

\section{Substitution for flat coeffects}
\label{sec:appendix-flat-cbn}
In Section~\ref{sec:flat-syntax-cbn}, we stated that, for a bottom-pointed flat coeffect
algebra (\ie~$\forall r \in \C \;.\; r \,\cgeq\, \cunit $), the call-by-name substitution 
preserves type if all operators of the flat coeffect algebra coincide (Lemma~\ref{thm:cbn-substitution-bot}).
This section provides the corresponding proof.

\begin{lemma*}[Bottom-pointed substitution]
In a bottom-pointed flat coeffect calculus with an algebra $(\C, \cseq, \cpar, \czip, \cunit, \czero, \cleq)$ 
where $\czip = \cseq = \cpar$ and the operation is also idempotent and commutative and
$\cclrd{r}\,\cleq\,\cclrd{r'} \Rightarrow \forall\cclrd{s}.\cclrd{r}\,\cseq\,\cclrd{s}\;\cleq\;\cclrd{r'}\,\cseq\,\cclrd{s}$ then:
%
\begin{equation*}
\begin{array}{l}
 \coctx{\Gamma}{\cclrd{S}} \vdash e_s : \tau_s \;\; \wedge \;\; 
 \coctx{\Gamma_1,  x : \tau_s, \Gamma_2}{ \cclrd{r}  } \vdash e_r : \tau_r\\
\quad \Rightarrow \;\; \coctx{\Gamma_1,\Gamma,\Gamma_2}{ \cclrd{r} \,\cseq\, \cclrd{S} } \vdash \subst{e_r}{x}{e_s} : \tau_r
\end{array}
\end{equation*}

\end{lemma*}
\begin{proof}
Assume that $\coctx{\Gamma}{\cclrd{S}} \vdash e_s : \tau_s$ and we are substituting a term $e_s$ for 
a variable $x$. Note that we use upper-case $\cclrd{S}$ to distinguish the coeffect of the expression
that is being substituted into an expression. Using structural induction over $\vdash$:

\paragraph{(var)} Given the following derivation using (\emph{var}):
\[
\inference
  { }
  {\coctx{\Gamma_1, y:\tau, \Gamma_2}{\cunit} \vdash y : \tau }
\]
There are two cases depending on whether $y$ is the variable $x$ or not:
\begin{itemize}
\item[--] If $y=x$ then also $\tau = \tau_s$ and thus $\subst{y}{x}{e_s} = e_s$. Using the assumption,
implicit weakening and the fact that $\cunit$ is a unit of $\cseq$:
\[
\inference
 {\inference 
  {\coctx{\Gamma}{\cclrd{S}} \vdash \subst{y}{x}{e_s} : \tau_s }
  {\coctx{\Gamma_1, \Gamma, \Gamma_2}{\cclrd{S}} \vdash \subst{y}{x}{e_s} : \tau } }
 {\coctx{\Gamma_1, \Gamma, \Gamma_2}{\cunit\,\cseq\,\cclrd{S}} \vdash \subst{y}{x}{e_s} : \tau } 
\]
\item[--] If $y\neq x$ then $\subst{y}{x}{e_s} = y$. Using the fact that $\cunit$ is the bottom element
and sub-coeffecting:
\[
\inference
  {\coctx{\Gamma_1, y:\tau, \Gamma_2}{\cunit} \vdash y : \tau }
  {\coctx{\Gamma_1, y:\tau, \Gamma_2}{\cunit\,\cseq\,\cclrd{S}} \vdash y : \tau }
\]
\end{itemize}

\paragraph{(const)} Similar to the (\emph{var}) case when the variable is not substituted.

\paragraph{(sub)} Given the following typing derivation using (\emph{sub}):
\[
\inference
  {\coctx{\Gamma_1,x:\tau_s,\Gamma_2}{\cclrd{r'}} \vdash e : \tau }
  {\coctx{\Gamma_1,x:\tau_s,\Gamma_2}{\cclrd{r}} \vdash e : \tau }\quad\quad(\cclrd{r'} \cleq \cclrd{r})
\]
From the induction hypothesis, we have that
$\coctx{\Gamma_1,\Gamma,\Gamma_2}{\cclrd{r'}\,\cseq\,\cclrd{S}} \vdash \subst{e}{x}{e_s} : \tau$.
The condition on $\cleq$ means that $\cclrd{r'}\,\cseq\,\cclrd{S}\;\cleq\;\cclrd{r}\,\cseq\,\cclrd{S}$
and so we can apply the (\emph{sub}) rule to obtain
$\coctx{\Gamma_1,\Gamma,\Gamma_2}{\cclrd{r}\,\cseq\,\cclrd{S}} \vdash \subst{e}{x}{e_s} : \tau$.

\paragraph{(abs)} Given the following typing derivation using (\emph{abs}):
\[
\inference
  {\coctx{\Gamma_1,x:\tau_s,\Gamma_2, y:\tau_1}{\cclrd{r}\;\czip\;\cclrd{s}} \vdash e : \tau_2}
  {\coctx{\Gamma_1,x:\tau_s,\Gamma_2}{\cclrd{r}} \vdash \lambda y.e : \tau_1 \xrightarrow{\cclrd{s}} \tau_2 }
\]
Assume w.l.o.g. that $x\neq y$. From the induction hypothesis, we have that:
\[
\coctx{\Gamma_1,\Gamma,\Gamma_2, y:\tau_1}{ \cclrd{r} \,\cseq\, \cclrd{s} } \vdash \subst{e}{x}{e_s} : \tau_2
\]
Now using the fact that $\czip\,=\,\cseq$, associativity and commutativity and (\emph{abs}):
\[
\inference
 {\inference
  {\coctx{\Gamma_1,\Gamma,\Gamma_2, y:\tau_1}{ (\cclrd{r} \,\czip\, \cclrd{s})\,\cseq\,\cclrd{S} } \vdash \subst{e}{x}{e_s} : \tau_2}
  {\coctx{\Gamma_1,\Gamma,\Gamma_2, y:\tau_1}{ (\cclrd{r} \,\cseq\, \cclrd{S})\,\czip\,\cclrd{s} } \vdash \subst{e}{x}{e_s} : \tau_2}}
 {\inference
  {\coctx{\Gamma_1,\Gamma,\Gamma_2}{\cclrd{r} \,\cseq\, \cclrd{S}} \vdash \lambda y.(\subst{e}{x}{e_s}) : \tau_1 \xrightarrow{\cclrd{s}} \tau_2}
  {\coctx{\Gamma_1,\Gamma,\Gamma_2}{\cclrd{r} \,\cseq\, \cclrd{S}} \vdash \subst{(\lambda y.e)}{x}{e_s} : \tau_1 \xrightarrow{\cclrd{s}} \tau_2} }
\]

\paragraph{(app)} Given the following typing derivation using (\emph{app}):
\[
\inference
  {\coctx{\Gamma_1,x:\tau_s,\Gamma_2}{\cclrd{r}} \vdash e_1 : \tau_1 \xrightarrow{\cclrd{t}} \tau_2 &
   \coctx{\Gamma_1,x:\tau_s,\Gamma_2}{\cclrd{s}} \vdash e_2 : \tau_1 }
  {\coctx{\Gamma_1,x:\tau_s,\Gamma_2}{\cclrd{r} \;\cpar\; (\cclrd{s} \,\cseq\, \cclrd{t})} \vdash e_1~e_2 : \tau_2}
\]
From the induction hypothesis, we have that:
\[
\begin{array}{l}
 \coctx{\Gamma_1, \Gamma, \Gamma_2}{\cclrd{r}\,\cseq\,\cclrd{S}} \vdash \subst{e_1}{x}{e_s} : \tau_1 \xrightarrow{\cclrd{t}} \tau_2 \\
 \coctx{\Gamma_1, \Gamma, \Gamma_2}{\cclrd{s}\,\cseq\,\cclrd{S}} \vdash \subst{e_2}{x}{e_s} : \tau_1
\end{array}\quad(*)
\]
Now using (\emph{app}) rule and the fact that $\cpar\,=\,\cseq$, associativity, commutativity and idempotence
(note that all three properties are needed):
\[
\inference
 { (*) }
 {\inference
   { \coctx{\Gamma_1,\Gamma,\Gamma_2}{(\cclrd{r}\,\cseq\,\cclrd{S}) \;\cpar\; ((\cclrd{s}\,\cseq\,\cclrd{S}) \,\cseq\, \cclrd{t})} 
       \vdash \subst{e_1}{x}{e_s}~\subst{e_2}{x}{e_s} : \tau_2}
   { \coctx{\Gamma_1,\Gamma,\Gamma_2}{(\cclrd{r} \;\cpar\; (\cclrd{s} \,\cseq\, \cclrd{t}))\;\cseq\;\cclrd{S}} 
       \vdash \subst{(e_1~e_2)}{x}{e_s} : \tau_2} }
\]

\paragraph{(let)} Given the following typing derivation using (\emph{let}):
\[
\inference
  { \coctx{\Gamma_1,x:\tau_s,\Gamma_2}{\cclrd{r}} \vdash e_1 : \tau_1 &
    \coctx{\Gamma_1,x:\tau_s,\Gamma_2, y:\tau_1}{\cclrd{s}} \vdash e_2 : \tau_2}
  {\coctx{\Gamma_1,x:\tau_s,\Gamma_2}{\cclrd{s} \;\cpar\; (\cclrd{s} \,\cseq\, \cclrd{r})} \vdash \kvd{let}~y=e_1~\kvd{in}~e_2 : \tau_2 }
\]
From the induction hypothesis, we have that:
\[
\begin{array}{l}
 \coctx{\Gamma_1, \Gamma, \Gamma_2}{\cclrd{r}\,\cseq\,\cclrd{S}} \vdash \subst{e_1}{x}{e_s} : \tau_1 \\
 \coctx{\Gamma_1, \Gamma, \Gamma_2, y:\tau_1}{\cclrd{s}\,\cseq\,\cclrd{S}} \vdash \subst{e_2}{x}{e_s} : \tau_2
\end{array}\quad(\dagger)
\]
Now using (\emph{let}) rule and similarly to the (\emph{app}) case:
\[
\inference
 { (\dagger) }
 { \inference
   { \coctx{\Gamma_1, \Gamma, \Gamma_2}{(\cclrd{s}\,\cseq\,\cclrd{S}) \;\cpar\; ((\cclrd{s}\,\cseq\,\cclrd{S}) \,\cseq\, (\cclrd{r}\,\cseq\,\cclrd{S}))} 
         \vdash \kvd{let}~y=\subst{e_1}{x}{e_s}~\kvd{in}~\subst{e_2}{x}{e_s} : \tau_2 } 
   { \coctx{\Gamma_1, \Gamma, \Gamma_2}{(\cclrd{s} \;\cpar\; (\cclrd{s} \,\cseq\, \cclrd{r})) \,\cseq\, \cclrd{S}} 
         \vdash \subst{(\kvd{let}~y=e_1~\kvd{in}~e_2)}{x}{e_s} : \tau_2 } }
\]
\end{proof}