### 2014-12-12 text/Chapter04.tex

%!TEX root = ../main.tex

%\begin{flushright}{\slshape    
%Even the most puritanical rationalist will be forced to stop arguing and \\
%use propaganda [...] because the psychological conditions have disappeared \\
%that allowed effective argument and therefore influence over the others.} \\ \medskip
%--- Paul Feyerabend, \emph{Against Method} \cite{philosophy-feyerabend}
%\end{flushright}
%\vspace{2em}

% ==================================================================================================

\chapter{The flat coeffect calculus} 
\label{ch:flat}

Successful programming language abstractions need to generalize a wide range of recurring
problems while capturing the key commonalities. These two aims are typically in opposition -- 
more general abstractions are less powerful, while less general abstractions cannot be
used as often.

In the previous chapter, we outlined a number of systems that capture how computations
access the environment in which they are executed. We identified two kinds of systems --
\emph{flat systems} capturing whole-context properties and \emph{structural systems} capturing 
per-variable properties. As we show in Chapter~\ref{ch:unified}, the systems can be unified 
using a single abstraction. This is useful when implementing and composing the systems, but such 
abstraction is  \emph{less powerful} -- \ie~its generality hides useful properties that we can see 
when we consider the systems separately. For this reason, we discuss \emph{flat coeffects} and
\emph{structural coeffects} separately in this and the next chapter.

\paragraph{Chapter structure and contributions}
\begin{itemize}
\item[--] We present a \emph{flat coeffect calculus} as a type system that is parameterized by a 
  \emph{flat coeffect algebra} (Section~\ref{sec:flat-calculus}). We show that the system can be 
  instantiated to obtain three of the systems discussed in previous section (implicit parameters, 
  liveness and data-flow).

\item[--] We give the equational theory of the calculus and discuss type-pre\-ser\-vation for call-by-name
  and call-by-value reduction (Section~\ref{sec:flat-syntax}). We also extend the calculus
  with subtyping and pairs (Section~\ref{sec:flat-exts}). These two sections motivate the laws
  of the flat coeffect algebra.

\item[--] We present the semantics of the calculus in terms of \emph{indexed comonads}, which are a
  generalization of comonads, a category-theoretical dual of monads (Section~\ref{sec:flat-semantics}).
  This provides a deeper insight into how (and why) the calculus works and shows an 
  intriguing link with effects.
\end{itemize}

% ==================================================================================================
%                                                                                      
%     ###           #                       #                 #       #                 
%      #            #                       #                 #                         
%      #    # ##   ####   # ##    ###    ## #  #   #   ###   ####    ##     ###   # ##  
%      #    ##  #   #     ##  #  #   #  #  ##  #   #  #   #   #       #    #   #  ##  # 
%      #    #   #   #     #      #   #  #   #  #   #  #       #       #    #   #  #   # 
%      #    #   #   #  #  #      #   #  #  ##  #  ##  #   #   #  #    #    #   #  #   # 
%     ###   #   #    ##   #       ###    ## #   ## #   ###     ##    ###    ###   #   # 
%                                                                                     
% ==================================================================================================

\section{Introduction}
\label{sec:flat-intro}

In the previous chapter, we looked at three important examples of systems that track whole-context 
properties. The type systems for whole-context liveness (Section~\ref{sec:applications-flat-live}) 
and whole-context data-flow (Section~\ref{sec:applications-flat-dataflow}) have a very similar 
structure. First, their lambda abstraction duplicates the requirements. Given a body with context 
requirements $\cclrd{r}$, the declaration site context \emph{as well as} the function arrow are
annotated with $\cclrd{r}$. Second, their application arises from the combination of \emph{sequential} 
and \emph{pointwise} composition.

The system for tracking of implicit parameters and similar (Section~\ref{sec:applications-flat-impl})
differ in two ways. In lambda abstraction, they split the context requirements between the 
declaration site and the call site and they use only a single operator on the indices, typically $\cup$.

Despite the differences, the systems fit the same unified framework. This becomes apparent when
we consider the categorical structure (Section~\ref{sec:flat-semantics}). However, rather than 
starting from the semantics, we first explain how the systems can be unified syntactically
(Section~\ref{sec:flat-calculus-lambda}) and then provide the semantics as a justification.

%---------------------------------------------------------------------------------------------------

\subsection{Related work}

The development in this chapter can be seen as a counterpart to the well-known development of 
\emph{effect systems} \cite{effects-gifford} and the use of \emph{monads} \cite{monad-notions}
in programming languages. The syntax and type system of the flat coeffect calculus follows 
asimilar style as effect systems \cite{effects-polymorphic,effects-talpin-et-al}, but differs
in the structure, as explained in the previous chapter, most importantly in lambda abstraction
(the relationship with monads is further discussed in Section~\ref{sec:flat-related}).

Wadler and Thiemann \cite{monads-effects-marriage} famously showed a correspondence between effect 
systems to monads (a topic that has been also discussed by Atkey \cite{monads-parameterised-notions}
and recently revisited by Vazou an Leijen \cite{monads-effects-remarriage}). This line of work
relates effectful functions $\tau_1 \xrightarrow{\sigma} \tau_2$ to monadic computations 
$\tau_1 \rightarrow M^\sigma \tau_2$. In this chapter, we show a similar correspondence between 
\emph{coeffect systems} and \emph{comonads}. However, due to the asymmetry of $\lambda$-calculus, 
defining the semantics in terms of comonadic computations is not a simple mechanical dualization
of the work on effect systems and monads.

The main purpose of the comonadic semantics presented in this chapter is to provide a motivation 
for the typing of the flat coeffect calculus. Our approach is inspired by the work of Uustalu and
Vene \cite{comonads-notions} who present the semantics of contextual computations (mainly for
data-flow) in terms of comonadic functions $C \tau_1 \rightarrow \tau_2$. Our \emph{indexed 
comonads} annotate the structure with information about the required context, \ie~$C^{\cclrd{r}} \tau_1 \rightarrow \tau_2$.
This is similar to the recent work on \emph{parameterized monads} by Katsumata \cite{monads-parametric}.

% ==================================================================================================
%                                                            
%       ###           ##                   ##                 
%      #   #           #                    #                 
%      #       ###     #     ###   #   #    #    #   #   ###  
%      #          #    #    #   #  #   #    #    #   #  #     
%      #       ####    #    #      #   #    #    #   #   ###  
%      #   #  #   #    #    #   #  #  ##    #    #  ##      # 
%       ###    ####   ###    ###    ## #   ###    ## #  ####  
%                                                           
% ==================================================================================================

\section{Flat coeffect calculus}
\label{sec:flat-calculus}

The flat coeffect calculus is defined in terms of \emph{flat coeffect algebra}, which defines
the structure of context annotations, such as $\cclrd{r}, \cclrd{s}, \cclrd{t}$. These can be
sets of implicit parameters, versions represented as integers or other values. The expressions of 
the calculus are those of the $\lambda$-calculus with \emph{let} binding; assuming $\alpha$ ranges 
over base types, the types of the calculus are defined as:
%
\begin{equation*}
\begin{array}{rcl}
e &::=& x \sep \lambda x.e \sep e_1~e_2 \sep \kvd{let}~x = e_1~\kvd{in}~e_2\\
\tau &::=& \alpha \sep \tau_1 \xrightarrow{\cclrd{r}} \tau_2
\end{array}
\end{equation*}
%
We discuss subtyping and pairs in Section~\ref{sec:flat-exts}. The type $\tau_1 \xrightarrow{\cclrd{r}} \tau_2$
represents a function from $\tau_1$ to $\tau_2$ that requires additional context $\cclrd{r}$.
It can be viewed as a pure function that takes $\tau_1$ \emph{with} or \emph{wrapped in} a 
context $r$.

In the categorical semantics, the function $\tau_1 \xrightarrow{\cclrd{r}} \tau_2$ is modelled
by a morphism $C^{\cclrd{r}} \tau_1 \rightarrow \tau_2$. However, the object $C^{\cclrd{r}}$
does not itself exist as a syntactical value. This is because we use comonads to define the 
\emph{semantics} rather than \emph{embedding} them into the language as in the meta-language
approaches (the distinction has been discussed in Section~\ref{sec:path-sem-langs}). The annotations
$\cclrd{r}$ are formed by an algebraic structure discussed next.

%---------------------------------------------------------------------------------------------------

\subsection{Reconciling lambda abstraction}
\label{sec:flat-calculus-lambda}

Recall the lambda abstraction rules for the implicit parameters system (annotating the context
with sets of required parameters) and the data-flow system (annotating the context with the
number of past required values):
%
\begin{equation*}
\tyrule{param}
  {\coctx{\Gamma, x:\tau_1}{\aclrd{r} \cup \aclrd{s}} \vdash e : \tau_2}
  {\coctx{\Gamma}{\aclrd{r}} \vdash \lambda x.e : \tau_1 \xrightarrow{\aclrd{s}} \tau_2 }
\;
\tyrule{df1}
  {\coctx{\Gamma, x:\tau_1}{\aclrd{n}} \vdash e : \tau_2}
  {\coctx{\Gamma}{\aclrd{n}} \vdash \lambda x.e : \tau_1 \xrightarrow{\aclrd{n}} \tau_2 }
\end{equation*}

In order to capture both systems using a single calculus, we need a way of unifying the two
systems. For the data-flow system, this can be achieved by over-approximating the number of 
required past elements:
%
\begin{equation*}
\tyrule{df2}
  {\coctx{\Gamma, x:\tau_1}{\textnormal{min}(\aclrd{n}, \aclrd{m})} \vdash e : \tau_2}
  {\coctx{\Gamma}{\aclrd{n}} \vdash \lambda x.e : \tau_1 \xrightarrow{\aclrd{m}} \tau_2 }
\end{equation*}
%
The rule (\emph{df1}) is admissible in a system that includes the (\emph{df2}) rule. 
Furthermore, if we include sub-typing rule (on annotations of functions) and sub-coeffecting rule (on 
annotations of contexts), then the reverse is also true -- because 
$\textit{min}(\aclrd{n}, \aclrd{m}) \leq \aclrd{m}$ and $\textit{min}(\aclrd{n}, \aclrd{m}) \leq \aclrd{n}$.
In other words (\emph{df1}) is more precise, but (\emph{df2}) gives a sound over-approximation
with a structure that can be unified with (\emph{param}).

%---------------------------------------------------------------------------------------------------

\subsection{Understanding flat coeffects}
\label{sec:flat-calculus-undestanding}

Before looking at the type system in Figure~\ref{fig:flat-types}, let us clarify how the rules
should be understood. The coeffect calculus provides both analysis of context dependence (type 
system) and semantics for context (how it is propagated). These two aspects provide different
ways of reading the judgements $\coctx{\Gamma}{\cclrd{r}} \vdash e : \tau$ and the typing rules
used to define it.

\begin{itemize}
\item \textsc{Analysis of context dependence.}
Syntactically, coeffect annotations $\cclrd{r}$ model \emph{context requirements}. This means
we can over-approximate them and require more than is actually needed at runtime.

Syntactically, the typing rules should be read top-down. In function application, the context 
requirements of multiple assumptions (arising from two sub-expressions) are \emph{merged}; in 
lambda abstraction, the requirements of a single expression (the body) are split between
the declaration site and the call site.

\item \textsc{Semantics of context passing.}
Semantically, coeffect annotations $\cclrd{r}$ mo\-del \emph{contextual capabilities}. This means
that we can throw away capabilities, if a sub-expression requires fewer than we 
currently have.

Semantically, the typing rules should be read bottom-up. In application, the capabilities 
provided to the term $e_1~e_2$ are \emph{split} between the two sub-expressions; in abstraction,
the capabilities provided by the call site and declaration site are \emph{merged} and passed
to the body.
\end{itemize}

The reason for this asymmetry follows from the fact that the context appears in a \emph{negative
position} in the semantic model (Section~\ref{sec:flat-semantics}). It means that we need to be
careful about using the words \emph{split} and \emph{merge}, because they can be read as meaning
exactly the opposite things. To disambiguate, we always use the term \emph{context requirements} 
when using the syntactic view, especially in the rest of Section~\ref{sec:flat-calculus}, and 
\emph{context capabilities} or just \emph{available context} when using the semantic view, 
especially in Section~\ref{sec:flat-semantics}.

%---------------------------------------------------------------------------------------------------

\subsection{Flat coeffect algebra}
\label{sec:flat-calculus-algebra}

To make the flat coeffect system general enough, the algebra consists of three operations.
Two of them, $\cseq$ and $\cpar$, represent the \emph{sequential} and \emph{pointwise} composition, 
which are mainly used in function application. The third operator, $\czip$ is used in lambda 
abstraction and represents \emph{splitting} of context requirements (or, semantically, \emph{merging} 
of available context capabilities).

In addition to the three operations, we also require two special values used to annotate
variable access and constant access and a relation that defines the ordering.

\begin{definition}
A \emph{\cclrd{flat coeffect algebra}} $(\C, \cseq, \cpar, \czip, \cunit, \czero, \cleq)$ is a set 
$\C$ together with elements $\cunit, \czero \in \C$, relation $\cleq$ and binary operations 
$\cseq, \cpar, \czip$ such that $(\C, \cseq, \cunit)$ is a monoid, $(\C, \cpar, \czero)$ is an
idempotent monoid, $(\C, \cleq)$ is a pre-order and $(\C, \czip)$ is a band (idempotent semigroup). 
That is, for all $r,s,t\in \C$: 
%
\begin{equation*}
\begin{array}{ccc}
r \;\cseq\; (s \;\cseq\; t) = (r \;\cseq\; s) \;\cseq\; t &&
\cunit \;\cseq\; r = r = r \;\cseq\; \cunit 
\\
r \;\cpar\; (s \;\cpar\; t) = (r \;\cpar\; s) \;\cpar\; t &
r\; \cpar\; r = r~~~~&
\czero \;\cpar\; r = r = r \;\cpar\; \czero 
\\
r \;\czip\; (s \;\czip\; t) = (r \;\czip\; s) \;\czip\; t &&
r\; \czip\; r = r \qquad
\\
\textnormal{if}~~r\; \cleq\; s ~~\textnormal{and}~~s\; \cleq\; t~~\textnormal{then}~~r\; \cleq\; t &&
t\; \cleq\; t \\[0.5em]
\end{array}
\end{equation*}
%
In addition, the following distributivity axioms hold:
\begin{align*}
\quad (\cclrd{r}\, \cpar\, \cclrd{s}) \;\cseq\; \cclrd{t} & = (\cclrd{r} \,\cseq\, \cclrd{t}) \;\cpar\; (\cclrd{s}\, \cseq\, \cclrd{t}) \\
\quad \cclrd{t} \;\cseq\; (\cclrd{r}\, \cpar\, \cclrd{s}) & = (\cclrd{t} \,\cseq\, \cclrd{r}) \;\cpar\; (\cclrd{t}\, \cseq\, \cclrd{s})
\end{align*}
\end{definition}

\noindent
In two of the three systems, some of the operators of the flat coeffect algebra coincide, but
in the data-flow system all three are distinct. Similarly, the two special elements  
coincide in some, but not all systems. The required laws are motivated by the aim to capture
common properties of the three examples, without unnecessarily restricting the system:

\begin{itemize}
\item The monoid $(\C, \cseq, \cunit)$ represents \emph{sequential} composition of (semantic)
functions. The laws of a monoid are required in order to form a category structure in the 
semantics (Section~\ref{sec:flat-semantics}).

\item The idempotent monoid $(\C, \cpar, \czero)$ represents \emph{pointwise} composition, 
\ie~the case when the same context is passed to multiple (independent) computations. The monoid 
laws guarantee that usual syntactic transformations on tuples and the unit value 
(Section~\ref{sec:flat-exts}) preserve the coeffect. Idempotence holds for all our examples
and allows us to unify the flat and structural systems in Chapter~\ref{ch:unified}.

\item For the $\czip$ operation, we require associativity and idempotence. The idempotence
requirement makes it possible to duplicate the coeffects and place the same requirement on both
call site and declaration site. Using the example from Section~\ref{sec:flat-calculus-lambda},
this guarantees that the rule (\emph{df1}) is not a special case, but can always be derived 
from (\emph{df2}). In some cases, the operator forms a monoid with the unit being the greatest 
element of the set $\C$. 
\end{itemize}

\noindent
It is worth noting that, in some of the systems, the operators $\cpar$ and $\czip$ are the least 
upper bound and the greatest lower bounds of a lattice. For example, in data-flow computations, they 
are \emph{max} and \emph{min} respectively. However, this duality does not hold for implicit parameters
(we discuss lattice-based formulation of coeffects in Section~\ref{sec:unified-impl-semilattice}.

Using the syntactic reading, the two operators represent \emph{merging} and \emph{splitting} of context 
requirements -- in the (\emph{abs}) rule, $\czip$ appears in the assumption and the combined context 
requirements of the body are split between two positions in the conclusions; in the (\emph{app}) rule, 
$\cpar$ appears in the conclusion and combines two context requirements from the assumptions.

\paragraph{Ordering.}

The flat coeffect algebra requires a pre-order relation $\cleq$, which is used to define 
sub-coeffecting rule of the type system. When the idempotent monoid $(\C, \cpar, \czero)$ also 
has the commutative property (\ie~forms a semi-lattice), the $\cleq$ relation can be defined as the
ordering of the semi-lattice:
%
\begin{equation*}
\cclrd{r} \;\cleq\; \cclrd{s} \;\Longleftrightarrow\; \cclrd{r} \;\cpar\; \cclrd{s} \;=\; \cclrd{s}
\end{equation*}
%
This definition is consistent with all three examples that motivate flat coeffect calculus, but
it cannot be used with the structural coeffects (where it fails for the bounded reuse 
calculus) and so we choose not to use it.

Furthermore, the $\cunit$ coeffect is often the top or the bottom element of the semi-lattice. 
As discussed in Section~\ref{sec:flat-syntax}, when this is the case, we are able to prove certain 
syntactic properties of the calculus.

%---------------------------------------------------------------------------------------------------

\begin{figure}[t]
\begin{equation*}
\tyrule{var}
  {x : \tau \in \Gamma}
  {\coctx{\Gamma}{\cunit} \vdash x : \tau }
\end{equation*}
\begin{equation*}
\tyrule{const}
  {c : \tau \in \Delta}
  {\coctx{\Gamma}{\czero} \vdash c : \tau }
\end{equation*}
\begin{equation*}
\tyrule{sub}
  {\coctx{\Gamma}{\cclrd{r'}} \vdash e : \tau }
  {\coctx{\Gamma}{\cclrd{r}} \vdash e : \tau }\quad\quad(\cclrd{r'} \cleq \cclrd{r})
\end{equation*}
\begin{equation*}
\tyrule{app}
  {\coctx{\Gamma}{\cclrd{r}} \vdash e_1 : \tau_1 \xrightarrow{\cclrd{t}} \tau_2 &
   \coctx{\Gamma}{\cclrd{s}} \vdash e_2 : \tau_1 }
  {\coctx{\Gamma}{\cclrd{r} \;\cpar\; (\cclrd{s} \,\cseq\, \cclrd{t})} \vdash e_1~e_2 : \tau_2}
\end{equation*}
\begin{equation*}
\tyrule{abs}
  {\coctx{\Gamma, x:\tau_1}{\cclrd{r}\;\czip\;\cclrd{s}} \vdash e : \tau_2}
  {\coctx{\Gamma}{\cclrd{r}} \vdash \lambda x.e : \tau_1 \xrightarrow{\cclrd{s}} \tau_2 }
\end{equation*}
\begin{equation*}
\tyrule{let}
  { \coctx{\Gamma}{\cclrd{r}} \vdash e_1 : \tau_1 &
    \coctx{\Gamma, x:\tau_1}{\cclrd{s}} \vdash e_2 : \tau_2}
  {\coctx{\Gamma}{\cclrd{s} \;\cpar\; (\cclrd{s} \,\cseq\, \cclrd{r})} \vdash \kvd{let}~x=e_1~\kvd{in}~e_2 : \tau_2 }
\end{equation*}

\figcaption{Type system for the flat coeffect calculus}
\label{fig:flat-types}
\end{figure}

%---------------------------------------------------------------------------------------------------

\subsection{Flat coeffect types}
\label{sec:flat-calculus-types}

The type system for flat coeffect calculus is shown in Figure~\ref{fig:flat-types}. Variables 
(\emph{var}) and constants (\emph{const}) are annotated with special values provided by the 
coeffect algebra. Following the top-down syntactic reading, the (\emph{sub}) rule allows us to 
treat an expression with fewer context requirements as an expression with more context requirements.

The (\emph{abs}) rule is defined as discussed in Section~\ref{sec:flat-calculus-lambda}. The
body is annotated with context requirements $\cclrd{r} \,\czip\, \cclrd{s}$, which are then split
between the context-requirements on the declaration site $\cclrd{r}$ and context-requirements on
the call site $\cclrd{s}$. Examples of the $\czip$ operator are discussed in the next section.

In function application (\emph{app}), context requirements of both expressions and the 
function are combined. As discussed in Chapter~\ref{ch:applications}, the pointwise composition
$\cpar$ is used to combine the context requirements of the expression representing a function 
$\cclrd{r}$ and the context requirements of the argument, sequentially composed with the 
context-requirements of the function $\cclrd{s}\, \cseq \,\cclrd{t}$.

The type system also includes a rule for let-binding. The rule is \emph{not} equivalent to the
typing derivation for $(\lambda x.e_2)~e_1$, but it corresponds to \emph{one} possible typing 
derivation. As we show in \ref{sec:flat-calculus-let}, the typing used in (\emph{let}) is more
precise than the general rule that can be derived from $(\lambda x.e_2)~e_1$. Additional 
constructs such as pairs are covered in Section~\ref{sec:flat-exts}.

%---------------------------------------------------------------------------------------------------

\subsection{Examples of flat coeffects}
\label{sec:flat-calculus-examples}

The flat coeffect calculus generalizes the flat systems discussed in 
Section~\ref{sec:applications-flat} of the previous chapter. We can instantiate it to a specific
use just by providing a flat coeffect algebra. The following summary defines the systems for implicit 
parameters, liveness and data-flow. For the latter two, the general calculus has a lambda abstraction
that is compatible with those discussed in Chapter~\ref{ch:flat}, but includes implicit sub-coeffecting.

\begin{example}[Implicit parameters]
Assuming \ident{Id} is a set of implicit parameter names, the flat coeffect algebra 
is formed by $(\mathcal{P}(\ident{Id}), \cup, \cup, \cup, \emptyset, \emptyset, \subseteq)$.
\end{example}

\noindent
For simplicity, we assume that all parameters have the same type $\rho$ and so the annotations only
track sets of names. The definition uses set union for all three operations. Both variables and
constants are are annotated with $\emptyset$ and the ordering is defined by $\subseteq$. The 
definition satisfies the flat coeffect algebra laws because $(S, \cup, \emptyset)$ is an idempotent, 
commutative monoid. The system has a single additional typing rule for accessing the value of a
parameter:
%
\begin{equation*}
\tyrule{param}
  { \ident{?p} \in \cclrd{c} }
  { \coctx{\Gamma}{\cclrd{c}} \vdash \ident{?p} : \rho }
\end{equation*}
%
The rule specifies that the accessed parameter $\ident{?p}$ needs to be in the set of required
parameters $\cclrd{c}$. As discussed earlier, we use the same type $\rho$ for all parameters, but
it is easy to define an extension that uses mappings from names to types.

\begin{example}[Liveness]
Let $\mathcal{L}=\{ \ident{L}, \ident{D} \}$ be a two-point lattice such that $\ident{D} \sqsubseteq \ident{L}$
with a join $\sqcup$ and meet $\sqcap$. The flat coeffect algebra for liveness is then formed by
$(\mathcal{L}, \sqcap, \sqcup, \sqcap, \ident{L}, \ident{D}, \sqsubseteq)$.
\end{example}

\noindent
As in Section~\ref{sec:applications-flat-live}, sequential composition $\cseq$ is modelled by 
the meet operation $\sqcap$ and pointwise composition $\cpar$ is modelled by join $\sqcup$. 
The two-point lattice is a commutative, idempotent monoid. Distributivity 
$(r \sqcup s) \sqcap t = (r \sqcap t) \sqcup (s \sqcap t)$ does not hold for \emph{every} 
lattice, but it trivially holds for the two-point lattice used here.

The definition uses join $\sqcup$ for the $\czip$ operator that is used by lambda abstraction.
This means that, when the body is live $\ident{L}$, both declaration site and call site are 
marked as live $\ident{L}$. When the body is dead $\ident{D}$, the declaration site and call site
can be marked as dead $\ident{D}$, or as live $\ident{L}$. The latter is less precise, but 
permissible over-approximation, which could otherwise be obtained via sub-typing.

\begin{example}[Data-flow]
In data-flow, context is annotated with natural numbers and the flat coeffect algebra is formed 
by $(\mathbb{N}, +, \mathit{max}, \mathit{min}, 0, 0, \leq)$.
\end{example}

\noindent
As discussed earlier, sequential composition $\cseq$ is represented by $+$ and pointwise 
composition $\cpar$ uses $\emph{max}$. For data-flow, we need a third separate operator for
lambda abstraction. Annotating the body with $\emph{min}(\cclrd{r}, \cclrd{s})$ ensures that
both call site and declaration site annotations are equal or greater than the annotation 
of the body. As with liveness, this allows over-approximation.

As required by the laws, $(\mathbb{N}, +, 0)$ and $(\mathbb{N}, \mathit{max}, 0)$ form monoids
and $(\mathbb{N}, \mathit{min})$ forms a band. Note that data-flow is our first example where 
$\cseq$ is not idempotent. The distributivity laws require the following to be the case:
$\mathit{max}(r,s) + t = \mathit{max}(r+t, s+t)$, which is easy to see. Finally, a simple 
data-flow language includes an additional rule for $\kvd{prev}$:
%
\begin{equation*}
\tyrule{prev}
  { \coctx{\Gamma}{\cclrd{c}} \vdash e : \tau }
  { \coctx{\Gamma}{\cclrd{c}+1} \vdash \kvd{prev}~e : \tau }
\end{equation*}
%
As a further example that was not covered earlier, it is also possible to combine liveness analysis
and data-flow. In the above calculus, $0$ denotes that we require the current value, but no previous
values. However, for constants, we do not even need the current value.

\begin{example}[Optimized data-flow]
In optimized data-flow, context is annotated with natural numbers extended with the $\bot$ element,
that is $\mathbb{N}_{\bot} = \{\bot, 0, 1, 2, 3, \ldots \}$ such that $\forall n \in \mathbb{N}. \bot \leq n$.
The flat coeffect algebra is $(\mathbb{N}_{\bot}, +, \mathit{max}, \mathit{min}, 0, \bot, \leq)$
where $m + n$ is $\bot$ whenever $m=\bot$ or $n=\bot$ and \emph{min}, \emph{max} treat $\bot$ as the
least element.
\end{example}

\noindent
Note that $(\mathbb{N}_{\bot}, +, 0)$ is a monoid for the extended definition of $+$; for the
bottom element $0 + \bot = \bot$ and for natural numbers $0 + n = n$. The structure 
$(\mathbb{N}, \emph{max}, \bot)$ is also a monoid, because $\bot$ is the least element and so
$\mathit{max}(n, \bot) = n$. Finally,  $(\mathbb{N}, \emph{min})$ is a band (the extended
\emph{min} is still idempotent and associative) and the distributivity law also holds
for $\mathbb{N}_{\bot}$.

%---------------------------------------------------------------------------------------------------

\subsection{Typing of let binding}
\label{sec:flat-calculus-let}

Recall the (\emph{let}) rule in Figure~\ref{fig:flat-types}. It annotates the expression 
$\kvd{let}~x=e_1~\kvd{in}~e_2$ with context requirements $\cclrd{s}\;\cpar\;(\cclrd{s}\,\cseq\,\cclrd{r})$.
This is a special case of typing an expression $(\lambda x.e_2)~e_1$, using the idempotence
of $\czip$ as follows:
%
\begin{equation*}
\tyrule{app}
  {\begin{array}{l}
   \vspace{-1.5em}
   \coctx{\Gamma}{\cclrd{r}} \vdash e_1 : \tau_1
   \end{array} &
   \tyruler{abs}
       { \coctx{\Gamma, x:\tau_1}{\cclrd{s}} \vdash e_2 : \tau_2 }
       { \coctx{\Gamma}{\cclrd{s}} \vdash \lambda x.e_2 : \tau_1 \xrightarrow{\cclrd{s}} \tau_2 } }
  { \coctx{\Gamma}{\cclrd{s}\;\cpar\;(\cclrd{s}\,\cseq\,\cclrd{r})} \vdash (\lambda x.e_2)~e_1 : \tau_2 }    
\end{equation*}
%
This design decision is similar to ML value restriction, but it works the other way round. Our
\emph{let} binding is more restrictive than the typing of abstraction-application, rather than
being more general. The choice is motivated by the fact that the typing obtained using the special 
rule for let-binding is more precise (with respect to sub-coeffecting) for all the examples consider 
in this chapter. Table~\ref{tab:flat-simplelet} shows how the coeffect annotations are simplified 
for our examples.

\begin{table}[!h]
\begin{center}
\begin{tabular}{ | l | c | c |}
\hline
& \textbf{\footnotesize Definition\hspace{1em}} & \textbf{\footnotesize Simplified\hspace{1em}} \\ \hline
\hspace{-1em}{\footnotesize Implicit parameters} & $\cclrd{s} \cup (\cclrd{s} \cup \cclrd{r})$ & $\cclrd{s} \cup \cclrd{r}$ \\ \hline
\hspace{-1em}{\footnotesize Liveness} & $\cclrd{s} \sqcap (\cclrd{s} \sqcup \cclrd{r})$ & $\cclrd{s}$ \\ \hline
\hspace{-1em}{\footnotesize Data-flow} & $\mathit{max}(\cclrd{s}, \cclrd{s} + \cclrd{r})$ & $\cclrd{s} + \cclrd{r}$ \\ \hline
\end{tabular}
\end{center}

\vspace{-0.5em}
\figcaption{Simplified annotation for let binding in sample flat calculi instances}
\label{tab:flat-simplelet}
\end{table}

\noindent
The simplified annotations directly follow from the definitions of particular flat coeffect 
algebras. It is perhaps somewhat unexpected that the annotation can be simplified in different
ways for different examples.

To see that the simplified annotations are more precise, assume that we used arbitrary 
splitting $\cclrd{s} = \cclrd{s_1}\,\czip\,\cclrd{s_2}$ rather than idempotence. The
``Definition'' column would use $\cclrd{s_1}$ and $\cclrd{s_2}$ for the first and second 
$\cclrd{s}$, respectively. The corresponding simplified annotation would have 
$\cclrd{s_1}\,\czip\,\cclrd{s_2}$ instead of $\cclrd{s}$. For all our systems, the 
simplified annotation (on the right) is more precise than the original (on the left):
%
\begin{equation*}
\begin{array}{rclll}
\cclrd{s_1} \cup (\cclrd{s_2} \cup \cclrd{r}) &\supseteq& (\cclrd{s_1} \cup \cclrd{s_2}) \cup \cclrd{r} 
  && \textnormal{(implicit parameters)}\\
\cclrd{s_1} \sqcap (\cclrd{s_2} \sqcup \cclrd{r}) &\sqsupseteq&  (\cclrd{s_1} \sqcap \cclrd{s_2}) 
  && \textnormal{(liveness)} \\
\mathit{max}(\cclrd{s_1}, \cclrd{s_2} + \cclrd{r}) &\geq& \mathit{min}(\cclrd{s_1}, \cclrd{s_2}) + \cclrd{r} 
  &\quad& \textnormal{(data-flow)} \\
\end{array}
\end{equation*}
%
In other words, the inequality states that using idempotence, we get a more precise typing.
This can be expressed using the abstract operators of the flat coeffect algebra as:
%
\begin{equation*}
\cclrd{s_1} \;\cpar\; (\cclrd{s_2} \,\cseq\, \cclrd{r}) \;\cgeq\; 
  (\cclrd{s_1} \,\czip\, \cclrd{s_2}) \;\cpar\; ((\cclrd{s_1} \,\czip\, \cclrd{s_2}) \,\cseq\, \cclrd{r})
\end{equation*}
%
This property cannot be proved from other properties of the flat coeffect algebra. To make
the flat coeffect system as general as possible, we do not \emph{in general} require it as
an additional axiom, although the above examples provide reasonable basis for requiring 
that the specialized annotation for let binding is the least possible annotation for the 
expression $(\lambda x.e_2)~e_1$.

% ==================================================================================================
%                                              
%      #####  #                                 
%        #    #                                 
%        #    # ##    ###    ###   # ##   #   # 
%        #    ##  #  #   #  #   #  ##  #  #   # 
%        #    #   #  #####  #   #  #      #  ## 
%        #    #   #  #      #   #  #       ## # 
%        #    #   #   ###    ###   #          # 
%                                         #   # 
%                                          ###                                                                         
% ==================================================================================================

\section{Categorical motivation}
\label{sec:flat-semantics}

The type system of flat coeffect calculus arises syntactically, as a generalization of the examples 
discussed in Chapter~\ref{ch:applications}, but we can also obtain it by looking at the categorical 
semantics of context-dependent computations. This is a direction that we explore in this section. 
Although the development presented here is interesting in its own, our main focus is \emph{using} 
categorical semantics to motivate and explain the design of flat coeffect calculus.

% --------------------------------------------------------------------------------------------------

\subsection{Categorical semantics}

As discussed in Section~\ref{sec:path-sem}, categorical semantics interprets terms as morphisms
in some category. For typed calculi, the semantics defined by $\sem{-}$ usually interprets typing 
judgements $x_1 \!:\! \tau_1 \ldots x_n \!:\! \tau_n \vdash e: \tau$ as morphisms 
$\sem{\tau_1 \times \ldots \times \tau_n} \rightarrow \sem{\tau}$.

As a best known example, Moggi \cite{monad-notions} showed that the semantics of various effectful 
computations can be captured uniformly using (\emph{strong}) \emph{monads}. In that 
approach, computations are interpreted as $\tau_1 \times \ldots \times \tau_n \rightarrow \mtyp{}{\tau}$
for some monad $\mtyp{}{}$. For example, $\mtyp{}{\alpha} = \alpha \cup \{ \bot \}$ models 
partiality (maybe monad), $\mtyp{}{\alpha} = \mathcal{P}(\alpha)$ models non-determinism (list 
monad) and side-effects can be modelled using $\mtyp{}{\alpha} = S \rightarrow (\alpha \times S)$ 
(state monad). Here, the structure of a strong monad provides necessary ``plumbing'' for composing 
monadic computations -- sequential composition and strength for lifting free variables into the
body of computation under a lambda abstraction.

Following similar approach to Moggi, Uustalu and Vene \cite{comonads-notions} showed that 
(\emph{monoidal}) \emph{comonads} uniformly capture the semantics of various kinds of context-dependent 
computations~\cite{comonads-notions}. For example, data-flow computations over non-empty lists
are modelled using the non-empty list comonad $\ident{NEList}\,\alpha = \alpha + (\alpha \times \ident{NEList}\,\alpha)$.

The monadic and comonadic model outlined here represents at most a binary analysis of effects or 
context-dependence. A function $\tau_1 \rightarrow \tau_2$ performs \emph{no} effects (requires no 
context) whereas $\tau_1 \rightarrow \mtyp{}{\tau_2}$ performs \emph{some} effects and
$\ctyp{}{\tau_1} \rightarrow \tau_2$ requires \emph{some} context\footnote{This is an 
over-simplification as we can use \eg~stacks of monad transformers and model functions with
two different effects using $\tau_1 \rightarrow M_1(M_2~\tau_2)$. However, monad transformers 
require defining complex system of lifting to be composable. Consequently, they are usually used
for capturing different kinds of impurities (exceptions, non-determinism, state), but not for
capturing fine-grained properties (\eg~a set of memory regions that may be accessed by a
stateful computation).}.

In the next section, we introduce \emph{indexed comonads}, which provide a more precise analysis 
and let us model computations with context requirements $\cclrd{r}$ as functions 
$\ctyp{\cclrd{r}}{\tau_1} \rightarrow \tau_2$ using an \emph{indexed comonad} $\ctyp{\cclrd{r}}{}$.

% --------------------------------------------------------------------------------------------------

\subsection{Introducing comonads}

In category theory, \emph{comonad} is a dual of \emph{monad}. As already outlined in 
Section~\ref{sec:path-sem}, we get a definition of a comonad by taking a definition of a monad 
and ``reversing the arrows''. More formally, one of the equivalent definitions of comonad 
looks as follows:

\begin{definition}
A \emph{comonad} over a category $\catc$ is a triple $(C, \ident{counit}, \ident{cobind})$ where:
\begin{compactitem}
\item $C$ is a mapping on objects (types) $C : \catc \rightarrow \catc$
\item $\ident{counit}$ is a mapping $\ctyp{}{\alpha} \rightarrow \alpha$ 
\item $\ident{cobind}$ is a mapping $(\ctyp{}{\alpha} \rightarrow \beta) 
  \rightarrow (\ctyp{}{\alpha} \rightarrow \ctyp{}{\beta})$
\end{compactitem}
such that, for all $f:\ctyp{}{\alpha} \rightarrow \beta$ and $g:\ctyp{}{\beta} \rightarrow \gamma$:
\begin{align}
\tag{\emph{left identity}}
  \ident{cobind}~\ident{counit} &= \idf{}
  \\
\tag{\emph{right identity}}
  \ident{counit} \circ \ident{cobind}~f &= f
  \\
\tag{\emph{associativity}}
  \ident{cobind}~(g \circ \ident{cobind}~f) &= (\ident{cobind}~g) \circ (\ident{cobind}~f)
\end{align}
\end{definition}

\noindent
From the functional programming perspective, we can see $\ctyp{}{}$ as a parametric data type such as
\ident{NEList}. The $\ident{counit}$ operations extracts a value $\alpha$ from a value that carries 
additional context $\ctyp{}{\alpha}$. The $\ident{cobind}$ operation turns a context-dependent function 
$\ctyp{}{\alpha} \rightarrow \beta$ into a function that takes a value with context, applies
the context-dependent function to value(s) in the context and then propagates the context.

As mentioned earlier, Uustalu and Vene \cite{comonads-notions} use comonads to model data-flow
computations. They describe infinite (coinductive) streams and non-empty lists as example comonads.

\begin{example}[Non-empty list]
A non-empty list is a recursive data-type defined as $\ident{NEList}\,\alpha = \alpha + (\alpha \times \ident{NEList}\,\alpha)$.
We write \kvd{inl} and \kvd{inr} for constructors of the left and right cases, respectively. The 
type \ident{NEList} forms a comonad together with the following \ident{counit} and \ident{cobind} mappings:
%
\begin{equation*}
\begin{array}{rclll}
\ident{counit}~l &\narrow{=}& h &\quad&\textnormal{when}~l=\kvd{inl}~h\\[-0.25em]
\ident{counit}~l &\narrow{=}& h &&\textnormal{when}~l=\kvd{inr}~(h, t)\\[0.5em]
\ident{cobind}\,f~l &\narrow{=}& \kvd{inl}~(f\,l) &&\textnormal{when}~l=\kvd{inl}~h\\[-0.25em]
\ident{cobind}\,f~l &\narrow{=}& \kvd{inr}~(f\,l,\;\ident{cobind}~f~t) &&\textnormal{when}~l=\kvd{inr}~(h, t)
\end{array}
\end{equation*}
\end{example}

\noindent
The \ident{counit} operation returns the head of the non-empty list. Note that it is crucial that
the list is \emph{non-empty}, because we always need to be able to obtain a value. The \ident{cobind}
operation defined here returns a list of the same length as the original where, for each element, the 
function $f$ is applied on a \emph{suffix} list starting from the element. Using a simplified
notation for list, the result of applying \ident{cobind} to a function that sums elements of a
list gives the following behaviour:
%
\begin{equation*}
\ident{cobind}~\ident{sum}~(7,6,5,4,3,2,1,0) = (28,21,15,10,6,3,1,0)
\end{equation*}
%
The fact that the function $f$ is applied to a \emph{suffix} is important in order to satisfy the
\emph{left identity} law, which requires that $\ident{cobind}~\ident{counit}~l = l$.

It is also interesting to examine some data types that do \emph{not} form a comonad. As already
mentioned, list $\ident{List}~\alpha = 1 + (\alpha \times \ident{List}~\alpha)$ is not a comonad,
because the \ident{counit} operation is not defined for the value $\kvd{inl}~()$. The \ident{Maybe} 
data type defined as $1 + \alpha$ is not a comonad for the same reason. However, if we consider 
flat coeffect calculus for liveness, it appears natural to model computations as functions 
$\ident{Maybe}~\tau_1 \rightarrow \tau_2$. To use such a model, we need to generalise comonads 
to \emph{indexed comonads}.

% --------------------------------------------------------------------------------------------------

\subsection{Generalising to indexed comonads}
\label{sec:flat-semantics-idx}

The flat coeffect algebra includes a monoid $(\C, \cseq, \cunit)$, which defines the behaviour of
sequential composition, where the annotation $\cunit$ represents a variable access. An indexed 
comonad is formed by a data type (object mapping) $\ctyp{\cclrd{r}}{\alpha}$ where the annotation 
$\cclrd{r}$ determines what context is required.

\begin{definition}
Given a monoid $(\C, \cseq, \cunit)$ with binary operator $\cseq$ and unit $\cunit$, an 
\emph{indexed comonad} over a category $\catc$ is a triple 
$(\ctyp{\cclrd{r}{}}, \ident{counit}_{\cunit}, \ident{cobind}_{\cclrd{r}, \cclrd{s}})$ where:

\begin{compactitem}
\item $\ctyp{\cclrd{r}}{}$ for all $\cclrd{r} \in \C$ is a family of object mappings 
\item $\ident{counit}_{\cunit}$ is a mapping $\ctyp{\cunit}{\alpha} \rightarrow \alpha$ 
\item $\ident{cobind}_{\cclrd{r}, \cclrd{s}}$ is a mapping $(\ctyp{\cclrd{r}}{\alpha} \rightarrow \beta) 
  \rightarrow (\ctyp{\cclrd{r}\cseq\cclrd{s}}{\alpha} \rightarrow \ctyp{\cclrd{s}}{\beta})$
\end{compactitem}
such that, for all $f:\ctyp{\cclrd{r}}{\alpha} \rightarrow \beta$ and $g:\ctyp{\cclrd{s}}{\beta} \rightarrow \gamma$:f
\begin{align}
\tag{\emph{left identity}}
  \ident{cobind}_{\cunit, \cclrd{s}}~\ident{counit}_{\cunit} &= \idf{}
  \\
\tag{\emph{right identity}}
  \ident{counit}_{\cunit} \circ \ident{cobind}_{\cclrd{r}, \cunit}~f &= f
  \\
\tag{\emph{associativity}}
\hspace{-10em}
  \ident{cobind}_{\cclrd{r}\cseq\cclrd{s},\cclrd{t}}~(g \circ \ident{cobind}_{\cclrd{r}, \cclrd{s}}~f) &= 
    (\ident{cobind}_{\cclrd{s}, \cclrd{t}}~g) \circ (\ident{cobind}_{\cclrd{r}, \cclrd{s}\cseq\cclrd{t}}~f)
\end{align}
\end{definition}

\noindent
Rather than defining a single mapping $\ctyp{}{}$, we are now defining a family of mappings 
$\ctyp{\cclrd{r}}{}$ indexed by a monoid structure. Similarly, the $\ident{cobind}_{\cclrd{r}, \cclrd{s}}$ 
operation is now formed by a \emph{family} of mappings for different pairs of indices 
$\cclrd{r}, \cclrd{s}$. To be fully precise, $\ident{cobind}$ is a family of natural transformations 
and we should include $\alpha, \beta$ as indices, writing $\ident{cobind}_{\cclrd{r},\cclrd{s}}^{\alpha, \beta}$.
For the purpose of this thesis, it is sufficient to omit the superscripts and treat $\ident{cobind}$ just 
as a family of mappings (rather than natural transformations). When this does not introduce ambiguity, 
we also occasionally omit the subscripts.

The $\ident{counit}$ operation is not defined for all $\cclrd{r} \in \C$, but only for 
the unit $\cunit$. We still include the unit as an index writing $\ident{counit}_{\cunit}$, 
but this is merely for symmetry and as a useful reminder to the reader. Crucially, this means that 
the operation is defined only for special contexts.

If we look at the indices in the laws, we can see that the left and right identity 
require $\cunit$ to be the unit of $\cseq$. Similarly, the associativity law implies the 
associativity of the $\cseq$ operator.

\paragraph{Composition.}
The co-Kleisli category that models sequential composition is formed by the unit arrow (provided
by $\ident{counit}$) together with the (associative) composition operation that composes computations 
with contextual requirements as follows:
%
\begin{equation*}
\begin{array}{ccl}
\textnormal{--}\, \hat{\circ} \,\textnormal{--}&\narrow{:}& (\ctyp{\cclrd{r}}{\tau_1} \rightarrow \tau_2) 
  \rightarrow (\ctyp{\cclrd{s}}{\tau_2} \rightarrow \tau_3) 
  \rightarrow (\ctyp{\cclrd{r} \cseq \cclrd{s}}{\tau_1} \rightarrow \tau_3) \\
g \, \hat{\circ} \, f &\narrow{=}& g \circ (\ident{cobind}_{\cclrd{r}, \cclrd{s}} f)
\end{array}
\end{equation*}
%
The composition $\hat{\circ}$ best expresses the intention of indexed comonads. Given two functions
with contextual requirements $\cclrd{r}$ and $\cclrd{s}$, their composition is a function that 
requires $\cclrd{r}\,\cseq\,\cclrd{s}$. The contextual requirements propagate \emph{backwards} and
are attached to the input of the composed function.

% --------------------------------------------------------------------------------------------------

\paragraph{Examples.}

Any comonad can be turned into an indexed comonad using a trivial monoid. However, indexed comonads
are more general and can be used with other data types, including indexed \ident{Maybe}.

\begin{example}[Comonads]
Any comonad $\ctyp{}{}$ is an indexed comonad with an index provided by a trivial monoid $(\{1\},\ast,1)$
where $1\ast 1 = 1$. The mapping $\ctyp{1}{}$ is the mapping $\ctyp{}{}$ of the underlying comonad. The
operations $\ident{counit}_1$ and $\ident{cobind}_{1,1}$ are defined by the operations $\ident{counit}$
and $\ident{cobind}$ of the comonad.
\end{example}

\begin{example}[Indexed option]
The \emph{indexed option comonad} is defined over a monoid $(\{ \ident{L},\ident{D} \}, \sqcup,\ident{L})$ 
where $\sqcup$ is defined as earlier, \ie~$\ident{L} = \cclrd{r} \sqcup \cclrd{s} \Longleftrightarrow \cclrd{r}=\cclrd{s}=\ident{L}$.
Assuming $1$ is the unit type inhabited by $()$, the mappings are defined as follows:
%
\begin{equation*}
\begin{array}{l}
\ctyp{\ident{L}}{\alpha} = \alpha\\
\ctyp{\ident{D}}{\alpha} = 1\\
\\
\ident{counit}_{\ident{L}} : \ctyp{\ident{L}}{\alpha} \rightarrow \alpha \\
\ident{counit}_{\ident{L}}~v = v\\
\end{array}
\qquad
\begin{array}{lcl}
\multicolumn{3}{l}{
  \ident{cobind}_{\cclrd{r},\cclrd{s}} ~:~ (\ctyp{\cclrd{r}}{\alpha} \rightarrow \beta) 
    \rightarrow (\ctyp{\cclrd{r}\sqcup\cclrd{s}}{\alpha} \rightarrow \ctyp{\cclrd{s}}{\beta}) }\\
\ident{cobind}_{\ident{L},\ident{L}}~f~x &\narrow{=}& f~x\\
\ident{cobind}_{\ident{L},\ident{D}}~f~() &\narrow{=}& ()\\
\ident{cobind}_{\ident{D},\ident{L}}~f~() &\narrow{=}& f~()\\
\ident{cobind}_{\ident{D},\ident{D}}~f~() &\narrow{=}& ()\\
\end{array}
\end{equation*}
\end{example}

\noindent
The \emph{indexed option comonad} models the semantics of the liveness coeffect system discussed in 
Section~\ref{sec:applications-flat-live}, where $\ctyp{\ident{L}}{\alpha} = \alpha$ models a live context 
and $\ctyp{\ident{D}}{\alpha}=1$ models a dead context which does not contain a value. The \ident{counit}
operation extracts a value from a live context. As in the direct model discussed in Chapter~\ref{ch:appendix},
the \ident{cobind} operation can be seen as an implementation of dead code elimination. The definition 
only evaluates $f$ when the result is marked as live and is thus required, and it only accesses $x$ if 
the function $f$ requires its input.

The indexed family $\ctyp{\cclrd{r}}{}$ in the above example is analogous to the \ident{Maybe}
(or option) data type $\ident{Maybe}\,\alpha = 1 + \alpha$. As mentioned earlier, this type does not 
permit (non-indexed) comonad structure, because $\ident{counit}~()$ is not defined. This is not a 
problem with indexed comonads, because live contexts are distinguished by the (type-level) coeffect 
annotation and \ident{counit} only needs to be defined on live contexts.

\begin{example}[Indexed product]
The semantics of implicit parameters is modelled by an indexed product comonad. We use a monoid
$(\mathcal{P}(\ident{Id}), \cup, \emptyset)$ where \ident{Id} is the set of (implicit parameter) names.
As previously, all parameters have the type $\rho$. The data type $\ctyp{\cclrd{r}}{\alpha}
= \alpha \times (\cclrd{r} \rightarrow \rho)$ represents a value $\alpha$ together with a function that
associates a parameter value $\rho$ with every implicit parameter name in $\cclrd{r} \subseteq \ident{Id}$.
The cobind and counit operations are defined as:
%
\begin{equation*}
\begin{array}{l}
\ident{counit}_{\emptyset} : \ctyp{\emptyset}{\alpha} \rightarrow \alpha \\[-0.25em]
\ident{counit}_{\emptyset}~(a, g) = a\\[0.5em]
\end{array}
\quad
\begin{array}{l}
\ident{cobind}_{\cclrd{r},\cclrd{s}} ~:~ (\ctyp{\cclrd{r}}{\alpha} \rightarrow \beta) 
    \rightarrow (\ctyp{\cclrd{r}\cup\cclrd{s}}{\alpha} \rightarrow \ctyp{\cclrd{s}}{\beta})\\[-0.25em]
\ident{cobind}_{\cclrd{r},\cclrd{s}}~f~(a,g) = (f(a,\restr{g}{\cclrd{r}}), \restr{g}{\cclrd{s}})\\[0.5em]
\end{array}
\end{equation*}
\end{example}

\noindent
In the definition, we use the notation $(a, g)$ for a pair containing the value of type $\alpha$
together with $g$, which is a function $\cclrd{r} \rightarrow \rho$. The \ident{counit} operation
takes a value and a function (with empty set as a domain), ignores the function and extracts the
value. The \ident{cobind} operation uses the restriction operation $\restr{g}{\cclrd{r}}$ to 
restrict the domain of $g$ to implicit parameters $\cclrd{r}$ and $\cclrd{s}$ in oder to get
implicit parameters required by the argument of $f$ and by the resulting computation, respectively
(\ie~semantically, it \emph{splits} the available context capabilities). The function $g$ passed 
to \ident{cobind} is initially defined on $\cclrd{r} \cup \cclrd{s}$ and so the restriction is 
valid in both cases.

The structure of \emph{indexed comonads} is sufficient to model sequential composition of 
computations that use a single variable (as discussed in Section~\ref{sec:path-sem}). To model 
full $\lambda$-calculus with lambda abstraction and multiple-variable contexts, we need 
additional operations introduced in the next section.

% --------------------------------------------------------------------------------------------------

\subsection{Flat indexed comonads}
\label{sec:flat-semantics-monoidal}

Because of the assymmetry of $\lambda$-calculus (discussed in Section~\ref{sec:applications-structure}),
the duality between monads and comonads can no longer help us with defining the additional structure
required to model full $\lambda$-calculus. In comonadic computations, additional information is attached
to the context. In application and lambda abstraction, the context is propagated differently than 
in effectful computations.

To model the effectful $\lambda$-calculus, Moggi~\cite{monad-notions} requires a \emph{strong} monad which
has an additional operation $\ident{strength} : \alpha\times\mtyp{}{\beta} \rightarrow \mtyp{}{(\alpha\times\beta)}$.
This allows lifting of free variables into an effectful computation. In Haskell, strength can be expressed
in the host language and so it is implicit.

To model $\lambda$-calculus with contextual properties, Uustalu and Vene \cite{comonads-notions} 
require \emph{lax semi-monoidal} comonad. This structure requires an additional monoidal operation:
%
\begin{equation*}
\ident{m} : \ctyp{}{\alpha} \times \ctyp{}{\beta} \rightarrow \ctyp{}{(\alpha \times \beta)}
\end{equation*}
%
The \ident{m} operation is needed in the semantics of lambda abstraction. It represents merging of 
contexts and is used to merge the context of the declaration site (containing free variables)
and the call site (containing bound variable). For example, for implicit parameters, this combines
the additional parameters defined in the two contexts.

The semantics of flat coeffect calculus requires operations for \emph{merging}, but also for
\emph{splitting} of contexts. In addition, we also need a lifting operation (similar to $\iota$ 
from Definition~\ref{def:flat-family}) to model sub-coeffecting.

\begin{definition}
Given a flat coeffect algebra $(\C, \cseq, \cpar, \czip, \cunit, \czero, \cleq)$,
a \emph{flat indexed comonad} is an indexed comonad over the monoid $(\C, \cseq, \cunit)$
equipped with families of operations $\ident{merge}_{\cclrd{r},\cclrd{s}}$, $\ident{split}_{\cclrd{r},\cclrd{s}}$ 
and $\ident{lift}_{\cclrd{r'},\cclrd{r}}$ where:
%
\begin{compactitem}
\item $\ident{merge}_{\cclrd{r},\cclrd{s}}$ is a family of mappings
  $\ctyp{\cclrd{r}}{\alpha} \times \ctyp{\cclrd{s}}{\beta} \rightarrow \ctyp{\cclrd{r}\czip\cclrd{s}}{(\alpha \times \beta)}$
\item $\ident{split}_{\cclrd{r},\cclrd{s}}$ is a family of mappings
  $\ctyp{\cclrd{r}\cpar\cclrd{s}}{(\alpha \times \beta)} \rightarrow \ctyp{\cclrd{r}}{\alpha} \times \ctyp{\cclrd{s}}{\beta}$
\item $\ident{lift}_{\cclrd{r'},\cclrd{r}}$ is a family of mappings
  $\ctyp{\cclrd{r'}}{\alpha} \rightarrow \ctyp{\cclrd{r}}{\alpha}$~ for all $\cclrd{r'}, \cclrd{r}$ such that $\cclrd{r}\,\cleq\,\cclrd{r'}$
\end{compactitem}
\end{definition}

\noindent
The $\ident{merge}_{\cclrd{r},\cclrd{s}}$ operation is the most interesting one. Given two comonadic
values with additional contexts specified by $\cclrd{r}$ and $\cclrd{s}$, it combineds them into a 
single value with additional context $\cclrd{r}\,\czip\,\cclrd{s}$. The $\czip$ operation often represents
\emph{greatest lower bound}\footnote{The $\czip$ and $\cpar$ operations are the greatest and least upper 
bounds in the liveness and data-flow examples, but not for implicit parameters. However, they remain useful 
as an informal analogy.}, elucidating the fact that merging may result in the loss of some parts of 
the contexts $\cclrd{r}$ and $\cclrd{s}$. We look at examples of this operation in the next section.

The $\ident{split}_{\cclrd{r},\cclrd{s}}$ operation splits a single comonadic value (containing a tuple)
into two separate values. Note that this does not simply duplicate the value, because the additional
context is also split. To obtain coeffects $\cclrd{r}$ and $\cclrd{s}$, the input needs to provide 
\emph{at least} $\cclrd{r}$ and $\cclrd{s}$, so the tags are combined using the $\cpar$, which is often 
the \emph{least upper-bound}\footnotemark[1].

Finally, $\ident{lift}_{\cclrd{r'}, \cclrd{r}}$ is a family of operations that ``forget'' some part of
a context. This models the sub-coeffecting operation and lets us, for example, forget some of the
available implicit parameters, or turn a live context (containing a value) into a dead context (empty).

\paragraph{Alternative definition.}
Although we do not demand this as a general law, in all our systems, it is the case that
$\cclrd{r} \;\cleq\; \cclrd{r}\,\cpar\,\cclrd{s}$ and $\cclrd{s} \;\cleq\; \cclrd{r}\,\cpar\,\cclrd{s}$.
This special case allows a simpler definition of \emph{indexed flat comonad} by expressing the 
\ident{split} operation in terms of lifting (sub-coeffecting) as follows:
%
\begin{equation*}
\begin{array}{rcl}
\ident{map}_{\cclrd{r}}~f &\narrow{=}& \ident{cobind}_{\cclrd{r}, \cclrd{r}}~(f\circ\ident{counit}_{\cunit}) \\
\ident{split}_{\cclrd{r}, \cclrd{s}}~c &\narrow{=}&
  ( \ident{map}_{\cclrd{r}}~\ident{fst}~(\ident{lift}_{\cclrd{r}\cpar\cclrd{s}, \cclrd{r}}~c), 
    \ident{map}_{\cclrd{s}}~\ident{snd}~(\ident{lift}_{\cclrd{r}\cpar\cclrd{s}, \cclrd{s}}~c) )
\end{array}
\end{equation*}
%
The $\ident{map}_{\cclrd{r}}$ operation is the mapping on arrows that corresponds to the object 
mapping $\ctyp{\cclrd{r}}{}$. The definition is dual to the standard definition of \ident{map} 
for monads in terms of \ident{bind} and \ident{unit}. The functions \ident{fst} and \ident{snd}
are first and second projections from a two-element pair. To define the 
$\ident{split}_{\cclrd{r}, \cclrd{s}}$ operation, we use the argument $c$ twice, use lifting
to throw away additional parts of the context and then transform the values in the 
context.

This alternative is valid for our examples, but we do not use it for two reasons. Firstly, this
would be the only place where our semantics uses a variable \emph{twice} (in this case $c$). 
Thus using an explicit \ident{split} means that the structure required by our semantics does not
need to provide variable duplication and our model could be embedded in linear or affine category.
Secondly, explicit \ident{split} is similar to the definition that is needed for structural 
coeffects in Chapter~\ref{ch:structural} and so it makes the connection between the two system 
easier to see.

\paragraph{Examples.}
All the examples of \emph{indexed comonads} discussed in Section~\ref{sec:flat-semantics-idx} can
be extended into \emph{flat indexed comonads}. Note that this is not a \emph{general} statement, 
because each example requires us to define additional operations, specific for the example.

\begin{example}[Monoidal comonads]
Just like indexed comonads generalise co\-monads, the additional structure of
flat indexed comonads generalises symmetric semimonoidal comonads of Uustalu 
and Vene \cite{comonads-notions}. The flat coeffect algebra is defined as $(\{1\}, \ast, \ast, \ast, 1, 1, =)$
where $1\ast1=1$ and $1=1$. The additional operation $\ident{merge}_{1,1}$ is provided by the 
monoidal operation called \ident{m} by Uustalu and Vene. The $\ident{split}_{1,1}$ operation 
is defined by duplication and $\ident{lift}_{1,1}$ is the identity function.
\end{example}

\begin{example}[Indexed option]
Flat coeffect algebra for liveness defines $\cpar$ and $\czip$ as $\sqcup$ and $\sqcap$, respectively 
and specifies that $\ident{D} \sqsubseteq \ident{L}$. Recall also that the object mapping is defined 
as $\ctyp{\ident{L}}{\alpha} = \alpha$ and $\ctyp{\ident{D}}{\alpha} = 1$. The additional operations 
of a flat indexed comonad are defined as follows:
%
\begin{equation*}
\begin{array}{rcl}
\ident{merge}_{\ident{L}, \ident{L}}~(a, b) &\narrow{=}& (a, b)\\
\ident{merge}_{\ident{L}, \ident{D}}~(a, ()) &\narrow{=}& ()\\
\ident{merge}_{\ident{D}, \ident{L}}~((), b) &\narrow{=}& ()\\
\ident{merge}_{\ident{D}, \ident{D}}~((), ()) &\narrow{=}& ()\\
\end{array}
\quad
\begin{array}{rcl}
\ident{split}_{\ident{L}, \ident{L}}~(a, b) &\narrow{=}& (a, b)\\
\ident{split}_{\ident{L}, \ident{D}}~(a, b) &\narrow{=}& (a, ())\\
\ident{split}_{\ident{D}, \ident{L}}~(a, b) &\narrow{=}& ((), b)\\
\ident{split}_{\ident{D}, \ident{D}}~() &\narrow{=}& ((), ()))\\
\end{array}
\quad
\begin{array}{rcl}
\ident{lift}_{\ident{L}, \ident{D}}~v &\narrow{=}& ()\\
\ident{lift}_{\ident{L}, \ident{L}}~v &\narrow{=}& v\\
\ident{lift}_{\ident{D}, \ident{D}}~() &\narrow{=}& ()\\
\end{array}
\end{equation*}
\end{example}

\noindent
Without the indexing, the \ident{merge} operations implements \emph{zip} on option values,
returning an option only when both values are present. The behaviour of the \ident{split} 
operation is partly determined by the indices. When the input is \emph{dead}, both values have 
to be dead (this is also the only solution of $\cclrd{\ident{D}} =\cclrd{r}\sqcap\cclrd{s}$), but when
the input is \emph{live}, the operation can perform implicit sub-coeffecting and drop one of
the values.

Explicit sub-coeffecting using the (\emph{sub}) rule is modelled by the \ident{lift} operation.
This can turn a \emph{live} value $v$ into a dead value $()$, or it can behave as identity.
The behaviour is, again, determined by the index.

\begin{example}[Indexed product]
For implicit parameters, both $\czip$ and $\cpar$ are the $\cup$ operation and the relation
$\cleq$ is formed by the subset relation $\subseteq$. Recall that the data type $\ctyp{\cclrd{r}}{\alpha}$
is $\alpha \times (\cclrd{r} \rightarrow \rho)$ where $\rho$ is the type of implicit parameter values.
The additional operations are defined as:
%
\begin{equation*}
\begin{array}{rcl}
\ident{split}_{\cclrd{r}, \cclrd{s}}~((a,b), g) &\narrow{=}& ((a, \restr{g}{\cclrd{r}}), (b, \restr{g}{\cclrd{s}}))\\
\ident{merge}_{\cclrd{r}, \cclrd{s}}~((a, f), (b, g)) &\narrow{=}& ((a, b), f \uplus g)\\
\ident{lift}_{\cclrd{r'}, \cclrd{r}}~(a, g) &\narrow{=}& (a, \restr{g}{\cclrd{r}})\\
\end{array}
\quad
\begin{array}{l}
\textnormal{where}~f \uplus g = \\[-0.25em]
\quad\restr{f}{\, \textit{dom}(f) \setminus \textit{dom}(g)} \cup g 
\end{array}
\end{equation*}
\end{example}

\noindent
The \ident{split} operation splits the tuple and restricts the function (representing available
implicit parameters) to the required sub-sets. This corresponds to the definition in terms
of \ident{lift}, which performs just the restriction. The \ident{merge} operation is more 
interesting. It uses the $\uplus$ operation that we defined when introducing implicit parameters
in Section~\ref{sec:applications-flat-impl}. It merges the values, preferring the definitions from
the right-hand side (call site) over left-hand side (declaration site). Thus the operation is not
symmetric.

\begin{example}[Indexed list]
Our last example provides the semantics of data-flow computations. The flat coeffect algebra 
is formed by $(\mathbb{N}, +, \mathit{max}, \mathit{min}, 0, 0, \leq)$. In a 
non-indexed version, the semantics is provided by a non-empty list. In the indexed semantics,
the index represents the number of available past values. The data type is then a pair of 
the current value, followed by $n$ past values. The mappings that form the flat indexed comonad 
are defined as follows:

\begin{equation*}
\begin{array}{l}
\ident{counit}_{0}\langle a_0 \rangle = a_0
\\[0.45em]
\ident{cobind}_{m, n}~f \langle a_0, \ldots a_{m+n} \rangle = \\[-0.25em]
\quad \langle f \langle a_0, \ldots, a_m \rangle, \ldots, f \langle a_{n}, \ldots, a_{m+n} \rangle \rangle
\\[0.45em]
\ident{merge}_{m, n} (\langle a_0, \ldots, a_m \rangle, \langle b_0, \ldots, b_n\rangle) = \\[-0.25em]
\quad \langle (a_0, b_0), \ldots, (a_{\mathit{min}(m,n)}, b_{\mathit{min}(m,n)}) \rangle
\\[0.45em]
\ident{split}_{m, n} \langle (a_0, b_0), \ldots, (a_{\mathit{max}(m,n)}, b_{\mathit{max}(m,n)}) \rangle = \\[-0.25em]
\quad (\langle a_0, \ldots, a_m \rangle, \langle b_0, \ldots, b_n\rangle)
\\[0.45em]
\ident{lift}_{n', n} \langle a_0, \ldots, a_{n'} \rangle = \qquad(\hspace{-0.95em}\textnormal{\footnotesize when}~n\leq n') \\[-0.25em]
\quad \langle a_0, \ldots, a_n \rangle
\end{array}
\begin{array}{l}
\hspace{-2em}\ctyp{n}{\alpha} = \underbrace{\alpha \times \ldots \times \alpha}_{(n+1)-\textnormal{times}}\\[11em]
~\\
\end{array}
\end{equation*}
\end{example}

\noindent
The reader is invited to check that the number of required past elements in each of the mappings
matches the number specified by the indices. The index specifies the number of \emph{past} elements
and so the list always contains at least one value. Thus \ident{counit} returns the element of a
singleton list.

The $\ident{cobind}_{m,n}$ operation requires $m + n$ elements in order to generate $n$ past results 
of the $f$ function, which itself requires $m$ past values. When combining two lists, 
$\ident{merge}_{m,n}$ behaves as \emph{zip} and produces a list that has the length of the shorter 
argument. When splitting a list, $\ident{split}_{m, n}$ needs the maximum of the required lengths. 
Finally, the lifting operation just drops some number of elements from a list.

% --------------------------------------------------------------------------------------------------

\subsection{Properties and related notions}

The flat indexed comonad structure is all we need to give the semantics of the flat coeffect 
calculus. Before doing so in Section~\ref{sec:flat-semantics-calculus}, we briefly consider 
additional properties and other categorical structures that have been proposed mainly in the 
context of monads and effects and we look how they relate to indexed comonads.

\paragraph{Shape preservation.}
Ordinary comonads have the \emph{shape preservation} property \cite{comonads-codo}. Intuitively, 
this means that the core comonad structure does not provide a way of modeling computations where
the additional context changes during the computation. For example, in the \ident{NEList} comonad, 
the length of the list stays the same after applying \ident{cobind}.

~

Indexed comonads are not restricted by this property of comonads. For example, given the indexed
product comonad, in the computation $\ident{cobind}_{\cclrd{r}, \cclrd{s}} f$, the shape of
the context changes from providing implicit parameters $\cclrd{r} \cup \cclrd{s}$  to providing
just implicit parameters $\cclrd{s}$.

\paragraph{Families of monads.}
When linking effect systems and monads, Wadler and Thiemann \cite{monad-notions} propose a
\emph{family of monads} as the categorical structure. The dual structure, \emph{family of
comonads}, is defined as follows.

\begin{definition}
\label{def:flat-family}
A \emph{family of comonads} is formed by triples $(\ctyp{\cclrd{r}}{}, \ident{cobind}_{\cclrd{r}}, 
  \ident{counit}_{\cclrd{r}})$ for all $\cclrd{r}$ such that each triple forms a comonad. Given 
$\cclrd{r}, \cclrd{r'}$ such that $\cclrd{r} \leq \cclrd{r'}$, there is also a mapping 
$\iota_{\cclrd{r'}, \cclrd{r}} : \ctyp{\cclrd{r'}}{} \rightarrow \ctyp{\cclrd{r}}{}$ satisfying
certain coherence conditions.
\end{definition}

\noindent
A \emph{family of comonads} is more restrictive than an \emph{indexed comonad}, because each of the data 
types needs to form a comonad separately. For example, our indexed option does not form a family of 
comonads (again, because \ident{counit} is not defined on $\ctyp{\ident{D}}{\alpha}=1$). However, given 
a family of comonads and indices such that $\cclrd{r} \leq \cclrd{r}\cseq\cclrd{s}$, we can define 
an indexed comonad. Briefly, to define $\ident{cobind}_{\cclrd{r},\cclrd{s}}$ of an indexed comonad, 
we use $\ident{cobind}_{\cclrd{r}\cseq\cclrd{s}}$ from the family, together with two lifting operations:
$\iota_{\cclrd{r}\cseq\cclrd{s}, \cclrd{r}}$ and $\iota_{\cclrd{r}\cseq\cclrd{s}, \cclrd{s}}$.

\paragraph{Parameteric effect monads.}
Parametric effect monads introduced by Katsumata \cite{monads-parametric} (independently to our 
indexed comonads) are closely related to our definition.  Although presented in a more general 
categorical framework (and using monads), the model defines the \ident{unit} operation only on the 
unit of a monoid and the \ident{bind} operation composes effect annotations using the provided monoidal 
structure.

% --------------------------------------------------------------------------------------------------

\subsection{Semantics of flat calculus}
\label{sec:flat-semantics-calculus}

In Section~\ref{sec:applications-flat}, we defined the semantics of concrete (flat) context-dependent
computations including implicit parameters, liveness and data-flow. Using the \emph{flat indexed 
comonad} structure, we can now define a single uniform semantics that is capable of capturing all 
our examples, as well as other computations that can be modelled by the structure.

\paragraph{Contexts and functions.}
The modelling of contexts and functions generalizes the earlier concrete examples. We use the 
family of mappings $\ctyp{\cclrd{r}}{}$ as an (indexed) data-type that wraps the product of 
free variables of the context and the arguments of functions:
%
\begin{equation*}
\begin{array}{rcl}
\sem{\coctx{x_1\!:\!\tau_1, \ldots, x_n\!:\!\tau_n}{ \cclrd{r} } \vdash e : \tau} 
  &:& \ctyp{\cclrd{r}}{(\tau_1 \times \ldots \times \tau_n)} \rightarrow \tau\\
\sem{\tau_1 \xrightarrow{\cclrd{r}} \tau_2} &=& \ctyp{\cclrd{r}}{\tau_1} \rightarrow \tau_2
\end{array}
\end{equation*}

% --------------------------------------------------------------------------------------------------

\begin{figure*}[t]

\begin{equation*}
\begin{array}{ll}
\sem{\coctx{\Gamma}{\cunit} \vdash x_i : \tau_i }~\ctx =
  \pi_i~(\ident{counit}_{\cunit}~\ctx) & (\emph{var})
\\[0.5em]
\sem{\coctx{\Gamma}{\czero} \vdash c_i : \tau }~\ctx =
  \delta~(c_i) & (\emph{const})
\\[0.5em]
\sem{\coctx{\Gamma}{\cclrd{r}} \vdash \lambda x.e : \tau_1 \xrightarrow{\cclrd{s}} \tau_2 }~\ctx = \lambda v.& (\emph{abs})\\[-0.25em]
  \hspace{3em}\sem{\coctx{\Gamma, x:\tau_1}{\cclrd{r}\;\czip\;\cclrd{s}} \vdash e : \tau_2}~(\ident{merge}_{\cclrd{r}, \cclrd{s}}~(\ctx, v))
\\[0.5em]
\sem{\coctx{\Gamma}{\cclrd{r} \;\cpar\; (\cclrd{s} \,\cseq\, \cclrd{t})} \vdash e_1~e_2 : \tau_2}~\ctx = & (\emph{app})\\[-0.25em]
  \hspace{2.5em}
  \begin{array}{l}  
  \kvd{let}~(\ctx_1, \ctx_2) = \ident{split}_{\cclrd{r}, \cclrd{s} \,\cseq\, \cclrd{t}}~
    (\ident{map}_{\cclrd{r} \;\cpar\; (\cclrd{s} \,\cseq\, \cclrd{t})}~(\lambda x.(x,x))~\ctx) \\[-0.25em]
  \kvd{in}~\sem{\coctx{\Gamma}{\cclrd{r}} \vdash e_1 : \tau_1 \xrightarrow{\cclrd{t}} \tau_2}~\ctx_1~
      (\ident{cobind}_{\cclrd{s}, \cclrd{t}}~\sem{\coctx{\Gamma}{\cclrd{s}} \vdash e_2 : \tau_1 }~\ctx_2)
  \end{array}    
\\[1.25em]
\sem{\coctx{\Gamma}{\cclrd{r}} \vdash e : \tau }~\ctx = & (\emph{sub})\\[-0.25em]
  \hspace{3em}\sem{\coctx{\Gamma}{\cclrd{r'}} \vdash e : \tau }~(\ident{lift}_{\cclrd{r}, \cclrd{r'}}~\ctx)~
    \hspace{7.5em}(\hspace{-0.7em}\textnormal{\footnotesize when}~\cclrd{r'} \leq \cclrd{r})
\end{array}
\end{equation*}

\figcaption{Categorical semantics of the flat coeffect calculus}
\label{fig:flat-semantics}
\end{figure*}

% --------------------------------------------------------------------------------------------------

\paragraph{Expressions.}
The definition of the semantics is shown in Figure~\ref{fig:flat-semantics}. For readability, we 
write the definitions in a simple programming language notation as opposed to the point-free 
categorical style. However, it can be equally written using just the operations of flat indexed 
comonad together with $i^{th}$ projection from a tuple represented by $\pi_i$, \emph{curry} and 
\emph{uncurry}, function composition, value duplication ($\Delta : A \rightarrow A \times A$) and 
function pairing (given $f:A\rightarrow B$ and $g:C\rightarrow D$ then $f\times g : A\times C \rightarrow B \times D$). 
These operations can be provided by \eg~a Cartesian-closed category.

The semantics of variable access and abstraction are the same as in the semantics of Uustalu and 
Vene \cite{comonads-notions}, modulo the indexing. The semantics of variable access (\emph{var}) uses 
$\ident{counit}_{\cunit}$ to extract product of free-variables from the context and then projection
$\pi_i$ to obtain the variable value. Abstraction (\emph{abs}) takes the context $\ctx$ and function argument 
$v$ and merges their additional contexts using $\ident{merge}_{\cclrd{r}, \cclrd{s}}$. Assuming
the context $\Gamma$ contains variables of types $\sigma_1, \ldots, \sigma_n$, this gives us a 
value $\ctyp{\cclrd{r}\czip\cclrd{s}}((\sigma_1 \times \ldots \times \sigma_n) \times \tau_1)$.
Assuming that $n$-element tuples are associated to the left, the wrapped context is equivalent to
$\sigma_1 \times \ldots \times \sigma_n \times \tau_1$, which can then be passed to the body of the
function.

The semantics of application is more complex. It first duplicates the free-variable product inside the 
context (using $\ident{map}_{\cclrd{r}}$ and duplication). Then it splits this context using 
$\ident{split}_{\cclrd{r}, \cclrd{s} \cpar \cclrd{t}}$. The two contexts contain the same variables
(as required by sub-expressions $e_1$ and $e_2$), but different coeffect annotations. The first
context (with index $\cclrd{r}$) is used to evaluate $e_1$, resulting in a function 
$\ctyp{\cclrd{t}}{\tau_1} \rightarrow \tau_2$. To obtain the result, we compose this with a function
created by applying $\ident{cobind}_{\cclrd{s}, \cclrd{t}}$ on the semantics of sub-expression
$e_2$, which is of type $\ctyp{\cclrd{s}\cseq\cclrd{t}}{\sigma_1 \times \ldots \times \sigma_n}
\rightarrow \ctyp{\cclrd{t}}{\tau_1}$.

Finally, constants (\emph{const}) are modelled by a global dictionary $\delta$ and sub-coeffecting
is interpreted by dropping additional context from the provided context $\ctx$ using 
$\ident{lift}_{\cclrd{r}, \cclrd{r'}}$ and providing it to the semantics of the assumption.

\paragraph{Properties.}

The categorical semantics can be used to embed context-dependent computations in functional 
programming languages, similarly to how monads provide a way of embedding effectful computations.
More importantly, it also provides validation for the design of the type system developed in 
Section~\ref{sec:flat-calculus-types}. As stated in the following theorem, the annotations in 
the type system match those of the semantic functions.

\vspace{2em}

\begin{theorem}[Correspondence]
\label{thm:flat-correspondence}
In all of the typing rules of the flat coeffect system, the context annotations $\cclrd{r}$ of typing 
judgements $\coctx{\Gamma}{\cclrd{r}} \vdash e : \tau$ and function types $\tau_1 \xrightarrow{\cclrd{r}} \tau_2$ 
correspond to the indices of mappings $\ctyp{\cclrd{r}}{}$ in the corresponding semantic function defined 
by $\sem{\coctx{\Gamma}{\cclrd{r}} \vdash e : \tau}$.
\end{theorem}
\begin{proof}
By analysis of the semantic rules in Figure~\ref{fig:flat-semantics}.
\end{proof}

\noindent
Thanks to indexing, the statement of the theorem is significantly stronger than for a 
non-indexed system, because it provides the justification for our choice of indices in the typing
rules. In particular, we can see that the annotations follow from the annotations on primitive 
functions that define the semantics. Also, each function defining the semantics uses a distinct 
operation of the coeffect algebra and so the type system is the most general possible definition
(within the comonadic framework we use).

Although the notion of indexed comonads presented in this section is novel and interesting 
in its own, the main reason for introducing it is to motivate the flat coeffect type system.
This is captured by the Theorem~\ref{thm:flat-correspondence}. In the next section, we shift our
focus from (categorical) semantics to syntactic properties of the calculus.

% ==================================================================================================
%                                                                   
%       ###                  #                    #       #          
%      #   #                 #                    #                  
%      #      #   #  # ##   ####    ###    ###   ####    ##     ###  
%       ###   #   #  ##  #   #         #  #   #   #       #    #   # 
%          #  #  ##  #   #   #      ####  #       #       #    #     
%      #   #   ## #  #   #   #  #  #   #  #   #   #  #    #    #   # 
%       ###       #  #   #    ##    ####   ###     ##    ###    ###  
%             #   #                                                  
%              ###                                                   
%
% ==================================================================================================

\section{Syntactic equational theory}
\label{sec:flat-syntax}

Each of the concrete coeffect calculi discussed in this chapter has a different notion of context,
much like various effectful languages have different notions of effects (such as exceptions or
mutable state). However, in all of the calculi, the context has a number of common properties that 
are captured by the \emph{flat coeffect algebra}. This means that there are equational properties
that hold for all of the coeffect systems. Further properties hold for systems where the context
satisfies additional properties.

In this section, we look at such shared syntactic properties. This accompanies the previous section, 
which provided a \emph{semantic} justification for the axioms of coeffect algebra with a 
\emph{syntactic} justification. Operationally, this section can also be viewed as providing a 
pathway to an operational semantics for two of our systems (implicit parameters and liveness),
which can be based on syntactic substitution. As we discuss later, the notion of context for
data-flow is more complex.

% --------------------------------------------------------------------------------------------------

\subsection{Syntactic properties}
\label{sec:flat-syntax-props}

Before discussing the syntactic properties of general coeffect calculus formally, it should be 
clarified what is meant by providing ``pathway to operational semantics'' in this section. We do
that by contrasting syntactic properties of coeffect systems with more familiar effect systems.
Assuming $\subst{e_1}{x}{e_2}$ is a standard capture-avoiding syntactic substitution, the following
equations define four syntactic reductions on the terms:
%
\begin{equation*}
\begin{array}{rllcl}
(\lambda x.e_1)~e_2 &{\reduce_{\ident{cbn}}}& \subst{e_1}{x}{e_2}   &&(\textit{call-by-name})\\
(\lambda x.e_1)~v   &{\reduce_{\ident{cbv}}}& \subst{e_1}{x}{v}     &\quad&(\textit{call-by-value})\\
(\lambda x.e_1)~e_2 &{\reduce_{\ident{seq}}}& \kvd{glet}~x=e_2~\kvd{in}~e_1  &&(\textit{internalized sequencing})\\
e &{\reduce_\eta}& \lambda x.e~x                                    &&(\textit{$\eta$-expansion})
\end{array}
\end{equation*}
%
The rules capture syntactic reductions that can be performed in a general calculus, without any 
knowledge of the specific notion of context. If the reductions preserve the type of the expression
(type preservation), then operational semantics can be defined as a repeated application of the 
rules, until a specified normal form (\ie~a value) is reached.

The first two equations define well-known call-by-name and call-by-value reductions. The \kvd{glet} 
reduction (inspired by Filinski \cite{monads-inaction}) models a system where program is reduced to 
a sequence of primitive operations, sequentially composed using the special \kvd{glet} construct. 
In the rest of the section, we briefly outline the interpretation of the four 
rules and then we focus on call-by-value (Section~\ref{sec:flat-syntax-cbv}) and call-by-name 
(Section~\ref{sec:flat-syntax-cbn}) in more details.

The focus of our work is on the general coeffect system and so we do not discuss the operational
semantics of the specific notions of context. However, some work in that area has been done
by Brunel et al. \cite{coeffects-quantitative}.

\paragraph{Call-by-name.} 
In call-by-name, the argument is syntactically substituted for all occurrences of a variable. It
can be used as the basis for operational semantics of purely functional languages. However, using
the rule in effectful languages breaks the \emph{type preservation} property. For example, consider
a language with effect system where functions are annotated with sets of effects such as $\{ \ident{write} \}$.
A function $\lambda x.y$ is effect-free:
%
\begin{equation*}
y\!:\!\tau_1 \vdash \lambda x.y : \tau_1 \xrightarrow{\emptyset} \tau_2 \;\&\; \emptyset
\end{equation*}
%
Substituting an expression $e$ with effects $\{ \ident{write} \}$ for $y$ changes the type of 
the function by adding latent effects (without changing the immediate effects):
%
\begin{equation*}
\vdash \lambda x.e : \tau_1 \xrightarrow{ \{ \ident{write} \} } \tau_2 \;\&\; \emptyset
\end{equation*}
%
Similarly to effect systems, substituting a context-dependent computation $e$ for a variable $y$ can
add latent coeffects to the function type. However, this is not the case for \emph{all} flat coeffect
calculi. For example, call-by-name reduction preserves types and coeffects for the implicit 
parameters system. This means that certain coeffect systems support call-by-name evaluation strategy
and could be embedded in purely functional language (such as Haskell).

\paragraph{Call-by-value.}

The call-by-value evaluation strategy is often used by effectful languages. Here, an argument is
first reduced to a \emph{value} before performing the substitution. In effectful languages, 
value is defined syntactically. For example, in the \emph{Effect} language \cite{monads-effects-marriage},
values are identifiers $x$ or functions $(\lambda x.e)$.

The notion of \emph{value} in coeffect systems differs from the usual syntactic understanding.
A function $(\lambda x.e)$ does not defer all context requirements of the body $e$ and may have
immediate context requirements. Thus we say that $e$ is a value if it is a value in the usual
sense \emph{and} has not immediate context requirements. We define this formally in 
Section~\ref{sec:flat-syntax-cbv}.

The call-by-value evaluation strategy preserves typing for a wide range of flat coeffect calculi, 
including all our three examples. However, it is rather weak -- in order to use it, the concrete 
semantics needs to provide a way for reducing context-dependent term $\coctx{\Gamma}{\cclrd{r}} \vdash e : \tau$ 
to a value, \ie~a term $\coctx{\Gamma}{\cunit} \vdash e' : \tau$ with no context requirements.

\paragraph{Internalized sequencing.}

The (\emph{internalized sequencing}) rule captures an operational semantics where the language
provides a construct representing \emph{sequential composition} and expressions can be reduced to 
a normal form, consisting of a sequence of context-dependent operations. We choose to write the 
sequencing operation as \kvd{glet} to emphasize that this is a separate primitive and not an 
ordinary syntactic \kvd{let}. The normal form looks as follows:
%
\begin{equation*}
\kvd{glet}~x_1 = e_1~\kvd{in}~\ldots~\kvd{glet}~x_n = e_n~\kvd{in}~e
\end{equation*}
%
Here, the expressions $e_1, \ldots, e_n, e$ do not contain further nested \kvd{glet} constructs.
This evaluation strategy provides context-dependent counterpart to operational view of monads
developed by Filinsky \cite{monads-inaction}. We discuss how expressions reduce to the normal form
in Section~\ref{sec:flat-syntax-norm}

The (\emph{internalized sequencing}) rule is useful when defining a concrete semantics for 
a language that provides constructs for explicitly providing the context. For example, consider
a language with implicit parameters where a parameter is defined by $\kvd{letdyn}~\ident{?p}=e_1~\kvd{in}~e_2$.
Semantics for such language would provide a reduction rule:
%
\begin{equation*}
\kvd{letdyn}~\ident{?p}=e_1~\kvd{in}~(\kvd{glet}~x = \ident{?p}~\kvd{in}~e)
 ~~\leadsto~~ \kvd{let}~x = e_1~\kvd{in}~e
\end{equation*}
%
The \kvd{glet} construct provides a way of sequentialising the context-require\-ments so
that they can be discharged by matching constructs providing the required context. As discussed
tfearlier, we focus on the general case and so we discuss when a flat coeffect calculus supports
this form of evaluation (rather than looking at semantics for concrete systems).

\paragraph{Local soundness and completeness.}
Two desirable properties of calculi, coined by Pfenning and Davies \cite{logic-modal-reconstruction},
are \emph{local soundness} and \emph{local completeness}. They guarantee that the rules which 
introduce a function arrow (lambda abstraction) and eliminate it (application) are sufficiently 
strong, but not too strong.

The local soundness property is witnessed by (call-by-name) $\beta$-reduction, which we discussed
already. The local completeness is witnessed by the $\eta$-expansion rule. We discuss the flat
coeffect algebra conditions under which the reduction holds in Section~\ref{sec:flat-syntax-cbn}.

% --------------------------------------------------------------------------------------------------

\subsection{Call-by-value evaluation}
\label{sec:flat-syntax-cbv}

As discussed in the previous section, call-by-value reduction can be used for most flat coeffect
calculi, but it provides a very weak general model \ie~the hard work of reducing context-dependent
term to a \emph{value} has to be provided for each system. Syntactic values are defined
in the usual way:
%
\begin{equation*}
\begin{array}{llll}
v\in\mathit{SynVal}   & v &::=& x \sep c \sep (\lambda x.e)\\
n\in\mathit{NonVal}   & n &::=& e_1\;e_2 \sep \kvd{let}~x=e_1~\kvd{in}~e_2\\
e\in\mathit{Expr}     & e &::=& v \sep n 
\end{array}
\end{equation*}
%
The syntactic form \emph{SynVal} captures syntactic values, but a context-dependency-free value in 
coeffect calculus cannot be defined purely syntactically, because a function $(\lambda x.e)$ does
not automatically defer all context requirements.

\begin{definition} 
An expression $e$ is a \emph{value}, written as $\textit{val}(e)$ if it is a syntactic value,
\ie~$e \in \mathit{SynVal}$ and it has no context-dependencies, \ie~$\coctx{\Gamma}{ \cunit } \vdash e : \tau$.
\end{definition}

\noindent
The call-by-value substitution substitutes a value, with context requirements $\cunit$, for a 
variable, whose access is also annotated with $\cunit$. Thus, it does not affect the type and
context requirements of the term:

\begin{lemma}[Call-by-value substitution]
\label{thm:flat-subst-cbv}
In a flat coeffect calculus with a coeffect algebra $(\C, \cseq, \cpar, \czip, \cunit, \czero, \cleq)$,
given a value $\coctx{\Gamma}{\cunit} \vdash v : \sigma$ and an expression
$\coctx{\Gamma,  x\!:\!\sigma}{ \cclrd{r}  } \vdash e : \tau$, then substituting $v$ for $x$ does not
change the type and context requirements, that is $\coctx{\Gamma}{ \cclrd{r} } \vdash \subst{e}{x}{v} : \tau$.
\end{lemma}
\begin{proof}
By induction over the type derivation, using the fact that $x$ and $v$ are annotated
with $\cunit$ and that $\Gamma$ is treated as a set in the flat calculus.
\end{proof}

\noindent
The substitution lemma~\ref{thm:flat-subst-cbv} holds for all flat coeffect systems. However, 
proving that call-by-value reduction preserves typing requires an additional constraint on the
flat coeffect algebra, which relates the $\czip$ and $\cpar$ operations. This is captured
by the (\emph{approximation}) property:

\begin{equation*}
\cclrd{r}\,\czip\,\cclrd{t}~\;\cleq~\;\cclrd{r}\,\cpar\,\cclrd{t}
\hspace{10em}(\textit{approximation})
\end{equation*}
%
Intuitively, this specifies that the $\czip$ operation (splitting of context requirements) 
under-approximates the actual context capabilities while the $\cpar$ operation (combining of
context requirements) over-approximates the actual context requirements.

The property holds for the three systems we consider -- for implicit parameters, this is 
an equality; for liveness and data-flow (which both use a lattice), the greatest lower bound 
is smaller than the least upper bound.

Assuming $\reduce_{\ident{cbv}}$ is call-by-value reduction that reduces the term 
$(\lambda x.e)\,v$ to a term $\subst{e}{x}{v}$, the soundness theorem is
stated as follows:

\begin{theorem}[Call-by-value soundness]
\label{thm:cbv-reduction}
In a flat coeffect system with the (\emph{approxi\-mation}) property, 
if $\coctx{\Gamma}{\cclrd{r}} \vdash e : \tau$ 
and $e \reduce_{\ident{cbv}} e'$ then $\coctx{\Gamma}{\cclrd{r}} \vdash e' : \tau$.
\end{theorem}
\begin{proof}
Consider the typing derivation for the term $(\lambda x.e)\,v$ before reduction:
\begin{equation*}
\inference
  { \inference
  { \inference
      {\coctx{\Gamma, x:\tau_1}{\cclrd{r}\;\czip\;\cclrd{t}} \vdash e : \tau_2}
      {\coctx{\Gamma}{\cclrd{r}} \vdash \lambda x.e : \tau_1 \xrightarrow{\cclrd{t}} \tau_2 } &
   \coctx{\Gamma}{\cunit} \vdash v : \tau_1 }
  {\coctx{\Gamma}{\cclrd{r} \;\cpar\; (\cunit \,\cseq\, \cclrd{t})} \vdash (\lambda x.e)\,v : \tau_2} }
  {\coctx{\Gamma}{\cclrd{r} \;\cpar\; \cclrd{t}} \vdash (\lambda x.e)\,v : \tau_2}
\end{equation*}
The final step simplifies the coeffect annotation using the fact that $\cunit$ is a unit of $\cseq$.
From Lemma~\ref{thm:flat-subst-cbv}, $\subst{e}{x}{v}$ has the same coeffect annotation as $e$.
As $\cclrd{r}\,\czip\,\cclrd{t}\;\cleq\;\cclrd{r}\,\cpar\,\cclrd{t}$, we can apply sub-coeffecting:
%
\begin{equation*}
\tyrule{sub}
 { \coctx{\Gamma}{\cclrd{r}\;\czip\;\cclrd{t}} \vdash \subst{e}{x}{v} : \tau_2 }
 { \coctx{\Gamma}{\cclrd{r}\;\cpar\;\cclrd{t}} \vdash \subst{e}{x}{v} : \tau_2 } 
\end{equation*}
%
Comparing the final conclusions of the above two typing derivations shows that 
the reduction preserves type and coeffect annotation.
\end{proof}

% --------------------------------------------------------------------------------------------------

\subsection{Call-by-name evaluation}
\label{sec:flat-syntax-cbn}

When reducing the expression $(\lambda x.e_1)~e_2$ using the call-by-name strategy, the 
sub-expression $e_2$ is substituted for all occurrences of the variable $v$ in an expression $e_1$. 
As discussed in Section~\ref{sec:flat-syntax-props}, the call-by-name strategy does not \emph{in 
general} preserve the type of a terms in coeffect calculi, but it does preserve the typing in two 
interesting cases.

The typing is preserved for different reasons in two of our systems, so we briefly review the 
concrete examples. Then, we prove the substitution lemma for two special cases of flat coeffects
(Lemma~\ref{thm:cbn-substitution-top} and Lemma~\ref{thm:cbn-substitution-bot}) and finally, we
state the conditions under which typing preservation hold for flat coeffect calculi 
(Theorem~\ref{thm:cbn-flat}).

\paragraph{Data-flow.} 
The type preservation property does not hold for data-flow. This case is similar to the example
shown earlier with effectful computations. As a minimal example, consider the substitution of 
a context-dependent expression $\kvd{prev}~z$ for a variable $y$ in a function $\lambda x.y$:
%
\begin{equation*}
\begin{array}{rclcl}
 \coctx{y\!:\!\tau_1,z\!:\!\tau_1}{0} &\narrow{\vdash}& \lambda x.y : \tau_1 \xrightarrow{0} \tau_2 & & (\textnormal{before}) \\
 \coctx{z\!:\!\tau_1}{1} &\narrow{\vdash}& \lambda x.\kvd{prev}~z : \tau_1 \xrightarrow{1} \tau_2 & & (\textnormal{after}) \\
\end{array}
\end{equation*}
%
After the substitution, the coeffect of the body is $1$. The rule for lambda abstraction requires
that $1=\mathit{min}(r,s)$ and so the least solution is to set both $r,s$ to $1$. The substitution
this affects the coeffects attached both to the function type and the overall context.

Semantically, the coeffect over-approximates the actual requirements -- at run-time, the code does not 
actually access a previous value of the argument $x$. This cannot be caputred by a flat coeffect 
system, but it can be captured using the structural system discussed in Chapter~\ref{ch:structural}.

\paragraph{Implicit parameters.} In data-flow, there is no typing for the resulting expression that
preserves the type of the function. However, this is not the case for all systems. Consider substituting
an implicit parameter access $\ident{?p}$ for a free variable $y$ under a lambda:
%
\begin{equation*}
\begin{array}{rclcl}
 \coctx{y\!:\!\tau_1}{\emptyset} &\narrow{\vdash}& \lambda x.y : \tau_1 \xrightarrow{\emptyset} \tau_2 & & (\textnormal{before}) \\
 \coctx{\emptyset}{ \{\ident{?p}\} } &\narrow{\vdash}& \lambda x.\ident{?p} : \tau_1 \xrightarrow{\emptyset} \tau_2 & & (\textnormal{after}) \\
\end{array}
\end{equation*}
%
The above shows one possible typing of the body -- one that does not change the coeffects of the 
function type and attaches all additional coeffects (implicit parameters) to the context. In case
of implicit parameters (and, more generally, systems with set-like annotations) this is always 
possible.

\paragraph{Liveness.} In liveness, the type preservation also holds, but for a different reason. Consider
substituting an arbitrary expression $e$ of type $\tau_1$ with coeffects $\cclrd{r}$ for a variable $y$:
%
\begin{equation*}
\begin{array}{rclcl}
 \coctx{y\!:\!\tau_1}{\ident{L}} &\narrow{\vdash}& \lambda x.y : \tau_1 \xrightarrow{\ident{L}} \tau_2 & & (\textnormal{before}) \\
 \coctx{\emptyset}{\ident{L}} &\narrow{\vdash}& \lambda x.e : \tau_1 \xrightarrow{\ident{L}} \tau_2 & & (\textnormal{after}) \\
\end{array}
\end{equation*}
%
In the original expression, both the overall context and the function type are annotated with \ident{L}, 
because the body contains a variable access. An expression $e$ can always be treated as being annotated
with $\ident{L}$ (because $\ident{L}$ is the top element of the lattice) and so we can also treat $e$
as being annotated with coeffects $\ident{L}$. As a result, substitution does not change the coeffect.

\paragraph{Reduction theorem.}
The above examples (implicit parameters and liveness) demonstrate two particular kinds of coeffect 
algebra for which typing preservation holds. Proving the type preservation separately provides 
more insight into how the systems work. We consider the two cases separately, but find a more general
formulation for both of them.

\begin{definition}
We call a flat coeffect algebra \emph{top-pointed} if $\cunit$ is the greatest (top) coeffect scalar
($\forall r \in \C \;.\; r \,\cleq\, \cunit $) and \emph{bottom-pointed} if it is the smallest (bottom) 
element ($\forall r \in \C \;.\; r \,\cgeq\, \cunit $).
\end{definition}

\noindent
Liveness is an example of top-pointed coeffects as variables are annotated with 
$\ident{L}$ and $\ident{D} \leq \ident{L}$, while implicit parameters and data-flow are examples
of bottom-pointed coeffects. For top-pointed flat coeffects, the substitution lemma holds without 
additional requirements:

\begin{lemma}[Top-pointed substitution]
\label{thm:cbn-substitution-top}
In a top-pointed flat coeffect calculus with an aglebra $(\C, \cseq, \cpar, \czip, \cunit, \czero, \cleq)$,
substituting an expression $e_s$ with arbitrary coeffects $\cclrd{s}$ for a variable $x$ in $e_r$ does
not change the coeffects of $e_r$:
%
\begin{equation*}
\begin{array}{l}
 \coctx{\Gamma}{\cclrd{s}} \vdash e_s : \tau_s \;\; \wedge \;\; 
   \coctx{\Gamma_1,  x : \tau_s, \Gamma_2}{ \cclrd{r}  } \vdash e_r : \tau_r\\
 \quad \Rightarrow \;\; \coctx{\Gamma_1,\Gamma,\Gamma_2}{ \cclrd{r} } \vdash \subst{e_r}{x}{e_s} : \tau_r
\end{array}
\end{equation*}
\end{lemma}

\begin{proof}
Using sub-coeffecting ($\cclrd{s} \,\cleq\, \cunit$) and a variation of Lemma~\ref{thm:flat-subst-cbv}.
\end{proof}
%
\noindent
As variables are annotated with the top element $\cunit$, we can substitute the term $e_s$ 
for any variable and use sub-coeffecting to get the original typing (because
$\cclrd{s} \,\cleq\, \cunit$).

In a bottom pointed coeffect system, substituting $e$ for $x$ increases the context 
requirements. However, if the system satisfies the strong condition that $\czip = \cseq = \cpar$ 
then the context requirements arising from the substitution can be associated with the context
$\Gamma$, leaving the context requirements of a function value unchanged. As a result, substitution 
does not break soundness as in effect systems. The requirement $\czip = \cseq = \cpar$ holds for 
our implicit parameters example (all three operators are a set union) and for other set-like 
coeffects. It allows the following substitution lemma:

\begin{lemma}[Bottom-pointed substitution]
\label{thm:cbn-substitution-bot}
In a bottom-pointed flat coeffect calculus with an algebra $(\C, \cseq, \cpar, \czip, \cunit, \czero, \cleq)$ 
where $\czip = \cseq = \cpar$ is an idempotent and commutative operation`  ' and
$\cclrd{r}\,\cleq\,\cclrd{r'} \Rightarrow \forall\cclrd{s}.\cclrd{r}\,\cseq\,\cclrd{s}\;\cleq\;\cclrd{r'}\,\cseq\,\cclrd{s}$ then:
%
\begin{equation*}
\begin{array}{l}
\coctx{\Gamma}{\cclrd{s}} \vdash e_s : \tau_s \;\; \wedge \;\; 
\coctx{\Gamma_1,  x : \tau_s, \Gamma_2}{ \cclrd{r}  } \vdash e_r : \tau_r\\
\quad \Rightarrow \;\; \coctx{\Gamma_1,\Gamma,\Gamma_2}{ \cclrd{r} \,\cseq\, \cclrd{s} } \vdash \subst{e_r}{x}{e_s} : \tau_r
\end{array}
\end{equation*}

\end{lemma}
\begin{proof}
By induction over $\vdash$, using the idempotent, commutative monoid structure to keep $\cclrd{s}$ with
the free-variable context. See Appendix~\ref{sec:appendix-flat-cbn}.
\end{proof}

\noindent
The flat system discussed here is \emph{flexible enough} to let us always re-associate new context 
requirements (arising from the substitution) with the free-vari\-able context. In contrast, the 
structural system discussed in Chapter~\ref{ch:structural} is \emph{precise enough} to keep the 
coeffects associated with individual variables, thus preserving typing in a complementary way.

The two substitution lemmas discussed above show that the call-by-name evaluation strategy can be 
used for certain coeffect calculi, including liveness and implicit parameters. Assuming
$\reduce_{\ident{cbn}}$ is the standard call-by-name reduction, the following soundness
theorem holds:

\begin{theorem}[Call-by-name soundness]
\label{thm:cbn-flat}
In a coeffect system that satisfies the conditions for Lemma~\ref{thm:cbn-substitution-top} or
Lemma~\ref{thm:cbn-substitution-bot}, if $\coctx{\Gamma}{\cclrd{r}} \vdash e : \tau$ 
and $e \rightarrow_{\ident{cbn}} e'$ then it is also the case that $\coctx{\Gamma}{\cclrd{r}} \vdash e' : \tau$.
\end{theorem}
\begin{proof}

For top-pointed coeffect algebra (using Lemma~\ref{thm:cbn-substitution-top}), the proof is similar
to the one in Theorem~\ref{thm:cbv-reduction}, using the facts that $\cclrd{s} \,\cleq\, \cunit$
and $r \,\czip\, t \;=\; r \,\cpar\, t$.
For bottom-pointed coeffect algebra, consider the typing derivation for the term
$(\lambda x.e_r)\,e_s$ before reduction:
%
\begin{equation*}
\inference
  { \inference
      {\coctx{\Gamma, x:\tau_s}{\cclrd{r}} \vdash e_r : \tau_r}
      {\coctx{\Gamma}{\cclrd{r}} \vdash \lambda x.e_r : \tau_s \xrightarrow{\cclrd{r}} \tau_r } &
   \coctx{\Gamma}{\cclrd{s}} \vdash e_s : \tau_s }
  {\coctx{\Gamma}{\cclrd{r} \;\cpar\; (\cclrd{s} \,\cseq\, \cclrd{r})} \vdash (\lambda x.e_r)\,e_s : \tau_r }
\end{equation*}
%
The derivation uses the idempotence of $\czip$ in the first step, followed by the 
(\emph{app}) rule. The type of the term after substitution, using Lemma~\ref{thm:cbn-substitution-bot} is:
%
\begin{equation*}
\inference
  { \coctx{\Gamma, x:\tau_s}{\cclrd{r}} \vdash e_r : \tau_r & 
    \coctx{\Gamma}{\cclrd{s}} \vdash e_s : \tau_s }
  { \coctx{\Gamma, x:\tau_r}{\cclrd{r}\;\cseq\;\cclrd{s}} \vdash \subst{e_r}{x}{e_s} : \tau_s }
\end{equation*}
%
From the assumptions of Lemma~\ref{thm:cbn-substitution-bot}, we know that $\cseq\,=\,\cpar$
and the operation is idempotent, so trivially:
$\cclrd{r}\;\cseq\;\cclrd{s} = \cclrd{r} \;\cpar\; (\cclrd{s} \,\cseq\, \cclrd{r})$
\end{proof}

\paragraph{Expansion theorem.}
The $\eta$-expansion (local completeness) is similar to $\beta$-reduction (local soundness) in that
it holds for some flat coeffect systems, but not for all. Out of the examples we discuss, it 
holds for implicit parameters, but does not hold for liveness and data-flow.

Recall that $\eta$-expansion turns $e$ into $\lambda x.e~x$. In the case of liveness, the 
expression $e$ may require no variables (both immediate and latent coeffects are marked as
$\ident{D}$). However, the resulting expression $\lambda x.e~x$ accesses a variable, 
marking the context and function argument as live. In case of data-flow, the immediate coeffects 
are made larger by the lambda abstraction -- the context requirements of the function value are 
imposed on the declaration site of the new lambda abstraction. We remedy this limitation in 
the next chapter.

However, $\eta$-expansion preserves the type for implicit parameters and, more generally,
for any flat coeffect algebra where $\cpar \,=\, \czip$. Assuming $\rightarrow_\eta$ is the 
standard $\eta$-reduction:

\begin{theorem}[Soundness of $\eta$-expansion]
In a bottom-pointed flat coeffect calculus with an algebra $(\C, \cseq, \cpar, \czip, \cunit, \czero, \cleq)$ 
where $\czip = \cpar$, if $\coctx{\Gamma}{\cclrd{r}} \vdash e : \tau_1 \xrightarrow{\cclrd{s}} \tau_2$ 
and $e \rightarrow_\eta e'$ then $\coctx{\Gamma}{\cclrd{r}} \vdash e' : \tau_1 \xrightarrow{\cclrd{s}} \tau_2 $.
\end{theorem}
\begin{proof}
The following derivation shows that $\lambda x.f~x$ has the same type as $f$:
\begin{equation*}
\inference
  { \inference
    { \coctx{\Gamma}{\cclrd{r}} \vdash f : \tau_1 \xrightarrow{\cclrd{s}} \tau_2 &
      \coctx{x:\tau_1}{\cunit} \vdash x:\tau_1 }
    { \coctx{\Gamma, x:\tau_1}{\cclrd{r} \;\cpar\; (\cunit \,\cseq\, \cclrd{s}) } \vdash f~x : \tau_2 } }
  { \inference
    { \inference
      { \coctx{\Gamma, x:\tau_1}{\cclrd{r} \;\cpar\; \cclrd{s} } \vdash f~x : \tau_2 }
      { \coctx{\Gamma, x:\tau_1}{\cclrd{r} \;\czip\; \cclrd{s} } \vdash f~x : \tau_2 } }
    { \coctx{\Gamma}{ \cclrd{r} } \vdash \lambda x.f~x : \tau_1 \xrightarrow{\cclrd{s}} \tau_2 } }
\end{equation*}
%
The derivation starts with the expression $e$ and derives the type for $\lambda x.e~x$. The
application yields context requirements $\cclrd{r}\,\cpar\,\cclrd{s}$. In order to recover the
original typing, this must be equal to $\cclrd{r}\,\czip\,\cclrd{s}$. Note that the derivation
is showing just one possible typing -- the expression $\lambda x.e~x$ has other types -- but
this is sufficient for showing type preservation.
\end{proof}

\noindent
In summary, flat coeffect calculi do not \emph{in general} permit call-by-name evaluation, but
there are several cases where call-by-name evaluation can be used. Among the examples we discuss,
these include liveness and implicit parameters. Moreover, for implicit parameters (and more
generally, any set-like flat coeffect algebra), the $\eta$-expansion holds as well, giving us
both local soundness and local completeness as coined by Pfenning and Davies \cite{logic-modal-reconstruction}.

% --------------------------------------------------------------------------------------------------

\subsection{Internalized sequencing}
\label{sec:flat-syntax-norm}

The call-by-value and call-by-name evaluation strategies discussed in the previous section are
the most important techniques for defining equational theory of flat coeffects. In this section, 
we briefly discuss another approach that follows the style of generic operational semantics 
designed by Filinski \cite{monads-inaction} for effectful computations.

The idea is to embed sequencing of context-dependent (effectful) computations as an explicit 
construct into the language and define a \emph{normal form} that consists of sequentially composed 
context-dependent expressions. The reduction to the normal form is generic for all coeffect systems 
(satisfying certain conditions), while the reduction of the normal form is provided by each 
concrete coeffect system. The extended language with the \kvd{glet} construct and its typing is 
defined as follows:
%
\begin{equation*}
\begin{array}{rcl}
e &::=& x \sep \lambda x.e \sep e_1~e_2 \sep \kvd{let}~x = e_1~\kvd{in}~e_2 \\
n &::=& e \sep \kvd{glet}~x=e~\kvd{in}~n\\
\tau &::=& T \sep \tau_1 \xrightarrow{\cclrd{r}} \tau_2
\end{array}
\end{equation*}
\begin{equation*}
\tyrule{glet}
  { \coctx{\Gamma}{\cclrd{s}} \vdash e_1 : \tau_1 &
    \coctx{\Gamma, x\!:\!\tau_1}{\cclrd{r}} \vdash e_2 : \tau_2 }
  { \coctx{\Gamma}{\cclrd{r}\;\cpar\;({\cclrd{s}\,\cseq\,\cclrd{r})}} \vdash \kvd{glet}~x=e_1~\kvd{in}~e_2 : \tau_2 }
\end{equation*}

The newly introduced syntactic form $n$ represents a normal form. The construct $\kvd{glet}~x=e~\kvd{in}~n$
models an explicit sequencing of an expression $e$, followed by an expression $n$. Note that \kvd{glet} can 
only contain further \kvd{glet} constructs in the body, but not in the argument. The typing of the \kvd{glet}
is the same as the typing of ordinary \kvd{let} construct -- as discussed in Section~\ref{sec:flat-calculus-types},
for the examples we consider, this gives a \emph{more precise} typing than the typing of the
term $(\lambda x.n)~e$.

The reduction rules that produce a normal form are shown in Figure~\ref{fig:flat-normal-form}. The
(\emph{eval}) rule introduces \kvd{glet} by reducing a redex $(\lambda x.e_2)~e_1$ to an expression
representing explicit sequencing of $e_1$ and $e_2$. The remaining two rules specify how \kvd{glet}
distributes across other constructs (\kvd{glet} and application). In (\emph{glet-app}), the \kvd{glet}
construct appearing inside an application is lifted to the top-level; similarly (\emph{glet-glet}) 
lifts a \kvd{glet} construct nested in the argument of another \kvd{glet}. The rules provide a way
for turning an expression into a normal form.

As with \emph{call-by-value} and \emph{call-by-name} strategies, the \emph{internalized sequencing}
strategy can only be used with coeffect systems satisfying certain conditions. The general form of
the conditions is summarized in Appendix~?. Here, we briefly consider our three examples.

% --------------------------------------------------------------------------------------------------

\begin{figure}[t]

\begin{equation*}
\tag{\emph{eval}}
(\lambda x.e_2)~e_1 ~~\leadsto~~ \kvd{glet}~x=e_1~\kvd{in}~e_2
\end{equation*}

\begin{equation*}
\tag{\emph{glet-app}}
\begin{array}{l}
  (\kvd{glet}~x=e_1~\kvd{in}~e_2)~e_3  ~~\leadsto~~\\ \qquad \kvd{glet}~x=e_1~\kvd{in}~(e_2~e_3)
\end{array}
\qquad x \notin \textit{fv}(e_3)\qquad
\end{equation*}

\begin{equation*}
\tag{\emph{glet-glet}}
\begin{array}{l}
  \kvd{glet}~x_2=(\kvd{glet}~x_1 = e_1~\kvd{in}~e_2)~\kvd{in}~e_3  ~~\leadsto~~\\
  \qquad \kvd{glet}~x_1=e_1~\kvd{in}~(\kvd{glet}~x_2=e_2~\kvd{in}~e_3)
\end{array}
\qquad x_1 \notin \textit{fv}(e_3)\qquad
\end{equation*}

\figcaption{Reduction to normal form.}
\label{fig:flat-normal-form}
\end{figure}

% --------------------------------------------------------------------------------------------------

\paragraph{Conditions.}
The (\emph{eval}) reduction can be safely applied for any flat coeffect calculus which satisfies the
condition that the typing of the \kvd{glet} expression (shown above) is equivalent or more precise
than typing of the expression $(\lambda x.e_2)~e_1$. More formally:

\begin{definition}
A flat coeffect algebra $(\C, \cseq, \cpar, \czip, \cunit, \czero, \cleq)$ is \emph{oriented} if for all
$\cclrd{s}, \cclrd{s_1}, \cclrd{s_2}, \cclrd{r} \in \C$ such that $\cclrd{s}=\cclrd{s_1}\,\czip\,\cclrd{s_2}$
it is the case that $\cclrd{s} \;\cseq\; (\cclrd{s} \,\cpar\, \cclrd{r}) \;\cleq\; \cclrd{s_1} \;\cseq\; (\cclrd{s_2} \,\cpar\, \cclrd{r})$.
\end{definition}

\noindent
As discussed in Section~\ref{sec:flat-calculus-let}, this condition is satisfied for all our
three examples, strictly for liveness and data-flow and by equality for implicit parameters.
For the (\emph{glet-app}) and (\emph{glet-glet}) rules, additional conditions arise from the
typing of the original and reduced expression. The results are summarized in the following two tables.

Assuming $\coctx{\Gamma}{\cclrd{r_1}} \vdash e_1 : \tau_1$ and
$\coctx{\Gamma}{\cclrd{r_2}} \vdash e_2 : \tau_3 \xrightarrow{\cclrd{s}} \tau_2$ and
$\coctx{\Gamma}{\cclrd{r_3}} \vdash e_3 : \tau_3$, the typings of the original and
reduced expressions in (\emph{glet-app}) rule are:

\begin{center}
\begin{tabular}{ | l | c | c | c |}
\hline
  & \footnotesize Before\hspace{1em} & \footnotesize After\hspace{1em}  & \footnotesize Satisfied\hspace{1em} \\ \hline
\hspace{-1em}{\footnotesize Parameters} 
  & $(\cclrd{r_2} \cup \cclrd{r_1}) \cup (\cclrd{r_3} \cup \cclrd{s}) $ 
  & $(\cclrd{r_2} \cup (\cclrd{r_3} \cup \cclrd{s})) \cup \cclrd{r_1} $ & \checkmark \\ \hline
\hspace{-1em}{\footnotesize Liveness} 
  & $\cclrd{r_2} \sqcap (\cclrd{r_3} \sqcup \cclrd{s})$ 
  & $\cclrd{r_2} \sqcap (\cclrd{r_3} \sqcup \cclrd{s})$ & \checkmark \\ \hline
\hspace{-1em}{\footnotesize Data-flow} 
  & $\mathit{max}(\cclrd{r_1} + \cclrd{r_2}, \cclrd{r_3}+\cclrd{s})$
  & $\cclrd{r_1}+\mathit{max}(\cclrd{r_2}, \cclrd{r_3} + \cclrd{s})$ & $\times$ \\ \hline
\end{tabular}
\end{center}

\noindent
Assuming $\coctx{\Gamma}{\cclrd{r_1}} \vdash e_1 : \tau_1$ and
$\coctx{\Gamma}{\cclrd{r_2}} \vdash e_2 : \tau_2$ and
$\coctx{\Gamma}{\cclrd{r_3}} \vdash e_3 : \tau_3$, the typings of the original and
reduced expressions in (\emph{glet-glet}) rule are:

\begin{center}
\begin{tabular}{ | l | c | c | c |}
\hline
  & \footnotesize Before\hspace{1em} & \footnotesize After\hspace{1em}  & \footnotesize Satisfied\hspace{1em} \\ \hline
\hspace{-1em}{\footnotesize Parameters} 
  & $\cclrd{r_3} + (\cclrd{r_2} + \cclrd{r_1}) $ 
  & $(\cclrd{r_3} + \cclrd{r_2}) + \cclrd{r_1}$ & \checkmark \\ \hline
\hspace{-1em}{\footnotesize Liveness} 
  & $\cclrd{r_3}$ 
  & $\cclrd{r_3}$ & \checkmark \\ \hline
\hspace{-1em}{\footnotesize Data-flow} 
  & $\cclrd{r_3} \cup (\cclrd{r_2} \cup \cclrd{r_1}) $ 
  & $(\cclrd{r_3} \cup \cclrd{r_2}) \cup \cclrd{r_1}$ & \checkmark \\ \hline
\end{tabular}
\end{center}

\noindent
This means that \emph{internalized sequencing} provides a basis for operational semantics of
liveness and implicit parameters, but (again) not for data-flow languages. For other coeffect 
systems, the conditions arising from the syntactic reductions have to be re-examined.

% ==================================================================================================
%
%      #####          #                            #                        
%      #              #                                                     
%      #      #   #  ####    ###   # ##    ###    ##     ###   # ##    ###  
%      ####    # #    #     #   #  ##  #  #        #    #   #  ##  #  #     
%      #        #     #     #####  #   #   ###     #    #   #  #   #   ###  
%      #       # #    #  #  #      #   #      #    #    #   #  #   #      # 
%      #####  #   #    ##    ###   #   #  ####    ###    ###   #   #  ####  
%
% ==================================================================================================

\section{Syntactic properties and extensions}
\label{sec:flat-exts}

The flat coeffect algebra introduced in Section~\ref{sec:flat-calculus} requires a number of laws.
The laws are required for three reasons -- to be able to define the categorical structure
in Section~\ref{sec:flat-semantics}, to prove equational properties in Section~\ref{sec:flat-syntax}
and finally, to guarantee intuitive syntactic properties for constructs such as 
$\lambda$-abstraction and pairs in context-aware calculi.

In this section, we look at the last point. We discuss what syntactic equivalences are permitted
by the properties of $\czip$ (Section~\ref{sec:flat-exts-lambda}) and we extend the calculus with 
pairs and units and discuss their syntactic properties (Section~\ref{sec:flat-exts-tup}). Before
doing that, the following section further develops subtyping relation for the calculus.

% --------------------------------------------------------------------------------------------------

\begin{figure}[t]

\begin{equation*}
\boxed{\coctx{\Gamma}{\cclrd{r}} \vdash e : \tau}
\end{equation*}

\begin{equation*}
\tyrule{typ}
  { \coctx{\Gamma}{\cclrd{r}} \vdash e : \tau & \tau <: \tau' }
  { \coctx{\Gamma}{\cclrd{r}} \vdash e : \tau' }
\quad
\end{equation*}
\begin{equation*}
\tyrule{sub}
  {\coctx{\Gamma}{\cclrd{r'}} \vdash e : \tau & \cclrd{r'} \cleq \cclrd{r} }
  {\coctx{\Gamma}{\cclrd{r}} \vdash e : \tau }
\end{equation*}

\begin{equation*}
\boxed{\tau <: \tau'}
\end{equation*}

\begin{equation*}
\tyrule{sub-trans}
  { \tau_1 <: \tau_2 & \tau_2 <: \tau_3 }
  { \tau_1 <: \tau_3  }
\end{equation*}
\begin{equation*}
\tyrule{sub-fun}
  { \tau_1' <: \tau_1 & \tau_2 <: \tau_2' & \cclrd{r'} \cgeq \cclrd{r} }
  { \tau_1 \xrightarrow{\cclrd{r}} \tau_2 <: \tau_1' \xrightarrow{\cclrd{r'}} \tau_2' }
\end{equation*}
\begin{equation*}
\tyrule{sub-refl}
  { }
  { \tau <: \tau }
\end{equation*}

\figcaption{Subtyping rules for flat coeffect calculus}
\label{fig:flat-types-sub}
\end{figure}

% --------------------------------------------------------------------------------------------------
~

~

\subsection{Subtyping for coeffects}
\label{sec:flat-exts-sub}

The typing rules discussed in Section~\ref{sec:flat-calculus-types} include sub-coeffecting rule
which makes it possible to treat an expression with smaller context requirements as an expression
with greater context requirements. In the corresponding categorical semantics, this means that
we can \emph{drop} some of the provided context. However, sub-coeffecting only affects the 
immediate coeffects attached to the free-variable context.

In Figure~\ref{fig:flat-types-sub}, we add sub-typing on function types, making it possible to treat
a function with smaller context requirements as a function with greater context requirements. 
The definition uses the standard reflexive and transitive $<:$ operator. As the (\emph{sub-fun})
shows, the function type is contra-variant in the input and co-variant in the output. The 
(\emph{typ}) rule allows using sub-typing on expressions in the coeffect calculus.

\paragraph{Semantics.} We follow the same approach as in Section~\ref{sec:flat-semantics} and use
categorical semantics to explain (and confirm) the design of the sub-typing rules. The semantics
of a judgement $\sem{\tau <: \tau'}$ is a function $\sem{\tau} \rightarrow \sem{\tau'}$. 
As shown in Figure~\ref{fig:flat-semantics-sub}, the seamntics of the sub-typing rule (\emph{typ}) 
then just composes the semantics of the original expression with the conversion produced by the 
semantics of the sub-typing judgement.

The rest of the Figure~\ref{fig:flat-semantics-sub} defines the semantics of the $<:$ operation. 
The reflexivity and transitivity are just the identity function and function composition,
respectively. The (\emph{sub-fun}) case is interesting -- recall that the semantics of functions:
%
\begin{equation*}
\sem{\tau_1' \xrightarrow{\cclrd{r'}} \tau_2'} ~=~ C^{\cclrd{r'}} \tau_1' \rightarrow \tau_2'
\end{equation*}
%
We build the transformation using an explicit lambda abstraction that takes the original function
$f$ as an argument. To build the required function, we first drop unnecessary context using 
$\ident{lift}_{\cclrd{r'}, \cclrd{r}}$ of type $C^{\cclrd{r'}} \tau_1' \rightarrow C^{\cclrd{r}} \tau_1'$,
then we use the $\ident{map}_{\cclrd{r}}$ function to transform the nested $\tau_1'$ to $\tau_1$. 
Finally we evaluate the original function $f$ and turn the resulting value of type $\tau_2$ into 
the required result of type $\tau_2'$.

% --------------------------------------------------------------------------------------------------

\begin{figure}[t]

\begin{equation*}
\begin{array}{lll}
\sem{\coctx{\Gamma}{\cclrd{r}} \vdash e : \tau' } = 
  \sem{\tau <: \tau'} \circ \sem{\coctx{\Gamma}{\cclrd{r}} \vdash e : \tau } &\qquad& (\emph{typ})
\\[1.75em]  
\sem{\tau <: \tau} = \ident{id} && (\emph{sub-refl})
\\[0.5em]
\sem{\tau_1 <: \tau_3} = \sem{\tau_2 <: \tau_3} \circ \sem{\tau_1 <: \tau_2} && (\emph{sub-trans})
\\[0.5em]
\sem{\tau_1 \xrightarrow{\cclrd{r}} \tau_2 <: \tau_1' \xrightarrow{\cclrd{r'}} \tau_2'} = \lambda f. && (\emph{sub-fun}) \\
\qquad\sem{\tau_2 <: \tau_2'} \circ f \circ \ident{map}_{\cclrd{r}}~\sem{\tau_1' <: \tau_1} 
  \circ \ident{lift}_{\cclrd{r'}, \cclrd{r}}
\end{array}
\end{equation*}

\figcaption{Semantics of subtyping for flat coeffects}
\label{fig:flat-semantics-sub}
\end{figure}

% --------------------------------------------------------------------------------------------------

\subsection{Properties of lambda abstraction}
\label{sec:flat-exts-lambda}

In Section~\ref{sec:flat-calculus-lambda}, we discussed how to reconcile two typings for
lambda abstraction -- for implicit parameters, the lambda function needs to split context 
requirements using $\cclrd{r} \cup \cclrd{s}$, but for data-flow and liveness it suffices to 
duplicate the requirement $\cclrd{r}$ of the body. We introduced the $\czip$ operation as 
a way of providing the additional abstraction.

In this section, we first identify coeffect calculi for which the simpler (duplicating) 
rule is sufficient. Then we look at syntactic transformations corresponding to other common 
properties of the $\czip$ operation.

\paragraph{Simplified abstraction.}
Recall that $(\C, \czip)$ is a band, that is, $\czip$  is idempotent and associative. The 
idempotence means that the context requirements of the body can be required from both the 
declaration site and the call site. Thus, the following (\emph{idabs}) typing is valid 
(for reference, it is shown side-by-side with the ordinary lambda abstraction rule):
%
\begin{equation*}
\tyrule{idabs}
  {\coctx{\Gamma, x:\tau_1}{\cclrd{r}} \vdash e : \tau_2}
  {\coctx{\Gamma}{\cclrd{r}} \vdash \lambda x.e : \tau_1 \xrightarrow{\cclrd{r}} \tau_2 }
\tyrule{abs}
  {\coctx{\Gamma, x:\tau_1}{\cclrd{r}\,\czip\,\cclrd{r}} \vdash e : \tau_2}
  {\coctx{\Gamma}{\cclrd{r}} \vdash \lambda x.e : \tau_1 \xrightarrow{\cclrd{r}} \tau_2 }
\end{equation*}
% 
To derive (\emph{idabs}), we use idempotence on the body annotation $\cclrd{r}\;=\;\cclrd{r}\;\czip\;\cclrd{r}$
and then use the standard (\emph{abs}) rule. So, (\emph{idabs}) follows from (\emph{abs}), 
but the other direction is not necessarily the case. The following condition identifies 
coeffect calculi where (\emph{abs}) can be derived from (\emph{idabs}).

\begin{definition}
A flat coeffect algebra $(\C, \cseq, \cpar, \czip, \cunit, \czero, \cleq)$ is \emph{strictly oriented} if for all
$\cclrd{s}, \cclrd{r} \in \C$ it is the case that $\cclrd{r} \,\czip\, \cclrd{s} \;\cleq\; \cclrd{r}$.
\end{definition}

\begin{remark}
\label{thm:flat-alt-abs}
For a flat coeffect calculus with a strictly oriented algebra, the standard (abs) rule can 
be derived from the (idfun) rule.
\end{remark}
\begin{proof}
The following derives the conclusion of (\emph{abs}) using (\emph{idabs}), sub-coeffecting, 
sub-typing and the fact that the algebra is \emph{stricly oriented}:

\begin{equation*}
\tyrule{typ}
  {\hspace{-5em} \tyrule{sub}
     {\hspace{-4em} \tyrule{idabs}  
        {\coctx{\Gamma, x:\tau_1}{\cclrd{r} \,\czip\, \cclrd{s}} \vdash e : \tau_2}
        {\coctx{\Gamma}{\cclrd{r} \,\czip\, \cclrd{s}} \vdash \lambda x.e : \tau_1 \xrightarrow{\cclrd{r} \,\czip\, \cclrd{s}} \tau_2 } }
     {\coctx{\Gamma}{\cclrd{r}} \vdash \lambda x.e : \tau_1 \xrightarrow{\cclrd{r} \,\czip\, \cclrd{s}} \tau_2} \;
           \textnormal{\footnotesize{($\cclrd{r} \;\cleq\; \cclrd{r} \,\czip\, \cclrd{s}$)}} \hspace{-5em} }
  {\coctx{\Gamma}{\cclrd{r}} \vdash \lambda x.e : \tau_1 \xrightarrow{\cclrd{s}} \tau_2} \; 
           \textnormal{\footnotesize{($\cclrd{r} \;\cleq\; \cclrd{r} \,\czip\, \cclrd{s}$)}}
\end{equation*}
\end{proof}

\noindent
The practical consequence of the Remark~\ref{thm:flat-alt-abs} is that, for strictly
oriented coeffect calculi (such as our liveness and data-flow computations), we can use
the (\emph{idabs}) rule and get an equivalent type system. This alternative formulation
removes the non-determinism of type checking that arises from the splitting of context 
requirements in the original (\emph{abs}) rule. Furthermore, for data-flow and liveness
the (\emph{idabs}) rule is more precise than (\emph{abs}).

\paragraph{Symmetry.}
The $\czip$ operation is idempotent and associative. In all of the three examples considered in 
this chapter, the operation is also \emph{symmetric}. To make our definitions more general, we
do not require this to be the case for \emph{all} flat coeffect systems. However, systems with
symmetric $\czip$ have the following property.

\begin{remark}
For a flat coeffect calculus such that $\cclrd{r}\,\czip\,\cclrd{s} = \cclrd{s}\,\czip\,\cclrd{r}$,
assuming that $\cclrd{r'}, \cclrd{s'}, \cclrd{t'}$ is a permutation of $\cclrd{r},\cclrd{s},\cclrd{t}$:
%
\begin{equation*}
\inference
  {\coctx{\Gamma, x:\tau_1, y:\tau_2}{\cclrd{r} \,\czip\, \cclrd{s} \,\czip\, \cclrd{t}} \vdash e : \tau_3}
  {\coctx{\Gamma}{\cclrd{r'}} \vdash \lambda x.\lambda y.e : 
      \tau_1 \xrightarrow{\cclrd{s'}} (\tau_2 \xrightarrow{\cclrd{t'}} \tau_3) }
\end{equation*} 
\end{remark}

\noindent
Intuitively, this means that the context requirements of a function with multiple arguments can be 
split arbitrarily between the declaration site and (multiple) call sites. In other words, it does
not matter how the context requirements are satisfied.

% --------------------------------------------------------------------------------------------------

\begin{figure}[t]
\begin{equation*}
\tyrule{pair}
  {\coctx{\Gamma}{\cclrd{r}} \vdash e_1 : \tau_1 & \coctx{\Gamma}{\cclrd{s}} \vdash e_2 : \tau_2}
  {\coctx{\cclrd{r}\,\cpar\,\cclrd{s}}{\Gamma} \vdash (e_1, e_2) : \tau_1 \times \tau_2 }
\end{equation*}
\begin{equation*}
\tyrule{proj}
  {\coctx{\Gamma}{\cclrd{r}} \vdash e : \tau_1 \times \tau_2 }
  {\coctx{\Gamma}{\cclrd{r}} \vdash \pi_i~e : \tau_{i} }
\end{equation*}
\begin{equation*}
\tyrule{unit}
  {}
  {\coctx{\Gamma}{\czero} \vdash () : \ident{unit} }
\end{equation*}

\figcaption{Typing rules for pairs and units}
\label{fig:flat-ext-types}
\end{figure}

% --------------------------------------------------------------------------------------------------

\subsection{Language with pairs and unit}
\label{sec:flat-exts-tup}

To show the key aspects of flat coeffect systems, the calculus introduced in Section~\ref{sec:flat-calculus} 
consists only of variables, abstraction, application and let binding. Here, we extend it 
with pairs and the unit value to sketch how it can be turned into a more complete programming 
language and to motivate the laws required about $\cpar$. The syntax of the language is extended as follows:
%
\begin{equation*}
\begin{array}{rcl}
e &::=& \ldots \sep () \sep e_1, e_2 \\
\tau &::=& \ldots \sep \ident{unit} \sep \tau \times \tau
\end{array}
\end{equation*}
%
The typing rules for pairs and the unit value are shown in Figure~\ref{fig:flat-ext-types}.
The unit value (\emph{unit}) is annotated with the $\czero$ coeffect (the same as other constants).
Pairs, created using the $(e_1, e_2)$ expression, are annotated with a coeffect that combines
the coeffects of the two sub-expressions using the \emph{pointwise} operator $\cpar$. The operator
models the case when the (same) available context is split and passed to two independent 
sub-expressions. Finally, the (\emph{proj}) rule is uninteresting, because $\pi_i$ can be 
viewed as a pure function.

\paragraph{Properties.}
Pairs and the unit value in a lambda calculus typically form a monoid. Assuming $\simeq$ is an
isomorphism that performs appropriate transformation on values, without affecting other 
properties (here, coeffects) of the expressions. The monoid laws then correspond to 
the requirement that $(e_1, (e_2, e_3)) \simeq ((e_1, e_2), e_3)$ (associativity) and the
requirement that $((), e) \simeq e \simeq (e, ())$ (unit).

Thanks to the properties of $\cpar$, the flat coeffect calculus obeys the monoid laws for pairs. 
In the following, we assume that \ident{assoc} is a pure function transforming a pair $(x_1, (x_2, x_3))$ 
to a pair $((x_1, x_2), x_3)$. We write $e \equiv e'$ when for all $\Gamma, \tau$ and $\cclrd{r}$,
it is the case that $\coctx{\Gamma}{\cclrd{r}} \vdash e : \tau$ if and only if
$\coctx{\Gamma}{\cclrd{r}} \vdash e' : \tau$.

\begin{theorem}
\label{thm:flat-tup-eq}
For a flat coeffect calculus with pairs and units, the following holds:
%
\begin{equation*}
\begin{array}{rclcl}
 \ident{assoc}~(e_1, (e_2, e_3)) &\equiv& ((e_1, e_2), e_3) &\qquad\qquad&(\emph{associativity}) \\
 \pi_1~(e, ()) &\equiv& e &&(\emph{right unit})\\
 \pi_2~((), e) &\equiv& e &&(\emph{left unit})\\
\end{array} 
\end{equation*}
\end{theorem}
\begin{proof}
Follows from the fact that $(\C, \cpar, \czero)$ is a monoid and \ident{assoc}, $\pi_1$ and
$\pi_2$ are pure functions (treated as contstants in the langauge).
\end{proof}

\noindent
The Theorem~\ref{thm:flat-tup-eq} motivates the requirement of the monoid structure 
$(\C, \cpar, \czero)$ of the flat coeffect algebra. We require only unit and associativity
laws. In our three examples, the $\cpar$ operator is also symmetric, which additionally
gives us the property that $(e_1, e_2) \simeq (e_2, e_1)$.

% ==================================================================================================
%                                                 
%      ####           ##            #                # 
%      #   #           #            #                # 
%      #   #   ###     #     ###   ####    ###    ## # 
%      ####   #   #    #        #   #     #   #  #  ## 
%      # #    #####    #     ####   #     #####  #   # 
%      #  #   #        #    #   #   #  #  #      #  ## 
%      #   #   ###    ###    ####    ##    ###    ## # 
%
% ==================================================================================================

\section{Related work}
\label{sec:flat-related}

Most of the related work leading to coeffects has already been discussed in Chapter~\ref{ch:pathways} 
and we covered work related to individual concepts throughout the chapter. In this section, we
do not repeat the discussion present elsewhere. Instead, we discuss one specific question that often 
arises when discussing coeffects and that is \emph{when is a coeffect (not) an effect?}

We start with a quick overview of the ways in which effects and coeffects differ and then
we briefly look at one (but illustrative) example where the two concepts overlap. We focus 
mainly on the equivalence between the \emph{categorical semantics}, which reveals the nature
of the computations -- rather than considering just the syntactic aspects of the type system.

\subsection{When is coeffect not a monad}
Coeffect systems differ from effect systems in three important ways:

\begin{itemize}
\item Semantically, coeffects capture different notions of computation. As demonstrated in 
  Chapter~\ref{ch:pathways}, coeffects track additional contextual properties required by a
  computation, many of which cannot be captured by a monad (\eg~liveness or data-flow).

\item Syntactically, coeffect calculi use a richer algebraic structure with pointwise composition,
  sequential composition and context merging ($\cpar, \cseq$, and $\czip$) while most effect systems
  only use a single operation for sequential composition (used by monadic bind).

\item Syntactically, the second difference is in the lambda abstraction (\emph{abs}). In 
  coeffect systems, the context requirements of the body can be split between (or duplicated
  at) declaration site and call site, while monadic effect systems always defer all effects.
\end{itemize}

\noindent
Despite the differences, our implicit parameters example can be also represented by a monad.
Semantically, the \emph{reader} monad is equivalent to the \emph{product} comonad. Syntactically,
we use the $\cup$ operation for all three operations of the coeffect algebra. However, to enable
splitting of implicit parameter requirements using the reader monad, we need to extend the 
monad structure and change the translation of monadic lambda abstraction.

% -------------------------------------------------------------------------------------------------

\subsection{When is coeffect a monad}
\label{sec:flat-related-monads}

Implicit parameters can be captured by a monad, but \emph{just} a monad is not enough.
Lambda abstraction in effect systems does not provide a way of splitting the context
requirements between declaration site and call site (or, semantically, combining the implicit 
parameters available in the scope where the function is deined and those specified by the caller).

\paragraph{Categorical relationship.}
Before looking at the necessary extensions, consider the two ways of modelling implicit 
parameters. We assume that the function $\cclrd{r} \rightarrow \sigma$ is a lookup function
for reading implicit parameter values that is defined on a set $\cclrd{r}$. The two definitions
are:
%
\begin{equation*}
\begin{array}{lll}
 \ctyp{\cclrd{r}}\tau = \tau \times (\cclrd{r} \rightarrow \sigma) &~\hspace{10em}~& (\emph{product comonad}) \\
 \mtyp{\cclrd{r}}\tau = (\cclrd{r} \rightarrow \sigma) \rightarrow \tau && (\emph{reader monad})
\end{array} 
\end{equation*}
%
The \emph{product comonad} simply pairs the value $\tau$ with the lookup function, while
the \emph{reader monad} is a function that, given a lookup function, produces a $\tau$ value.
As noted by Orchard \cite{comonads-vs-monads}, when used to model computation semantics, the 
two representations are equivalent:
%
\begin{remark}
Computations modelled as $\ctyp{\cclrd{r}}{\tau_1}\rightarrow\tau_2$ using the product comonad
are isomorphic to computations modelled as $\tau_1\rightarrow\mtyp{\cclrd{r}}{\tau_2}$ using the
reader monad via currying/uncurrying isomoprhism.
\end{remark}
\begin{proof}
The isomorphism is demonstrated by the following equation:
\begin{equation*}
\begin{array}{lll}
 \ctyp{\cclrd{r}}{\tau_1} \rightarrow \tau_2 &\narrow{=}&
 (\tau_1 \times (\cclrd{r} \rightarrow \sigma)) \rightarrow \tau_2 \\
 &\narrow{=}& \tau_1 \rightarrow ((\cclrd{r} \rightarrow \sigma) \rightarrow \tau_2) =
 \tau_1 \rightarrow \mtyp{\cclrd{r}}{\tau_2} \hspace{23em}\qedhere \\
\end{array}
\end{equation*}
\end{proof}

\noindent
This equivalence holds for monads and comonads (as well as \emph{indexed} monads
and comonads), but it does not extend to \emph{flat} indexed comonads which also provide
the $\ident{merge}_{\cclrd{r}, \cclrd{s}}$ operation to model context merging.

\paragraph{Delaying effects in monads.} 
In the syntax of the language, the above difference is manifested by the (\emph{abs}) rules for 
monadic effect systems and comonadic coeffect systems. The following listing shows the two rules 
side-by-side, using the effect system notation for both of them:
%
\begin{equation*}
\tyrule{cabs}
  { \Gamma, x\!:\!\tau_1 \vdash e : \tau_2 \;\&\; \cclrd{r}\,\cclrd{\cup}\,\cclrd{s} }
  { \Gamma \vdash \lambda x.e : \tau_1 \xrightarrow{\cclrd{s}} \tau_2 \;\&\; \cclrd{r}}
\quad
\tyrule{mabs}
  { \Gamma, x\!:\!\tau_1 \vdash e : \tau_2 \;\&\; \cclrd{r}\,\cclrd{\cup}\,\cclrd{s}}
  { \Gamma \vdash \lambda x.e : \tau_1 \xrightarrow{\cclrd{r}\,\cclrd{\cup}\,\cclrd{s}} \tau_2 \;\&\; \cclrd{\emptyset} }
\qquad
\end{equation*}
%
In the comonadic (\emph{cabs}) rule, the implicit parameters of the body are split. However,
the monadic rule (\emph{mabs}) places all requirements on the call site. This follows from the
fact that monadic semantics uses the \ident{unit} operation in the interpretation of lambda abstraction:
%
\begin{equation*}
\sem{\lambda x.e} \;=\; \ident{unit}~(\lambda x. \sem{e})
\end{equation*}
%
The type of $\ident{unit}$ is $\alpha\rightarrow\mtyp{\alpha}{\cclrd{\emptyset}}$, but in this specific
case, the $\alpha$ is instantiated to be $\tau_1\rightarrow\mtyp{\cclrd{r \cup s}}{\tau_2}$ and so this
use of \ident{unit} has a type:
%
\begin{equation*}
\ident{unit}~:~(\tau_1\rightarrow\mtyp{\cclrd{r \cup s}}{\tau_2}) \rightarrow \mtyp{\cclrd{\emptyset}}(\tau_1\rightarrow\mtyp{\cclrd{r \cup s}}{\tau_2})
\end{equation*}
%
In order to split the implicit parameters of the body ($\cclrd{r \cup s}$ on the left-hand side) between 
the declaration site ($\cclrd{\emptyset}$ on the outer $\mtyp{}$ on the right-hand side) and the 
call site ($\cclrd{r \cup s}$ on the inner $\mtyp{}$ on the right-hand side), we need an operation
(which we call \ident{delay}) with the following signature:
%
\begin{equation*}
\ident{delay}_{\cclrd{r}, \cclrd{s}}~:~(\tau_1\rightarrow\mtyp{\cclrd{r \cup s}}{\tau_2}) \rightarrow \mtyp{\cclrd{r}}(\tau_1\rightarrow\mtyp{\cclrd{s}}{\tau_2})
\end{equation*}
%
The operation reveals the difference between effects and coeffects -- intuitively, given a function
with effects $\cclrd{r \cup s}$, it should execute the effects $\cclrd{r}$ when wrapping the 
function, \emph{before} the function actually perforsm the effectful operation with the effects.
The remaining effects $\cclrd{s}$ are delayed as usual, while effects $\cclrd{r}$ are removed
from the effect annotation of the body.

Another important aspect of the signature is that the function needs to be indexed by the coeffect
annotations $\cclrd{r}, \cclrd{s}$. The indices determine how the input context requirements
$\cclrd{r \cup s}$ are split -- and thus guarantee determinism of the function at run-time.

The operation cannot be implemented in a useful way for most standard monads, but the 
reader monad is, indeed, an exception. It is not difficult to see how it can be implemented
when we expand the definitions of $\mtyp{\cclrd{r}}\tau$:
%
\begin{equation*}
\ident{delay}_{\cclrd{r}, \cclrd{s}}~:~
(\tau_1\rightarrow(\cclrd{r \cup s} \rightarrow \sigma) \rightarrow \tau_2) \rightarrow 
((\cclrd{r} \rightarrow \sigma) \rightarrow \tau_1\rightarrow(\cclrd{s} \rightarrow \sigma) \rightarrow \tau_2)
\end{equation*}

\paragraph{Restricting coeffects in comonads.} 
As just demonstrated, we can extend monads so that the reader monad is capable of capturing
the semantics of implicit parameters, including the splitting of implicit parameter requirements
in lambda abstraction. Can we also go the other way round and \emph{restrict} the comonadic
semantics so that all requirements are delayed as in the (\emph{mabs}) rule, thus modelling 
fully dynamically scoped parameters?

This is, indeed, possible. Recall that the semantics of lambda abstraction in the flat
coeffect calculus is modelled using $\ident{merge}_{\cclrd{r}, \cclrd{s}}$. The operation takes
two contexts (wrapped in an indexed comonad $\ctyp{\cclrd{r}}\alpha$), combines their carried 
values and additional contextual information (implicit parameters). To obtain the (\emph{mabs}) 
rule, we can restrict the first parameter, which corresponds to the declaration site context:
%
\begin{equation*}
\begin{array}{rcll}
 \ident{merge}_{\cclrd{r}, \cclrd{s}} &\narrow{:}& \ctyp{\cclrd{r}}\alpha \times \ctyp{\cclrd{s}}\beta \rightarrow \ctyp{\cclrd{r \cup s}}(\alpha \times \beta)
 &\hspace{4em}(\textit{normal}) \\
 \ident{merge}_{\cclrd{r}, \cclrd{s}} &\narrow{:}& \ctyp{\cclrd{\emptyset}}\alpha \times \ctyp{\cclrd{s}}\beta \rightarrow \ctyp{\cclrd{s}}(\alpha \times \beta)
 &\hspace{4em}(\textit{restricted})
\end{array} 
\end{equation*}
%
In the (\emph{restricted}) version of the operation, the declaration site context requires
no implicit parameters and so all implicit parameters have to be satisfied by the call site.
The semantics using the restricted version corresponds to the (\emph{mabs}) rule shown above.

The idea of restricting the operations of the coeffect calculus semantics could be used more
generally. We could allow any of the coeffect algebra operations $\cseq, \czip, \cpar$ to be
\emph{partial} and thus the restricted (fully dynamically-scoped) version of implicit parameters
could be obtained just by changing the definition of $\czip$. Similarly, we could obtain \eg~a
fully lexically-scoped version of the system. The ability to restrict operations to partial 
functions has been used in the semantics of effectful computations by Tate~\cite{effects-producer-semantics}.

% ==================================================================================================
%                                                                             
%        ###                         ##                    #                        
%       #   #                         #                                             
%       #       ###   # ##    ###     #    #   #   ###    ##     ###   # ##    ###  
%       #      #   #  ##  #  #   #    #    #   #  #        #    #   #  ##  #  #     
%       #      #   #  #   #  #        #    #   #   ###     #    #   #  #   #   ###  
%       #   #  #   #  #   #  #   #    #    #  ##      #    #    #   #  #   #      # 
%        ###    ###   #   #   ###    ###    ## #  ####    ###    ###   #   #  ####  
%                                                                             
% ==================================================================================================                                                                             
\section{Conclusions}

This chapter presented the \emph{flat coeffect calculus} -- a unified system for tracking 
\emph{whole-context} properties of computations, that is properties related to the 
execution environment or the enitre context in which programs are executed.
This is the first of the three \emph{coeffect calculi} developed in this thesis.

The flat coeffect calculus is parameterized by a \emph{flat coeffect algebra} that captures
the structure of the information tracked by the type system. We instantiated the system to 
capture three specific systems, namely liveness, data-flow and implicit parameters. However,
the system is more general and an capture numerous other applications outlined in Section~\ref{sec:applications-flat}.

Next, we introduced the notion of \emph{flat indexed comonad}, which generalizes of comonads
and adds additional operations needed to provide categorical semantics of the flat coeffect 
calculus. The indices of the flat indexed comonad operations correspond to the
coeffect annotations in the type system and provide a foundation for the design of the 
calculus.

Finally, we discussed the equational theory for flat coeffect calculus. Although each 
concrete instance of flat coeffect calculus models different notion of context, there are
syntactic properties that hold for all flat coeffect systems satisfying certain additional
conditions. In particular, two \emph{typing preservation} theorems prove that the operational
semantics for two classes of flat coeffect calculi (including liveness and implicit parameters) 
can be based on the standard call-by-name reduction.

In the upcoming chapter, we move from \emph{flat} coeffect calculi, tracking whole-context
properties to \emph{structural} coeffect calculi, tracking per-variable information, thus
covering systems from the second half of Chapter~\ref{ch:applications}.