### 2014-12-14 text/Chapter06.tex
Finishing proofreading (Chapter 6)

%!TEX root = ../main.tex

\chapter{Unified coeffect languages} 
\label{ch:unified}

%===================================================================================================

The main goal of this thesis is to provide a \emph{unified} calculus for tracking context 
dependence. We have not achieved this goal yet. In Chapter~\ref{ch:applications}, we identified 
two kinds of contextual properties that we further covered separately. Flat coeffects in 
Chapter~\ref{ch:flat} track whole-context properties, while structural coeffects, covered in 
Chapter~\ref{ch:structural}, track per-variable properties. In this chapter, we unify the two 
notions.

Although the results presented in this thesis are mainly of a theoretical nature, we indeed 
believe that coeffects should be integrated in main-stream programming languages. This is addressed
in the second part of the chapter, where we outline possible approach for a practical implementation
of coeffect systems. Finally, we conclude the chapter with a brief discussion of an alternative
approach to coeffect system, based on the meta-language style.

\paragraph{Chapter structure and contributions}
\begin{itemize}
\item We start by unifying the two kinds of coeffects discussed so far. In 
  Section~\ref{sec:unified-unified}, we introduce the \emph{unified coeffect calculus}, which 
  generalizes flat and structural systems and can be instantiated to track both flat and structural 
  properties.

\item We outline one possible approach for practical implementations of coeffects 
  (Section~\ref{sec:unified-impl}). The technique is based on embedding comonadic computations
  via a lightweight syntactic extension and using rich type systems for capturing the structure of 
  coeffect algebra.

\item Finally, we discuss an alternative approach to defining coeffect systems based on the
  meta-language style, which highlights the relationship between our work and related work arising 
  from modal logics such as CMTT \cite{logic-cmtt} (Section~\ref{sec:unified-meta}).
\end{itemize}

\noindent
The unified coeffect calculus presented in the first part of the chapter is a novel result that
completes the theory developed in this thesis. The next two sections present important future and
related work, respectively.

% ==================================================================================================

\section{The unified coeffect calculus}
\label{sec:unified-unified}

The type systems of the flat coeffect calculus (Figure~\ref{fig:flat-types}) and the structural 
coeffect calculus (Figure~\ref{fig:struct-types}) differ in a number of ways. Understanding the 
differences is the key to reconciling the two systems:

\begin{itemize}
\item The structural coeffect calculus contains explicit rules for context manipulation
  (weakening, contraction, exchange). In the flat coeffect calculus, these rules are not defined
  explicitly, but are admissible.

\item In the structural coeffect calculus, the variable context is treated as a vector
  and is annotated with a vector of (scalar) coeffects. In the flat coeffect calculus,
  the variable context is a finite partial function (from names to types) and is annotated with 
  a single (scalar) coeffect.

\item In the flat coeffect calculus, we distinguish between \emph{splitting} context requirements
  and \emph{merging} of context requirements ($\cpar$ and $\czip$, respectively). In the structural
  coeffect calculus, the operations (which model appending and splitting vectors) are invertible 
  and so the structural coeffect algebra uses the same tensor product $\atimes$.
\end{itemize}

\noindent
In the unified calculus presented in this section, we address the three differences as follows.
We include explicit rules for context manipulation in the calculus; in systems that arise from flat
calculi, the structural rules do not change coeffects and can thus be applied freely. We generalize 
the structure of coeffect annotations using the notion of a ``container'' which can be specialized 
to obtain a single annotation or a vector of annotations. This could potentially be used to capture
other structures such as trees in bunched typing \cite{substruct-bunched}. Finally, we distinguish 
between splitting and merging of context requirements (using the notation $\apar$ and $\azip$, 
respectively). For structural coeffect calculi, the two operators coincide, but for flat
coeffect calculi, they provide the needed flexibility.

\subsection{Shapes and containers}
Our generalization of coeffect structure using a \emph{coeffect container} is based on containers of 
Abbott et al. \cite{types-containers}. Interestingly, containers have later been linked
to comonads by Ahman et al. \cite{comonads-containers}. Intuitively a container describes data
types such as lists, trees or streams. A container is formed by shapes (\eg~lengths of 
lists)\footnote{Shapes can also be intuitively thought of as sizes, but this is not fully precise 
as some containers may have multiple shapes of the same length, \eg~differently ballanced trees.}.
For every shape, we can obtain a set of positions in the container (\eg~offsets in a list of a
specified length). More formally:

\begin{definition}
A \emph{container} consists of a pair $(S,P)$, usually written as $S \triangleleft P$. 
Here, $S$ is a set of shapes and $P$ is a shape-indexed family $P : S \rightarrow \ident{Set}$ 
of positions.
\end{definition}

\noindent
Well-known examples of containers include lists, non-empty lists, (un\-boun\-ded) streams,
trees and the singleton data type (which contains exactly one element). Containers relevant 
to our work are lists and singleton data types:

\begin{itemize}
\item The container representing lists is given by $S \triangleleft P$ where 
  shapes are integers $S = \ident{Nat}$ (lengths of a list). The set of positions for a 
  given length $n$ is the set of indices $P(n)= \{ 1 \ldots n \}$.

\item The contianer representing the singleton data type is given by $S \triangleleft P$ where
  shapes are given by a singleton set $S = \{ \ast \}$ and the set of positions for the
  shape $\ast$ contains exactly one position. To follow the intuition based on offset or 
  index, we write the single position as $0$, that is $P(\ast) = \{ 0 \}$.
\end{itemize}

\noindent
In the unified coeffect calculus, the structure of coeffect annotations is defined by a 
container with additional operations (discussed later) that links it with the free-variable 
context $\Gamma$.

\subsection{Structure of coeffects}
In the structural coeffect calculus, the coeffect annotation was formed by a vector of coeffect
scalars. The annotations in the unified coeffect calculus are similar, but a \emph{vector} is 
replaced with a more general \emph{container}. The primitive coeffect annotations in the unified 
calculus are formed of \emph{coeffect scalars}. The coeffect scalars are equipped with the same
structure as in structural coeffect calculus (Definition~\ref{def:structural-scalar}). In this 
section, we refer to it as \emph{\cclrd{unified coeffect scalar}} (and we repeat the definition 
below). Then we define \emph{\sclrd{unified coeffect containers}} which determines how coeffect 
scalar values are attached to the free-variable context. Finally, we define the 
\emph{\aclrd{unified coeffect algebra}} which consists of shape-indexed coeffect scalar values.

As in the structural coeffect calculus, the contexts in the unified calculus are annotated with 
shape-indexed coeffects, written as $\coctx{\Gamma}{\aclrd{\textbf{r}}} \vdash e : \tau$; 
functions take just a single input parameter and so are annotated with scalar coeffect values 
$\sigma \xrightarrow{\cclrd{r}} \tau$.

\paragraph{Coeffect scalars.}
The following definition of the coeffect scalar structure repeats the 
Definition~\ref{def:structural-scalar} from the previous chapter.

\begin{definition}
A \emph{\cclrd{unified coeffect scalar}} $(\C, \cseq, \cpar, \cunit, \czero, \cleq)$ is a set 
$\C$ together with elements $\cunit, \czero \in \C$, relation $\cleq$ and binary operations 
$\cseq, \cpar$ such that $(\C, \cseq, \cunit)$ and $(\C, \cpar, \czero)$ are monoids and
$(\C, \cleq)$ is a pre-order. That is, for all $r,s,t\in \C$:
%
\begin{equation*}
\begin{array}{ccr}
r \;\cseq\; (s \;\cseq\; t) = (r \;\cseq\; s) \;\cseq\; t  &
\cunit \;\cseq\; r = r = r \;\cseq\; \cunit &
\textnormal{(monoid)}   
\\
r \;\cpar\; (s \;\cpar\; t) = (r \;\cpar\; s) \;\cpar\; t &
\czero \;\cpar\; r = r = r \;\cpar\; \czero &
\textnormal{(monoid)}   
\\
\textnormal{if}~~r\; \cleq\; s ~~\textnormal{and}~~s\; \cleq\; t~~\textnormal{then}~~r\; \cleq\; t&
t\; \cleq\; t &
\textnormal{(pre-order)}   
\\
\end{array}
\end{equation*}
%
The following distributivity axioms hold:
\begin{align*}
\quad (\cclrd{r}\, \cpar\, \cclrd{s}) \;\cseq\; \cclrd{t} & = (\cclrd{r} \,\cseq\, \cclrd{t}) \;\cpar\; (\cclrd{s}\, \cseq\, \cclrd{t}) \\
\quad \cclrd{t} \;\cseq\; (\cclrd{r}\, \cpar\, \cclrd{s}) & = (\cclrd{t} \,\cseq\, \cclrd{r}) \;\cpar\; (\cclrd{t}\, \cseq\, \cclrd{s})
\end{align*}
%
In addition, we require the following two properties of $\czero, \cpar$ and $\cleq$:
\begin{equation*}
\begin{array}{l}
 \quad \cclrd{r}\,\cleq\,\cclrd{r'} \; \Rightarrow \; \forall\cclrd{s}.\;(\cclrd{r}\,\cseq\,\cclrd{s})\;\cleq\;(\cclrd{r'}\,\cseq\,\cclrd{s}) \\
 \quad \czero\;\cleq\;(\czero\,\cseq\,\cclrd{r})
\end{array}
\end{equation*}
\end{definition}

\noindent
As previously, the monoid $(\C,\cseq,\cunit)$ models sequential composition; the laws guarantee
an underlying category structre; $\cunit$ and $\czero$ represent an accessed and unused variable, 
respectively.

The $\cpar$ operation models combining of context requirements arising from multiple parts of a 
program. The meaning depends on the coeffect container. The operation can either combine requirements
of individual variables (structural coeffects) or requirements attached to the whole context of 
multiple sub-exp\-ressions (flat coeffects).

\paragraph{Coeffect containers.}
The coeffect container is a container that determines how are scalar coeffects attached to 
free-variable contexts. In addition to a container $\SHP$ formed by shapes and shape-indexed positions, 
the coeffect container provides a mapping that returns the shape corresponding to a free-variable context.

The mapping between the shape of the variable context and the shape of the coeffect annotation
is not necessarily bijective. For example, coeffect annotations in flat systems have just a single 
shape $\SH = \{ \ast \}$.

In the coeffect judgment $\coctx{\Gamma}{\aclrd{\textbf{r}}} \vdash e : \tau$, the coeffect annotation
$\aclrd{\textbf{r}}$ is drawn from the set of coeffect scalars $\C$ indexed by the shape corresponding
to $\Gamma$. We write $\sclrd{s} = \slen{\Gamma}$ for the shape corresponding to $\Gamma$. The operation 
$\SP(\sclrd{s})$ returns a \emph{set} of positions and so we can write $\aclrd{\textbf{r}} \in 
  \SP(\sclrd{s}) \rightarrow \C$  as a mapping from positions (defined by the shape) to scalar coeffects. 
We usually write this as the exponent $\aclrd{\textbf{r}} \in \C^{\SP(\sclrd{s})}$.
The coeffect container is also equipped with an operation that appends shapes (when we concatenate
variable contexts) and two special shapes in $\SH$ representing the empty context and the singleton 
context.

\begin{definition}
A \emph{\sclrd{coeffect container}} structure $(\SHP, \stimes, \sempty, \sunit, \slen{-})$ 
comprises a container $\SHP$ with a binary operation $\stimes$ on $\SH$ for appending shapes, a 
mapping from free-variable contexts to shapes $\slen{\Gamma} \in \SH$, and elements $\sempty,\sunit \in
\SH$ such that $(\SH, \stimes, \sempty)$ is a monoid.

The elements $\sempty$ and $\sunit$ represent the shapes of the empty and the singleton free-variable 
context, respectively. The $\stimes$ operation corresponds to concatentation of free-variable contexts.
Given $\Gamma_1$ and $\Gamma_2$ such that $\sclrd{s_1}=\slen{\Gamma_1}, \sclrd{s_2}=\slen{\Gamma_2}$, 
we require that $\sclrd{s_1}\,\stimes\,\sclrd{s_2}=\slen{\Gamma_1, \Gamma_2}$. 
\end{definition}

\noindent
As discussed earlier, we use two kinds of coeffect containers that describe the structure of vectors 
(for structural coeffects) and the shape of trivial singleton container (for flat coeffects):

\begin{example}
\label{ex:unified-struct-shape}
Structural coeffect container is defined as $\sclrd{(\SHP, {|}\textnormal{--}{|}, +, 0, 1)}$
where $\SH=\mathbb{N}$ and $\SP(\sclrd{n})=\{ 1 \ldots n \}$. The shape mapping ${|}\Gamma{|}$ returns the 
number of variables in $\Gamma$. Empty and singleton contexts are annotated with $0$ and $1$, 
respectively, and shapes of combined contexts are added so that ${|}\Gamma_1, \Gamma_2{|} = 
{|}\Gamma_1{|} + {|}\Gamma_2{|}$.

Therefore, a coeffect annotation is a \emph{vector} 
$\aclrd{\textbf{r}} \in \C^{\SP(\sclrd{n})}$ and assigns a coeffect scalar $\aclrd{\textbf{r}}(i) \in \C$ 
for each position (corresponding to a variable $x_i$ in the context).
\end{example}

\begin{example}
\label{ex:unified-flat-shape}
Flat coeffect container is defined as $\sclrd{(\SHP, {|}-{|}, \diamond, \sflat, \sflat)}$.
The container is defined as a singleton data type $\SH = \{\sflat \}$ and $\SP(\sflat) = \{ 0 \}$
with a constant function ${|}\Gamma{|}=\sflat$ and a trivial operation $\sflat\,\diamond\,\sflat = \sflat$.

That is, there is a single shape
$\sflat$ with a single position and all free-variable contexts have the same singleton shape.
Therefore, a coeffect annotation is drawn from $\C^{\{\sflat\}}$ which is isomorphic to $\C$
and so a coeffect scalar $\cclrd{r}\in \C$ is associated with every free-variable context.
\end{example}

\begin{example}
Similarly to the previous example, we can also define a coeffect container with \emph{no} positions,
\ie~$\sclrd{(\SHP, {|}-{|}, \diamond, \sflat, \sflat)}$ where $\SH = \{\sflat \}$, $\SP(\sflat) = \emptyset$,
${|}\Gamma{|}=\sflat$ and $\sflat\,\diamond\,\sflat = \sflat$. 
This reduces our system to the simply-typed $\lambda$-calculus with no context annotations, because
$\SP(\sflat)=\emptyset$ and so coeffect annotations would be from the empty set $\C^\emptyset$.
\end{example}

\paragraph{Unified coeffect algebra.}
The unified coeffect calculus annotates typing judgments with shape-indexed coeffect annotations.
The \emph{\aclrd{unified coeffect algebra}} combines a coeffect scalar and a coeffect container
to define shape-indexed coeffects and operations for manipulating these.

The definition here reconciles the third point discussed in Section~\ref{sec:unified-unified} --
the fact that flat coeffects use separate operations for splitting and merging ($\cpar$ and $\czip$)
while structural coeffects use the tensor $\atimes$. In the unified calculus, we use two operators
that can, however, coincide.

\begin{definition}
Given a \cclrd{unified coeffect scalar} $(\C, \cseq, \cpar, \cunit, \czero, \cleq)$ and a 
\sclrd{coeffect container} $(\SHP, \slen{-}, \stimes, \sempty, \sunit)$ a \emph{\aclrd{unified 
coeffect algebra}} extends the two structures with $(\azip, \apar, \azero)$ where $\azero \in \C^{\SP(\sempty)}$
is a coeffect annotation for the empty context and $\azip, \apar$ are families of operations that 
combine coeffect annotations indexed by shapes. That is $\forall \sclrd{n},\sclrd{m} \in \SH$:
%
\begin{equation*}
\apar_{\sclrd{m},\sclrd{n}}, \azip_{\sclrd{m},\sclrd{n}} ~:~ \C^{\SP \sclrd{m}} \times \C^{\SP \sclrd{n}} \rightarrow \C^{\SP (\sclrd{m}\,\stimes\,\sclrd{n})}
\end{equation*}
\end{definition}

\noindent
A coeffect algebra induces the following three additional operations:
%
\begin{equation*}
\begin{array}{rclcrcl}
 \alift{-} &\narrow{:}& \C \rightarrow \C^{\SP(\sunit)}  \\
 \alift{\cclrd{x}} &\narrow{=}& \lambda \_ . \cclrd{x}   \\
\\[-0.5em]
 \aseq_{\sclrd{m}} &\narrow{:}&  \C \times \C^{\SP(\sclrd{m})} \rightarrow \C^{\SP(\sclrd{m})}   \\
 \cclrd{r}\, \aseq\, \aclrd{\textbf{s}} &\narrow{=}& \lambda \sclrd{s} . \cclrd{r}\, \cseq\, \aclrd{\textbf{s}}(\sclrd{s}) \\
\\[-0.5em]
 \alen{-} &\narrow{:}& \C^{\SP(\sclrd{m})} \rightarrow \sclrd{m}\\
 \alen{\aclrd{\textbf{r}}} &\narrow{=}& \sclrd{m}
\end{array}
\end{equation*}
%
The $\alift{-}$ operation lifts a scalar coeffect to a shape-indexed coeffect that is indexed by 
the shape of a singleton context. The $\aseq_{\sclrd{m}}$ operation is a left multiplication of a 
vector by a scalar. As we always use bold face for vectors and ordinary face for scalars (as well 
as a distinct colour), using the same symbol is not ambiguous. We also tend to omit the subscript
$\sclrd{m}$ and write~$\aseq$.

Finally, we define $\alen{-}$ as an operation that returns the shape of a given shape-indexed 
coeffect. The only purpose is to simplify notation, as we tend to avoid subscripts, but often need 
to specify that shapes of variable context and coeffect match, 
\eg~$\alen{\aclrd{\textbf{r}}} = \slen{\Gamma}$.

\paragraph{Splitting and merging coeffects.}
The operators $\apar$ and $\azip$ combine shape-indexed coeffects associated with two contexts. 
For example, assume we have $\Gamma_1$ and $\Gamma_2$ with coeffects $\aclrd{\textbf{r}} \in 
\C^{\SP(\sclrd{m})}$ and $\aclrd{\textbf{s}} \in \C^{\SP(\sclrd{n})}$. In the structural system, the 
context shapes $\sclrd{m}, \sclrd{n}$ denote the number of variables in the two contexts. The 
combined context $\Gamma_1, \Gamma_2$ has a shape $\sclrd{m}\,\stimes\,\sclrd{n}$ and the combined 
coeffects $\aclrd{\textbf{r}} \,\azip\, \aclrd{\textbf{s}}, \aclrd{\textbf{r}} \,\apar\, \aclrd{\textbf{s}} 
\in \C^{\SP(\sclrd{m} \stimes \sclrd{n})}$ are indexed by that shape.

For structural coeffect systems such as bounded reuse, both $\apar$ and $\azip$ are
just the tensor product $\times$ of vectors. For flat coeffect systems, the operations
can be defined independently, letting $\azip=\czip$ and $\apar=\cpar$.

The difference between $\azip$ and $\apar$ is clarified by the semantics (Sec\-tion~\ref{sec:unified-semantics}), 
where $\aclrd{\textbf{r}}\,\azip\,\aclrd{\textbf{s}}$ is an annotation of the \emph{codomain} of 
a morphism that merges the capabilities provided by two contexts (in the syntactic 
reading, splits the context requirements), while $\aclrd{\textbf{r}}\,\apar\,\aclrd{\textbf{s}}$ 
is an annotation of the \emph{domain} of a morphism that splits the capabilities of a single 
context into two parts (in the syntactic reading, merges their context requirements). 
Syntactically, this means that we always use $\azip$ in the rule \emph{assumptions} and 
$\apar$ in \emph{conclusions}.

% --------------------------------------------------------------------------------------------------

\newcommand{\tystmt}[2]{ (\text{\footnotesize{#1}})~~{#2} }
\newcommand{\ctxtrans}[3]{ #2 \rightsquigarrow #1, #3 }
\newcommand{\ctxtransnl}[3]{ \begin{array}{l} #2 \rightsquigarrow\\ #1, #3 \end{array}}

\begin{figure}[t]

{\small a.) Syntax-driven typing rules:}
\begin{equation*}
\tyrule{var}
  { }
  {\coctx{x\!:\!\tau}{\alift{\cunit}} \vdash x : \tau }
\end{equation*}
\begin{equation*}
\tyrule{const}
  {c \!:\! \tau \in \Delta}
  {\coctx{()}{\azero} \vdash c : \tau }
\end{equation*}
\begin{equation*}
\tyrule{app}
  {\coctx{\Gamma_1}{\aclrd{\textbf{r}}} \vdash e_1 : \tau_1 \xrightarrow{\cclrd{t}} \tau_2 &
   \coctx{\Gamma_2}{\aclrd{\textbf{s}}} \vdash e_2 : \tau_1 }
  {\coctx{\Gamma_1, \Gamma_2}{\aclrd{\textbf{r}\;\apar\;(\cclrd{t} \,\aseq\, \aclrd{\textbf{s}})}} \vdash e_1~e_2 : \tau_2}
\end{equation*}
\begin{equation*}
\tyrule{abs}
  {\coctx{\Gamma, x\!:\!\tau_1}{\aclrd{\textbf{r}}\,\azip\,\alift{\cclrd{s}}} \vdash e : \tau_2}
  {\coctx{\Gamma}{\aclrd{\textbf{r}}} \vdash \lambda x.e : \tau_1 \xrightarrow{\cclrd{s}} \tau_2 }
\end{equation*}
\begin{equation*}
\tyrule{let}
  { \coctx{\Gamma_1}{\aclrd{\textbf{r}}} \vdash e_1 : \tau_1 &
    \coctx{\Gamma_2, x\!:\!\tau_1}{\aclrd{\textbf{s}}\;\azip\;\alift{\cclrd{t}}} \vdash e_2 : \tau_2}
  {\coctx{\Gamma_1, \Gamma_2}{(\cclrd{t} \,\aseq\, \aclrd{\textbf{r}}) \;\apar\; \aclrd{\textbf{s}}} \vdash \kvd{let}~x=e_1~\kvd{in}~e_2 : \tau_2 }
\end{equation*}

\vspace{1.5em}
{\small b.) Structural rules for context manipulation:}
\begin{equation*}
\tyrule{sub}
  {\coctx{\Gamma_1,x\!:\!\tau_1,\Gamma_2}{\aclrd{\textbf{r}} \azip \alift{\cclrd{s'}} \azip \aclrd{\textbf{q}}} \vdash e : \tau}
  {\coctx{\Gamma_1,x\!:\!\tau_1,\Gamma_2}{\aclrd{\textbf{r}} \apar \alift{\cclrd{s}} \apar \aclrd{\textbf{q}}} \vdash e : \tau}
\quad
(\cclrd{s'} \cleq \cclrd{s})
\end{equation*}
\begin{equation*}
\tyrule{weak}
  {\coctx{\Gamma}{ \aclrd{\textbf{r}} } \vdash e : \tau}
  {\coctx{\Gamma,x \!:\! \tau_1}{\aclrd{\textbf{r}} \apar \alift{ \czero }} \vdash e : \tau} 
\end{equation*}
\begin{equation*}
\tyrule{exch}
  {\coctx{\Gamma_1,x\!:\!\tau_1,y\!:\!\tau_2,\Gamma_2}{\aclrd{\textbf{r}} \azip \alift{\cclrd{s}} \azip \alift{\cclrd{t}} \azip \aclrd{\textbf{q}}} \vdash e : \tau}
  {\coctx{\Gamma_1,y\!:\!\tau_2,x\!:\!\tau_1,\Gamma_2}{\aclrd{\textbf{r}} \apar \alift{\cclrd{t}} \apar \alift{\cclrd{s}} \apar \aclrd{\textbf{q}}} \vdash e : \tau}
\quad
\begin{array}{l}
 \slen{\Gamma_1} = \alen{\aclrd{\textbf{r}}}\\[-0.25em]
 \slen{\Gamma_2} = \alen{\aclrd{\textbf{s}}}
\end{array}
\end{equation*}
\begin{equation*}
\hspace{2.5em} 
\tyrule{contr}
  {\coctx{\Gamma_1,y\!:\!\tau_1,z\!:\!\tau_1,\Gamma_2}{\aclrd{\textbf{r}} \azip \alift{\cclrd{s}} \azip \alift{\cclrd{t}} \azip \aclrd{\textbf{q}}} \vdash e : \tau}
  {\coctx{\Gamma_1,x\!:\!\tau_1,\Gamma_2}{\aclrd{\textbf{r}} \apar \alift{\cclrd{s} \,\cpar\, \cclrd{t}} \apar \aclrd{\textbf{q}}} \vdash \subst{e}{z,y}{x} : \tau}
~
\begin{array}{l}
 \slen{\Gamma_1} = \alen{\aclrd{\textbf{r}}}\\[-0.25em]
 \slen{\Gamma_2} = \alen{\aclrd{\textbf{s}}}
\end{array}
\end{equation*}

\vspace{0.5em}
\figcaption{Type system for the unified coeffect calculus}
\label{fig:unif-types}
\end{figure}

% --------------------------------------------------------------------------------------------------

\subsection{Unified coeffect type system}
\label{sec:unified-types}
The unified coeffect system in Figure~\ref{fig:unif-types} resembles the structural type system
shown in Figure~\ref{fig:struct-types}. Rather than explaining the rules one-by-one, we focus on
how the unified system differs from the structural system.

The type system for the unified coeffect calculus is parameterized by a coeffect scalar 
$(\C, \cseq, \cpar, \cunit, \czero, \cleq)$ together with a coeffect algebra $(\azip, \apar, \azero)$ 
and the derived constructs $\alift{-}$, $\alen{-}$ and $\aseq$. 
As in the structural system, free-variable contexts $\Gamma$ are treated as vectors modulo duplicate
use of variables -- the associativity is built-in. The order of variables matters, but can be changed 
using the exchane rule. The context annotations $\aclrd{\textbf{r}}, \aclrd{\textbf{s}}, \aclrd{\textbf{t}}$ 
are shape-indexed coeffects (rather than simple vectors as before). As before, functions are
annotated with coeffects scalars.

\paragraph{Syntax-driven rules.} 
The (\emph{var}) rule is syntactically the same as in the structural system, but it should be read
differently. The $\alift{-}$ operation does not create a \emph{vector}, but a \emph{shape-indexed
coeffect} that returns $\cunit$ for all positions in the singleton shape $\sunit$. 
The (\emph{const}) rule annotates empty context with a special annotation of the shape $\sempty$.

In a structural system, the two annotations correspond to a singleton and empty vector, respectively.
However, for a singleton shape with one position, the annotations are equivalent to annotating
variables with a scalar $\cunit$ and constants with a scalar $\czero$.

In the (\emph{app}), (\emph{abs}) and (\emph{let}) rules, the only change from the structural
system is that the vector concatentation $\atimes$ is now replaced with context splitting/merging 
of the unified coeffect algebra. As already mentioned, we use splitting of context requirements $\azip$
in rule \emph{assumptions} and merging of context requirements $\apar$ in rule \emph{conclusions}.

Note that we use the terms \emph{merging} and \emph{splitting} in the syntactic (top-down) sense. As 
discussed in Chapter~\ref{ch:flat}, Section~\ref{sec:flat-calculus-undestanding}, we can also read 
the rules in semantic (bottom-up) sense, in which case assumptions merge available contextual 
information and conclusions split available contextual information.

\paragraph{Structural rules.}
The merging/splitting operations in the structural rules are changed in the same way as in the
rules above. It is also worth noting that structural coeffect system used vectors of multiple 
elements. For example, $\alift{\cclrd{s},\cclrd{t}}$ and $\alift{\cclrd{t},\cclrd{s}}$ in the (\emph{exch}) 
rule denoted two-element vectors. In the unified system, this is replaced with merging/splitting
of two lifted scalars: $\alift{\cclrd{s}} \azip \alift{\cclrd{t}}$ and $\alift{\cclrd{t}} \apar \alift{\cclrd{s}}$.

In structural systems, the two notations mean the same thing -- we are simply concatenating or splitting two 
singleton vectors. However, this generalization allows us to capture flat coeffect systems as well.
The lifting operation in flat systems simply returns the lifted scalar and operators $\azip$ and
$\apar$ correspond to operations on coeffect scalars. As discussed in Section~\ref{sec:unified-flat},
thanks to the properties of coeffect scalars, contraction, weakening and exchange that do not 
affect the coeffect annotation are admissible for all flat systems embedded in the unified 
coeffect calculus.

% --------------------------------------------------------------------------------------------------

\subsection{Structural coeffects}
\label{sec:unified-structural}

The unified coeffect system uses a general notion of context shape, but it has been designed with 
structural and flat systems in mind. In this and the next section, we show how it captures 
the two coeffect systems from Chapter~\ref{ch:flat} and Chapter~\ref{ch:structural}.

The unified calculus closely resembles the structural system and so using it to model structural
systems is easy -- given a coeffect scalar, we use the coeffect container that describes a \emph{vector} 
of annotations (Example~\ref{ex:unified-struct-shape}) and define a coeffect algebra formed by
a vector (free monoid) of scalars.

\begin{definition}
\label{def:unified-struct}
Given a coeffect scalar $(\C, \cseq, \cpar, \cunit, \czero, \cleq)$ a \emph{structural coeffect system} 
is defined by:

\begin{itemize}
\item A coeffect container $\sclrd{(\SHP, {|}\textnormal{--}{|}, +, 0, 1)}$ where $\SH=\mathbb{N}$ and
  $\SP(\sclrd{n})=\{ 1 \ldots n \}$ and  ${|}x_1\!:\!\tau_1,\ldots,x_n\!:\!\tau_n{|}=n$.

\item A coeffect algebra $\aclrd{(\times, \times, \epsilon)}$ where $\times$ and $\epsilon$ are
  shape-indexed versions of the binary operation and the unit of a free monoid over $\C$.
  That is $\epsilon : \C^{\SP(\sclrd{0})}$ is an empty vector and 
  $\times : \C^{\SP(\sclrd{n})} \times \C^{\SP(\sclrd{m})} \rightarrow \C^{\SP(\sclrd{n + m})}$ 
  appends vectors.
\end{itemize}
\end{definition}

\noindent
The definition is valid since the shape operations form a monoid
$\sclrd{(\mathbb{N}, +, 0)}$ and $\slen{\textnormal{--}}$ (calculating the length of
a list) is a monoid homomorphism from the free monoid to the monoid of
shapes.

\paragraph{Properties.}
An important property of the unified system is that, when used in a structural way as discussed
above, it gives calculi with the same properties as the structural system described in 
Chapter~\ref{ch:structural}. This can be easily seen by comparing the Figure~\ref{fig:unif-types}
with the Figure~\ref{fig:struct-types} and using the free monoid interpretation of the unified
coeffect algebra.

\begin{remark}
The system described in Definition~\ref{def:unified-struct} is equivalent to the structural coeffect
system described in Figure~\ref{fig:struct-types}. That is, a typing derivation using a structural
coeffect embedded in the unified system is valid if and only if the corresponding derivation is
valid in the structural system.
\end{remark}

\noindent
Using the above definition, our unified coeffect system can capture per-variable coeffect properties
discussed in Section~\ref{sec:applications-structural}. This includes the system for bounded reuse 
(which is only used in the structural form) and precise tracking of per-variable dataflow and liveness.

% --------------------------------------------------------------------------------------------------

\subsection{Flat coeffects}
\label{sec:unified-flat}

The same unified coeffect system can be used to capture systems that track whole-context (flat) 
coeffects such as implicit parameters. This is achieved using a singleton-shaped container for
coeffect annotations. The resulting system has explicit structural rules (and is syntactically
different from the standard flat coeffect system), but we show that they are equivalent.

Flat coeffect systems are characterised by a singleton set of shapes (Example~\ref{ex:unified-flat-shape}).
In this setting, the context annotations $\C^{ \SP(\sflat) } = \{0\} \rightarrow \C$ are equivalent 
to coeffect scalars $\C$. In addition to the coeffect scalar structure, we need to define $\apar$ 
and $\azip$. Our examples of flat coeffects use $\cpar$ (merging of scalar coeffects) for $\apar$ 
(merging of shaped coeffect annotations). The $\azip$ operation (corresponding to $\czip$
in flat coeffect calculus) needs to be provided explicitly.

\begin{definition}
\label{def:unified-flat}
Given a coeffect scalar $(\C, \cseq, \cpar, \cunit, \czero, \cleq)$ and a binary operation 
$\cclrd{\wedge} : \C \times \C \rightarrow \C$ such that $(\cclrd{r \wedge s}) \,\cleq\, (\cclrd{r}\,\cpar\,\cclrd{s})$, 
the unified coeffect algebra modelling a flat coeffect systems consists of:
%
\begin{itemize}{}
\item A flat coeffect container $\sclrd{(\SHP, {|}-{|}, \diamond, \sflat, \sflat)}$ as defined in 
  Example~\ref{ex:unified-flat-shape}.

\item A flat coeffect algebra $(\czip, \cpar, \czero)$, \ie~$\apar = \cpar$ 
  and $\azero = \czero$ with an\\ additional binary operation $\azip = \czip$. 
\end{itemize}
\end{definition}

\noindent
Intuitively, the requirement $(\cclrd{r}\,\czip\,\cclrd{s}) \;\cleq\; (\cclrd{r}\,\cpar\,\cclrd{s})$,
which could be also written as $(\aclrd{r}\,\azip\,\aclrd{s}) \;\aclrd{\leq}\; (\aclrd{r}\,\apar\,\aclrd{s})$,
denotes that splitting context requirements and then re-combining them preserves all the requirements
from the assumptions. The system may be imprecise and conclusions can overapproximate assumptions, but
it cannot lose requirements. This is fundamental for showing that exchange and contraction are admissible
in the unified system.

\paragraph{Properties.}
To show that the typing of flat properties in the unified system is equivalent to the typing in the
flat system, we show that a valid typing judgement in the first system is a valid typing judgement in 
the second system and vice versa.

In one direction, we show that the unified system (capturing flat properties) permits weakening, 
contraction and exchange rules that do not change the coeffect annotations. This guarantees that
a valid judgement in flat system is also valid in the unified system.

\begin{lemma}
\label{thm:unified-weak}
A unified coeffect calculus capturing flat properties admits weakening that does not change the
coeffect annotation.
\end{lemma}
\begin{proof}
The rule is admissible using the following derivation:
\begin{equation*}
\inference
  { \inference
      {\coctx{\Gamma}{ \cclrd{r} } \vdash e : \tau}
      {\coctx{\Gamma,x \!:\! \tau_1}{\cclrd{r} \;\apar\; \alift{ \czero }} \vdash e : \tau} }
  { \inference
      {\coctx{\Gamma,x \!:\! \tau_1}{\cclrd{r} \,\cpar\, \czero} \vdash e : \tau}      
      {\coctx{\Gamma,x \!:\! \tau_1}{\cclrd{r}} \vdash e : \tau} }
\end{equation*}

\noindent
We write $\cclrd{r}$ rather than $\aclrd{\textbf{r}}$, because we are tracking flat properties.
The first step is an application of the (\emph{weak}) rule. Next, we use the fact that
$\apar=\cpar$ and $\alift{\czero}=\czero$, which is the unit of the monoid $(\C, \cpar, \czero)$.
\end{proof}

\begin{lemma}
A unified coeffect calculus capturing flat properties admits exchange that does not change the
coeffect annotation.
\end{lemma}
\begin{proof}
We use the idempotence of $\czip$ and $\cpar$ together with the (\emph{exch}) rule:
\begin{equation*}
\inference
  { \inference
      { \inference
          {\coctx{\Gamma_1,x\!:\!\tau_1,y\!:\!\tau_2,\Gamma_2}{\cclrd{r}} \vdash e : \tau}
          {\coctx{\Gamma_1,x\!:\!\tau_1,y\!:\!\tau_2,\Gamma_2}{\cclrd{r} \,\czip\, \cclrd{r} \,\czip\, \cclrd{r} \,\czip\, \cclrd{r}} \vdash e : \tau} }
      { \coctx{\Gamma_1,x\!:\!\tau_1,y\!:\!\tau_2,\Gamma_2}{\cclrd{r} \azip \alift{\cclrd{r}} \azip \alift{\cclrd{r}} \azip \cclrd{r}} \vdash e : \tau } }
  { \inference 
      {\coctx{\Gamma_1,y\!:\!\tau_2,x\!:\!\tau_1,\Gamma_2}{\cclrd{r} \apar \alift{\cclrd{r}} \apar \alift{\cclrd{r}} \apar \cclrd{r}} \vdash e : \tau}
      {\inference
         {\coctx{\Gamma_1,y\!:\!\tau_2,x\!:\!\tau_1,\Gamma_2}{\cclrd{r}\,\cpar\,\cclrd{r}\,\cpar\,\cclrd{r}\,\cpar\,\cclrd{r}} \vdash e : \tau}
         {\coctx{\Gamma_1,y\!:\!\tau_2,x\!:\!\tau_1,\Gamma_2}{\cclrd{r}} \vdash e : \tau} } }
\end{equation*}

\noindent
Using idempotence, we first duplicate the annotation $\cclrd{r}$ to get a coeffect in the form
$\cclrd{r} \azip \alift{\cclrd{r}} \azip \alift{\cclrd{r}} \azip \cclrd{r}$ as required by the
assumption of the (\emph{exch}) rule. Note that $\alift{-}$ can be added freely as
$\alift{\cclrd{r}}$ is equivalent to $\cclrd{r}$. After applying (\emph{exch}), we use 
idempotence of $\apar$.
\end{proof}

\begin{lemma}
A unified coeffect calculus capturing flat properties admits contraction that does not change the
coeffect annotation.
\end{lemma}
\begin{proof}
Similarly to exchange, the proof uses idempotence of $\czip$ and $\cpar$:
\begin{equation*}
\inference
  { \inference
      {\coctx{\Gamma_1,y\!:\!\tau_1,z\!:\!\tau_1,\Gamma_2}{\cclrd{r} \czip\, \cclrd{r} \czip\, \cclrd{r} \czip\, \cclrd{r}} \vdash e : \tau}
      {\coctx{\Gamma_1,y\!:\!\tau_1,z\!:\!\tau_1,\Gamma_2}{\cclrd{r} \azip \alift{\cclrd{r}} \azip \alift{\cclrd{r}} \azip \cclrd{r}} \vdash e : \tau} }
  { \inference
      {\coctx{\Gamma_1,x\!:\!\tau_1,\Gamma_2}{\cclrd{r} \apar \alift{\cclrd{r} \,\cpar\, \cclrd{r}} \apar \cclrd{r}} \vdash \subst{e}{z,y}{x} : \tau}
      {\coctx{\Gamma_1,x\!:\!\tau_1,\Gamma_2}{\cclrd{r} \,\cpar\, \cclrd{r} \,\cpar\, \cclrd{r} \,\cpar\, \cclrd{r}} \vdash \subst{e}{z,y}{x} : \tau} }
\begin{array}{r} \\[2.5em] \qedhere \end{array}
\end{equation*}
\end{proof}

\noindent
In the last two cases, we need to turn the coeffect into a form that is required by the exchange
and contraction rules. Aside from idempotence, we could use the unit property and obtain 
\eg~$\czero\,\czip\,\cclrd{r}\,\czip\,\czero\,\czip\,\czero$. However, this appraoch does not 
work because $\cpar$ and $\czip$ may have different unit elements (in fact, we do not ever require
the existence of a unit for $\czip$).

In the other direction, we need to show that any valid judgement in the unified system (tracking
flat properties) is also valid in the flat system. The syntax-directed rules are the same in both
systems, but we need to show that any use of (explicit) weakening, contraction and exchange can be
derived in the flat system.

\begin{lemma}
Weakening, as defined in a unified coeffect calculus capturing flat properties, is admissible in 
the flat coeffect calculus.
\end{lemma}
\begin{proof}
Similar to the proof in Lemma~\ref{thm:unified-weak}. The property follows from the fact that 
$\bot=\alift{\czero}$ is the unit of $\cpar$.
\end{proof}

\begin{lemma}
\label{thm:unified-rev-contr}
Contraction, as defined in a unified coeffect calculus capturing flat properties, is admissible in 
the flat coeffect calculus.
\end{lemma}
\begin{proof}
The application of (\emph{contr}) rule has the following form:
\begin{equation*}
\inference
 { \inference
    {\coctx{\Gamma_1,y\!:\!\tau_1,z\!:\!\tau_1,\Gamma_2}{\cclrd{r} \,\czip\, \cclrd{s} \,\czip\, \cclrd{t} \,\czip\, \cclrd{q}} \vdash e : \tau}
    {\coctx{\Gamma_1,y\!:\!\tau_1,z\!:\!\tau_1,\Gamma_2}{\cclrd{r} \,\azip\, \alift{\cclrd{s}} \,\azip\, \alift{\cclrd{t}} \,\azip\, \cclrd{q}} \vdash e : \tau} }
 { \inference
    {\coctx{\Gamma_1,x\!:\!\tau_1,\Gamma_2}{\cclrd{r} \,\apar\, \alift{\cclrd{s} \,\cpar\, \cclrd{t}} \,\apar\, \cclrd{q}} \vdash \subst{e}{z,y}{x} : \tau}
    {\coctx{\Gamma_1,x\!:\!\tau_1,\Gamma_2}{\cclrd{r} \,\cpar\, \cclrd{s} \,\cpar\, \cclrd{t} \,\cpar\, \cclrd{q}} \vdash \subst{e}{z,y}{x} : \tau} } 
\end{equation*}
%
From Definition~\ref{def:unified-flat}, we know that $(\cclrd{r} \,\czip\, \cclrd{s} \,\czip\, \cclrd{t} \,\czip\, \cclrd{q}) 
\;\cleq\; (\cclrd{r} \,\cpar\, \cclrd{s} \,\cpar\, \cclrd{t} \,\cpar\, \cclrd{q})$. Thus, the judgement
 can be derived using the (\emph{sub}) rule of flat coeffect calculus.
\end{proof}

\begin{lemma}
\label{thm:unified-rev-exch}
Exchange, as defined in a unified coeffect calculus capturing flat properties, is admissible in 
the flat coeffect calculus.
\end{lemma}
\begin{proof}
The application of (\emph{exch}) rule has the following form:
\begin{equation*}
\inference
 {\inference
    {\coctx{\Gamma_1,x\!:\!\tau_1,y\!:\!\tau_2,\Gamma_2}{\cclrd{r} \,\czip\, \cclrd{s} \,\czip\, \cclrd{t} \,\czip\, \cclrd{q}} \vdash e : \tau}
    {\coctx{\Gamma_1,x\!:\!\tau_1,y\!:\!\tau_2,\Gamma_2}{\cclrd{r} \,\azip\, \alift{\cclrd{s}} \,\azip\, \alift{\cclrd{t}} \,\azip\, \cclrd{q}} \vdash e : \tau} }
 {\inference
    {\coctx{\Gamma_1,y\!:\!\tau_2,x\!:\!\tau_1,\Gamma_2}{\cclrd{r} \,\apar\, \alift{\cclrd{t}} \,\apar\, \alift{\cclrd{s}} \,\apar\, \cclrd{q}} \vdash e : \tau}
    {\coctx{\Gamma_1,y\!:\!\tau_2,x\!:\!\tau_1,\Gamma_2}{\cclrd{r} \,\cpar\, \cclrd{t} \,\cpar\, \cclrd{s} \,\cpar\, \cclrd{q}} \vdash e : \tau} }
\end{equation*}
%
From idempotence of $\czip$, we get that $\cclrd{r} \,\czip\, \cclrd{s} \,\czip\, \cclrd{t} \,\czip\, \cclrd{q} =
\cclrd{r} \,\czip\, \cclrd{t} \,\czip\, \cclrd{s} \,\czip\, \cclrd{q}$. Thus the judgement can be derived using
(\emph{sub}) as in contraction.
\end{proof}

\noindent
A consequence of the equivalence discussed above is that the unified coeffect system can capture 
all properties that can be captured by the flat coeffect system -- including implicit parameters,
rebindable resources, Haskell type classes (discussed by Orchard~\cite{comonads-dom-thesis}),
data-flow and variable liveness.

%--------------------------------------------------------------------------------------------------

\subsection{Semanitcs of unified calculus}
\label{sec:unified-semantics}

The semantics of the unified calculus can be obtained by generalizing the semantics of the structural
calculus in the same way as we generalized the type system. However, it is worth exploring the type 
signatures of the operations of the underlying (comonad-based) structure.

The semantics of the unified coeffect calculus follows the same pattern as the semantics of the
structural calculus (Section~\ref{sec:struct-semantics}) with a number of differences. The 
index $\aclrd{\textbf{r}}$ of the object mapping $\ctyp{\aclrd{\textbf{r}}}$ is an element of 
the structure $\C^{\SP(\sclrd{n})}$ for a specified container $\SH \triangleleft \SP$ and a shape 
$\sclrd{n}$ determined as $\sclrd{n} = \slen{\Gamma}$. To match the type system discussed in 
Section~\ref{sec:unified-types}, we need to modify the domains and codomains of the \ident{merge}, 
\ident{split} and \ident{dup} operations as follows.
%
\begin{equation*}
\begin{array}{lcl}
 \ident{merge}_{\aclrd{\textbf{r}},\aclrd{\textbf{s}}} &\narrow{:}& 
  \ctyp{\aclrd{\textbf{r}}}{\alpha} \times \ctyp{\aclrd{\textbf{s}}}{\beta} \rightarrow \ctyp{\aclrd{\textbf{r}}\azip\aclrd{\textbf{s}}}{(\alpha \xtimes \beta)} \\
 \ident{split}_{\aclrd{\textbf{r}},\aclrd{\textbf{s}}} &\narrow{:}& 
  \ctyp{\aclrd{\textbf{r}}\apar\aclrd{\textbf{s}}}{(\alpha \xtimes \beta)} \rightarrow \ctyp{\aclrd{\textbf{r}}}{\alpha} \times \ctyp{\aclrd{\textbf{s}}}{\beta} \\
 \ident{dup}_{\cclrd{r}, \cclrd{s}} &\narrow{:}& 
  \ctyp{\alift{\cclrd{r} \cpar \cclrd{s}}}{\alpha} \rightarrow \ctyp{\alift{\cclrd{r}}\,\azip\,\alift{\cclrd{s}}}{(\alpha \xtimes \alpha)} \\
\end{array}
\end{equation*}
%
The key difference is that the vector concatenation $\atimes$ is replaced with $\apar$ in the
\emph{domain} of the operations (in \ident{split}) and with $\azip$ in the codomain
(in \ident{merge}, but also in \ident{dup} which models contraction).
This means that the splitting and merging the context can change the coeffect annotation:
%
\begin{equation*}
\ident{merge}_{\aclrd{\textbf{r}},\aclrd{\textbf{s}}} \circ \ident{split}_{\aclrd{\textbf{r}},\aclrd{\textbf{s}}} ~:~
  \ctyp{\aclrd{\textbf{r}}\apar\aclrd{\textbf{s}}}{(\alpha \xtimes \beta)}
  \rightarrow \ctyp{\aclrd{\textbf{r}}\azip\aclrd{\textbf{s}}}{(\alpha \xtimes \beta)} 
\end{equation*}
%
For flat coeffect calculi where $(\aclrd{r}\,\azip\,\aclrd{s}) \;\aclrd{\leq}\; (\aclrd{r}\,\apar\,\aclrd{s})$,
splitting and then merging a context may lose some of the avilable contextual capabilities
(and increase context requirements). This illuminates why our proofs of Lemma~\ref{thm:unified-rev-contr} 
and Lemma~\ref{thm:unified-rev-exch} relied on sub-coeffecting in the flat system.

\paragraph{Related and future work.}
The semantics presented here serves main\-ly to inform the design of the type system. For this 
reason, we keep the semantics simple and close to the concrete notions used in the type system
such as flat, structural and unified coeffect algebras. In particular, the indexed comonad 
structure consists of object mapping $\ctyp{r}{}$ that is indexed by a corresponding algebra:

\begin{compactitem}
\item[--] In the flat coeffect calculus, $\ctyp{\cclrd{r}}{}$ is indexed by scalar coeffects $\cclrd{r} \in \C$
\item[--] In the structural coeffect calculus, the object mapping $\ctyp{\aclrd{\textbf{r}}}{}$ is 
  indexed by vectors of scalars $\aclrd{\textbf{r}} \in \bigcup_{\sclrd{m}\in \sclrd{\mathbb{N}}} \C^{\sclrd{m}}$
\item[--] In unified coeffect calculus with a coeffect container $\SH\,\sclrd{\triangleleft}\,\SP$,
  the mapping $\ctyp{\aclrd{\textbf{r}}}{}$ is indexed by shape-indexed scalars 
  $\aclrd{\textbf{r}} \in \bigcup_{\sclrd{s}\in \SH} \C^{\SP(\sclrd{m})}$
\end{compactitem}

\noindent
In this treatment, the indices are formed by objects that are modelled outside of category theory.
A fully categorical semantics can be found in a joint paper with Orchard \cite{coeffects-icfp14}. 
There, the index (coeffect scalars and shape-indexed scalars) are modelled as categories too.
The definition is written using the following notation:

\begin{compactitem}
\item[--] $[\mathbb{C}, \mathbb{D}]$ is a category of functors between categories $\mathbb{C}$ and $\mathbb{D}$
\item[--] $B^A$ is an exponential object $A \Rightarrow B$ in a Cartesian-closed category
\end{compactitem}

\noindent
In the categorical semantics of coeffect calculi, coeffect scalars are modelled as a
category of scalars $\cclrd{\mathbb{I}}$ such that $\textit{obj}(\cclrd{\mathbb{I}}) = \C$.
The structure of the context (indexed comonad) over the input then becomes a functor
indexed by the free-variable context shape $\slen{\Gamma}$ and coeffect annotations
$\aclrd{\textbf{r}}$, \ie~$\ctyp{\slen{\Gamma}}_{\aclrd{\textbf{r}}}$:
%
\begin{equation*}
\sem{\coctx{\Gamma}{\aclrd{R}} \vdash e : \tau} : \ctyp{\slen{\Gamma}}_{\aclrd{R}}
\sem{\Gamma} \rightarrow \sem{\tau}
\end{equation*}
%
The structure $\ctyp{}$ can be thought of as a dependent product of functors $\ctyp{\sclrd{n}}$ 
over possible shapes $\sclrd{n} \in \SH$ where:
%
\begin{equation*}
\ctyp{\sclrd{n}} : \cclrd{\mathbb{I}}^{\SP(\sclrd{n})} \rightarrow [\mathbb{C}^{\SP(\sclrd{n})}, \mathbb{C}]
\end{equation*}
%
For a fixed context shape $\sclrd{n}$ the functor 
$\ctyp{\sclrd{n}} : \cclrd{\mathbb{I}}^{\SP(\sclrd{n})} \rightarrow [\mathbb{C}^{\SP(\sclrd{n})}, \mathbb{C}]$ 
maps a coeffect indexed by positions $\SP(\sclrd{n})$ to a functor from a context 
$\mathbb{C}^{\sclrd{n}}$ to an object in $\mathbb{C}$. That is, given a coeffect annotation 
(matching the shape of the context), we get a functor $\in [\mathbb{C}^{\sclrd{n}},\mathbb{C}]$. 
From a programming perspective, this functor defines a data structure that models the additional
context provided to the program. The shape of this data structure depends on the coeffect
annotation $\cclrd{\mathbb{I}}^{\sclrd{n}}$. For example, in bounded reuse, the annotation defines
the number of values needed for each variable and the functor will be 
formed by lists of length matching the required number.

In the model above, shapes and positions are still treated as ordinary sets. Indeed, in the 
joint paper \cite{coeffects-icfp14}, we treat shapes as just sets of sets. In this chapter, we 
instead define \emph{coeffect container} as an extension of containers, which have a well-defined 
categorical structure \cite{semantics-containers}. A future work is adapting the semantics of
the unified coeffect calculus so that it is based on the semantics of containers. However, note 
that context is \emph{not} a (simple) container, because containers are used as indices (coeffect
annotations) of the object that wrap the free-variable context.

% =================================================================================================
%                                                                   
%      ####                         #       #                   ##   
%      #   #                        #                            #   
%      #   #  # ##    ###    ###   ####    ##     ###    ###     #   
%      ####   ##  #      #  #   #   #       #    #   #      #    #   
%      #      #       ####  #       #       #    #       ####    #   
%      #      #      #   #  #   #   #  #    #    #   #  #   #    #   
%      #      #       ####   ###     ##    ###    ###    ####   ###  
%                                                                
% =================================================================================================

\section{Towards practical coeffects}
\label{sec:unified-impl}

As discussed earlier, the main focus of this thesis is the development of the much needed 
\emph{theory of context-aware programming languages} and so discussing the details of a practical 
implementation of coeffect tracking is beyond the scope of the thesis. However, this section
briefly outlines one possible pathway towards this goal.

Many of the examples of contextual computation that we discussed earlier have been implemented as 
a single-purpose programming language feature (\eg~implicit parameters \cite{app-implicit-parameters} 
or distributed computations \cite{app-distributed-ml5,app-distributed-links}). However, 
the main contribution of this thesis is that it captures
\emph{multiple} different notions of context-aware computations using just a \emph{single}
common abstraction. For this reason, we advocate that future practical implementations of
coeffects should not be single-purpose language features, but rather reusable abstractions that
can be instantiated with a concrete \emph{coeffect algebra} specified by the user.

In order to do this, programming languages need to provide two features; one that allows 
embedding of context-aware computations themselves in programs akin to the ``do'' notation
in Haskell (Section~\ref{sec:unified-impl-embed}) and one that allows tracking of the 
contextual information in the type system (Section~\ref{sec:unified-impl-types}). To make
constraint resolution in type inference easier, we also discuss how the coeffect algebra 
can be simplified (Section~\ref{unified-impl-semilattice})

% -------------------------------------------------------------------------------------------------

\subsection{Embedding contextual computations}
\label{sec:unified-impl-embed}

The embedding of \emph{contextual} computations into programming languages can follow the successful
model of \emph{effectful} computations. In purely functional programming languages such as Haskell, 
effectful computations are embedded by \emph{implementing} their model within the language 
and inserting the necessary (monadic) plumbing. This is made easier by the ``do'' notation or
monad comprehensions \cite{other-haskell98,monads-compre}, which insert the monadic operations 
automatically.

The recently proposed ``codo'' notation \cite{comonads-codo} provides similar automation for
context-aware computations based on comonads. The notation follows the semantics of our 
flat coeffect calculus (Chapter~\ref{ch:flat}). Extending the ``codo'' notation to support calculi based 
on the structural coeffect calculus (Chapter~\ref{ch:structural}) is an interesting future work -- 
this would require explicitly manipulating individual context variables and application of structural
rules, which is not needed in flat coeffects.

In ML-like languages, effects (and many coeffects) are built-in into the language semantics, 
but they can still benefit from a special notation for explicitly marking effectful (coeffectful) 
blocks of code. In F\#, this is done using \emph{computation expressions}  \cite{app-computation-zoo}
that differ from the ``do'' notation in two ways. First, they support wider range of 
constructs, making it possible to wrap existing F\# code within a block without other changes.
Second, they support other abstractions including monads, applicative functors and monad
transformers. It would be interesting to see if computation expressions can be extended to
support programming with computations based on flat/structural indexed comonads.

More lightweight syntax for effectful computation can be obtained using
techniques that automatically insert the necessary monadic plumbing (without
special syntax). This has been done for effecful computations \cite{monads-lightweight-ml}
and a similar approach would be worth exploring for coeffects.

% -------------------------------------------------------------------------------------------------

\subsection{Coeffect annotations as types}
\label{sec:unified-impl-types}

The other aspect of practical implementation of coeffects is that of tracking the context requirements
(coeffect annotations) in the type system. To achieve this (without resorting to a single-purpose
language feature) the type system needs to be able to capture various kinds of coeffect algebras. 
The structures required in this thesis include sets (with union or interesection), natural numbers (with 
addition, maximum, minimum and multiplication), two-point lattice (for liveness) and free monoids 
(vectors of annotations).

The work \cite{effects-embedding} on embedding effect systems in Haskell demonstrated that the
recent additions to the Haskell type system provide enough power to implement structures such
as sets at the type level. Using these to embed coeffect systems in Haskell is one fruitful 
future direction for applied coeffects.

In dependently-typed programming languages such as Agda or Idris \cite{other-agda,other-idris},
the embedding of coeffects can be implemented more directly (as terms implementing sets or lattices
can be lifted to the type level). However, we believe that coeffect tracking does not require full 
dependent types and can be made accessible in more mainstream languages. Dependent ML 
\cite{types-dependent-ml} provides an interesting example of a language with some dependent typing 
(\eg~computations with integers) which is still close to its non-dependently-typed predecessor ML.

Another approach for embedding computations into the type system has been pioneered by F\#
\emph{type providers} \cite{app-inforich}. Technically, type providers are compiler extensions 
that generate types based on external data sources or other information in order to provide easy 
access to data or services. A similar approach could be used for embedding \emph{algberas} such
as coeffect algebras into the type system. An \emph{algebra provider} would then be a library that 
specifies the objects of the algebra, its equational laws and generalization rules for type inference.
This could provide an easy-to-use way of embedding coeffect tracking in pragmatic languages
such as F\#. It is worth noting that the mechanism could also subsume F\# units of measure
\cite{types-units-of-measure}; these could be alternatively provided via one such \emph{algebra
provider}.

% -------------------------------------------------------------------------------------------------

\subsection{Alternative formulation using coeffect lattice}
\label{sec:unified-impl-semilattice}

Compared to typical effect systems based on sets, coeffect systems have richer structure consisting
of four operations ($\cseq,\czip,\cpar$ and a relation $\cleq$). This allows capturing interesting 
properties, but it makes working with coeffect annotations difficult -- it compilcates tasks
such as type generalization and constraint solving in type inference. Here, we look at one possible 
simplification of the structure.

\begin{itemize}
\item As mentioned in Section~\ref{sec:flat-calculus-algebra}, when the monoid $(\C,\cpar,\czero)$ is a
semi-lattice, the $\cleq$ relation can often be expressed in terms of the $\cpar$ operation 
($\cclrd{r} \;\cleq\; \cclrd{s} \;\Longleftrightarrow\; \cclrd{r} \;\cpar\; \cclrd{s} \;=\; \cclrd{s}$).
This does not give a definition equivalent to our, but it is a simplification that is consistent 
with all examples in this thesis.

\item Another approach that works for most (but not all) systems we discussed is to require the structure
$(\C,\cpar,\czip)$ to form a lattice. This is consistent with all our structural coeffect systems as
well as flat systems for liveness and data-flow, but not with the system for tracking of implicit
parameters (or sets of rebindable resources).
\end{itemize}

\noindent
The system for tracking implicit parameters does not fit the lattice-based model, because it uses the 
same operation $\cup$ for both $\cpar$ and $\czip$.

This also means that the system for implicit parameters cannot be used with a system where lambda 
abstraction duplicates the context requirements (Section~\ref{sec:flat-exts-lambda}).

\paragraph{Lattice-based coeffects.} Putting implicit parameters aside, there are many other coeffect
systems for which we could use the following definition of coeffect scalars (for simplicity, we only 
consider flat systems):

\begin{definition}
A \emph{\cclrd{flat coeffect lattice}} $(\C, \cseq, \cleq, \cunit, \czero)$ is a set 
$\C$ together with elements $\cunit, \czero \in \C$ and an ordering $\cleq$ such 
that $(\C, \cseq, \cunit)$ is a monoid and $(\C, \cleq)$ is a lattice with the least 
element $\czero$.

Assuming $\cpar$ and $\czip$ are the \emph{least upper bound} and \emph{greatest lower bound}
of the lattice, we require the following distributivity axioms:
%
\begin{align*}
\quad (\cclrd{r}\, \cpar\, \cclrd{s}) \;\cseq\; \cclrd{t} & = (\cclrd{r} \,\cseq\, \cclrd{t}) \;\cpar\; (\cclrd{s}\, \cseq\, \cclrd{t}) \\
\quad \cclrd{t} \;\cseq\; (\cclrd{r}\, \cpar\, \cclrd{s}) & = (\cclrd{t} \,\cseq\, \cclrd{r}) \;\cpar\; (\cclrd{t}\, \cseq\, \cclrd{s})
\end{align*}
\end{definition}

\noindent
The definition provides all the operations of a flat coeffect algebra, but it defines a more rigid 
structure. In particular, $\czip$ and $\cpar$ are both defined in terms of $\cleq$. This simplifies 
the system as the operations of the underlying indexed comonad structure can be expressed using just
$\cseq$ and $\cleq$:
%
\begin{equation*}
\begin{array}{lcll}
 \ident{counit}_{\cunit} &\narrow{:}& 
    \ctyp{\cunit}{\alpha} \rightarrow \alpha \\
 \ident{cobind}_{\cclrd{r}, \cclrd{s}} &\narrow{:}& 
    (\ctyp{\cclrd{r}}{\alpha} \rightarrow \beta) \rightarrow (\ctyp{\cclrd{r}\cseq\cclrd{s}}{\alpha} \rightarrow \ctyp{\cclrd{s}}{\beta}) \\
\ident{merge}_{\cclrd{r},\cclrd{s},\cclrd{t}} &\narrow{:}& 
    \ctyp{\cclrd{r}}{\alpha} \times \ctyp{\cclrd{s}}{\beta} \rightarrow \ctyp{\cclrd{t}}{(\alpha \times \beta)} &
    (\cclrd{t}\,\cleq\,\cclrd{r}, \cclrd{t}\,\cleq\,\cclrd{s}) \\
\ident{split}_{\cclrd{r},\cclrd{s},\cclrd{t}} &\narrow{:}& 
    \ctyp{\cclrd{t}}{(\alpha \times \beta)} \rightarrow \ctyp{\cclrd{r}}{\alpha} \times \ctyp{\cclrd{s}}{\beta} &
    (\cclrd{r}\,\cleq\,\cclrd{t}, \cclrd{s}\,\cleq\,\cclrd{t}) \\
\ident{lift}_{\cclrd{r},\cclrd{s}} &\narrow{:}& 
    \ctyp{\cclrd{r}}{\alpha} \rightarrow \ctyp{\cclrd{s}}{\alpha} & 
    (\cclrd{s}\,\cleq\,\cclrd{r})
\end{array}
\end{equation*}
%
Note that we still need \emph{two distinct} operators in the system. Analysis of our examples shows
that the operator used for sequential composition $\cseq$ is not necessarilly related in any way to 
the ordering on the lattice provided by $\cleq$.

However, the operations $\ident{merge}$, $\ident{split}$ and $\ident{lift}$ can now all
use ordering. For example, the \ident{merge} operation previously returned
a result with greatest lower bound $\cclrd{r}\;\czip\;\cclrd{s}$. Now, we require any
element that is smaller than $\cclrd{r}$ and $\cclrd{s}$. This adds an implicit sub-coeffecting
to the operation -- for any $\cclrd{t}$ such that $\cclrd{t}\;\cleq\;\cclrd{r}$ and 
$\cclrd{t}\;\cleq\;\cclrd{s}$, it is also the case that $\cclrd{t}\;\cleq\;(\cclrd{r}\,\czip\,\cclrd{s})$.
This means that we can compose the original \ident{merge} operation which returns greatest lower
bound with $\ident{lift}$ to get the (less precise) operation required here.

% -------------------------------------------------------------------------------------------------

\begin{figure}[t]
{\small a.) Littice-based flat systems (selected rules):}

\begin{equation*}
\tyrule{abs}
  {\coctx{\Gamma, x:\tau_1}{\cclrd{t}} \vdash e : \tau_2}
  {\coctx{\Gamma}{\cclrd{r}} \vdash \lambda x.e : \tau_1 \xrightarrow{\cclrd{s}} \tau_2 }
~~(\cclrd{t}\,\cleq\,\cclrd{r}, \cclrd{t}\,\cleq\,\cclrd{s})
\end{equation*}
\begin{equation*}
\tyrule{app}
  {\coctx{\Gamma}{\cclrd{r}} \vdash e_1 : \tau_1 \xrightarrow{\cclrd{t}} \tau_2 &
   \coctx{\Gamma}{\cclrd{s}} \vdash e_2 : \tau_1 }
  {\coctx{\Gamma}{\cclrd{u}} \vdash e_1~e_2 : \tau_2}
~~(\cclrd{r}\,\cleq\,\cclrd{u}, (\cclrd{s} \,\cseq\, \cclrd{t})\,\cleq\,\cclrd{u})
\end{equation*}

\vspace{1.5em}
{\small b.) Littice-based structural systems (selected rules):}

\begin{equation*}
\tyrule{contr}
  {\coctx{\Gamma_1,y\!:\!\tau_1,z\!:\!\tau_1,\Gamma_2}{\aclrd{\textbf{r}} \atimes \alift{\cclrd{s},\cclrd{t}} \atimes \aclrd{\textbf{q}}} \vdash e : \tau}
  {\coctx{\Gamma_1,x\!:\!\tau_1,\Gamma_2}{\aclrd{\textbf{r}} \atimes \alift{\cclrd{u}} \atimes \aclrd{\textbf{q}}} \vdash \subst{e}{z,y}{x} : \tau}
~~(\cclrd{s}\,\cleq\,\cclrd{u}, \cclrd{t}\,\cleq\,\cclrd{u})
\end{equation*}

\figcaption{Flat and structural coeffects using lattice-based formulation}
\label{fig:unified-lattice-types}
\end{figure}

% -------------------------------------------------------------------------------------------------

\paragraph{Lattice-based type system.}
\hspace{-1em}
Using lattice-based coeffect structure leads to the typing rules shown in Figure~\ref{fig:unified-lattice-types}.
In the flat coeffect system, this affects both application and abstraction -- in abstraction,
the context-requirements of the body should be smaller than the context available on the 
call site and declaration site. In application, we require more than what is required by
the sub-expressions.

Structural coeffect systems only require least upper bound $\cpar$ in the (\emph{contr}) rule.
This can be rewritten similarly to the flat (\emph{app}) rule --  when contracting two variables, the
context requirements of the resulting variable are greater than the requirements associated with
each of the original variables.

In smmary, the lattice-based coeffect system has some advantages over the design used in this
thesis. In particular, it uses better-known structures and is simpler. However, we do not use
it as it rules out one of the important motivating examples (implicit parameters) and, even with
the simplification, solving constraints with $\cseq$ and $\cleq$ is still an open question.

% =================================================================================================
%                                                                                             
%    #   #          #                    ##                                                    
%    #   #          #                     #                                                    
%    ## ##   ###   ####    ###            #     ###   # ##    ## #  #   #   ###    ## #   ###  
%    # # #  #   #   #         #  #####    #        #  ##  #  #  #   #   #      #  #  #   #   # 
%    #   #  #####   #      ####           #     ####  #   #   ##    #   #   ####   ##    ##### 
%    #   #  #       #  #  #   #           #    #   #  #   #  #      #  ##  #   #  #      #     
%    #   #   ###     ##    ####          ###    ####  #   #   ###    ## #   ####   ###    ###  
%                                                            #   #                #   #        
%                                                             ###                  ###         
% =================================================================================================

\section{Coeffect meta-language}
\label{sec:unified-meta}

In Section~\ref{sec:path-sem-langs}, we discussed two ways of using monads in programming
language semantics as introduced by Moggi \cite{monad-notions}. The first approach is to use
monads in the \emph{semantics} of an effectful language. The second approach is to extend the
language with (additional) monadic constructs that can then be used for writing effectful
monads explicitly.

In this thesis, we focused on the first approach. In both flat and structural coeffect calculi,
the term language is that of a simply-typed $\lambda$-calculus, and we used (flat or structural)
indexed comonads to give the semantics for the language and to derive type system for it.

In this section, we briefly discuss the other technique. That is, we embed indexed
comonads into a $\lambda$-calculus as additional constructs. To do that, we introduce the
type constructor $\ctyp{\cclrd{r}}{\tau}$ which represents a value $\tau$ wrapped in additional
context (semantically, this corresponds to an indexed comonad) and we add language constructs
that correspond to the operations of indexed comonads.

This section provides a brief sketch of the coeffect meta-language in order to highlight the relationship
between coeffects and closely related work on contextual modal type theory (CMTT) \cite{logic-cmtt}.
Developing the system further is an interesting future research direction.

% -------------------------------------------------------------------------------------------------

\subsection{Coeffects and contextual modal type theory}

As discussed in Section~\ref{sec:path-sem-contextdep}, context-aware computations are related
to modal logics -- comonads have been used to model the $\square$ modality and as a basis for
meta-languages that include $\square$ as a type constructor
\cite{logic-intuitionistic-modal,logic-modal-reconstruction,logic-intuitionistic-modal,logic-cmtt}.
Nanevski et al. \cite{logic-cmtt} extend an S4 term language to a contextual modal type theory (CMTT).
From the perspective of this thesis, CMTT can be viewed as a \emph{meta-language} version of our
coeffect calculus.

\paragraph{Context in CMTT and coeffects.}
Aside from the fact that coeffect calculi use comonads for \emph{semantics} and CMTT embeds
comonads (the $\square$ modality) into the meta-language, there are two important differences.

Firstly, the \emph{context} in CMTT is a set of variables required by a computation, which
makes CMTT useful for meta-programming and staged computations. In coeffect calculi, the context
requirements are formed by an abstract coeffect algebra, which is more general and can capture
variable contexts, but also integers, two-point lattices, etc.

Secondly, CMTT uses different intuitive understanding of the comonad (type constructor) and
the associated operations. In categorical semantics of coeffect calculi, the $\ctyp{\cclrd{r}}{\tau}$
constructor refers to a value of type $\tau$ \emph{together} with additional context specified
by the coeffect annotation $\cclrd{r}$ (\eg~a list of past values or additional resources).

In contrast in CMTT\footnote{To avoid using different notations, we write $\ctyp{\Psi}{\tau}$ instead of the 
original $[\Psi]{\tau}$} the type $\ctyp{\Psi}{\tau}$ models a value that \emph{requires} the 
context $\Psi$ in order to produce value $\tau$. This also changes the interpretation of the 
two operations of a comonad:
\begin{equation*}
\begin{array}{rcl}
 \ident{counit} &\narrow{:}& \ctyp{\cunit}{\alpha} \rightarrow \alpha\\
 \ident{cobind} &\narrow{:}& (\ctyp{\cclrd{r}}{\alpha} \rightarrow \beta) \rightarrow 
    \ctyp{\cclrd{r}\cseq\cclrd{s}}{\alpha} \rightarrow \ctyp{\cclrd{s}}{\beta}
\end{array}
\end{equation*}
Two readings of the signatures are possible, but they give quite different meanings to the operations:

\begin{itemize}
\item \emph{Coeffect interpretation.} The \ident{counit} operation extracts a value 
  and does not require any additional context; the \ident{cobind} operation requires
  context $\cclrd{r}\,\cseq\,\cclrd{s}$, uses the part specified by $\cclrd{r}$ to 
  evaluate the function, ending with a value $\beta$ together with remaining context
  $\cclrd{s}$.
\item \emph{CMTT interpretation.} The \ident{counit} operation evaluates a computation 
  that requires no additional context to obtain a $\alpha$ value; given a function that
  turns a computation requiring context $\cclrd{r}$ into a value $\beta$, the \ident{cobind}
  operation can turn a computation that requires context $\cclrd{r}\,\cseq\,\cclrd{s}$ 
  into a computation that requires just $\cclrd{s}$ and contains $\beta$ (a part of the
  context requirements is eliminated by the function).
\end{itemize}

\noindent
Although the different reading does not affect formal properties of the systems, it is important
to understand the difference when discussing the two systems as they provide a different 
intuition.

The sketch of a coeffect meta-language in the following section attempts to bridge the gap
between coeffects and CMTT. Just like CMTT, it embeds comonads as language constructs, but
it annotates them with a (flat) coeffect algebra, thus it generalizes CMTT which tracks only
sets of variables. Future work on the coeffect meta-language would thus be an interesting 
development for both coeffect systems and CMTT.

% -------------------------------------------------------------------------------------------------

\subsection{Coeffect meta-language}

The coeffect meta-langauge could be designed using both flat and structural indexed comonads.
For simplicity, this section only discusses the flat variant. The syntax of types and terms
of the langauge includes the type constructor $\ctyp{\cclrd{r}}{\tau}$ and four additional
language constructs:
%
\begin{equation*}
\begin{array}{rcl}
 \tau &\narrow{::=}& \alpha \sep \tau_1 \rightarrow \tau_2 \sep \ctyp{\cclrd{r}}{\tau}\\
 e &\narrow{::=}& v \sep \lambda x.e \sep e_1~e_2 \sep !e    \\[-0.25em]
  &        \sep& \kvd{let\;box}~x=e_1~\kvd{in}~e_2           \\[-0.25em]
  &        \sep& \kvd{split}~e_1~\kvd{into}~x,y~\kvd{in}~e_2 \\[-0.25em]
  &        \sep& \kvd{merge}~e_1, e_2~\kvd{into}~x~\kvd{in}~e_2 
\end{array}
\end{equation*}
%
The $!e$ and $\kvd{let\;box}$ constructs correspond to the \ident{counit} and \ident{cobind}
operation of the comonad. To define meta-langauge for flat indexed comonads, we also include 
\kvd{split} and \kvd{merge} that embed the corresponding operations.

\paragraph{Types for coeffect meta-language.}
The type system for the language is shown in Figure~\ref{fig:conclusions-cml}. The first part
shows the usual typing rules for simply-typed $\lambda$-calculus. For simplicity, we omit typing 
rules for pairs, but those need to be present as the \kvd{merge} operation works on tuples.

The second part of the typing rules is more interesting. The (\emph{counit}) operation extracts
a value from a comonadic context and corresponds to variable access in coeffect calculi. The
\kvd{let\;box} construct (\emph{cobind}) takes an input $e_1$ with context $\cclrd{r}\,\cseq\,\cclrd{s}$
and a computation that turns a variable $x$ with a context $\cclrd{r}$ into a value $\tau_2$.
The result is a computation that produces a $\tau_2$ value with the remaining context specified
by $\cclrd{s}$. Note that the expressions $e_2$ and $e_1$ correspond to the first and second
arguments of the \ident{cobind} operation. The keyword \kvd{let\;box} is chosen following 
CMTT\footnote{The rule is similar to the \ident{letbox} rule for ICML \cite[p. 14]{logic-cmtt},
although it differs because of our generalization of comonads where \ident{bind} composes
coeffect annotations rather than requiring the same annotation everywhere.}.

The \kvd{split} and \kvd{merge} constructs follow a similar pattern. They both apply some 
transformation on one or two values in a context and then add the new value as a fresh variable
to the variable context. We omit sub-coeffecting, but it could be easily added following 
the method used elsewhere.

% -------------------------------------------------------------------------------------------------

\begin{figure}[t]

{\small a.) Typing rules for the simply typed $\lambda$-calculus}

\begin{equation*}
\tyrule{var}
  {x : \tau \in \Gamma}
  {\Gamma \vdash x : \tau }
\end{equation*}
\begin{equation*}
\tyrule{app}
  {\Gamma \vdash e_1 : \tau_2 \rightarrow \tau_1 &
   \Gamma \vdash e_2 : \tau_2 }
  {\Gamma \vdash e_1~e_2 : \tau_1 }
\end{equation*}
\begin{equation*}
\tyrule{abs}
  {\Gamma, x:\tau_1 \vdash e : \tau_2}
  {\Gamma \vdash \lambda x.e : \tau_1 \rightarrow \tau_2 }
\end{equation*}

\vspace{1em}
{\small b.) Additional typing rules arising from \emph{flat indexed comonads}}

\begin{equation*}
\tyrule{cobind}
  {\Gamma \vdash e_1 : \ctyp{\cclrd{r}\,\cseq\,\cclrd{s}}{\tau_1} & \Gamma, x : \ctyp{\cclrd{r}}{\tau_1} \vdash e_2 : \tau_2 }
  {\Gamma \vdash \kvd{let\;box}~x=e_1~\kvd{in}~e_2 : \ctyp{\cclrd{s}}{\tau_2}}
\end{equation*}
\begin{equation*}
\tyrule{counit}
  {\Gamma \vdash e : \ctyp{\cunit}{\tau}}
  {\Gamma \vdash !e : \tau}
\end{equation*}
\begin{equation*}
\tyrule{split}
  {\Gamma \vdash e_1 : \ctyp{\cclrd{r}\,\cpar\,\cclrd{s}}{\tau_1} &
   \Gamma, x:\ctyp{\cclrd{r}}{\tau_1}, y:\ctyp{\cclrd{s}}{\tau_1} \vdash e_2 : \tau_2 }
  {\Gamma \vdash \kvd{split}~e_1~\kvd{into}~x,y~\kvd{in}~e_2 : \tau_2 }
\end{equation*}
\begin{equation*}
\tyrule{merge}
  {\Gamma \vdash e_1 : \ctyp{\cclrd{r}}{\tau_1} &
   \Gamma \vdash e_2 : \ctyp{\cclrd{s}}{\tau_2} \\
   \Gamma, x:\ctyp{\cclrd{r}\,\czip\,\cclrd{s}}{(\tau_1 \times \tau_2)} \vdash e_2 : \tau_3 }
  {\Gamma \vdash \kvd{merge}~e_1, e_2~\kvd{into}~x~\kvd{in}~e_2 : \tau_3 }
\end{equation*}

\figcaption{Type system for the (flat) coeffect meta-language}
\label{fig:conclusions-cml}
\end{figure}

% -------------------------------------------------------------------------------------------------

\subsection{Embedding flat coeffect calculus}

The \emph{meta-language} approach of embedding comonads in a language is more general than the
\emph{semantics} approach. This thesis focuses on a narrower use that better guides the design of
a type system for context-aware programming languages.

However, it is worth demonstrating that the (flat) coeffect calculus can be embedded in the 
meta-language described in the previous section. This may be desirable, \eg~when using the 
meta-language for reasoning about context-aware computations. We briefly consider the embedding 
as it illuminates the relationship between coeffects and CMTT (unfortunately it is not possible 
to embed coeffect calculi in CMTT because of the more general annotations structure).

Given a typing judgement $\coctx{\Gamma}{\cclrd{r}} \vdash e : \tau$ in the flat coeffect calculus,
we define $\sem{\coctx{\Gamma}{\cclrd{r}} \vdash e : \tau}_v$ as its emebdding in the coeffect
meta-language. Note that the translation is indexed by $v$, which is a name of variable used 
to represent the entire variable context of the source language. The translation is defined in
Figure~\ref{fig:conclusions-embed}. The embedding resembles the semantics discussed in 
Section~\ref{sec:flat-semantics}. This is not surprising as the meta-language directly mirrors
the operations of an indexed comonad.

% -------------------------------------------------------------------------------------------------

\begin{figure}[t]
\begin{equation*}
\begin{array}{l}
 \sem{ \coctx{\Gamma}{\cunit}\vdash x_i : \tau_i }_v = \pi_i (!v) \\[0.25em]
 \sem{ \coctx{\Gamma}{\cclrd{r}\;\cpar\;(\cclrd{s}\,\cseq\,\cclrd{t})} \vdash e_1~e_2 : \tau_2 }_v =\\[-0.25em]
   \qquad \kvd{split}~v~\kvd{into}~v_s, v_{rt}~\kvd{in} \\[-0.25em]
   \qquad \sem{ \coctx{\Gamma}{\cclrd{s}} \vdash e_1 : \tau_1 \xrightarrow{\cclrd{t}} \tau_2 }_{v_s}
      ~(\kvd{let\;box}~v_r = v_{rt}~\kvd{in}~\sem{ \coctx{\Gamma}{\cclrd{r}} \vdash e_2 : \tau_1}_{v_r}) \\[0.25em]
 \sem{ \coctx{\Gamma}{\cclrd{r}} \vdash \lambda x.e:\tau_1 \xrightarrow{\cclrd{s}} \tau_2 }_v = \lambda x. \\[-0.25em]
   \qquad \kvd{merge}~v,x~\kvd{into}~v_{rs}~\kvd{in}~\sem{ \coctx{(\Gamma,x:\tau_1)}{\cclrd{r}\,\czip\,\cclrd{s}} \vdash e : \tau_2 }_{v_{rs}}
\end{array}
\end{equation*}
\figcaption{Embedding flat coeffect calculus in coeffect meta-language }
\label{fig:conclusions-embed}
\end{figure}

% ==================================================================================================
%                                                                             
%        ###                         ##                    #                        
%       #   #                         #                                             
%       #       ###   # ##    ###     #    #   #   ###    ##     ###   # ##    ###  
%       #      #   #  ##  #  #   #    #    #   #  #        #    #   #  ##  #  #     
%       #      #   #  #   #  #        #    #   #   ###     #    #   #  #   #   ###  
%       #   #  #   #  #   #  #   #    #    #  ##      #    #    #   #  #   #      # 
%        ###    ###   #   #   ###    ###    ## #  ####    ###    ###   #   #  ####  
%                                                                             
% ==================================================================================================

\section{Conclusions}

In this chapter, we completed our quest of presenting a \emph{unified} system for tracking contextual
properties and then we discussed two important topics of future and related work that require 
in-depth discussion.

The key technical contribution of this chapter is the \emph{unified coeffect system} 
(Section~\ref{sec:unified-unified}), which unifies the flat and structural system presented
in the previous two chapters. To achieve this, we introduced \emph{coeffect container}, which
determines how are coeffect annotations attached to variable contexts. We then discussed two
instances of the structure that capture flat and structural properties.

In the rest of the chapter, we discussed two topis of future and related work. First
(Section~\ref{sec:unified-impl}), we considered pathways to practical implementations of
coeffect systems, including a discussion of a novel lattice-based coeffect system, which is simpler,
but cannot capture all our motivating examples. Finally, we discussed how our work relates to 
meta-language based on comonads (Section~\ref{sec:unified-meta}). We present a \emph{coeffect 
meta-language} that follows similar style to CMTT, but is based on indexed comonads. Next, we 
discussed how to embed flat coeffect systems in this coeffect meta-language, which elucidates the 
relationship between our work and CMTT.