### 2015-06-16 response.tex
Add first response draft

% ==================================================================================================
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
% ==================================================================================================

\documentclass[
		twoside,openright,titlepage,numbers=noenddot,headinclude,%1headlines,
                footinclude=true,cleardoublepage=empty,
                BCOR=10mm,paper=a4,fontsize=10pt, % Binding correction, paper type and font size
                ngerman,american, % Languages
                ]{scrreprt}

\input{classicthesis-config}
\usepackage{semantic}
\usepackage{paralist}
\usepackage{etoolbox}
\usepackage{stmaryrd}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm}

\newcounter{mythmcounter}
\newcounter{mydefcounter}
\newcounter{myexpcounter}
\newtheorem{example}[myexpcounter]{Example}
\newtheorem{remark}[mythmcounter]{Remark}
\newtheorem{definition}[mydefcounter]{Definition}
\newtheorem{theorem}[mythmcounter]{Theorem}
\newtheorem{lemma}[mythmcounter]{Lemma}
\newtheorem{proposition}[mythmcounter]{Proposition}
\newtheorem*{lemma*}{Lemma}

% Formatting for source code & types
\definecolor{cmtclr}{rgb}{0.0,0.6,0.0}
\definecolor{kvdclr}{rgb}{0.0,0.0,0.6}
\definecolor{strclr}{rgb}{0.5,0.1,0.0}
\definecolor{prepclr}{rgb}{0.0,0.0,0.0}

\newcommand{\sep}[0]{\; | \;}
\newcommand{\kvd}[1]{\textnormal{\textcolor{kvdclr}{\ttfamily #1}}}
\newcommand{\str}[1]{\textnormal{\textcolor{strclr}{\ttfamily "#1"}}}
\newcommand{\prepk}[1]{\textnormal{\textcolor{prepclr}{\bfseries\sffamily #1}}}
\newcommand{\prepi}[1]{\textnormal{\textcolor{prepclr}{\ttfamily #1}}}
\newcommand{\ident}[1]{\textnormal{\sffamily #1}}
\newcommand{\cmt}[1]{\textit{\sffamily\textcolor{cmtclr}{#1}}}

\newcommand{\ctyp}[2]{C^{#1}#2}
\newcommand{\mtyp}[2]{M^{#1}#2}
\newcommand{\subst}[3]{#1[#2 \leftarrow #3]}

% Typing rule and typing statement (single line)
\newcommand{\tyrule}[3]{ \inference[\footnotesize{(\emph{#1})}~~]{#2}{#3} }
\newcommand{\tyruler}[3]{ \inference{#2}{#3}~~\footnotesize{(\emph{#1})} }

% Coeffect algebra
\definecolor{aclr}{rgb}{0.6,0.3,0.0}
\newcommand{\aclrd}[1]{\textcolor{aclr}{#1}}

% Coeffect algebra
\definecolor{cclr}{rgb}{0.0,0.5,0.0}
\newcommand{\cclrd}[1]{\textcolor{cclr}{#1}}

\newcommand{\czero}{ \textcolor{cclr}{ \mathsf{ign} } }
\newcommand{\cunit}{ \textcolor{cclr}{ \mathsf{use} } }
\newcommand{\cseq}{ \textcolor{cclr}{ \circledast }}
\newcommand{\cpar}{ \textcolor{cclr}{\oplus} }
\newcommand{\czip}{ \textcolor{cclr}{\wedge} } % only used by flat coeffects
\newcommand{\cleq}{ \textcolor{cclr}{\leq} }
\newcommand{\cgeq}{ \textcolor{cclr}{\geq} }

\newcommand{\atimes}{ \textcolor{aclr}{\times} }
\newcommand{\aseq}{ \textcolor{aclr}{\circledast}} % Should be the same as \cseq
\newcommand{\aparstr}{ \textcolor{aclr}{\oplus}} % Should be the same as \cseq

% Unified

\newcommand{\SH}{ {\textcolor{sclr}{S}} }
\newcommand{\SP}{ {\textcolor{sclr}{P}} }
\newcommand{\SHP}{ {\textcolor{sclr}{S\triangleleft P}} }
\newcommand{\sempty}{ {\textcolor{sclr}{ \hat{0}}} }
\newcommand{\sunit}{ {\textcolor{sclr}{\hat{1}}} }
\newcommand{\stimes}{ {\textcolor{sclr}{\diamond}} }
\newcommand{\sflat}{\star}

\newcommand{\azero}{ \textcolor{aclr}{\bot} }
\newcommand{\aunit}{ \alift{\cunit} }
\newcommand{\aweak}{ \alift{\czero} }

\newcommand{\azip}{ \textcolor{aclr}{ \mathbin{\rotatebox[origin=c]{-90}{$\ltimes$}} } }
\newcommand{\apar}{ \textcolor{aclr}{ \mathbin{\rotatebox[origin=c]{90}{$\ltimes$}} } }

\newcommand{\xtimes}{\hat{\times}}

\newcommand{\ctx}{\textit{ctx}}

% Ordinary context where colouring is done by hand
\newcommand{\coctx}[2]{ #1 \,\text{\scriptsize @}\, {{\text{$#2$}}}  }

% Simple coeffect systems in the introduction
\newcommand{\cons}{\hspace{-0.15em}\times\hspace{-0.25em}} %\sctxclrd{+\hspace{-0.55em}+}}

\newcommand{\alift}[1]{ \textcolor{aclr}{\langle} #1 \textcolor{aclr}{\rangle}}

\definecolor{sclr}{rgb}{0.2,0,0.8}
\newcommand{\sclrd}[1]{ \textcolor{sclr}{#1} }

\newcommand{\C}{ {\textcolor{cclr}{\mathcal{C}}} }
\newcommand{\M}{ {\textcolor{cclr}{\mathcal{M}}} }

\newcommand{\alen}[1]{{\aclrd{\textit{len}(}#1\aclrd{)}}}
\newcommand{\slen}[1]{{\sclrd{\textit{len}(}#1\sclrd{)}}}
\newcommand{\xlen}[1]{{\textit{len}(#1)}}

\newcommand{\reduce}{\longrightarrow}

\newcommand{\restr}[2]{#1|_{#2}}
\newcommand{\narrow}[1]{\hspace{-0.5em} #1 \hspace{-0.65em}~}

\definecolor{todoclr}{rgb}{0.8,0,0.5}
\newcommand{\todo}[1]{\textcolor{todoclr}{ \begin{quotation} \noindent\textbf{TODO:} #1 \end{quotation} }}

\newcommand{\sem}[1]{\llbracket #1 \rrbracket}
\newcommand{\catc}{\mathcal{C}}
\newcommand{\idf}[1]{ {\textnormal{\sffamily id}}_{#1} }

\newcommand{\figcaption}[1]{
  \vspace{0.5em}
  \noindent\makebox[\linewidth]{\rule{\textwidth}{0.5pt}}
  \vspace{-2em}
  \caption{#1}
}
% Half line spacing around align
\newcommand{\zerodisplayskips}{%
  \setlength{\abovedisplayskip}{0.5em}
  \setlength{\belowdisplayskip}{0.5em}
  \setlength{\abovedisplayshortskip}{0.5em}
  \setlength{\belowdisplayshortskip}{0.5em}}
\appto{\normalsize}{\zerodisplayskips}
\appto{\small}{\zerodisplayskips}
\appto{\footnotesize}{\zerodisplayskips}

% half line spacing around compactitem
\setlength{\pltopsep}{0.5em}
\setlength{\plitemsep}{0.3em}

\patchcmd{\thebibliography}{\chapter*}{\section*}{}{}

\begin{document}

\frenchspacing % Reduces space after periods to make text more compact
\raggedbottom % Makes all pages the height of the text on that page
\selectlanguage{american} % Select your default language - e.g. american or ngerman
\pagenumbering{roman} % Roman page numbering prior to the start of the thesis content (i, ii, iii, etc)
\pagestyle{plain} % Suppress headers for the pre-content pages

% ==================================================================================================

\chapter*{Proposal for thesis corrections}
\renewcommand*\thesection{\arabic{section}}

This document addresses some of the concerns discussed in the joint report on a viva of the thesis
``Context-aware programming languages'' and to proposes an outline for corrections that would
resolve the issues discussed in the report.

The main concern presented in the report is related to the use of ``coeffect-directed semantics''
in the thesis. We start by discussing additional background that provides a better perspective on
this approach as well as why it was employed in the thesis (Section~\ref{sec:background}). Next,
we turn our attention to the issue of \emph{implicit parameters} (Section~\ref{sec:ambiguity}) --
this problem (one of the three problems used throughout the thesis) demonstrates the ambiguity
of type checking that arises in \emph{flat coeffect calculus} and relates it with prior work in
languages such as Haskell. Finally, the last section (Section~\ref{sec:proposal}) outlines four
concrete steps that address the reviewer's concerns.

\section{Background}
\label{sec:background}

Perhaps the most important objection raised in the report is that the semantics presented throughout
the thesis is given based on a concrete typing derivation. In other words, the meaning of a term
depends on the concrete typing assigned by a compiler\footnote{The thesis does not clearly and
explicitly state this, leading to potential confusion -- this can be clarified, both in the
text and typographically (by writing the semantics over a derivation rule, rather than over
a judgement) to avoid confusion.} In this section, we discuss additional programming language
background\footnote{Which will be added in the final version of the thesis.} that provides
additional grounds for specifying the meaning of programs in this way.

\subsection{Type-directed semantics in practice}
\label{sec:background-practice}

One of the aims of this thesis is to model how real-world functional-first programming languages
could be extended with the support for context-aware programming and so the work is done in
the context of languages such as F\# \cite{app-fsharp-spec} and Scala \cite{app-scala-spec}.

In theoretically-focused programming languages, it is possible to define the meaning of a term
without considering the typing, but there are many examples where this is not the case in languages
such as F\#:
%
\begin{equation*}
\begin{array}{l}
  \kvd{let}~\ident{nextTwo}~\ident{rnd}~=~(\ident{rnd.Next}(), \ident{rnd.Next}()) \\
  \kvd{let}~\ident{add}~\ident{a}~\ident{b}~=~\ident{a}+\ident{b}
\end{array}
\end{equation*}
%
In the first example, the meaning of $\ident{rnd.Next}()$ depends on the type of \ident{rnd}.
This can be any .NET class with the \ident{Next} method, but the F\# compiler can only compile
the code if it knows the concrete type. Thus, the meaning depends the typing derivation.

In the second example, the meaning of $+$ depends on the types of the operands (it can be an
addition or string concatenation). The F\# compiler can compile the function
without additional information, because the type inference has a defaulting mechanism (using
\ident{int}), but if the program contains e.g.~$\ident{add}~\str{hi}~\str{there}$, the compiler
will use \ident{string}.

As demonstrated in this section, it is not uncommon in practical programming languages that the
meaning of a term depends on its typing derivation. In F\#, this is addressed in two ways. In case
of \ident{nextTwo}, the compiler rejects the program if it does not know the type of \ident{rnd}
(an explicit type annotation would be required to disambiguate). In case of \ident{add}, the
compiler uses a defaulting mechanism to pick one of the avialable typing derivations. We return to
these two options when discussing the resolution of ambiguities in \emph{coeffect calculi} in
Section~\ref{sec:ambiguity}.

\subsection{Type-directed semantics in theory}

In a number of places in the thesis, we give the semantics of based on typing derivation (both for
the concrete languages in Chapter~3 and for the coeffect calculi in Chapters~4 and 5). The notation
used in the thesis is (using simple let-binding as an example):
%
\begin{equation*}
\begin{array}{l}
  \sem{x:\tau \vdash \kvd{let}~x_1=e_2~\kvd{in}~e_2 : \tau_2} = \\
  \quad \sem{x_1:\tau_1 \vdash e_2 : \tau_2} \circ \sem{x:\tau \vdash e_1 : \tau_1}
\end{array}
\end{equation*}
%
This notation is somewhat ambiguous as it suggests that we are defining a function from a typing
judgement for the expression on the left-hand side. However, this should be read as giving an
interpretation of the term based on the typing derivation.

It is worth noting that giving an interpretation of a term based on its typing derivation is not
an unusual construction used only in the thesis. In fact, Moggi in his seminal paper on monads uses
the same approach and writes \cite[p7]{monad-notions} (emphasis mine):

\begin{quotation}
\noindent
An interpretation $\sem{-}$ of the language in a category $C$ is parametric in an interpretation
of the symbols in the signature and is defined by \emph{induction on the derivation} of
well-formedness for (types,) terms and equations.
\end{quotation}

\noindent
The notation used by Moggi is unambiguous (and we intend to adopt the same clear notation in the
revised version of the thesis). Using the same example as above:
%
\begin{equation*}
  \begin{array}{r}
    x:\tau \vdash e_1 : \tau_1 \\
    x_1:\tau_1 \vdash e_2 : \tau_2 \\ \hline
    x:\tau \vdash \kvd{let}~x_1=e_2~\kvd{in}~e_2 : \tau_2
  \end{array}
  \begin{array}{l}
    = g_1 \\
    = g_2 \\
    = g_2 \circ g_1
  \end{array}
\end{equation*}
%
In systems that have unique typing derivation (such as the language of Moggi), the fact that the
semantics is defined by induction over a typing derivation means merely that we only define
semantics for well-typed prorgams (as there is no other typing derivation). However, there are two
possible directions that can be taken from this simple starting point.

We can extend the type system with non-syntax-driven rules (for example, with sub-effecting or
explicit structural rules like exchange, weakening and contraction). As a result, the language
loses uniqueness of typing, but the semantics of terms is equivalent for all typing derivations.
This is the \emph{coherence property} that has been proved for certain systems
\cite{semantics-bltres}.

However, when we consider programming languages where the meaning of a term \emph{depends} on a
typing derivation (as discussed in Section~\ref{sec:background-practice}), we can no longer
prove the coherence property -- because the semantics of a term is different for different
typing derivations. As further discussed in Section~\ref{sec:ambiguity}, this is exactly the case
of the flat coeffect calculus for implicit parameters presented in the thesis. Before discussing
implicit parameters, the following subsection provides a more general argument for why this is the
correct approach for context-aware computations.

\subsection{Why type-directed semantics for coeffects?}

One aim of the thesis is to make explicit some of the context-dependencies that are often implicit
in today's programming languages. For example, a language today might provide a construct for
accessing a resource provided by the environment such as $\kvd{access}~\ident{Date}$ (typically,
this is just a library call). Operationally, the meaning of the meaning of the term depends on
the context -- if \ident{Date} is available on the client-side, it is accessed from there; if
\ident{Date} is available on the server-side, it is accessed from there instead and if \ident{Date}
is not available, the program raises an exception.

We aim to make the programming language aware of these contextual dependencies through the use
of type-and-coeffect systems. Our calculi lift some of the operational properties into the type
system as coeffects. This makes the existing context-dependencies more visible -- and as a
consequence, the semantics can only be defined for a concrete typing derivation. An alternative
approach is to make the language \emph{more explicit}, but we do not attempt to do this and
instead follow an approach similar to F\#\footnote{This demarcation for the work
in the thesis is not clearly discussed in the text and will be clarified.} as outlined in
Section~\ref{sec:background-practice}.

\section{Resolving ambiguity}
\label{sec:ambiguity}

The ambiguity in typing derivation that is ``a problem'' in the thesis is best demonstrated
using implicit parameters. Here, we look at how Haskell resolves this (and how the same
problem appears elsewhere).

The presented thesis does not attempt to resolve the ambiguity and so it provides a more
interesting insight. We discuss why this is the case.

\subsection{Haskell implicit parameters}

If we try to understand implicit parameters in a more general way, then we can see that the meaning
indeed depends on the typing derivation. (Haskell resolves this by preferring one specific typing derivation.)

Haskell implicit parameters

\subsection{Exceptions}
Exceptions

\subsection{Counter example}

\subsection{Summary}

Haskell does one thing, exceptions do (strictly) another thing.
(Maybe for good reasons, but we should view this more generally)

\section{Proposal}
\label{sec:proposal}

Coeffect-directed meaning makes sense.

The thesis does not use category theory to give a semantics in the traditional sense with
traiditional properties (and this needs to be said clearly)

\subsection{Translation}

The "coeffect-directed semantics" should not be seen as a "semantics of a term", but it should
instead be seen as a translation that - given a typing derivation - produces a new term that
unambiguously defines the meaning. The chapter 3 does so using direct denotational style; the
chapters 4 and 5 show that we can abstract common structures used in these definitions.

Doing so also explains why our type system (about which we also prove syntactic properties) is
appropriately constructed (the structure of the typing follows from the structure of this
coeffect-directed semantics - and BTW this was the hard thing to figure out when working on the thesis!)

As a result, we can see the "coeffect-directed semantics" more as a translation from a type-and-coeffect
derivation to an unambiguous term language (which I'm happy to define more explicitly and precisely
in a syntactic style, as opposed to categorical style).

\subsection{Unique typing derivation}

>> 1. The problem occurs because the type derivation in all the proposed
>> systems is not unique, due to features such as sub-typing

After doing the changes outlined above, this still remains a valid point. In Haskell's implicit
parameters, this is resolved by choosing one canonical typing derivation (I'll give an example how
this works). The same approach can be used to correct this in the thesis - that is, I would like
to give an algorithmic formulation of the type system that produces a canonical typing derivation.
(There are three sources of the problem - subtyping, lambda abstraction and structural system; I
intend to make subtyping an explicit term, add externally provided function for resolving ambiguities
in lambda abstraction and have algorithmic typing rules for the structural systems).

\subsection{Lack of equational structure}

>> 3. The co-monadic structures proposed in the categorical semantics
>> lack an equational structure

This is correct - and it follows from the fact that the categorical structures are used to motivate
the design of the type system, rather than - which is more common - to provide a reasoning
framework about coeffectful programs.

\subsection{Generalization}

>> 4. It is unclear as to whether the concrete denotational semantics
>> are instances of the categorical framework.

This is a valid point and I fully accept it. I intend to add sections showing that this is, indeed, the case.

% ==================================================================================================

\section{Summary}

In practice, people use monads as common structure - not for proving program properties.
This thesis aims to find a similar common structure in context-aware programming.

This is done in two ways pursued in parallel - by finding a type system with desirable syntactic
properties and by finding a way to structure the translational semantics using a common structure.

Finding common structure.

yo

\bibliographystyle{abbrv}
\bibliography{Bibliography}

\end{document}