### 2016-03-06 main.tex
Move stuff in chapter 4, improve introductions

%%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[
		twoside,openright,titlepage,numbers=noenddot,headinclude,%1headlines,
                footinclude=true,cleardoublepage=empty,
                BCOR=10mm,paper=a4,fontsize=10pt, % Binding correction, paper type and font size
                ngerman,american, % Languages
                ]{scrreprt}

% Includes the file which contains all the document configurations and packages - make sure to edit this file
\input{classicthesis-config}
\usepackage{semantic}
\usepackage{paralist}
\usepackage{etoolbox}
\usepackage{stmaryrd}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm}
%\usepackage{catchfilebetweentags}

\newcounter{mythmcounter}
\newcounter{mydefcounter}
\newcounter{myexpcounter}
\newtheorem{example}[myexpcounter]{Example}
\newtheorem{remark}[mythmcounter]{Remark}
\newtheorem{definition}[mydefcounter]{Definition}
\newtheorem{theorem}[mythmcounter]{Theorem}
\newtheorem{lemma}[mythmcounter]{Lemma}
\newtheorem{proposition}[mythmcounter]{Proposition}
\newtheorem*{lemma*}{Lemma}

% Formatting for source code & types
\definecolor{cmtclr}{rgb}{0.0,0.6,0.0}
\definecolor{kvdclr}{rgb}{0.0,0.0,0.6}
\definecolor{strclr}{rgb}{0.5,0.1,0.0}
\definecolor{prepclr}{rgb}{0.0,0.0,0.0}

\newcommand{\sep}[0]{\; | \;}
\newcommand{\kvd}[1]{\textnormal{\textcolor{kvdclr}{\ttfamily #1}}}
\newcommand{\str}[1]{\textnormal{\textcolor{strclr}{\ttfamily "#1"}}}
\newcommand{\prepk}[1]{\textnormal{\textcolor{prepclr}{\bfseries\sffamily #1}}}
\newcommand{\prepi}[1]{\textnormal{\textcolor{prepclr}{\ttfamily #1}}}
\newcommand{\ident}[1]{\textnormal{\sffamily #1}}
\newcommand{\cmt}[1]{\textit{\sffamily\textcolor{cmtclr}{#1}}}

\newcommand{\ctyp}[2]{C^{#1}#2}
\newcommand{\mtyp}[2]{M^{#1}#2}
\newcommand{\subst}[3]{#1[#2 \leftarrow #3]}

% Typing rule and typing statement (single line)
\newcommand{\tyrule}[3]{ \inference[\footnotesize{(\emph{#1})}~~]{#2}{#3} }
\newcommand{\tyruler}[3]{ \inference{#2}{#3}~~\footnotesize{(\emph{#1})} }

% Coeffect algebra
\definecolor{aclr}{rgb}{0.6,0.3,0.0}
\newcommand{\aclrd}[1]{\textcolor{aclr}{#1}}

% Coeffect algebra
\definecolor{cclr}{rgb}{0.0,0.5,0.0}
\newcommand{\cclrd}[1]{\textcolor{cclr}{#1}}

% Red color
\definecolor{rclr}{rgb}{0.75,0.0,0.0}
\newcommand{\rclrd}[1]{\textcolor{rclr}{#1}}

\newcommand{\czero}{ \textcolor{cclr}{ \mathsf{ign} } }
\newcommand{\cunit}{ \textcolor{cclr}{ \mathsf{use} } }
\newcommand{\cseq}{ \textcolor{cclr}{ \circledast }} 
\newcommand{\cpar}{ \textcolor{cclr}{\oplus} }
\newcommand{\czip}{ \textcolor{cclr}{\wedge} } % only used by flat coeffects
\newcommand{\cleq}{ \textcolor{cclr}{\leq} }
\newcommand{\cgeq}{ \textcolor{cclr}{\geq} }

\newcommand{\atimes}{ \textcolor{aclr}{\times} }
\newcommand{\aseq}{ \textcolor{aclr}{\circledast}} % Should be the same as \cseq 
\newcommand{\aparstr}{ \textcolor{aclr}{\oplus}} % Should be the same as \cseq

% Unified

\newcommand{\SH}{ {\textcolor{sclr}{S}} }
\newcommand{\SP}{ {\textcolor{sclr}{P}} }
\newcommand{\SHP}{ {\textcolor{sclr}{S\triangleleft P}} }
\newcommand{\sempty}{ {\textcolor{sclr}{ \hat{0}}} }
\newcommand{\sunit}{ {\textcolor{sclr}{\hat{1}}} }
\newcommand{\stimes}{ {\textcolor{sclr}{\diamond}} }
\newcommand{\sflat}{\star}

\newcommand{\azero}{ \textcolor{aclr}{\bot} }
\newcommand{\aunit}{ \alift{\cunit} }
\newcommand{\aweak}{ \alift{\czero} }

\newcommand{\azip}{ \textcolor{aclr}{ \mathbin{\rotatebox[origin=c]{-90}{$\ltimes$}} } }
\newcommand{\apar}{ \textcolor{aclr}{ \mathbin{\rotatebox[origin=c]{90}{$\ltimes$}} } }

\newcommand{\xtimes}{\hat{\times}}

\newcommand{\ctx}{\textit{ctx}}

% Ordinary context where colouring is done by hand 
\newcommand{\coctx}[2]{ #1 \,\text{\scriptsize @}\, {{\text{$#2$}}}  }

% Simple coeffect systems in the introduction
\newcommand{\cons}{\hspace{-0.15em}\times\hspace{-0.25em}} %\sctxclrd{+\hspace{-0.55em}+}}

\newcommand{\alift}[1]{ \textcolor{aclr}{\langle} #1 \textcolor{aclr}{\rangle}}

\definecolor{sclr}{rgb}{0.2,0,0.8}
\newcommand{\sclrd}[1]{ \textcolor{sclr}{#1} }

\newcommand{\C}{ {\textcolor{cclr}{\mathcal{C}}} }
\newcommand{\M}{ {\textcolor{cclr}{\mathcal{M}}} }

\newcommand{\alen}[1]{{\aclrd{\textit{len}(}#1\aclrd{)}}}
\newcommand{\slen}[1]{{\sclrd{\textit{len}(}#1\sclrd{)}}}
\newcommand{\xlen}[1]{{\textit{len}(#1)}}

\newcommand{\reduce}{\longrightarrow}

\newcommand{\restr}[2]{#1|_{#2}}
\newcommand{\narrow}[1]{\hspace{-0.5em} #1 \hspace{-0.65em}~}

\definecolor{todoclr}{rgb}{0.8,0,0.5}
\newcommand{\todo}[1]{\textcolor{todoclr}{ \begin{quotation} \noindent\textbf{TODO:} #1 \end{quotation} }}

\newcommand{\sem}[1]{\llbracket #1 \rrbracket}
\newcommand{\catc}{\mathcal{C}}
\newcommand{\idf}[1]{ {\textnormal{\sffamily id}}_{#1} }

\newcommand{\figcaption}[1]{
  \vspace{0.5em}
  \noindent\makebox[\linewidth]{\rule{\textwidth}{0.5pt}} 
  \vspace{-2em}
  \caption{#1}
}

% For definining semantics of things over a typing derivation
\newcommand{\semdeff}[4]{
\begin{array}{rcl}
  #1 & = & #3 \\ \cline{1-1}\cline{3-3}
  #2 & = & #4 \\
\end{array}  
}
\newcommand{\semdefff}[6]{
\begin{array}{rcl}
  #1 & = & #4 \\
  #2 & = & #5 \\ \cline{1-1}\cline{3-3}
  #3 & = & #6 \\
\end{array}  
}
\newcommand{\semdef}[2]{
\begin{array}{rcl}
  ~ & ~ & ~ \\ \cline{1-1}\cline{3-3}
  #1 & = & #2 \\
\end{array}  
}

% Half line spacing around align
\newcommand{\zerodisplayskips}{%
  \setlength{\abovedisplayskip}{0.5em}
  \setlength{\belowdisplayskip}{0.5em}
  \setlength{\abovedisplayshortskip}{0.5em}
  \setlength{\belowdisplayshortskip}{0.5em}}
\appto{\normalsize}{\zerodisplayskips}
\appto{\small}{\zerodisplayskips}
\appto{\footnotesize}{\zerodisplayskips}

% half line spacing around compactitem
\setlength{\pltopsep}{0.5em}
\setlength{\plitemsep}{0.3em}

\begin{document}

\frenchspacing % Reduces space after periods to make text more compact

\raggedbottom % Makes all pages the height of the text on that page

\selectlanguage{american} % Select your default language - e.g. american or ngerman

%\renewcommand*{\bibname}{new name} % Uncomment to change the name of the bibliography
%\setbibpreamble{} % Uncomment to include a preamble to the bibliography - some text before the reference list starts

\pagenumbering{roman} % Roman page numbering prior to the start of the thesis content (i, ii, iii, etc)

\pagestyle{plain} % Suppress headers for the pre-content pages

%----------------------------------------------------------------------------------------
%	PRE-CONTENT THESIS PAGES
%----------------------------------------------------------------------------------------

%\include{text/Titlepage} % Main title page

%\cleardoublepage\include{text/Declaration} % Publications from the thesis page

%\cleardoublepage\include{text/Abstract} % Abstract page

\pagestyle{scrheadings} % Show chapter titles as headings

\cleardoublepage\include{text/Contents} % Contents, list of figures/tables/listings and acronyms

\pagenumbering{arabic} % Arabic page numbering for thesis content (1, 2, 3, etc)
%\setcounter{page}{90} % Uncomment to manually start the page counter at an arbitrary value (for example if you wish to count the pre-content pages in the page count)

\cleardoublepage % Avoids problems with pdfbookmark

%----------------------------------------------------------------------------------------
%	THESIS CONTENT - CHAPTERS
%----------------------------------------------------------------------------------------

% \ctparttext{You can put some informational part preamble text here. Illo principalmente su nos. Non message \emph{occidental} angloromanic da. Debitas effortio simplificate sia se, auxiliar summarios da que, se avantiate publicationes via. Pan in terra summarios, capital interlingua se que. Al via multo esser specimen, campo responder que da. Le usate medical addresses pro, europa origine sanctificate nos se.} % Text on the Part 1 page describing  the content in Part 1

% \part{Some Kind of Manual} % First part of the thesis

% \include{Chapters/Chapter01} % Chapter 1

% \cleardoublepage % Empty page before the start of the next part

%------------------------------------------------

% \ctparttext{You can put some informational part preamble text here. Illo principalmente su nos. Non message \emph{occidental} angloromanic da. Debitas effortio simplificate sia se, auxiliar summarios da que, se avantiate publicationes via. Pan in terra summarios, capital interlingua se que. Al via multo esser specimen, campo responder que da. Le usate medical addresses pro, europa origine sanctificate nos se.} % Text on the Part 2 page describing the content in Part 2

% \part{The Showcase} % Second part of the thesis

%\include{text/Chapter01}
%\include{text/Chapter02}
%\include{text/Chapter03}

% ==================================================================================================

\ctparttext{In this part, we capture the similarities between the concrete context-aware langauges
  presented in the previous chapter. We also develop the key novel technical contributions of
  the thesis. We define a \emph{flat coeffect type system} (Chapter~\ref{ch:flat}) that is 
  parameterized by a \emph{coeffect algebra} and a mechanism for choosing unique typing derivation. 
  We instantiate a coeffect type system with a concrete coeffect algebra and procedure for choosing 
  unique typing derivation for three languages to capture dataflow, implicit parameters and liveness.

  The type system is complemented with a translational semantics for coeffect-based context-aware
  programming languages (Chapter~\ref{ch:semantics}). The semantics is inspired by a categorical
  model based on \emph{indexed comonads} and it translates source context-aware program into a 
  target program in a simple functional language with comonadically-inspired primitives. We give 
  concrete definition of the primitives for dataflow, implicit parameters and liveness and 
  present a syntactic safety proof for these three languages.

  The following page provides a detailed overview of the content of Chapters~\ref{ch:flat} and
  Chapters~\ref{ch:semantics}, highlighting the split between general definitions and properties
  (about the coeffect calculus) and concrete definitions and properties (about concrete
  context-awre language). The Chapter~\ref{ch:structural} mirrors the same development for 
  \emph{structural coeffect systems}.
}
\setcounter{part}{1}
\part{Coeffect calculi}

\begin{center}
\begin{tabular}{ | l | l | l |}
\hline
\multicolumn{3}{ | c | }{\spacedlowsmallcaps{Chapter 4}} \\ \hline \hline
\hspace{7.1em} & \textsc{Coeffect calculus} \hspace{2.5em} & \textsc{Language-specific} \hspace{6.9em} \\ \hline
\textsc{Syntax} 
  & \hspace{-0.5em}\begin{tabular}{l} Coeffect $\lambda$-calculus \\[-0.3em] (Section~\ref{sec:flat-calculus-syntax}) \end{tabular}
  & \hspace{-0.5em}\begin{tabular}{l} Extensions such as \\[-0.3em] \ident{?param} and \kvd{prev} \\[-0.3em] (Section~\ref{sec:flat-calculus-examples})  \end{tabular} \\ \hline
\textsc{Type system} 
  & \hspace{-0.5em}\begin{tabular}{l} Abstract coeffect \\[-0.3em] algebra (Section~\ref{sec:flat-calculus-algebra})  \end{tabular}
  & \hspace{-0.5em}\begin{tabular}{l} Concrete instances \\[-0.3em] of the coeffect algebra \\[-0.3em] (Section~\ref{sec:flat-calculus-examples})  \end{tabular} \\ 
\hline
  & \hspace{-0.5em}\begin{tabular}{l} Coeffect type system \\[-0.3em] parameterized by \\[-0.3em] the coeffect algebra \\[-0.3em] (Section~\ref{sec:flat-calculus-types})  \end{tabular}
  & \hspace{-0.5em}\begin{tabular}{l} Typing for language- \\[-0.3em] specific extensions \\[-0.3em] (Section~\ref{sec:flat-calculus-examples})  \end{tabular} \\ 
\hline
  & 
  & \hspace{-0.5em}\begin{tabular}{l} Procedure for determining \\[-0.3em] a unique typing derivation \\[-0.3em]  (Section~\ref{sec:flat-unique})  \end{tabular} \\ \hline
\textsc{Properties} 
  & \hspace{-0.5em}\begin{tabular}{l} Syntactic properties \\[-0.3em] of coeffect $\lambda$-calculus \\[-0.3em] (Section~\ref{sec:flat-syntax})  \end{tabular}
  & \hspace{-0.5em}\begin{tabular}{l} Uniqueness of the above \\[-0.3em]  (Section~\ref{sec:flat-unique})  \end{tabular} \\ \hline
\end{tabular}
\end{center}

~

\begin{center}
\begin{tabular}{ | l | l | l |}
\hline
\multicolumn{3}{ | c | }{\spacedlowsmallcaps{Chapter 5}} \\ \hline \hline
& \textsc{Coeffect calculus} & \textsc{Language-specific} \\ \hline
\textsc{Categorical} 
  & \hspace{-0.5em}\begin{tabular}{l} Indexed comonads \\[-0.3em] (Sectiuon~\ref{sec:semantics-flat-idx}) \end{tabular}
  & \hspace{-0.5em}\begin{tabular}{l} Examples including indexed \\[-0.3em] product, list and maybe \\[-0.3em] comonads (Section~\ref{sec:semantics-flat-monoidal}) \end{tabular} \\ 
\hline
  & \hspace{-0.5em}\begin{tabular}{l} Categorical semantics \\[-0.4em] of coeffect $\lambda$-calculus \\[-0.3em] (Section~\ref{sec:semantics-flat-calculus}) \end{tabular}
  & \\ \hline
\textsc{Translational} 
  & \hspace{-0.5em}\begin{tabular}{l} Functional target \\[-0.3em] language (Section~\ref{sec:semantics-translation-target}) \end{tabular}
  &  \\ 
\hline
  & \hspace{-0.5em}\begin{tabular}{l} Translation from coeffect \\[-0.3em] $\lambda$-calculus to target \\[-0.3em]  language (Section~\ref{sec:semantics-translation-transl}) \end{tabular}
  & \hspace{-0.5em}\begin{tabular}{l} Translation for language-\\[-0.3em] specific extensions (\kvd{prev}, \ident{?p}) \\[-0.3em] (Sections~\ref{sec:semantics-proofs-df} and \ref{sec:semantics-proofs-impl}) \end{tabular} \\ \hline
\textsc{Operational} 
  & \hspace{-0.5em}\begin{tabular}{l} Abstract comonadically- \\[-0.3em] inspired primitives \\[-0.3em] (Section~\ref{sec:semantics-translation-transl}) \end{tabular}
  & \hspace{-0.5em}\begin{tabular}{l} Concrete reduction rules for \\[-0.3em] comonadically-inspired primitives \\[-0.3em] (Sections~\ref{sec:semantics-proofs-df} and \ref{sec:semantics-proofs-impl}) \end{tabular} \\
\hline
  & 
  & \hspace{-0.5em}\begin{tabular}{l} Reduction rules for language-\\[-0.3em] specific extensions (\kvd{prev}, \ident{?p}) \\[-0.3em] (Sections~\ref{sec:semantics-proofs-df} and \ref{sec:semantics-proofs-impl}) \end{tabular} \\ 
\hline
  & \hspace{-0.5em}\begin{tabular}{l} Sketch of generalized \\[-0.3em] syntactic soundness \\[-0.3em] (Section~\ref{sec:semantics-generalising}) \end{tabular}
  & \hspace{-0.5em}\begin{tabular}{l} Syntactic soundness \\[-0.3em] (Sections~\ref{sec:semantics-proofs-df} and \ref{sec:semantics-proofs-impl}) \end{tabular} \\ \hline
\end{tabular}
\end{center}

% ==================================================================================================

\include{text/Chapter04}
\include{text/Chapter05}
%\include{text/Chapter06}
%\include{text/Chapter07}
%\include{text/Chapter08}

%----------------------------------------------------------------------------------------
%	THESIS CONTENT - APPENDICES
%----------------------------------------------------------------------------------------

% \part{Appendix} % New part of the thesis for the appendix

% \include{Chapters/Chapter0B} % Appendix B - empty template

%----------------------------------------------------------------------------------------
%	POST-CONTENT THESIS PAGES
%----------------------------------------------------------------------------------------

\cleardoublepage\include{Bibliography} % Bibliography
% \cleardoublepage\include{FrontBackMatter/Colophon} % Colophon
% \cleardoublepage\include{FrontBackMatter/Declaration} % Declaration

%----------------------------------------------------------------------------------------

%\appendix
%\include{text/AppendixA}
%\include{text/AppendixB}

\end{document}