### 2016-03-06 text/Chapter05.tex

% ==================================================================================================

\chapter{Semantics of flat coeffects}
\label{ch:semantics}

The \emph{flat coeffect calculus} introduced in the previous chapter uniformly captures a number
of context-aware systems discussed in Chapter~\ref{ch:applications}. The coeffect calculus 
can be seen as a \emph{language framework} that simplifies the construction of concrete 
\emph{domain-specific} coeffect lnguages. In the previous chapter, we discussed how it provides a  
type system that tracks the required context. In this chapter, we show that the langauge framework 
also provides a way for defining the semantics of concrete domain-specific coeffect languages,  
guides their implementation and simplifies safety proofs.

This is done using a \emph{comonadically-inspired translation}. We translate a program written 
using the coeffect calculus into a simple functional language with additional coeffect-specific 
comonadically-inspired primitives that implement the concrete notion of context-awareness.

We use comonads in a syntactic way, following the example of Haskell's use of monads. The 
translation is the same for all coeffect languages, but the safety depends on the concrete
coeffect-specific comonadically-inspired primitives. We prove the soundness of two concrete
coeffect calculi (dataflow and implicit parameters). We note that the proof crucially relies on a 
relationship between coeffect annotations (provided by the type system) and the comonadically-inspired
primitives (defining the semantics), which makes it easy to extend it to other concrete 
context-aware langauges.

\paragraph{Chapter structure and contributions}

\begin{itemize}
\item We introduce \emph{indexed comonads}, a generalization of comonads, a category-theoretical 
  dual of monads (Section~\ref{sec:semantics-theory}) and we discuss how they provide semantics
  for coeffect calculus. This provides an insight into how (and why) the calculus works and shows 
  an intriguing link with effects and monads.

\item We use indexed comonads to guide our \emph{translational semantics} of coeffect calculus
  (Section~\ref{sec:semantics-translation}). We define a simple functional programming language
  (with type system and operational semantics) and prove that well-typed programs in the language
  do not get stuck. We extend it with uninterpreted comonadically-inspired primitives and
  define a translation that turns well-typed context-aware coeffect programs into programs of 
  our functional language.

\item For two of the coeffect calculi discussed earlier (dataflow and implicit parameters),
  we give reduction rules for the comonadically-inspired primitives and we extend the progress
  and preservation proofs, showing that well-typed programs produced by translation from two
  coeffect languages do not get stuck (Section~\ref{sec:semantics-proofs})

\item We leave the details to future work, but we note that the proof for concrete coeffect language 
  (dataflow and implicit parameters) can be generalized -- rather than reconsidering progress and 
  preservation of the whole target language, we rely just on the correctness of the coeffect-specific 
  comonadically-inspired primitives and abstraction mechanism provided by languages 
  such as ML and Haskell (Section~\ref{sec:semantics-related}).
\end{itemize}

% ==================================================================================================
%
%    ###
%     #  #    # ##### #####   ####  #####  #    #  ####  ##### #  ####  #    #
%     #  ##   #   #   #    # #    # #    # #    # #    #   #   # #    # ##   #
%     #  # #  #   #   #    # #    # #    # #    # #        #   # #    # # #  #
%     #  #  # #   #   #####  #    # #    # #    # #        #   # #    # #  # #
%     #  #   ##   #   #   #  #    # #    # #    # #    #   #   # #    # #   ##
%    ### #    #   #   #    #  ####  #####   ####   ####    #   #  ####  #    #
%
% ==================================================================================================

\section{Introduction and safety}
\label{sec:semantics-intro}

This chapter links together all the different technical developments presented in this thesis.
We take the flat coeffect calculus introduced in Chatper~\ref{ch:flat}, define its \emph{abstract 
comonadic semantics} and use it to define a translation that gives a \emph{concrete operational
semantics} to a number of concrete context-aware languages. The type system is used to guarantee 
that the resulting programs are correct. Finally, the development in this chapter is closely 
mirrored by the implementation presented in Chapter~\ref{ch:impl}, which implements the 
translation together with an interpreter for the target language.

The key claim of this thesis is that writing context-aware programs using coeffects is easier and
safer. In this chapter, we substantiate the claim by showing that programs written in the coeffect
calculus and evaluated using the translation provided here do not ``go wrong''.

To provide an intuition, consider two context-aware programs. The first calls a function that
adds two implicit parameters in a context where one of them is defined. The second calculates
the difference between the current and the previous value in a dataflow computation.
For comparison, we show the code written in a coeffect dataflow language (on the left) and 
using standard ML-like libraries (on the right):
%
\begin{equation*}
\begin{array}{lcl}
\begin{array}{l}
\kvd{let}~\ident{add}=\kvd{fun}~\ident{x}\rightarrow\\[-0.2em]
 \quad \ident{?one} + \ident{?two}~\kvd{in}\\[-0.2em]
\kvd{let}~\ident{?one} = 10~\kvd{in}\\[-0.2em] 
\ident{add}~0
\end{array} 
&\quad&
\begin{array}{l}
\kvd{let}~\ident{add}=\kvd{fun}~\ident{x}~\ident{params}\rightarrow\\[-0.2em]
 \quad \ident{lookup}~\str{one}~\ident{params}~+ \\[-0.2em]
 \quad\quad \ident{lookup}~\str{two}~\ident{params}~\kvd{in}\\[-0.2em]
\ident{add}~0~(\ident{add}~\str{one}~10~\ident{params})
\end{array} 
\\[3.5em]
\begin{array}{l}
\kvd{let}~\ident{diff}=\kvd{fun}~\ident{x}\rightarrow\\[-0.2em]
 \quad \ident{x}-\kvd{prev}~\ident{x}
\end{array} 
&&
\begin{array}{l}
\kvd{let}~\ident{diff}=\kvd{fun}~\ident{x}\rightarrow\\[-0.2em]
 \quad \ident{List.head}~\ident{x}-\ident{List.head}~(\ident{List.tail}~\ident{x})
\end{array} 
\end{array} 
\end{equation*}
%
The \ident{add} function (on the left) has a type $\ident{int}\xrightarrow{\cclrd{ \{ \ident{?one},\ident{?two} \} }}\ident{int}$.
We call it in a context containing $\cclrd{\ident{?one}}$ and so the coeffect of the program is
$\cclrd{ \{ \ident{?two} \} }$. The safety property for implicit parameters
(Theorem~\ref{thm:semantics-sound-impl}) guarantees that, when executed in a context that
provides a value for the implicit parameter $\cclrd{\ident{?one}}$, the program reduces to 
a value of an correct type (or never terminates).

If we wrote the code without coeffects (on the right), we could use a dynamic map to pass around
a dictionary of parameters (the \ident{lookup} function obtains a value and \ident{add} adds a new
assignment to the map). In that case, the type of \ident{add} is just $\ident{int}\rightarrow\ident{int}$
and so the user does not know which implicit parameters it will need.

Similarly, the \ident{diff} function can be implemented in terms of lists (on the right) as
a function of type $\ident{num}~\ident{list}\rightarrow\ident{num}$. The function fails for input 
lists containing only zero or one elements and this is not reflected in the type and it is not 
enforced by the type checker.

Using coeffects (on the left), the function has a type $\ident{num}\xrightarrow{\cclrd{1}}\ident{num}$ 
meaning that it requires one past value (in addition to the current value). The safety property
for dataflow (Theorem~\ref{thm:semantics-sound-df}) shows that, when called with a context that
contains the required number of past values as captured by the coeffect type system, the function 
does not get stuck.

In summary, coeffect type system, captures certain runtime 
demands of context-aware programs and (as we show in this chapter), eliminates common errors 
related to working with context.

% ==================================================================================================
%                
%    #######
%       #    #    # ######  ####  #####  #   #
%       #    #    # #      #    # #    #  # #
%       #    ###### #####  #    # #    #   #
%       #    #    # #      #    # #####    #
%       #    #    # #      #    # #   #    #
%       #    #    # ######  ####  #    #   #                              
%
% ==================================================================================================

\section{Categorical motivation}
\label{sec:semantics-theory}

The type system of flat coeffect calculus arises syntactically, as a generalization of the examples 
discussed in Chapter~\ref{ch:applications}, but we can also obtain it by looking at the categorical 
semantics of context-dependent computations. This is a direction that we explore in this section. 
Although the development presented here is interesting in its own, our main focus is \emph{using} 
categorical semantics to motivate and explain the translation discussed in 
Section~\ref{sec:semantics-translation}.

% --------------------------------------------------------------------------------------------------

\subsection{Comonads are to coeffects what monads are to effects}

The development in this chapter closely follows the example of effectful computations. 
Effect systems provide a type system for tracking effects and monadic translation can be used as 
a basis for implementing effectful domain-specific languages (e.g.~through the do-notation in 
Haskell).

The correspondence between effect system and monads has been pointed out by Wadler and Thiemann 
\cite{monads-effects-marriage} and further explored by Atkey \cite{monads-parameterised-notions}
and Vazou and Leijen \cite{monads-effects-remarriage}). This line of work relates effectful 
functions $\tau_1 \xrightarrow{\sigma} \tau_2$ to monadic computations 
$\tau_1 \rightarrow \mtyp{\sigma} \tau_2$. In this chapter, we show a similar correspondence between 
\emph{coeffect systems} and \emph{comonads}. However, due to the asymmetry of $\lambda$-calculus, 
defining the semantics in terms of comonadic computations is not a simple mechanical dualisation
of the work on effect systems and monads.

Our approach is inspired by the work of Uustalu and Vene \cite{comonads-notions} who present the 
semantics of contextual computations (mainly for dataflow) in terms of comonadic functions 
$\ctyp{}\tau_1 \rightarrow \tau_2$. Our \emph{indexed comonads} annotate the structure with 
information about the required context, \ie~$\ctyp{\cclrd{r}} \tau_1 \rightarrow \tau_2$.
This is similar to the recent development on monads and effects by Katsumata \cite{monads-parametric}
who parameterizes monads in a similar way to our parameterization of comonads.

% --------------------------------------------------------------------------------------------------

\subsection{Categorical semantics}

As discussed in Section~\ref{sec:path-sem}, categorical semantics interprets terms as morphisms
in some category. For typed calculi, the semantics defined by $\sem{-}$ usually interprets typing 
judgements $x_1 \!:\! \tau_1 \ldots x_n \!:\! \tau_n \vdash e: \tau$ as morphisms 
$\sem{\tau_1 \times \ldots \times \tau_n} \rightarrow \sem{\tau}$.

As a best known example, Moggi \cite{monad-notions} showed that the semantics of various effectful 
computations can be captured uniformly using (\emph{strong}) \emph{monads}. In that 
approach, computations are interpreted as $\tau_1 \times \ldots \times \tau_n \rightarrow \mtyp{}{\tau}$
for some monad $\mtyp{}{}$. For example, $\mtyp{}{\alpha} = \alpha \cup \{ \bot \}$ models 
partiality (maybe monad), $\mtyp{}{\alpha} = \mathcal{P}(\alpha)$ models non-determinism (list 
monad) and side-effects can be modelled using $\mtyp{}{\alpha} = S \rightarrow (\alpha \times S)$ 
(state monad). Here, the structure of a strong monad provides necessary ``plumbing'' for composing 
monadic computations -- sequential composition and strength for lifting free variables into the
body of computation under a lambda abstraction.

Following similar approach to Moggi, Uustalu and Vene \cite{comonads-notions} showed that 
(\emph{monoidal}) \emph{comonads} uniformly capture the semantics of various kinds of context-dependent 
computations~\cite{comonads-notions}. For example, data-flow computations over non-empty lists
are modelled using the non-empty list comonad $\ident{NEList}\,\alpha = \alpha + (\alpha \times \ident{NEList}\,\alpha)$.

The monadic and comonadic model outlined here represents at most a binary analysis of effects or 
context-dependence. A function $\tau_1 \rightarrow \tau_2$ performs \emph{no} effects (requires no 
context) whereas $\tau_1 \rightarrow \mtyp{}{\tau_2}$ performs \emph{some} effects and
$\ctyp{}{\tau_1} \rightarrow \tau_2$ requires \emph{some} context\footnote{This is an 
over-simplification as we can use \eg~stacks of monad transformers and model functions with
two different effects using $\tau_1 \rightarrow M_1(M_2~\tau_2)$. However, monad transformers 
require defining complex system of lifting to be composable. Consequently, they are usually used
for capturing different kinds of impurities (exceptions, non-determinism, state), but not for
capturing fine-grained properties (\eg~a set of memory regions that may be accessed by a
stateful computation).}.

In the next section, we introduce \emph{indexed comonads}, which provide a more precise analysis 
and let us model computations with context demands $\cclrd{r}$ as functions 
$\ctyp{\cclrd{r}}{\tau_1} \rightarrow \tau_2$ using an \emph{indexed comonad} $\ctyp{\cclrd{r}}{}$.

% --------------------------------------------------------------------------------------------------

\subsection{Introducing comonads}

In category theory, \emph{comonad} is a dual of \emph{monad}. As already outlined in
Chapter~\ref{ch:pathways}, we obtain a definition of a comonad by taking a definition of a monad 
and ``reversing the arrows''. More formally, one of the equivalent definitions of comonad 
looks as follows (repeated from Section~\ref{sec:path-sem}):

\begin{definition}
A \emph{comonad} over a category $\catc$ is a triple $(C, \ident{counit}, \ident{cobind})$ where:
\begin{compactitem}
\item $C$ is a mapping on objects (types) $C : \catc \rightarrow \catc$
\item $\ident{counit}$ is a mapping $\ctyp{}{\alpha} \rightarrow \alpha$ 
\item $\ident{cobind}$ is a mapping $(\ctyp{}{\alpha} \rightarrow \beta) 
  \rightarrow (\ctyp{}{\alpha} \rightarrow \ctyp{}{\beta})$
\end{compactitem}
such that, for all $f:\ctyp{}{\alpha} \rightarrow \beta$ and $g:\ctyp{}{\beta} \rightarrow \gamma$:
\begin{align}
\tag{\emph{left identity}}
  \ident{cobind}~\ident{counit} &= \idf{}
  \\
\tag{\emph{right identity}}
  \ident{counit} \circ \ident{cobind}~f &= f
  \\
\tag{\emph{associativity}}
  \ident{cobind}~(g \circ \ident{cobind}~f) &= (\ident{cobind}~g) \circ (\ident{cobind}~f)
\end{align}
\end{definition}

\noindent
From the functional programming perspective, we can see $\ctyp{}{}$ as a parametric data type such as
\ident{NEList}. The $\ident{counit}$ operations extracts a value $\alpha$ from a value that carries 
additional context $\ctyp{}{\alpha}$. The $\ident{cobind}$ operation turns a context-dependent function 
$\ctyp{}{\alpha} \rightarrow \beta$ into a function that takes a value with context, applies
the context-dependent function to value(s) in the context and then propagates the context.

As mentioned earlier, Uustalu and Vene \cite{comonads-notions} use comonads to model data-flow
computations. They describe infinite (coinductive) streams and non-empty lists as example comonads.

\begin{example}[Non-empty list]
A non-empty list is a recursive data-type defined as $\ident{NEList}\,\alpha = \alpha + (\alpha \times \ident{NEList}\,\alpha)$.
We write \kvd{inl} and \kvd{inr} for constructors of the left and right cases, respectively. The 
type \ident{NEList} forms a comonad together with the following \ident{counit} and \ident{cobind} mappings:
%
\begin{equation*}
\begin{array}{rclll}
\ident{counit}~l &\narrow{=}& h &\quad&\textnormal{when}~l=\kvd{inl}~h\\[-0.25em]
\ident{counit}~l &\narrow{=}& h &&\textnormal{when}~l=\kvd{inr}~(h, t)\\[0.5em]
\ident{cobind}\,f~l &\narrow{=}& \kvd{inl}~(f\,l) &&\textnormal{when}~l=\kvd{inl}~h\\[-0.25em]
\ident{cobind}\,f~l &\narrow{=}& \kvd{inr}~(f\,l,\;\ident{cobind}~f~t) &&\textnormal{when}~l=\kvd{inr}~(h, t)
\end{array}
\end{equation*}
\end{example}

\noindent
The \ident{counit} operation returns the head of the non-empty list. Note that it is crucial that
the list is \emph{non-empty}, because we always need to be able to obtain a value. The \ident{cobind}
operation defined here returns a list of the same length as the original where, for each element, the 
function $f$ is applied on a \emph{suffix} list starting from the element. Using a simplified
notation for list, the result of applying \ident{cobind} to a function that sums elements of a
list gives the following behaviour:
%
\begin{equation*}
\ident{cobind}~\ident{sum}~(7,6,5,4,3,2,1,0) = (28,21,15,10,6,3,1,0)
\end{equation*}
%
The fact that the function $f$ is applied to a \emph{suffix} is important in order to satisfy the
\emph{left identity} law, which requires that $\ident{cobind}~\ident{counit}~l = l$.

It is also interesting to examine some data types that do \emph{not} form a comonad. As already
mentioned, list $\ident{List}~\alpha = 1 + (\alpha \times \ident{List}~\alpha)$ is not a comonad,
because the \ident{counit} operation is not defined for the value $\kvd{inl}~()$. The \ident{Maybe} 
data type defined as $1 + \alpha$ is not a comonad for the same reason. However, if we consider 
flat coeffect calculus for liveness, it appears natural to model computations as functions 
$\ident{Maybe}~\tau_1 \rightarrow \tau_2$. To use such a model, we need to generalise comonads 
to \emph{indexed comonads}.

% --------------------------------------------------------------------------------------------------

\subsection{Generalising to indexed comonads}
\label{sec:semantics-flat-idx}

The flat coeffect algebra includes a monoid $(\C, \cseq, \cunit)$, which defines the behaviour of
sequential composition, where the annotation $\cunit$ represents a variable access. An indexed 
comonad is formed by a data type (object mapping) $\ctyp{\cclrd{r}}{\alpha}$ where the annotation 
$\cclrd{r}$ determines what context is required.

\begin{definition}
Given a monoid $(\C, \cseq, \cunit)$ with binary operator $\cseq$ and unit $\cunit$, an 
\emph{indexed comonad} over a category $\catc$ is a triple 
$(\ctyp{\cclrd{r}{}}, \ident{counit}_{\cunit}, \ident{cobind}_{\cclrd{r}, \cclrd{s}})$ where:

\begin{compactitem}
\item $\ctyp{\cclrd{r}}{}$ for all $\cclrd{r} \in \C$ is a family of object mappings 
\item $\ident{counit}_{\cunit}$ is a mapping $\ctyp{\cunit}{\alpha} \rightarrow \alpha$ 
\item $\ident{cobind}_{\cclrd{r}, \cclrd{s}}$ is a mapping $(\ctyp{\cclrd{r}}{\alpha} \rightarrow \beta) 
  \rightarrow (\ctyp{\cclrd{r}\cseq\cclrd{s}}{\alpha} \rightarrow \ctyp{\cclrd{s}}{\beta})$
\end{compactitem}
such that, for all $f:\ctyp{\cclrd{r}}{\alpha} \rightarrow \beta$ and $g:\ctyp{\cclrd{s}}{\beta} \rightarrow \gamma$:f
\begin{align}
\tag{\emph{left identity}}
  \ident{cobind}_{\cunit, \cclrd{s}}~\ident{counit}_{\cunit} &= \idf{}
  \\
\tag{\emph{right identity}}
  \ident{counit}_{\cunit} \circ \ident{cobind}_{\cclrd{r}, \cunit}~f &= f
  \\
\tag{\emph{associativity}}
\hspace{-10em}
  \ident{cobind}_{\cclrd{r}\cseq\cclrd{s},\cclrd{t}}~(g \circ \ident{cobind}_{\cclrd{r}, \cclrd{s}}~f) &= 
    (\ident{cobind}_{\cclrd{s}, \cclrd{t}}~g) \circ (\ident{cobind}_{\cclrd{r}, \cclrd{s}\cseq\cclrd{t}}~f)
\end{align}
\end{definition}

\noindent
Rather than defining a single mapping $\ctyp{}{}$, we are now defining a family of mappings 
$\ctyp{\cclrd{r}}{}$ indexed by a monoid structure. Similarly, the $\ident{cobind}_{\cclrd{r}, \cclrd{s}}$ 
operation is now formed by a \emph{family} of mappings for different pairs of indices 
$\cclrd{r}, \cclrd{s}$. To be fully precise, $\ident{cobind}$ is a family of natural transformations 
and we should include $\alpha, \beta$ as indices, writing $\ident{cobind}_{\cclrd{r},\cclrd{s}}^{\alpha, \beta}$.
For the purpose of this thesis, it is sufficient to omit the superscripts and treat $\ident{cobind}$ just 
as a family of mappings (rather than natural transformations). When this does not introduce ambiguity, 
we also occasionally omit the subscripts.

The $\ident{counit}$ operation is not defined for all $\cclrd{r} \in \C$, but only for 
the unit $\cunit$. We still include the unit as an index writing $\ident{counit}_{\cunit}$, 
but this is merely for symmetry and as a useful reminder to the reader. Crucially, this means that 
the operation is defined only for special contexts.

If we look at the indices in the laws, we can see that the left and right identity 
require $\cunit$ to be the unit of $\cseq$. Similarly, the associativity law implies the 
associativity of the $\cseq$ operator.

\paragraph{Composition.}
The co-Kleisli category that models sequential composition is formed by the unit arrow (provided
by $\ident{counit}$) together with the (associative) composition operation that composes computations 
with contextual demands as follows:
%
\begin{equation*}
\begin{array}{ccl}
\textnormal{--}\, \hat{\circ} \,\textnormal{--}&\narrow{:}& (\ctyp{\cclrd{r}}{\tau_1} \rightarrow \tau_2) 
  \rightarrow (\ctyp{\cclrd{s}}{\tau_2} \rightarrow \tau_3) 
  \rightarrow (\ctyp{\cclrd{r} \cseq \cclrd{s}}{\tau_1} \rightarrow \tau_3) \\
g \, \hat{\circ} \, f &\narrow{=}& g \circ (\ident{cobind}_{\cclrd{r}, \cclrd{s}} f)
\end{array}
\end{equation*}
%
The composition $\hat{\circ}$ best expresses the intention of indexed comonads. Given two functions
with contextual demands $\cclrd{r}$ and $\cclrd{s}$, their composition is a function that 
requires $\cclrd{r}\,\cseq\,\cclrd{s}$. The contextual demands propagate \emph{backwards} and
are attached to the input of the composed function.

% --------------------------------------------------------------------------------------------------

\paragraph{Examples.}

Any comonad can be turned into an indexed comonad using a trivial monoid. However, indexed comonads
are more general and can be used with other data types, including indexed \ident{Maybe}.

\begin{example}[Comonads]
Any comonad $\ctyp{}{}$ is an indexed comonad with an index provided by a trivial monoid $(\{1\},\ast,1)$
where $1\ast 1 = 1$. The mapping $\ctyp{1}{}$ is the mapping $\ctyp{}{}$ of the underlying comonad. The
operations $\ident{counit}_1$ and $\ident{cobind}_{1,1}$ are defined by the operations $\ident{counit}$
and $\ident{cobind}$ of the comonad.
\end{example}

\begin{example}[Indexed option]
The \emph{indexed option comonad} is defined over a monoid $(\{ \ident{L},\ident{D} \}, \sqcup,\ident{L})$ 
where $\sqcup$ is defined as earlier, \ie~$\ident{L} = \cclrd{r} \sqcup \cclrd{s} \Longleftrightarrow \cclrd{r}=\cclrd{s}=\ident{L}$.
Assuming $1$ is the unit type inhabited by $()$, the mappings are defined as follows:
%
\begin{equation*}
\begin{array}{l}
\ctyp{\ident{L}}{\alpha} = \alpha\\
\ctyp{\ident{D}}{\alpha} = 1\\
\\
\ident{counit}_{\ident{L}} : \ctyp{\ident{L}}{\alpha} \rightarrow \alpha \\
\ident{counit}_{\ident{L}}~v = v\\
\end{array}
\qquad
\begin{array}{lcl}
\multicolumn{3}{l}{
  \ident{cobind}_{\cclrd{r},\cclrd{s}} ~:~ (\ctyp{\cclrd{r}}{\alpha} \rightarrow \beta) 
    \rightarrow (\ctyp{\cclrd{r}\sqcup\cclrd{s}}{\alpha} \rightarrow \ctyp{\cclrd{s}}{\beta}) }\\
\ident{cobind}_{\ident{L},\ident{L}}~f~x &\narrow{=}& f~x\\
\ident{cobind}_{\ident{L},\ident{D}}~f~() &\narrow{=}& ()\\
\ident{cobind}_{\ident{D},\ident{L}}~f~() &\narrow{=}& f~()\\
\ident{cobind}_{\ident{D},\ident{D}}~f~() &\narrow{=}& ()\\
\end{array}
\end{equation*}
\end{example}

\noindent
The \emph{indexed option comonad} models the semantics of the liveness coeffect system discussed in 
Section~\ref{sec:applications-flat-live}, where $\ctyp{\ident{L}}{\alpha} = \alpha$ models a live context 
and $\ctyp{\ident{D}}{\alpha}=1$ models a dead context which does not contain a value. The \ident{counit}
operation extracts a value from a live context. As in the direct model discussed in Chapter~\ref{ch:appendix},
the \ident{cobind} operation can be seen as an implementation of dead code elimination. The definition 
only evaluates $f$ when the result is marked as live and is thus required, and it only accesses $x$ if 
the function $f$ requires its input.

The indexed family $\ctyp{\cclrd{r}}{}$ in the above example is analogous to the \ident{Maybe}
(or option) data type $\ident{Maybe}\,\alpha = 1 + \alpha$. As mentioned earlier, this type does not 
permit (non-indexed) comonad structure, because $\ident{counit}~()$ is not defined. This is not a 
problem with indexed comonads, because live contexts are distinguished by the (type-level) coeffect 
annotation and \ident{counit} only needs to be defined on live contexts.

\begin{example}[Indexed product]
The semantics of implicit parameters is modelled by an indexed product comonad. We use a monoid
$(\mathcal{P}(\ident{Id}), \cup, \emptyset)$ where \ident{Id} is the set of (implicit parameter) names.
As previously, all parameters have the type $\rho$. The data type $\ctyp{\cclrd{r}}{\alpha}
= \alpha \times (\cclrd{r} \rightarrow \rho)$ represents a value $\alpha$ together with a function that
associates a parameter value $\rho$ with every implicit parameter name in $\cclrd{r} \subseteq \ident{Id}$.
The cobind and counit operations are defined as:
%
\begin{equation*}
\begin{array}{l}
\ident{counit}_{\emptyset} : \ctyp{\emptyset}{\alpha} \rightarrow \alpha \\[-0.25em]
\ident{counit}_{\emptyset}~(a, g) = a\\[0.5em]
\end{array}
\quad
\begin{array}{l}
\ident{cobind}_{\cclrd{r},\cclrd{s}} ~:~ (\ctyp{\cclrd{r}}{\alpha} \rightarrow \beta) 
    \rightarrow (\ctyp{\cclrd{r}\cup\cclrd{s}}{\alpha} \rightarrow \ctyp{\cclrd{s}}{\beta})\\[-0.25em]
\ident{cobind}_{\cclrd{r},\cclrd{s}}~f~(a,g) = (f(a,\restr{g}{\cclrd{r}}), \restr{g}{\cclrd{s}})\\[0.5em]
\end{array}
\end{equation*}
\end{example}

\noindent
In the definition, we use the notation $(a, g)$ for a pair containing the value of type $\alpha$
together with $g$, which is a function $\cclrd{r} \rightarrow \rho$. The \ident{counit} operation
takes a value and a function (with empty set as a domain), ignores the function and extracts the
value. The \ident{cobind} operation uses the restriction operation $\restr{g}{\cclrd{r}}$ to 
restrict the domain of $g$ to implicit parameters $\cclrd{r}$ and $\cclrd{s}$ in oder to get
implicit parameters required by the argument of $f$ and by the resulting computation, respectively
(\ie~semantically, it \emph{splits} the available context capabilities). The function $g$ passed 
to \ident{cobind} is initially defined on $\cclrd{r} \cup \cclrd{s}$ and so the restriction is 
valid in both cases.

The structure of \emph{indexed comonads} is sufficient to model sequential composition of 
computations that use a single variable (as discussed in Section~\ref{sec:path-sem}). To model 
full $\lambda$-calculus with lambda abstraction and multiple-variable contexts, we need 
additional operations introduced in the next section.

% --------------------------------------------------------------------------------------------------

\subsection{Flat indexed comonads}
\label{sec:semantics-flat-monoidal}

Because of the asymmetry of $\lambda$-calculus (discussed in Section~\ref{sec:applications-structure}),
the duality between monads and comonads can no longer help us with defining the additional structure
required to model full $\lambda$-calculus. In comonadic computations, additional information is attached
to the context. In application and lambda abstraction, the context is propagated differently than 
in effectful computations.

To model the effectful $\lambda$-calculus, Moggi~\cite{monad-notions} requires a \emph{strong} monad which
has an additional operation $\ident{strength} : \alpha\times\mtyp{}{\beta} \rightarrow \mtyp{}{(\alpha\times\beta)}$.
This allows lifting of free variables into an effectful computation. In Haskell, strength can be expressed
in the host language and so it is implicit.

To model $\lambda$-calculus with contextual properties, Uustalu and Vene \cite{comonads-notions} 
require \emph{lax semi-monoidal} comonad. This structure requires an additional monoidal operation:
%
\begin{equation*}
\ident{m} : \ctyp{}{\alpha} \times \ctyp{}{\beta} \rightarrow \ctyp{}{(\alpha \times \beta)}
\end{equation*}
%
The \ident{m} operation is needed in the semantics of lambda abstraction. It represents merging of 
contexts and is used to merge the context of the declaration site (containing free variables)
and the call site (containing bound variable). For example, for implicit parameters, this combines
the additional parameters defined in the two contexts.

The semantics of flat coeffect calculus requires operations for \emph{merging}, but also for
\emph{splitting} of contexts.

\begin{definition}
Given a flat coeffect algebra $(\C, \cseq, \cpar, \czip, \cunit, \czero, \cleq)$,
a \emph{flat indexed comonad} is an indexed comonad over the monoid $(\C, \cseq, \cunit)$
equipped with families of operations $\ident{merge}_{\cclrd{r},\cclrd{s}}$, $\ident{split}_{\cclrd{r},\cclrd{s}}$ where:
%
\begin{compactitem}
\item $\ident{merge}_{\cclrd{r},\cclrd{s}}$ is a family of mappings
  $\ctyp{\cclrd{r}}{\alpha} \times \ctyp{\cclrd{s}}{\beta} \rightarrow \ctyp{\cclrd{r}\czip\cclrd{s}}{(\alpha \times \beta)}$
\item $\ident{split}_{\cclrd{r},\cclrd{s}}$ is a family of mappings
  $\ctyp{\cclrd{r}\cpar\cclrd{s}}{(\alpha \times \beta)} \rightarrow \ctyp{\cclrd{r}}{\alpha} \times \ctyp{\cclrd{s}}{\beta}$
\end{compactitem}
\end{definition}

\noindent
The $\ident{merge}_{\cclrd{r},\cclrd{s}}$ operation is the most interesting one. Given two comonadic
values with additional contexts specified by $\cclrd{r}$ and $\cclrd{s}$, it combines them into a 
single value with additional context $\cclrd{r}\,\czip\,\cclrd{s}$. The $\czip$ operation often represents
\emph{greatest lower bound}\footnote{The $\czip$ and $\cpar$ operations are the greatest and least upper 
bounds in the liveness and data-flow examples, but not for implicit parameters. However, they remain useful 
as an informal analogy.}, elucidating the fact that merging may result in the loss of some parts of 
the contexts $\cclrd{r}$ and $\cclrd{s}$. We look at examples of this operation in the next section.

The $\ident{split}_{\cclrd{r},\cclrd{s}}$ operation splits a single comonadic value (containing a tuple)
into two separate values. Note that this does not simply duplicate the value, because the additional
context is also split. To obtain coeffects $\cclrd{r}$ and $\cclrd{s}$, the input needs to provide 
\emph{at least} $\cclrd{r}$ and $\cclrd{s}$, so the tags are combined using the $\cpar$, which is often 
the \emph{least upper-bound}\footnotemark[1].

\paragraph{Semantics of sub-coeffecting.}
Although we do not include sub-coeffecting in the core flat coeffect calculus, it is an interesting 
extension to consider. Semantically, sub-coeffecting drops some of the available contextual 
capabilities (drops some of the implicit parameters or some of the past values). This can be
modelled by adding a (family of) lifting operation(s):
%
\begin{itemize}
 \item $\ident{lift}_{\cclrd{r'},\cclrd{r}}$ is a family of mappings
   $\ctyp{\cclrd{r'}}{\alpha} \rightarrow \ctyp{\cclrd{r}}{\alpha}$~ for all $\cclrd{r'}, \cclrd{r}$ such that $\cclrd{r}\,\cleq\,\cclrd{r'}$
\end{itemize}
%
Although we do not demand this as a general law, in all our systems, it is the case that
$\cclrd{r} \;\cleq\; \cclrd{r}\,\cpar\,\cclrd{s}$ and $\cclrd{s} \;\cleq\; \cclrd{r}\,\cpar\,\cclrd{s}$.
This special case allows a simpler definition of \emph{indexed flat comonad} by expressing the 
\ident{split} operation in terms of lifting (sub-coeffecting) as follows:
%
\begin{equation*}
\begin{array}{rcl}
\ident{map}_{\cclrd{r}}~f &\narrow{=}& \ident{cobind}_{\cclrd{r}, \cclrd{r}}~(f\circ\ident{counit}_{\cunit}) \\
\ident{split}_{\cclrd{r}, \cclrd{s}}~c &\narrow{=}&
  ( \ident{map}_{\cclrd{r}}~\ident{fst}~(\ident{lift}_{\cclrd{r}\cpar\cclrd{s}, \cclrd{r}}~c), 
    \ident{map}_{\cclrd{s}}~\ident{snd}~(\ident{lift}_{\cclrd{r}\cpar\cclrd{s}, \cclrd{s}}~c) )
\end{array}
\end{equation*}
%
The $\ident{map}_{\cclrd{r}}$ operation is the mapping on arrows that corresponds to the object 
mapping $\ctyp{\cclrd{r}}{}$. The definition is dual to the standard definition of \ident{map} 
for monads in terms of \ident{bind} and \ident{unit}. The functions \ident{fst} and \ident{snd}
are first and second projections from a two-element pair. To define the 
$\ident{split}_{\cclrd{r}, \cclrd{s}}$ operation, we use the argument $c$ twice, use lifting
to throw away additional parts of the context and then transform the values in the 
context.

This alternative is valid for our examples, but we do not use it for three reasons. First, it
requires making sub-coeffecting a part of the core definition. Second, this would be the only 
place where our semantics uses a variable \emph{twice} (in this case $c$). Thus using an explicit 
\ident{split} means that the structure required by our semantics does not need to provide variable 
duplication and our model could be embedded in linear or affine category. Finally, explicit 
\ident{split} is similar to the definition that is needed for structural coeffects in 
Chapter~\ref{ch:structural} and it makes the connection between the two easier to see.

\paragraph{Examples.}
All the examples of \emph{indexed comonads} discussed in Section~\ref{sec:semantics-flat-idx} can
be extended into \emph{flat indexed comonads}. Note that this is not true \emph{in general}, 
because each example requires us to define additional operations, specific for the example.

\begin{example}[Monoidal comonads]
Just like indexed comonads generalise co\-monads, the additional structure of
flat indexed comonads generalises symmetric semimonoidal comonads of Uustalu 
and Vene \cite{comonads-notions}. The flat coeffect algebra is defined as $(\{1\}, \ast, \ast, \ast, 1, 1, =)$
where $1\ast1=1$ and $1=1$. The additional operation $\ident{merge}_{1,1}$ is provided by the 
monoidal operation called \ident{m} by Uustalu and Vene. The $\ident{split}_{1,1}$ operation 
is defined by duplication.
\end{example}

\begin{example}[Indexed option]
\label{thm:semantics-indexed-opt}

Flat coeffect algebra for liveness defines $\cpar$ and $\czip$ as $\sqcup$ and $\sqcap$, respectively 
and specifies that $\ident{D} \sqsubseteq \ident{L}$. Recall also that the object mapping is defined 
as $\ctyp{\ident{L}}{\alpha} = \alpha$ and $\ctyp{\ident{D}}{\alpha} = 1$. The additional operations 
of a flat indexed comonad are defined as follows:
%
\begin{equation*}
\begin{array}{rcl}
\ident{merge}_{\ident{L}, \ident{L}}~(a, b) &\narrow{=}& (a, b)\\
\ident{merge}_{\ident{L}, \ident{D}}~(a, ()) &\narrow{=}& ()\\
\ident{merge}_{\ident{D}, \ident{L}}~((), b) &\narrow{=}& ()\\
\ident{merge}_{\ident{D}, \ident{D}}~((), ()) &\narrow{=}& ()\\
\end{array}
\qquad
\begin{array}{rcl}
\ident{split}_{\ident{L}, \ident{L}}~(a, b) &\narrow{=}& (a, b)\\
\ident{split}_{\ident{L}, \ident{D}}~(a, b) &\narrow{=}& (a, ())\\
\ident{split}_{\ident{D}, \ident{L}}~(a, b) &\narrow{=}& ((), b)\\
\ident{split}_{\ident{D}, \ident{D}}~() &\narrow{=}& ((), ()))\\
\end{array}
\end{equation*}
\end{example}

\noindent
Without the indexing, the \ident{merge} operations implements \emph{zip} on option values,
returning an option only when both values are present. The behaviour of the \ident{split} 
operation is partly determined by the indices. When the input is \emph{dead}, both values have 
to be dead (this is also the only solution of $\cclrd{\ident{D}} =\cclrd{r}\sqcap\cclrd{s}$), but when
the input is \emph{live}, the operation can perform implicit sub-coeffecting and drop one of
the values.

\begin{example}[Indexed product]
\label{thm:semantics-indexed-prod}

For implicit parameters, both $\czip$ and $\cpar$ are the $\cup$ operation and the relation
$\cleq$ is formed by the subset relation $\subseteq$. Recall that the data type $\ctyp{\cclrd{r}}{\alpha}$
is $\alpha \times (\cclrd{r} \rightarrow \rho)$ where $\rho$ is the type of implicit parameter values.
The additional operations are defined as:
%
\begin{equation*}
\begin{array}{rcl}
\ident{split}_{\cclrd{r}, \cclrd{s}}~((a,b), g) &\narrow{=}& ((a, \restr{g}{\cclrd{r}}), (b, \restr{g}{\cclrd{s}}))\\
\ident{merge}_{\cclrd{r}, \cclrd{s}}~((a, f), (b, g)) &\narrow{=}& ((a, b), f \uplus g)\\
\end{array}
\quad
\begin{array}{l}
\textnormal{where}~f \uplus g = \\[-0.25em]
\quad\restr{f}{\, \textit{dom}(f) \setminus \textit{dom}(g)} \cup g 
\end{array}
\end{equation*}
\end{example}

\noindent
The \ident{split} operation splits the tuple and restricts the function (representing available
implicit parameters) to the required sub-sets. The \ident{merge} operation is more 
interesting. It uses the $\uplus$ operation that we defined when introducing implicit parameters
in Section~\ref{sec:applications-flat-impl}. It merges the values, preferring the definitions from
the right-hand side (call site) over left-hand side (declaration site). Thus the operation is not
symmetric.

\begin{example}[Indexed list]
\label{thm:semantics-indexed-list}

Our last example provides the semantics of data-flow computations. The flat coeffect algebra 
is formed by $(\mathbb{N}, +, \mathit{max}, \mathit{min}, 0, 0, \leq)$. In a 
non-indexed version, the semantics is provided by a non-empty list. In the indexed semantics,
the index represents the number of available past values. The data type is then a pair of 
the current value, followed by $n$ past values. The mappings that form the flat indexed comonad 
are defined as follows:
%
\begin{equation*}
\begin{array}{l}
\ident{counit}_{0}\langle a_0 \rangle = a_0
\\[0.45em]
\ident{cobind}_{m, n}~f \langle a_0, \ldots a_{m+n} \rangle = \\[-0.25em]
\quad \langle f \langle a_0, \ldots, a_m \rangle, \ldots, f \langle a_{n}, \ldots, a_{m+n} \rangle \rangle
\\[0.45em]
\ident{merge}_{m, n} (\langle a_0, \ldots, a_m \rangle, \langle b_0, \ldots, b_n\rangle) = \\[-0.25em]
\quad \langle (a_0, b_0), \ldots, (a_{\mathit{min}(m,n)}, b_{\mathit{min}(m,n)}) \rangle
\\[0.45em]
\ident{split}_{m, n} \langle (a_0, b_0), \ldots, (a_{\mathit{max}(m,n)}, b_{\mathit{max}(m,n)}) \rangle = \\[-0.25em]
\quad (\langle a_0, \ldots, a_m \rangle, \langle b_0, \ldots, b_n\rangle)
\end{array}
\begin{array}{l}
\hspace{-2em}\ctyp{n}{\alpha} = \underbrace{\alpha \times \ldots \times \alpha}_{(n+1)-\textnormal{times}}\\[8em]
~\\
\end{array}
\end{equation*}
\end{example}

\noindent
The reader is invited to check that the number of required past elements in each of the mappings
matches the number specified by the indices. The index specifies the number of \emph{past} elements
and so the list always contains at least one value. Thus \ident{counit} returns the element of a
singleton list.

The $\ident{cobind}_{m,n}$ operation requires $m + n$ elements in order to generate $n$ past results 
of the $f$ function, which itself requires $m$ past values. When combining two lists, 
$\ident{merge}_{m,n}$ behaves as \emph{zip} and produces a list that has the length of the shorter 
argument. When splitting a list, $\ident{split}_{m, n}$ needs the maximum of the required lengths. 
Finally, the lifting operation just drops some number of elements from a list.

% --------------------------------------------------------------------------------------------------

\begin{figure*}[t]
The semantics is defined over a typing derivation:
%  
\begin{equation*}
\hspace{-3em}
\begin{array}{ll}
\hspace{5.2em}\semdef
  {\coctx{\Gamma}{\cunit} \vdash x_i : \tau_i }
  {\pi_i \circ \ident{counit}_{\cunit}} 
& (\emph{var})
\\[1em]
\hspace{4.9em}\semdef
  {\coctx{\Gamma}{\czero} \vdash n : \ident{num} }
  {\ident{const}~n} 
& (\emph{num})
\\[1.5em]
\hspace{2.2em}\semdeff
  {\coctx{\Gamma, x:\tau_1}{\cclrd{r}\;\czip\;\cclrd{s}} \vdash e : \tau_2}
  {\coctx{\Gamma}{\cclrd{r}} \vdash \lambda x.e : \tau_1 \xrightarrow{\cclrd{s}} \tau_2 }
  {f}
  {f \circ \ident{curry}~\ident{merge}_{\cclrd{r}, \cclrd{s}} }
& (\emph{abs})
\\[2em]
\semdefff
  {\coctx{\Gamma}{\cclrd{r}} \vdash e_1 : \tau_1 \xrightarrow{\cclrd{t}} \tau_2}
  {\coctx{\Gamma}{\cclrd{s}} \vdash e_2 : \tau_1 }
  {\begin{array}{l}\coctx{\Gamma}{\cclrd{r} \;\cpar\; (\cclrd{s} \,\cseq\, \cclrd{t})} \vdash e_1~e_2 : \tau_2\\[-0.3em]~\end{array}}
  {f}
  {g}
  {\hspace{-0.5em}\begin{array}{l} 
  \ident{app}~\circ~f\hspace{-0.15em}\times\hspace{-0.15em}(\ident{cobind}_{\cclrd{s}, \cclrd{t}}~g)~\circ~\ident{split}_{\cclrd{r}, \cclrd{s} \,\cseq\, \cclrd{t}} \\[-0.3em]
  ~~\circ~\ident{map}_{\cclrd{r} \;\cpar\; (\cclrd{s} \,\cseq\, \cclrd{t})}~\ident{dup}\\[-0.3em]
  \end{array}\hspace{-0.5em}}   
& (\emph{app})
\end{array}
\end{equation*}

~

Assuming the following auxiliary operations:
%
\begin{equation*}
\begin{array}{rcl}
  \ident{map}_\cclrd{r}}~f &=& \ident{cobind}_{\cunit, \cclrd{r}}~(f\circ\ident{counit}_\cunit) \\[-0.3em]
  \ident{const}~v &=& \lambda x.v \\[-0.3em]
  \ident{curry}~f~x~y &=& \lambda f.\lambda x.\lambda y.f~(x, y)\\[-0.3em]
  \ident{dup}~x &=& (x, x)\\[-0.3em]
  f \times g &=& \lambda (x, y).(f~x, g~y)\\[-0.3em]
  \ident{app}~(f, x) &=& f~x
\end{array}
\end{equation*}

\figcaption{Categorical semantics of the flat coeffect calculus}
\label{fig:semantics-flat}
\end{figure*}

% --------------------------------------------------------------------------------------------------

\subsection{Semantics of flat calculus}
\label{sec:semantics-flat-calculus}

In Section~\ref{sec:applications-flat}, we defined the semantics of concrete (flat) context-dependent
computations including implicit parameters, liveness and data-flow. Using the \emph{flat indexed 
comonad} structure, we can now define a single uniform semantics that is capable of capturing all 
our examples, as well as other computations that can be modelled by the structure.

As discussed in Section~\ref{sec:flat-unique}, different typing derivations of coeffect programs 
may have different meaning (\eg~when working with implicit parameters) and so the semantics is
defined over a \emph{typing derivation} rather than over an \emph{term}. To assign a 
semantics to a term, we choose its unique coeffect-specific typing derivation as defined in 
Section~\ref{sec:flat-unique}.

\paragraph{Contexts and types.}
The modelling of contexts and functions generalizes the earlier concrete examples. We use the 
family of mappings $\ctyp{\cclrd{r}}{}$ as an (indexed) data-type that wraps the product of 
free variables of the context and the arguments of functions:
%
\begin{equation*}
\begin{array}{rcl}
\sem{\coctx{x_1\!:\!\tau_1, \ldots, x_n\!:\!\tau_n}{ \cclrd{r} } \vdash e : \tau} 
  &:& \ctyp{\cclrd{r}}{(\tau_1 \times \ldots \times \tau_n)} \rightarrow \tau\\
\sem{\tau_1 \xrightarrow{\cclrd{r}} \tau_2} &=& \ctyp{\cclrd{r}}{\tau_1} \rightarrow \tau_2
\end{array}
\end{equation*}

\paragraph{Expressions.}
The definition of the semantics is shown in Figure~\ref{fig:semantics-flat}. For consistency
with earlier work \cite{comonads-notions,comonads-dom-thesis}, the definitions use a point-free 
categorical notation. The semantics uses a number of auxiliary definitions that can be expressed
in a Cartesian-closed category such as currying \ident{curry}, value duplication \ident{dup},
function pairing (given $f:A\rightarrow B$ and $g:C\rightarrow D$ then $f\times g : A\times C \rightarrow B \times D$)
and application \ident{app}. We will embed the definitions in a simple programming language 
later (Section~\ref{sec:semantics-translation}).

The semantics of variable access and abstraction are the same as in the semantics of Uustalu and 
Vene \cite{comonads-notions}, modulo the indexing. The semantics of variable access (\emph{var}) uses 
$\ident{counit}_{\cunit}$ to extract a product of free variables and projection
$\pi_i$ to obtain the variable value. Abstraction (\emph{abs}) is interpreted as a curried
function that takes the declaration-site context and a function argument, merges them using
$\ident{merge}_{\cclrd{r}, \cclrd{s}}$ and passes the result to the semantics of the body $f$. 
Assuming the context $\Gamma$ contains variables of types $\sigma_1, \ldots, \sigma_n$, this gives 
us a value $\ctyp{\cclrd{r}\czip\cclrd{s}}((\sigma_1 \times \ldots \times \sigma_n) \times \tau_1)$.
Assuming that $n$-element tuples are associated to the left, the wrapped context is equivalent to
$\sigma_1 \times \ldots \times \sigma_n \times \tau_1$, which can then be passed to the body of the
function.

The semantics of application (\emph{app}) first duplicates the free-variable product inside the 
context (using $\ident{map}_{\cclrd{r}}$ and duplication). Then it splits this context using 
$\ident{split}_{\cclrd{r}, \cclrd{s} \cpar \cclrd{t}}$. The two contexts contain the same variables
(as required by sub-expressions $e_1$ and $e_2$), but different coeffect annotations. The first
context (with index $\cclrd{r}$) is used to evaluate $e_1$ using the semantic function $f$. The 
result is a function $\ctyp{\cclrd{t}}{\tau_1} \rightarrow \tau_2$. The second context
(with index $\cclrd{s}\cseq\cclrd{t}$) is used to evaluate $e_2$ and using the semantic function $g$
and wrap it with context required by the function $e_1$ by applying $\ident{cobind}_{\cclrd{s}, \cclrd{t}}$.
The \ident{app} operation than applies the function (first element) on the argument (second element).
Finally, numbers (\emph{num}) become constant functions that ignore the context.

\paragraph{Properties.}
We the categorical semantics  in Section~\ref{sec:semantics-translation} to define a translation 
that embeds context-dependent computations in a functional programming langauge, similarly to how 
monads and the do notation provide a way of embedding effectful computations in Haskell.

An important property of the translation is that it respects the coeffect annotations provided 
by the type system. The annotations of the semantic functions match the annotations in the typing
judgement and so the semantics is well-defined. 
This provides a further validation for the design of the type system developed in 
Section~\ref{sec:flat-calculus-types} -- if the coeffect annotations for (\emph{app}) and (\emph{abs}) 
were different, we would not be able to provide a well-defined semantics using flat indexed comonads.

Informally, the following states that if we see the semantics as a translation, the resulting code
is well-typed. We revisit the property in Lemma~\ref{thm:semantics-welltyped} once we define
the target langauge and its typing.

\begin{lemma}[Correspondence]
\label{thm:flat-correspondence}
In the semantics defined in Figure~\ref{fig:semantics-flat}, the context annotations $\cclrd{r}$ of 
typing judgements $\coctx{\Gamma}{\cclrd{r}} \vdash e : \tau$ and function types 
$\tau_1 \xrightarrow{\cclrd{r}} \tau_2$  on the left-hand side correspond to the indices of mappings 
$\ctyp{\cclrd{r}}{}$ in the corresponding semantic function on the right-hand side.
\end{lemma}
\begin{proof}
By analysis of the semantic rules in Figure~\ref{fig:semantics-flat}.
\end{proof}

\noindent
Thanks to indexing, the correspondence provides more guarantees than for a non-indexed system.
In the semantics, we not only know which values are comonadic, but we also know what contextual
information they are required to provide. In Section~\ref{sec:semantics-generalising}, we note
that this lets us generalize the proofs about concrete languages discussed in this chapter to
a more general setting.

The semantics is also a generalization of the concrete semantics given when introducing
context-aware programming languages in Chapter~\ref{ch:applications}.

\begin{theorem}[Generalization]
\label{thm:flat-generalization}

Consider a typing derivation obtained according to the rules for finding unique typing derivations
as specified in Section~\ref{sec:flat-unique} for a coeffect language with liveness, dataflow or 
implicit parameters.

The semantics obtained by instantiating the rules in Figure~\ref{fig:semantics-flat} with the
concrete operations defined in Example~\ref{thm:semantics-indexed-opt},
Example~\ref{thm:semantics-indexed-prod} or Example~\ref{thm:semantics-indexed-list} is the
same as the one defined in Figure~\ref{fig:applications-flat-livsem},
Figure~\ref{fig:applications-flat-implsem} and Figure~\ref{fig:applications-flat-dfsem},
respectively.
\end{theorem}
\begin{proof}
Simple expansion of the definitions for the unique typing derivation.
\end{proof}

% ==================================================================================================
%
%      #######
%      #       #    # #####  ###### #####  #####  # #    #  ####
%      #       ##  ## #    # #      #    # #    # # ##   # #    #
%      #####   # ## # #####  #####  #    # #    # # # #  # #
%      #       #    # #    # #      #    # #    # # #  # # #  ###
%      #       #    # #    # #      #    # #    # # #   ## #    #
%      ####### #    # #####  ###### #####  #####  # #    #  ####
%
% ==================================================================================================

\begin{figure}[t]
\paragraph{Language syntax}  
\begin{equation*}
\begin{array}{rcl}
  v &  = & n \sep \lambda x.e \sep (v_1, \ldots, v_n) \\
  e &  = & x \sep n \sep \pi_i~e \sep (e_1, \ldots, e_n) \sep e_1~e_2 \sep \lambda x.e \\
  \tau &  = & \ident{num} \sep \tau_1 \times \ldots \times \tau_n \sep \tau_1 \rightarrow \tau_2 \\
  C  & = & (v_1, \ldots, v_{i-1}, \_, e_{i+1}, \ldots e_n) \sep v~\_ \sep \_~e \sep \pi_i~\_   
\end{array}
\end{equation*}

~
\paragraph{Reduction rules}
\begin{equation*}
\begin{array}{rll}
  \footnotesize{(\emph{fn})}  & (\lambda x.e)~v \rightarrow e[x\leftarrow v] \\[0.25em]
  \footnotesize{(\emph{prj})} & \pi_i (v_1, \ldots, v_n) \rightarrow v_i\\[0.25em]
  \footnotesize{(\emph{ctx})}  & C[e] \rightarrow C[e']  & (\textnormal{when}~e\rightarrow e')\\[0.25em]
\end{array}
\end{equation*}

~
\paragraph{Typing rules}
\begin{equation*}
\begin{array}{l}
\tyrule{var}
  { x:\tau \in \Gamma }
  { \Gamma \vdash x:\tau }
\\[1.25em]
\tyrule{num}
  { ~ }
  { \Gamma \vdash n:\ident{num} }
\\[1.25em]
\tyrule{abs}
  { \Gamma, x\!:\!\tau_1 \vdash e : \tau_2 }
  { \Gamma \vdash \lambda x.e : \tau_1 \rightarrow \tau_2 }
\\[1.25em]
\tyrule{app}
  { \Gamma \vdash e_1 : \tau_1 \rightarrow \tau_2  & \Gamma \vdash e_2 : \tau_1 }
  { \Gamma \vdash e_1~e_2 : \tau_2 }
\\[1.25em]
\tyrule{proj}
  { \Gamma \vdash e : \tau_1 \times \ldots \tau_i \times \ldots \times \tau_n }
  { \Gamma \vdash \pi_i~e:\tau_i }
\\[1.25em]
\tyrule{tup}
  { \forall i\in\{1 \ldots n\}.~ \Gamma \vdash e_i : \tau_i }
  { \Gamma \vdash (e_1, \ldots, e_n) : \tau_1 \times \ldots \times \tau_n }
\end{array}
\end{equation*}

\caption{Common syntax and reduction rules of the target langauge}
\label{fig:semantics-target}
\end{figure}

% --------------------------------------------------------------------------------------------------

\section{Embedding coeffect languages}
\label{sec:semantics-translation}

Although the notion of indexed comonads presented in the previous section is novel and interesting 
in its own, the main reason for introducing it is that we can view it as a translation that provides 
embedding of context-aware domain-specific languages in a simple target functional language.
In this section, we follow the example of effects and monads and we use the semantics to define a 
translation akin to the do notation in Haskell.

A context-aware \emph{source} program written using a concrete context-aware domain-specific 
language (capturing dataflow, implicit parameters or other kinds of context awareness) with
domain-specific language extensions (the \kvd{prev} keyword, or the \ident{?impl} syntax)
is translated to a \emph{target} langauge that is not context-aware. The target language is
a small functional language consisting of:
%
\begin{itemize}
  \item Simple functional subset formed by lambda calculus with support for tuples and numbers.
  \item Comonadically-inspired primitives corresponding to \emph{counit}, \emph{cobind} and
    other operations of flat indexed comonads.
  \item Additional primitives that model contextual operations of each concrete coeffect language
    (\emph{prev} for the \kvd{prev} keyword, \emph{lookup} for the \ident{?p} syntax and
    \emph{letimpl} for the $\kvd{let}~\ident{?p}=\ldots$ notation).
\end{itemize}
%
The syntx, typing and reduction rules of the first part (simple functional language) are used by 
all concrete coeffect domain-specific langauges. The syntax and typing rules of the second part 
(comonadically-inspired) primitives are also shared by all coeffect DSLs, however the \emph{reduction
rules} for the comonadically-inspired primitives differ -- they capture the concrete notion of
context. Finally, the third part (domain-specific primitives) will differ for each coeffect DSL.

% --------------------------------------------------------------------------------------------------

\subsection{Functional target langauge}
\label{sec:semantics-translation-target}

The target langauge for the translation is a simply typed lambda calculus with integers and tuples.
We include integers as an example of a concrete type. Tuples are needed by the translation, which
keeps a tuple of variable assignments. Encoding those without tuples would be possible, but 
cumbersome. In this section, we define the common parts of the language without the 
comonadically-inspired primitives.

The syntax of the target programming langauge is shown in Figure~\ref{fig:semantics-target}. 
The values include numbers $n$, tuples and function values. The expressions include variables $x$, 
values, lambda abstraction and application and operations on tuples. We do not need recursion or 
other data types (although a realistic programming language would include them). 
In what follows, we also use the following syntactic sugar for let binding:
%
\begin{equation*}
\begin{array}{rcl}
  \kvd{let}~x=e_1~\kvd{in}~e_2 &=& (\lambda x.e_2)~e_1
\end{array}
\end{equation*}
%
Finally, $C[e]$ defines the context in which sub-expressions are evaluated. Together with the
evaluation rules shown in  Figure~\ref{fig:semantics-target}, this captures the standard 
call-by-name semantics of the common parts of the target language. The (standard) typing rules 
for the common expressions of the target langauge are also shown in Figure~\ref{fig:semantics-target}.

% --------------------------------------------------------------------------------------------------

\subsection{Safety of functional target language}

The functional subset of the langauge described so far models a simple ML-like langauge. We 
choose call-by-value over call-by-name for no particular reason and Haskell-like language would
work equally well.

The subset of the language introduced so far is type-safe in the standard sense that
``well-typed programs do not get stuck''. Although standard, we outline the important parts
of the proof for the functional subset here, before we extend it to concrete context-aware
languages in Section~\ref{sec:semantics-proofs}.

We use the standard syntactic approach to type safety introduced by Milner \cite{syntactic-sml}.
Following Wright, Felleisen and Pierce \cite{syntactic-tapl,syntactic-approach}, we prove the
type preservation property (reduction does not change the type of an expression) and the progress 
property (a well-typed expression is either a value or can be further reduced).

\begin{lemma}[Canonical forms]
\label{thm:semantics-fp-canon}
For all $e, \tau$, if $\vdash e : \tau$ and $e$ is a value then:
\begin{enumerate}
  \item If $\tau=\ident{num}$ then $e = n$ for some $n \in \mathbb{Z}$
  \item If $\tau=\tau_1 \rightarrow \tau_2$ then $e = \lambda x.e'$ for some $x, e'$
  \item If $\tau=\tau_1\times\ldots\times\tau_n$ then $e = (v_1, \ldots, v_n)$ for some $v_i$
\end{enumerate}  
\end{lemma}
\begin{proof}
  For (1), the last typing rule must have been (\emph{num}); for (2), it must have been
  (\emph{abs}) and for (3), the last typing rule must have been (\emph{tup})
\end{proof}

\begin{lemma}[Preservation under substitution]
\label{thm:semantics-fp-pres-subst}
For all $\Gamma, e, e' \tau, \tau'$, if $\Gamma, x:\tau \vdash e : \tau'$ and $\Gamma \vdash e' : \tau$ 
then $\Gamma \vdash e[x \leftarrow e'] : \tau$.
\end{lemma}
\begin{proof}
By induction over the derivation of $\Gamma, x:\tau \vdash e : \tau'$. 
\end{proof}

\begin{theorem}[Type preservation]
\label{thm:semantics-fp-pres}
  If $\Gamma \vdash e : \tau$ and $e \rightarrow e'$ then $\Gamma \vdash e' : \tau$
\end{theorem}
\begin{proof}
  Rule induction over $\rightarrow$.

\vspace{0.25em}\noindent\hangindent=0.6cm 
Case (\emph{fn}): $e = (\lambda x.e_0)~v$, from Lemma~\ref{thm:semantics-fp-pres-subst}
  it follows that $\Gamma \vdash e_0[x \leftarrow v] : \tau$.

\vspace{0.25em}\noindent\hangindent=0.6cm 
Case (\emph{prj}): $e = \pi_i(v_1, \ldots, v_n)$ and so the last applied typing rule must have been
  (\emph{tup}) and $\Gamma \vdash (v_1, \ldots, v_n) : \tau_1 \times\ldots\times \tau_n$ and
  $\tau = \tau_i$. After applyng (\emph{prj}) reduction, $e' = v_i$ and so $\Gamma \vdash e' : \tau_i$.

\vspace{0.25em}\noindent\hangindent=0.6cm 
Case (\emph{ctx}): By induction hypothesis, the type of the reduced sub-expression does not change
  and the last used rule in the derivation of $\Gamma \vdash e : \tau$ also applies on $e'$
  giving $\Gamma \vdash e' : \tau$.  
\end{proof}

\begin{theorem}[Progress]
\label{thm:semantics-fp-prog}  
  If $\vdash e : \tau$ then either $e$ is a value or there exits $e'$ such that $e \rightarrow e'$
\end{theorem}
\begin{proof}
  By rule induction over $\vdash$.

\vspace{0.25em}\noindent\hangindent=0.6cm 
Case (\emph{num}): $e = n$ for some $n$ and so $e$ is a value.

\vspace{0.25em}\noindent\hangindent=0.6cm 
Case (\emph{abs}): $e = \lambda x.e'$ for some $x, e'$, which is a value.

\vspace{0.25em}\noindent\hangindent=0.6cm 
Case (\emph{var}): This case cannot occur, because $e$ is a closed expression.

\vspace{0.25em}\noindent\hangindent=0.6cm 
Case (\emph{app}): $e = e_1~e_2$ which is not a value. By induction, $e_1$ is either
a value or it can reduce. If it can reduce, apply (\emph{ctx}) reduction with context $\_~e$. Otherwise
consider $e_2$. If it can reduce, apply (\emph{ctx}) with context $v~\_$. If both are values,
Lemma~\ref{thm:semantics-fp-canon} guarantees that $e_1 = \lambda x.e_1'$ and so we can apply
reduction (\emph{fn}).

\vspace{0.25em}\noindent\hangindent=0.6cm 
Case (\emph{proj}): $e = \pi_i e_0$ and $\tau=\tau_1\times\ldots\tau_n$. If $e_0$ can be reduced, 
  apply (\emph{ctx}) with context $\pi_i~\_$. Otherwise from Lemma~\ref{thm:semantics-fp-canon}, 
  we have that $e_0=(v_1, \ldots, v_n)$ and we can apply reduction (\emph{prj}).

\vspace{0.25em}\noindent\hangindent=0.6cm 
Case (\emph{tup}): $e = (e_1, \ldots, e_n)$. If all sub-expressions are values, then $e$ is also
  a value. Otherwise, we can apply reduction using (\emph{ctx}) with a context $(v_1, \ldots, v_{i-1}, \_, e_{i+1}, \ldots, e_n)$.
\end{proof}

\begin{theorem}[Safety]
  If $\Gamma \vdash e : \tau$ and $e \rightarrow^* e'$ then either $e'$ is a value or there
  exists $e''$ such that $e' \rightarrow e''$
\end{theorem}
\begin{proof}
  Rule induction over $\rightarrow^*$ using Theorem~\ref{thm:semantics-fp-pres} and Theorem~\ref{thm:semantics-fp-prog}.
\end{proof}

% --------------------------------------------------------------------------------------------------

\begin{figure}[t]
\paragraph{Language syntax}
Given $(\C, \cseq, \cpar, \czip, \cunit, \czero, \cleq)$, extend the syntax:

\begin{equation*}
\begin{array}{rcl}
  e &  =  &\ldots \sep\ident{cobind}_{\cclrd{s},\cclrd{r}}~e_1~e_2 \sep \ident{counit}_{\cunit}~e \sep \ident{merge}_{\cclrd{r},\cclrd{s}}~e 
         \sep \ident{split}_{\cclrd{r},\cclrd{s}}~e \sep \ident{lift}_{\cclrd{r},\cclrd{r'}}~e \\
  \tau &  = &\ldots \sep \ctyp{\cclrd{r}}{\tau} \\
  C  & = &\ldots \sep \ident{cobind}_{\cclrd{s},\cclrd{r}}~\_~e \sep \ident{cobind}_{\cclrd{s},\cclrd{r}}~v~\_ \sep \ident{counit}_{\cunit}~\_ \\
    && \hspace{1.4em} \sep \ident{merge}_{\cclrd{r},\cclrd{s}}~\_
           \sep \ident{split}_{\cclrd{r},\cclrd{s}}~\_ \sep \ident{lift}_{\cclrd{r},\cclrd{r'}}~\_ \\  
\end{array}
\end{equation*}

~

~

\paragraph{Typing rules}
Given $(\C, \cseq, \cpar, \czip, \cunit, \czero, \cleq)$, add the typing rules:

\begin{equation*}
\begin{array}{l}
\tyrule{counit}
  {\Gamma \vdash e : \ctyp{\cunit}{\tau}}
  {\Gamma \vdash \ident{counit}_{\cunit}~e : \tau} \\
\\
\tyrule{cobind}
  {\Gamma \vdash e_1 : \ctyp{\cclrd{r}}{\tau_1} \rightarrow \tau_2 & \Gamma \vdash e_2 : \ctyp{\cclrd{r}\cseq\cclrd{s}}{\tau_1} }
  {\Gamma \vdash \ident{cobind}_{\cclrd{r}, \cclrd{s}}~e_1~e_2 : \ctyp{\cclrd{s}}{\tau_2}} \\
  \\
\tyrule{merge}
  {\Gamma \vdash  e : \ctyp{\cclrd{r}}{\tau_1} \times \ctyp{\cclrd{s}}{\tau_2} } 
  {\Gamma \vdash  \ident{merge}_{\cclrd{r},\cclrd{s}}~e : \ctyp{\cclrd{r}\czip\cclrd{s}}{(\tau_1 \times \tau_2)} }\\
    \\
\tyrule{split}
  {\Gamma \vdash  e : \ctyp{\cclrd{r}\cpar\cclrd{s}}{(\tau_1 \times \tau_2)} }
  {\Gamma \vdash  \ident{split}_{\cclrd{r},\cclrd{s}}~e : \ctyp{\cclrd{r}}{\tau_1} \times \ctyp{\cclrd{s}}{\tau_2}} \\
\end{array}
\end{equation*}

\caption{Comonadically-inspired extensions for the target langauge}
\label{fig:semantics-ext}
\end{figure}

% --------------------------------------------------------------------------------------------------

\subsection{Comonadically-inspired translation}
\label{sec:semantics-translation-transl}

In Section~\ref{sec:semantics-theory}, we presented the semantics of the flat coeffect calculus in 
terms of indexed comonads. We treated the semantics as denotational -- interpreting the meaning of 
a given typing derivation of a program in terms of category theory.

In this chapter, we use the same structure in a different way. Rather than treating the rules as
\emph{denotation} in categorical sense, we treat them as \emph{translation} from a source 
domain-specific coeffect language into a target language with comonadically-inspired primitives
described in the previous section.

\paragraph{Language extension.}
Given a coeffect language with a flat coeffect algebra $(\C, \cseq, \cpar, \czip, \cunit, \czero, \cleq)$,
we first extend the language syntax and typing rules with terms that correspond to the 
comonadically-inspired operations. This is done in the same way for all concrete coeffect 
domain-specific languages and so we give the additional syntax, evaluation context and 
typing rules in Figure~\ref{fig:semantics-ext} for all languages we consider later in 
Section~\ref{sec:semantics-proofs}.

The new type $\ctyp{\cclrd{r}}$ represents an indexed comonad, which is left abstract for now.
The additional expressions such as $\ident{counit}_\cunit$ and $\ident{cobind}_{\cclrd{r,s}}$ 
correspond to the operations of indexed comonads. Note that the we embed the coeffect annotations 
into the target language -- these are known when translationg a typed term from a source language
and they will be useful when proving that sufficient context (as specified by the coeffect
annotations) is available.

The figure defines the syntax and the typing rules, but it does not define the reduction rules.
Those -- together with the values for a concrete notion of context -- will be defined separately
for each individual coeffect language.

\paragraph{Contexts and types.}
The interpretation of contexts and types in the category now becomes a translation from types and
contexts in the source language into the types of the target language:
%
\begin{equation*}
\begin{array}{rcl}
\sem{\coctx{x_1\!:\!\tau_1, \ldots, x_n\!:\!\tau_n}{\cclrd{r}}} &=& \ctyp{\cclrd{r}}{(\sem{\tau_1} \times \ldots \times \sem{\tau_n})} \\[0.5em]
\sem{\tau_1 \xrightarrow{\cclrd{r}} \tau_2} & = & \ctyp{\cclrd{r}}{\sem{\tau_1}} \rightarrow \sem{\tau_2} \\[-0.2em]
\sem{\ident{num}} & = & \ident{num} \\
\end{array}
\end{equation*}

\noindent
Here, a context becomes a comonadically-inspired data type wrapping a tuple of variable values 
and a coeffectful function is translated into an ordinary function in the target language
with a comonadically-inspired data type wrapping the input type.

% --------------------------------------------------------------------------------------------------

\begin{figure*}[t]

The translation is defined over a typing derivation:
%
\begin{equation*}
\hspace{-3em}
\begin{array}{ll}
\hspace{4.8em}\semdef
  {\coctx{\Gamma}{\cunit} \vdash x_i : \tau_i }
  {\lambda\ctx.\pi_i~(\ident{counit}_{\cunit}~\ctx)} 
& (\emph{var})
\\[1em]
\hspace{4.3em}\semdef
  {\coctx{\Gamma}{\czero} \vdash n : \ident{num} }
  {\lambda\ctx.n} 
& (\emph{num})
\\[1.5em]
\hspace{1.5em}\semdeff
  {\coctx{\Gamma, x_i:\tau_1}{\cclrd{r}\;\czip\;\cclrd{s}} \vdash e : \tau_2}
  {\hspace{-1em}\begin{array}{l}  
    \coctx{\Gamma}{\cclrd{r}} \vdash \lambda x_i.e : \tau_1 \xrightarrow{\cclrd{s}} \tau_2
    \\[-0.3em]~\\[-0.3em]~ \\[-0.3em]~
  \end{array} }
  {f}
  {\hspace{-1em}\begin{array}{l}  
    \lambda \ctx.\lambda v.\\[-0.3em]
    \quad \kvd{let}~\ident{assoc} = \lambda x.\\[-0.3em]
    \quad\quad (\pi_1(\pi_1~x), \ldots, \pi_{i-1}(\pi_1~x), \pi_2~x)\\[-0.3em]
    \quad f~(\ident{map}_{\cclrd{r}\czip\cclrd{s}}~\ident{assoc}~(\ident{merge}_{\cclrd{r}, \cclrd{s}}~(\ctx, v)))
   \end{array} }
& (\emph{abs})
\\[4em]
\semdefff
  {\coctx{\Gamma}{\cclrd{r}} \vdash e_1 : \tau_1 \xrightarrow{\cclrd{t}} \tau_2}
  {\coctx{\Gamma}{\cclrd{s}} \vdash e_2 : \tau_1 }
  {\begin{array}\coctx{\Gamma}{\cclrd{r} \;\cpar\; (\cclrd{s} \,\cseq\, \cclrd{t})} \vdash e_1~e_2 : \tau_2\\[-0.3em]~\\[-0.3em]~\\[-0.3em]~\end{array}}
  {f}
  {g}
  {\hspace{-1em}\begin{array}{l}  
  \lambda\ctx.~~\\[-0.3em]
    \quad \kvd{let}~\ctx_0 = \ident{map}_{\cclrd{r} \;\cpar\; (\cclrd{s} \,\cseq\, \cclrd{t})}~\ident{dup}~\ctx\\[-0.3em]
    \quad \kvd{let}~(\ctx_1, \ctx_2) = \ident{split}_{\cclrd{r}, \cclrd{s} \,\cseq\, \cclrd{t}}~\ctx_0\\[-0.3em]
    \quad f~\ctx_1~
      (\ident{cobind}_{\cclrd{s}, \cclrd{t}}~g~\ctx_2)
  \end{array} }   
& (\emph{app})
\end{array}
\end{equation*}

~

~

Assuming the following auxiliary operations:
%
\begin{equation*}
\begin{array}{rcl}
  \ident{map}_\cclrd{r}}~f &=& \ident{cobind}_{\cunit, \cclrd{r}}~(\lambda x.f~({\ident{counit}_\cunit}~x)) \\[-0.3em]
  \ident{dup} &=& \lambda x. (x, x)
\end{array}
\end{equation*}

\figcaption{Translation from a flat DSL to a comonadically-inspired target language}
\label{fig:semantics-translation}
\end{figure*}

% --------------------------------------------------------------------------------------------------

\paragraph{Expressions.}
The rules shown in Figure~\ref{fig:semantics-translation} define how expressions of the source
language are translated into the target language. The rules are very similar to those shown earlier
in Figure~\ref{fig:semantics-flat}. The result is now written as source code in the
target programming language rather than as composition of morphisms in a category. However,
thanks to the equivalence between $\lambda$-calculus and category theory, both interpretations 
are equivalent.

One change from Figure~\ref{fig:semantics-flat} is that we are now more explicit about the tuple
that contains variable assignments. Previously, we assumed that the tuple is appropriately 
reassociated. Now, we perform the reassociation explicitly. We keep a flat tuple of variables, so
given $\Gamma=x_1\!:\!\tau_1,\ldots,x_n\!:\!\tau_n$, the tuple has a type $\tau_1\times\ldots\times\tau_n$.
In (\emph{var}), we access a variable using $\pi$, but in (\emph{abs}), the \ident{merge} operation
produces a tuple $(\tau_1\times\ldots\times\tau_{i-1})\times\tau_i$ that we turn into a flat
tuple $\tau_1\times\ldots\times\tau_{i-1}\times\tau_i$ using the \ident{assoc} function.

\paragraph{Properties.} 
The most important property of the translation is that it produces well-typed programs in the
target language. This is akin to the correspondence property of the semantics discussed earlier
(Theorem~\ref{thm:flat-correspondence}), but now it has more obvious practical consequences.

In Section~\ref{sec:semantics-proofs}, we will prove safety properties of well-typed programs in the 
target language. Thanks to the fact that the translation produces a well-typed program (as complex
as it might be) means that we are also proving safety of well-typed programs in the source 
context aware languages.

\begin{theorem}[Well-typedness of the translation]
\label{thm:semantics-welltyped}
Given a typing derivation for a well-typed closed expression $\coctx{}{\cclrd{r}} \vdash e : \tau$
written in a context-aware programming langugae that is translated to the target langauge as:
%  
\begin{equation*}  
  \semdeff
    {(\ldots)\qquad~}
    {\coctx{\Gamma}{\cclrd{r}} \vdash e : \tau }
    {(\ldots)}
    { ~~f }
\end{equation*}  
%
Then $f$ is well-typed, \ie~in the target langauge: $\vdash f : \sem{\coctx{\Gamma}{\cclrd{r}}} \rightarrow \sem{\tau}$.
\end{theorem}
\begin{proof}
By rule induction over the derivation of the translation. Given a judgement
$\coctx{x_1\!:\!\tau_1 \ldots x_n\!:\!\tau_n}{\cclrd{c}} \vdash e : \tau$, the translation
constructs a function of type $\ctyp{\cclrd{c}}(\sem{\tau_1} \times\ldots\times \sem{\tau_n}) \rightarrow \sem{\tau}$.

\vspace{0.5em}\noindent\hangindent=0.6cm 
Case (\emph{var}): $\cclrd{c}=\cunit$ and $\tau = \tau_i$ and so $\pi_i ({\ident{counit}_\cunit}~\ctx)$ is well-typed.

\vspace{0.25em}\noindent\hangindent=0.6cm 
Case (\emph{num}): $\tau = \ident{num}$ and so the body $n$ is well-typed.

\vspace{0.25em}\noindent\hangindent=0.6cm 
Case (\emph{abs}): The type of $\ctx$ is $\ctyp{\cclrd{r}}(\ldots)$ and the type of $v$ is $\ctyp{\cclrd{s}}\tau_1$,
  calling $\ident{merge}_{\cclrd{r},\cclrd{s}}$ and reassociating produces
  $\ctyp{\cclrd{r}\czip\cclrd{s}}(\ldots)$ as expected by $f$.

\vspace{0.25em}\noindent\hangindent=0.6cm 
Case (\emph{app}): After applying $\ident{split}_{\cclrd{r}, \cclrd{s}\cseq\cclrd{t}}$, the types of 
  $\ctx_1, \ctx_2$ are $\ctyp{\cclrd{r}}(\ldots)$ and $\ctyp{\cclrd{s}\cseq\cclrd{t}}(\ldots)$, respectively.
  $g$ requires $\ctyp{\cclrd{s}}(\ldots)$ and so the result of $\ident{cobind}_{\cclrd{s},\cclrd{t}}$ is
  $\ctyp{\cclrd{t}}\tau_1$ as required by $f$.
\end{proof}

% ==================================================================================================
%
%       #####
%      #     #   ##   ###### ###### ##### #   #
%      #        #  #  #      #        #    # #
%       #####  #    # #####  #####    #     #
%            # ###### #      #        #     #
%      #     # #    # #      #        #     #
%       #####  #    # #      ######   #     #
%
% ==================================================================================================

\section{Safety of context-aware languages}
\label{sec:semantics-proofs}

The previous two languages provide a general structrue that we now use to prove the safety of 
concrete context-aware programming languages based on the coeffect language framework. We
consider a langauge for dataflow computations (Section~\ref{sec:semantics-proofs-df}) and
for implicit parameters (Section~\ref{sec:semantics-proofs-impl}). In both cases, we extend
the progress and preservation theorems of the functional subset of the target language, but the
approach can be generalised as discussed in Section~\ref{sec:semantics-generalising}.

% --------------------------------------------------------------------------------------------------

\subsection{Coeffect langauge for dataflow}
\label{sec:semantics-proofs-df}

The types of the comonadically-inspired operations are the same for each concrete coeffect DSL, but
each DSL introduces its own \emph{values} of type $\ctyp{\cclrd{r}}\tau$ and also its own reduction
rules that define how comonadically-inspired operations evaluate.

We first consider dataflow computations. As discussed earlier in the
semantics of dataflow, the indexed comonad for a context with $n$ past values carries $n+1$ 
values. When reducing translated programs, the comonadic values will not be directly manipulated by the
user code. In a programming language, it could be seen as an \emph{abstract data type} whose only
operations are the comonadically-inspired ones defined earlier, together with an additional 
\emph{domain-specific} operation that models the \kvd{prev} construct.

The Figure~\ref{fig:semantics-ext-df} extensionds the target language with syntax, typing rules,
additional translation rule and reductions for modelling dataflow computations. We introduce a 
new kind of values written as $\ident{Df}\langle v_0, \ldots, v_n\rangle$
and a matching kind of expressions. We specify how the \kvd{prev} keyword is translated into a
$\ident{prev}_\cclrd{r}$ operation of the target language nd we also add a typing rule
(\emph{df}) that checks the types of the elements of the stream and also guarantees 
that the number of elements in the stream matches the number in the coeffect.
The additional reduction rules mirror the semantics that we discussed earlier when talking about
indexed dataflow comonad.

% --------------------------------------------------------------------------------------------------

\begin{figure}[t]
\paragraph{Language syntax}

\begin{equation*}
\begin{array}{rcl}
v &=& \ldots \sep \ident{Df}\langle v_0, \ldots, v_n\rangle \\
e &=& \ldots \sep \ident{Df}\langle e_0, \ldots, e_n\rangle \sep \ident{prev}_{\cclrd{n}}~e\\
C &=& \ldots \sep \ident{prev}_{\cclrd{n}}~\_ \sep \ident{Df}\langle v_0, \ldots, v_{i-1}, \_, e_{i+1} \ldots, e_n\rangle
\end{array}  
\end{equation*}

\vspace{1.25em}
\paragraph{Typing rules}

\begin{equation*}
\tyrule{df}
  { \forall i\in\{ 0 \ldots n \}. ~ \Gamma \vdash e_i : \tau}
  { \Gamma \vdash \ident{Df}\langle e_0, \ldots, e_n \rangle : \ctyp{\cclrd{n}}{\tau} }
\end{equation*}
\begin{equation*}
\tyrule{prev}
  { \Gamma \vdash e : \ctyp{\cclrd{n+1}}{\tau} }
  { \Gamma \vdash \ident{prev}_{\cclrd{n}}~e : \ctyp{\cclrd{n}}{\tau} }
\end{equation*}

\vspace{1.25em}
\paragraph{Translation}

\begin{equation*}
\semdeff
  { \coctx{\Gamma}{\cclrd{n}} \vdash e : \tau }
  { \coctx{\Gamma}{\cclrd{n+1}} \vdash \kvd{prev}~e : \tau }
  {f}
  {\lambda ctx.\ident{prev}_{\cclrd{n}}~\ctx }
\end{equation*}

\vspace{1.25em}
\paragraph{Reduction rules}

\begin{equation*}
\begin{array}{rl}
{\footnotesize(\emph{counit})} & \begin{array}{l}
\ident{counit}_{\cclrd{0}}(\ident{Df} \langle v_0 \rangle)   \rightarrow v_0
\end{array}
\\[0.5em]
{\footnotesize(\emph{cobind})} & \begin{array}{l}
\ident{cobind}_{\cclrd{m}, \cclrd{n}}~f~(\ident{Df}\langle v_0, \ldots v_{m+n} \rangle) \rightarrow
\\[-0.25em]
\quad (\ident{Df}\langle f (\ident{Df}\langle v_0, \ldots, v_m \rangle), \ldots, f (\ident{Df}\langle v_{n}, \ldots, v_{m+n} \rangle) \rangle)
\end{array}
\\[0.5em]
{\footnotesize(\emph{merge})} & \begin{array}{l}
\ident{merge}_{\cclrd{m}, \cclrd{n}} ((\ident{Df}\langle v_0, \ldots, v_m \rangle), (\ident{Df}\langle v'_0, \ldots, v'_n\rangle)) \rightarrow
\\[-0.25em]
\quad (\ident{Df}\langle (v_0, b_0), \ldots, (v_{\mathit{min}(m,n)}, v'_{\mathit{min}(m,n)}) \rangle)
\end{array}
\\[0.5em]
{\footnotesize(\emph{split})} & \begin{array}{l}
\ident{split}_{\cclrd{m}, \cclrd{n}} (\ident{Df}\langle (v_0, b_0), \ldots, (v_{\mathit{max}(m,n)}, b_{\mathit{max}(m,n)}) \rangle)  \rightarrow
\\[-0.25em]
\quad \ident{Df}\langle v_0, \ldots, v_m \rangle, (\ident{Df}\langle b_0, \ldots, b_n\rangle
\end{array}
\\[0.5em]
{\footnotesize(\emph{prev})} & \begin{array}{l}
\ident{prev}_{\cclrd{n}} (\ident{Df}\langle v_0, \ldots, v_n, v_{n+1} \rangle) \rightarrow
\\[-0.25em]
\quad \ident{Df}\langle v_0, \ldots, v_n \rangle
\end{array}
\end{array}
\end{equation*}

\caption{Additional constructs for modelling dataflow in the target langauge}
\label{fig:semantics-ext-df}
\end{figure}

% --------------------------------------------------------------------------------------------------

\paragraph{Properties.}
Now consider a target langauge consisting of the core (ML-subset) defined by the syntax,
reduction rules and typing rules given in Figure~\ref{fig:semantics-target} and comonadically-inspired 
primtives defined in Figure~\ref{fig:semantics-ext} and also concrete notion of comonadically-inspired 
value and reduction rules for dataflow as defined in Figure~\ref{fig:semantics-ext-df}.

In order to prove type safety, we first extend the canonical forms lemma
(Lemma~\ref{thm:semantics-fp-canon}) and the preservation under substitution lemma
(Lemma~\ref{thm:semantics-fp-pres-subst}). Those need to consider the new (\emph{df}) 
and (\emph{prev}) typing rules and substitution under the newly introduced expression forms 
$\ident{Df}\langle\ldots\rangle$ and $\ident{prev}_{\cclrd{n}}$. We show that the translation 
rule for \kvd{prev} produces well-typed expressions. Finally, we extend the type preservation 
(Theorem~\ref{thm:semantics-fp-pres}) and progress (Theorem~\ref{thm:semantics-fp-prog}).

\begin{theorem}[Well-typedness of the \kvd{prev} translation]
\label{thm:semantics-df-welltyped}
Given a typing derivation for a well-typed closed expression $\coctx{}{\cclrd{r}} \vdash e : \tau$,
the translated program $f$ obtained using the rules in Figure~\ref{fig:semantics-translation}
and Figure~\ref{fig:semantics-ext-df} is well-typed, \ie~in the target langauge: $\vdash f : \sem{\coctx{\Gamma}{\cclrd{r}}} \rightarrow \sem{\tau}$.
\end{theorem}
\begin{proof}
By rule induction over the derivation of the translation.

\vspace{0.5em}\noindent\hangindent=0.6cm 
Case (\emph{var, num, abs, app}): As before.

\vspace{0.5em}\noindent\hangindent=0.6cm 
Case (\emph{prev}): Type of $\ctx$ is $\ctyp{\cclrd{n+1}}\tau$ and so we can apply the (\emph{prev}) rule.
\end{proof}

\begin{lemma}[Canonical forms]
\label{thm:semantics-df-canon}
For all $e, \tau$, if $\vdash e : \tau$ and $e$ is a value then:
\begin{enumerate}
  \item If $\tau=\ident{num}$ then $e = n$ for some $n \in \mathbb{Z}$
  \item If $\tau=\tau_1 \rightarrow \tau_2$ then $e = \lambda x.e'$ for some $x, e'$
  \item If $\tau=\tau_1\times\ldots\times\tau_n$ then $e = (v_1, \ldots, v_n)$ for some $v_i$
  \item If $\tau=\ctyp{\cclrd{n}}\tau_1$ then $e = \ident{Df}\langle v_0, \ldots v_n\rangle$ for some $v_i$
\end{enumerate}  
\end{lemma}
\begin{proof}
  (1,2,3) as before; for (4) the last typing rule must have been (\emph{df}).
\end{proof}

\begin{lemma}[Preservation under substitution]
\label{thm:semantics-df-pres-subst}
For all $\Gamma, e, e' \tau, \tau'$, if $\Gamma, x:\tau \vdash e : \tau'$ and $\Gamma \vdash e' : \tau$ 
then $\Gamma \vdash e[x \leftarrow e'] : \tau$.
\end{lemma}
\begin{proof}
  By induction over the derivation of $\Gamma, x:\tau \vdash e : \tau'$ as before, with new
  cases for $\ident{Df}\langle\ldots\rangle$ and $\ident{prev}_{\cclrd{n}}$.
\end{proof}

\begin{theorem}[Type preservation]
\label{thm:semantics-df-pres}
  If $\Gamma \vdash e : \tau$ and $e \rightarrow e'$ then $\Gamma \vdash e' : \tau$
\end{theorem}
\begin{proof}
  Rule induction over $\rightarrow$.

\vspace{0.25em}\noindent\hangindent=0.6cm 
Case (\emph{fn}, \emph{prj}, \emph{ctx}): As before, using Lemma~\ref{thm:semantics-df-pres-subst} for (\emph{fn}).

\vspace{0.25em}\noindent\hangindent=0.6cm 
Case (\emph{counit}): $e = \ident{counit}_{\cclrd{0}}(\ident{Df} \langle v_0 \rangle)$. The last rule in the type
  derivation of $e$ must have been (\emph{counit}) with $\Gamma \vdash \ident{Df} \langle v_0 \rangle : \ctyp{\cclrd{0}}\tau$
  and therefore $\Gamma \vdash v_0 : \tau$.

\vspace{0.25em}\noindent\hangindent=0.6cm 
Case (\emph{cobind}): $e = \ident{cobind}_{\cclrd{m}, \cclrd{n}}~f~(\ident{Df}\langle v_0, \ldots v_{m+n} \rangle)$. The
  last rule in the type derivation of $e$ must have been (\emph{cobind}) with a type $\tau = \ctyp{\cclrd{n}}\tau_2$ and assumptions
  $\Gamma \vdash f : \ctyp{\cclrd{m}}{\tau_1} \rightarrow \tau_2$ and $\Gamma \vdash \ident{Df}\langle v_0, \ldots v_{m+n} \rangle : \ctyp{\cclrd{m+n}}{\tau}$.
  The reduced expression has a type $\ctyp{\cclrd{n}}\tau_2$:
\begin{equation*}
\inference
  { \inference 
      { \Gamma \vdash f : \ctyp{\cclrd{m}}{\tau_1} \rightarrow \tau_2 & \forall i\in 0 \ldots n.~\Gamma\vdash \ident{Df}\langle v_i, \ldots, v_{i+m} \rangle :  \ctyp{\cclrd{m}}{\tau_1}}  
      { \forall i\in 0 \ldots n.~\Gamma\vdash f (\ident{Df}\langle v_i, \ldots, v_{i+m} \rangle) : \tau_2} }
  { \Gamma \vdash \ident{Df}\langle f (\ident{Df}\langle v_0, \ldots, v_m \rangle), \ldots, f (\ident{Df}\langle v_{n}, \ldots, v_{m+n} \rangle) \rangle : \ctyp{\cclrd{n}}\tau_2 }
\end{equation*}

\vspace{0.25em}\noindent\hangindent=0.6cm 
Case (\emph{merge}, \emph{split}, \emph{next}): Similar. In all three cases, the last typing rule in the derivation of $e$
  guarantees that the stream contains a sufficient number of elements of correct type.
\end{proof}

\begin{theorem}[Progress]
\label{thm:semantics-df-prog}  
  If $\vdash e : \tau$ then either $e$ is a value or there exits $e'$ such that $e \rightarrow e'$
\end{theorem}
\begin{proof}
  By rule induction over $\vdash$.

\vspace{0.25em}\noindent\hangindent=0.6cm 
Case (\emph{num,abs,var,app,proj,tup}): As before, using the adapted canonical forms lemma
  (Lemma~\ref{thm:semantics-df-canon}) for (\emph{app}) and (\emph{proj}).

\vspace{0.25em}\noindent\hangindent=0.6cm 
Case (\emph{counit}): $e = {\ident{counit}_\cunit}~e_1$. If $e_1$ is not a value, it can be reduced 
  using (\emph{ctx}) with context ${\ident{counit}_\cunit}~\_$, otherwise it is a value. From Lemma~\ref{thm:semantics-df-canon},
  $e_1 = \ident{Df}\langle v\rangle$ and so we can apply (\emph{counit}) reduction rule.

\vspace{0.25em}\noindent\hangindent=0.6cm 
Case (\emph{cobind}): $e = \ident{cobind}_{\cclrd{m}, \cclrd{n}}~e_1~e_2$. If $e_1$ is not a value,
  reduce using (\emph{ctx}) with context ${\ident{cobind}_{\cclrd{m},\cclrd{n}}}~\_~e$. If $e_2$ is
  not a value reduce using (\emph{ctx}) with context ${\ident{cobind}_{\cclrd{m},\cclrd{n}}}~v~\_$.
  If both are values, then from Lemma~\ref{thm:semantics-df-canon}, we have that 
  $e_2 = \ident{Df}\langle v_0, \ldots v_{m+n} \rangle$ and so we can apply the (\emph{cobind})
  reduction.

\vspace{0.25em}\noindent\hangindent=0.6cm 
Case (\emph{merge}): $e = \ident{merge}_{\cclrd{m}, \cclrd{n}} e_1$. If $e_1$ is not a value, 
  reduce using (\emph{ctx}) with context $e = \ident{merge}_{\cclrd{m}, \cclrd{n}}~\_$. If $e_1$ is
  a value, it must be a pair of streams $(\ident{Df}\langle v_0, \ldots, v_m \rangle, \ident{Df}\langle v'_0, \ldots, v'_n\rangle)$
  using Lemma~\ref{thm:semantics-df-canon} and it can reduce using (\emph{merge}) reduction.

\vspace{0.25em}\noindent\hangindent=0.6cm 
Case (\emph{df}): $e = \ident{Df}\langle e_0, \ldots, e_n\rangle$. If $e_i$ is not a value
  then reduce using (\emph{ctx}) with context $\ident{Df}\langle v_0, \ldots, v_{i-1}, \_, e_{i+1} \ldots, e_n\rangle$.
  Otherwise, $e_0, \ldots, e_n$ are values and so $\ident{Df}\langle e_0, \ldots, e_n\rangle$
  is also a value.

\vspace{0.25em}\noindent\hangindent=0.6cm 
Case (\emph{split}, \emph{prev}): Similar. Either sub-expression is not a value, or the type 
  guarantees that it is a stream with correct number of elements to enable the (\emph{split}) 
  or (\emph{prev}) reduction, respectively.
\end{proof}

\begin{theorem}[Safety]
  If $\Gamma \vdash e : \tau$ and $e \rightarrow^* e'$ then either $e'$ is a value or there
  exists $e''$ such that $e' \rightarrow e''$
\end{theorem}
\begin{proof}
  Rule induction over $\rightarrow^*$ using Theorem~\ref{thm:semantics-df-pres} and Theorem~\ref{thm:semantics-df-prog}.
\end{proof}

% --------------------------------------------------------------------------------------------------

\subsection{Coeffect langauge for implicit parameters}
\label{sec:semantics-proofs-impl}

As discussed earlier (Example~\ref{thm:semantics-indexed-prod}), implicit parameters can be modelled 
by an indexed product comonad, which annotates a value with additional context -- in our case, a 
mapping from implicit parameter names to their values. In this section, we embed this model into
the target language.

As with dataflow computations, we take the core functional subset (Figure~\ref{fig:semantics-target}) 
with comonadically-inspired extensions (Figure~\ref{fig:semantics-ext}) and we specify a new kind
of values of type $\ctyp{\cclrd{r}}\tau$ and domain-specific reduction rules that specify how
the operations propagate and access the context containing implicit parameter bindings.
Again, the $\ctyp{\cclrd{r}}\tau$ values can be seen as an \emph{abstract data type}, which are 
never manipulated directly.

\paragraph{Domain-specific extensions.}

The Figure~\ref{fig:semantics-ext-impl} shows extensions to the target language for modelling
implicit parameters. A comonadic value with a coeffect $\cclrd{ \{ \ident{?p}_1, \ldots, \ident{?p}_n \} }}$
is modelled by a new kind of value written as 
$\ident{Impl}(v, \{ \ident{?p}_1 \mapsto v_1, \ldots \ident{?p}_n \mapsto v_n \})$ which contains
a value $v$ together with implicit parameter assignments for all the parameters specified in the
coeffect. We add a corresponding kind of expression with its typing rule (\emph{impl}).

There are also two domain-specific operations for working with implicit parameters. The
$\ident{lookup}_{\cclrd{\ident{?p}}}$ operation reads a value of an implicit parameter and
the $\ident{letimpl}_{\cclrd{\ident{?p}}, \cclrd{r}}$ operation adds a mapping assigning a value
to an implicit parameter \ident{?p}. The typing rule (\emph{lookup}) specifies that the accessed
parameter need to be a part of the context and the rule (\emph{letimpl}) specifies that the
\ident{letimpl} operation extends the context with a new implicit parameter binding.

The new translation rules specify how implicit parameter access, written as \ident{?p}, and
implicit parameter binding, written as $\kvd{let}~\ident{?p}~=e_1~\kvd{in}~e_2$ are translated to the
target language. The first one is straightforward. The binding is similar to the translation for
function application -- we split the context, evaluate $e_1$ using the first part of the context
$\ctx_1$ and then add the new binding to the remaining context $\ctx_2$.

Finally, Figure~\ref{fig:semantics-ext-impl} also defines the reduction rules. The (\emph{lookup})
rule accesses an implicit parameter and (\emph{letimpl}) adds a new binding. The reduction rules 
for comonadically-inspired primitives closely model the product comonad discussed in 
Example~\ref{thm:semantics-indexed-prod}. Reductions for (\emph{cobind}) and (\emph{split}) 
restrict the set of available implicit parameters according to the annotations and 
(\emph{merge}) combines them, preferring the values from the call-site.

For the semantics of implicit parameter programs that we consider, the preference of call-site
bindings over declaration-site bindings in (\emph{merge}) does not matter. The unique typing 
derivations for implicit parameter coeffects obtained in Section~\ref{sec:flat-unique} always
split implicit parameters into \emph{disjoint sets}.

% --------------------------------------------------------------------------------------------------

\begin{figure}[h!]  
\paragraph{Language syntax}

\begin{equation*}
\begin{array}{rcl}
v &=& \ldots \sep \ident{Impl}(v, \{ \ident{?p}_1\mapsto v_1, \ldots, \ident{?p}_n\mapsto v_n\}) \\
e &=& \ldots \sep \ident{Impl}(e, \{\ident{?p}_1\mapsto e_1, \ldots, \ident{?p}_n\mapsto e_n\}) \\
  && \hspace{1.4em} \sep \ident{lookup}_{\cclrd{\ident{?p}}}~e \sep \ident{letimpl}_{\cclrd{\ident{?p}}, \cclrd{r}}~e_1~e_2\\
C &=& \ldots \sep \ident{lookup}_{\cclrd{\ident{?p}}}~\_ \sep \ident{letimpl}_{\cclrd{\ident{?p}}, \cclrd{r}}~\_~e \sep \ident{letimpl}_{\cclrd{\ident{?p}}, \cclrd{r}}~v~\_ \\
  && \hspace{1.4em} \sep \ident{Impl}(\_, \{ \ident{?p}_1\mapsto e_1, .., \ident{?p}_n\mapsto e_n\})\\
  && \hspace{1.4em} \sep \ident{Impl}(v, \{ \ident{?p}_1 \mapsto v_1, .., \ident{?p}_{i-1} \mapsto v_{i-1}, \ident{?p}_i \mapsto \_, \ident{?p}_{i+1} \mapsto v_{i+1}, .. \ident{?p}_n \mapsto e_n\})
\end{array}  
\end{equation*}

\vspace{1em}
\paragraph{Typing rules}

\begin{equation*}
  \hspace{-2em}
\tyrule{impl}
  { \Gamma \vdash e:\tau & \forall i\in\{ 1 \ldots n \}. ~ \Gamma \vdash e_i : \ident{num}}
  { \Gamma \vdash \ident{Impl}(e, \{ \ident{?p}_1\mapsto e_1, \ldots, \ident{?p}_n\mapsto e_n \}) : \ctyp{\cclrd{ \{ \ident{?p}_1, \ldots, \ident{?p}_n \}}}{\tau} }
\end{equation*}
\begin{equation*}
  \hspace{-2em}
\tyrule{lookup}
  { \Gamma \vdash e : \ctyp{\cclrd{ \{\ident{?p}\} }}{\tau} }
  { \Gamma \vdash \ident{lookup}_{\cclrd{\ident{?p}}}~e : \ident{num} }
\end{equation*}
\begin{equation*}
  \hspace{-2em}
\tyrule{letimpl}
  { \Gamma \vdash e_1 : \ident{num} &
    \Gamma \vdash e_2 : \ctyp{\cclrd{ \{ \ident{?p}_1, \ldots, \ident{?p}_n \} }}{\tau} }
  { \Gamma \vdash \ident{letimpl}_{\cclrd{\ident{?p}}, \cclrd{ \{ \ident{?p}_1, \ldots, \ident{?p}_n \} }}~e_1~e_2 : 
      \ctyp{\cclrd{ \{ \ident{?p}_1, \ldots, \ident{?p}_n, \ident{?p} \} }}{\tau} }
\end{equation*}

\vspace{1em}
\paragraph{Translation}

\begin{equation*}
\begin{array}{rl}  
(\footnotesize{\emph{lookup}}) &
\semdef
  { \coctx{\Gamma}{\cclrd{ \{\ident{?p}\} }} \vdash \ident{?p} : \ident{num} }
  {\lambda ctx.\ident{lookup}_{\cclrd{\ident{?p}}}~\ctx }
\\[2em]
\begin{array}{l}(\footnotesize{\emph{letimpl}})\\[1.25em]~\end{array}&
\semdefff
  { \coctx{\Gamma}{\cclrd{r}} \vdash e_1 : \tau_1}
  { \coctx{\Gamma}{\cclrd{s}} \vdash e_2 : \tau_2}
  {\hspace{-1em}\begin{array}{l}  
      \coctx{\Gamma}{\cclrd{r} \;\cclrd{\cup\; (\cclrd{s} \setminus \{ \ident{?p}) \}} }\\[-0.3em]
        \quad \vdash \kvd{let}~\ident{?p}=e_1\\[-0.3em]
        \quad\qquad \kvd{in}~e_2 : \tau_2 \\[-0.3em]
        ~ \\[-0.3em]
   \end{array}}
  {f}
  {g}
  {\hspace{-1em}\begin{array}{l}  
  \lambda\ctx.~~\\[-0.3em]
    \quad \kvd{let}~\ctx_0 = \ident{map}_{\cclrd{r} \;\cclrd{\cup}\; \cclrd{(s\setminus\{\ident{?p}\})}}~\ident{dup}~\ctx\\[-0.3em]
    \quad \kvd{let}~(\ctx_1, \ctx_2) = \ident{split}_{\cclrd{r}, \cclrd{(s\setminus\{\ident{?p}\})}}~\ctx_0\\[-0.3em]
    \quad g~(\ident{letimpl}_{\cclrd{\ident{?p}}, \cclrd{(s\setminus\{\ident{?p}\})}}~(f~\ctx_1)~\ctx_2)
  \end{array} }   
\end{array}
\end{equation*}

\vspace{1em}
\paragraph{Reduction rules}

\begin{equation*}
\begin{array}{rl}
{\footnotesize(\emph{counit})} & \begin{array}{l}
\ident{counit}_{\cclrd{ \emptyset }}~(\ident{Impl} (v, \ldots))   \rightarrow v
\end{array}
\\[0.5em]
{\footnotesize(\emph{cobind})} & \begin{array}{l}
\ident{cobind}_{\cclrd{r}, \cclrd{s}}~f~(\ident{Impl}(v, \{ \ident{?p}_1 \mapsto v_1, \ldots, \ident{?p}_{n}\mapsto v_n \})) \rightarrow
\\[-0.25em]
\quad \ident{Impl}(f~(\ident{Impl}(v, \{ \ident{?p}_i \mapsto v_i \;|\; \ident{p}_i \in \cclrd{r} \} )), \{ \ident{?p}_i \mapsto v_i \;|\; \ident{p}_i \in \cclrd{s} \} )
\end{array}
\\[1em]
{\footnotesize(\emph{merge})} & \begin{array}{l}
\ident{merge}_{\cclrd{r}, \cclrd{s}} 
              (~\ident{Impl}(v, \{ \ident{?p}_1 \mapsto v_1, \ldots, \ident{?p}_{n}\mapsto v_n \})  \\[-0.25em]
\hspace{4.4em}    \ident{Impl}(v', \{ \ident{?p}'_1 \mapsto v'_1, \ldots, \ident{?p}'_{n}\mapsto v'_n \})~) \rightarrow \\[-0.25em]
\quad \ident{Impl}((v, v'), \{ \ident{?p}_i \mapsto v_i \;|\; \ident{?p}\in\cclrd{r \setminus s}  \} \cup \{ \ident{?p}'_i \mapsto v'_i \;|\; \ident{?p}'\in\cclrd{s}  \})
\end{array}
\\[1.75em]
{\footnotesize(\emph{split})} & \begin{array}{l}
\ident{split}_{\cclrd{r}, \cclrd{s}} (\ident{Impl}((v,v'), \{ \ident{?p}_1 \mapsto v_1, \ldots, \ident{?p}_{n}\mapsto v_n \}))  \rightarrow
\\[-0.25em]
\quad \ident{Impl}(v, \{ \ident{?p}_i \mapsto v_i \;|\; \ident{p}_i \in \cclrd{r} \} ), \ident{Impl}(v', \{ \ident{?p}_i \mapsto v_i \;|\; \ident{p}_i \in \cclrd{s} \} )
\end{array}
\\[1em]
{\footnotesize(\emph{letimpl})} & \begin{array}{l}
\ident{letimpl}_{\cclrd{\ident{?p}}, \cclrd{r}}~v'~(\ident{Impl}(v, \{ \ident{?p}_1 \mapsto v_1, \ldots, \ident{?p}_{n}\mapsto v_n \}))  \rightarrow
\\[-0.25em]
\quad \ident{Impl}(v, \{ \ident{?p}_i \mapsto v_i\;|\; \ident{?p}_i \in \cclrd{r}, \ident{?p}_i \neq \ident{?p} \} \cup \{ \ident{?p} \mapsto v' \} )
\end{array}
\\[1em]
{\footnotesize(\emph{lookup})} & \begin{array}{l}
\ident{lookup}_{\cclrd{\ident{?p}_i}} (\ident{Impl}(v, \{\ident{?p}_i\mapsto v_i\}) \rightarrow v_i
\end{array}
\end{array}
\end{equation*}

\vspace{-0.5em}
\caption{Additional constructs embedding implicit parameters into the language}
\label{fig:semantics-ext-impl}
\end{figure}

% --------------------------------------------------------------------------------------------------

\newpage
~
\newpage
\paragraph{Properties.}

The core functional language subset (Figure~\ref{fig:semantics-target}) with the comonadically-inspired 
extensions (Figure~\ref{fig:semantics-ext}) and the domain-specific extensions for implicit 
parameters defined in Figure~\ref{fig:semantics-ext-impl} is safe. Furthermore, the 
translation from coeffect calculus produces well-typed programs. Together, these two properties 
also guarantee safety of the source coeffect calculus, when used with implicit parameters.

As for dataflow computations, we prove safety by extending the preservation
(Theorem~\ref{thm:semantics-fp-pres}) and progress (Theorem~\ref{thm:semantics-fp-prog})
for the core functional subset of the language, but it is worth noting that the key parts of
the proofs are centered around the new reduction rules for comonadically-inspired primitives
and newly defined \ident{Impl} values. These do not interact with the rest of the language in 
any unexpected ways.

\begin{theorem}[Well-typedness of the implicit parameters translation]
\label{thm:semantics-impl-welltyped}
Given a typing derivation for a well-typed closed expression $\coctx{}{\cclrd{r}} \vdash e : \tau$,
the translated program $f$ obtained using the rules in Figure~\ref{fig:semantics-translation}
and Figure~\ref{fig:semantics-ext-impl} is well-typed, \ie~in the target langauge: $\vdash f : \sem{\coctx{\Gamma}{\cclrd{r}}} \rightarrow \sem{\tau}$.
\end{theorem}
\begin{proof}
By rule induction over the derivation of the translation.

\vspace{0.5em}\noindent\hangindent=0.6cm 
Case (\emph{var, num, abs, app}): As before.

\vspace{0.5em}\noindent\hangindent=0.6cm 
Case (\emph{lookup}): The type of $\ctx$ has a coeffect  $\cclrd{ \{\ident{?p}\} }$
  which includes the parameter \ident{?p} as required in order to use the (\emph{lookup}) typing rule.

\vspace{0.5em}\noindent\hangindent=0.6cm 
Case (\emph{letimpl}): The type of $\ctx$ matches with the input type of
  $\ident{map}_{  \cclrd{r \cup (s\setminus\{\ident{?p}\})} }$. After duplication and splitting
  the context, $\ctx_1$ and $\ctx_2$ have types $\ctyp{\cclrd{r}}(\ldots)$ and
  $\ctyp{\cclrd{s\setminus\{\ident{?p}\}}}(\ldots)$, respectively. This matches with the expected
  types of $\ident{f}$ and $\ident{letimpl}$. The context returned by $\ident{letimpl}$ then 
  matches the one required by $\ident{g}$.
\end{proof}

\begin{lemma}[Canonical forms]
\label{thm:semantics-impl-canon}
For all $e, \tau$, if $\vdash e : \tau$ and $e$ is a value then:
\begin{enumerate}
  \item If $\tau=\ident{num}$ then $e = n$ for some $n \in \mathbb{Z}$
  \item If $\tau=\tau_1 \rightarrow \tau_2$ then $e = \lambda x.e'$ for some $x, e'$
  \item If $\tau=\tau_1\times\ldots\times\tau_n$ then $e = (v_1, \ldots, v_n)$ for some $v_i$
  \item If $\tau=\ctyp{\cclrd{ \{ \ident{?p}_1, \ldots, \ident{?p}_n \} }}\tau_1$ then $e = \ident{Impl}(v, \{ \ident{?p}_1 \mapsto v_1, \ldots, \ident{?p}_n \mapsto v_n \} )$
\end{enumerate}  
\end{lemma}
\begin{proof}
  (1,2,3) as before; for (4) the last typing rule must have been (\emph{impl}).
\end{proof}

\begin{lemma}[Preservation under substitution]
\label{thm:semantics-impl-pres-subst}
For all $\Gamma, e, e' \tau, \tau'$, if $\Gamma, x:\tau \vdash e : \tau'$ and $\Gamma \vdash e' : \tau$ 
then $\Gamma \vdash e[x \leftarrow e'] : \tau$.
\end{lemma}
\begin{proof}
  By induction over the derivation of $\Gamma, x:\tau \vdash e : \tau'$ as before, with new
  cases for $\ident{Impl}(e, \{ \ldots \})$, $\ident{lookup}_{\cclrd{\ident{?p}}}$ and 
  $\ident{letimpl}_{\cclrd{\ident{?p}}, \cclrd{r}}$.
\end{proof}

\begin{theorem}[Type preservation]
\label{thm:semantics-impl-pres}
  If $\Gamma \vdash e : \tau$ and $e \rightarrow e'$ then $\Gamma \vdash e' : \tau$
\end{theorem}
\begin{proof}
  Rule induction over $\rightarrow$.

\vspace{0.25em}\noindent\hangindent=0.6cm 
Case (\emph{fn}, \emph{prj}, \emph{ctx}): As before, using Lemma~\ref{thm:semantics-impl-pres-subst} for (\emph{fn}).

\vspace{0.25em}\noindent\hangindent=0.6cm 
Case (\emph{counit}): $e = \ident{counit}_{\cclrd{0}}(\ident{Impl}(v, \{\;\}))$. The last rule in the type
  derivation of $e$ must have been (\emph{counit}) with $\Gamma \vdash \ident{Impl}(v, \{ \; \}) : \ctyp{\cclrd{\emptyset}}\tau$
  which, in turn, required that $\Gamma \vdash v : \tau$.

\newpage

\vspace{0.25em}\noindent\hangindent=0.6cm 
Case (\emph{cobind}): $e = \ident{cobind}_{\cclrd{r}, \cclrd{s}}~f~(\ident{Impl}(v, \{ \ident{?p}_1\mapsto v_1, \ldots, \ident{?p}_n\mapsto v_n  \}))$.  
  The last rule in the type derivation of $e$ must have been (\emph{cobind}) with a type $\tau = \ctyp{\cclrd{s}}\tau_2$ and assumptions
  $\Gamma \vdash \ident{Impl}(v, \{ \ident{?p}_1\mapsto v_1, \ldots, \ident{?p}_n\mapsto v_n  \}) : \ctyp{\cclrd{r}}{\tau}$ and
  $\Gamma \vdash f : \ctyp{\cclrd{r}}{\tau_1} \rightarrow \tau_2$.
  The reduced expression has a type $\ctyp{\cclrd{s}}\tau_2$:
\begin{equation*}
\inference
  { \inference 
      { \Gamma \vdash f : \ctyp{\cclrd{r}}{\tau_1} \rightarrow \tau_2 & 
        \Gamma \vdash \ident{Impl}(v, \{ \ident{?p}_i \mapsto v_i \;|\; \ident{p}_i \in \cclrd{r} \} ) :  \ctyp{\cclrd{r}}{\tau_1}}  
      { \Gamma \vdash f~(\ident{Impl}(v, \{ \ident{?p}_i \mapsto v_i \;|\; \ident{p}_i \in \cclrd{r} \} )) : \tau_2} }
  { \Gamma \vdash \ident{Impl}(f~(\ident{Impl}(v, \{ \ident{?p}_i \mapsto v_i \;|\; \ident{p}_i \in \cclrd{r} \} )), \{ \ident{?p}_i \mapsto v_i \;|\; \ident{p}_i \in \cclrd{s} \}) : \ctyp{\cclrd{s}}\tau_2 }
\end{equation*}

\vspace{0.25em}\noindent\hangindent=0.6cm 
Case (\emph{lookup}): $e = \ident{lookup}_{\cclrd{\ident{?p}_i}}(\ident{Impl}(v, \{\; \ldots, \ident{?p}_i \mapsto v_i \ldots \}))$. The last rule in the type
  derivation must have been (\emph{lookup}) with $\tau = \ident{num}$ and an assumption $\Gamma\vdash\ident{Impl}(v, \{\; \ldots, \ident{?p}_i\mapsto v_i \ldots \}) : \ctyp{\cclrd{ \{\ldots,\ident{?p}_i,\ldots\} }}{\tau}$,
  which requires $\Gamma \vdash v_i : \ident{num}$.

\vspace{0.25em}\noindent\hangindent=0.6cm 
Case (\emph{merge}, \emph{split}, \emph{letimpl}): Similar. In all three cases, the last typing rule in the derivation of $e$
  guarantees that all values of all implicit parameters that are required for the reduction are available.
\end{proof}

\begin{theorem}[Progress]
\label{thm:semantics-impl-prog}  
  If $\vdash e : \tau$ then either $e$ is a value or there exits $e'$ such that $e \rightarrow e'$
\end{theorem}
\begin{proof}
  By rule induction over $\vdash$.

\vspace{0.25em}\noindent\hangindent=0.6cm 
Case (\emph{num,abs,var,app,proj,tup}): As before, using the adapted canonical forms lemma
  (Lemma~\ref{thm:semantics-impl-canon}) for (\emph{app}) and (\emph{proj}).

\vspace{0.25em}\noindent\hangindent=0.6cm 
Case (\emph{counit}): $e = {\ident{counit}_\cunit}~e_1$. If $e_1$ is not a value, it can be reduced 
  using (\emph{ctx}) with context ${\ident{counit}_\cunit}~\_$, otherwise it is a value. From Lemma~\ref{thm:semantics-df-canon},
  $e_1 = \ident{Impl}(v, \{\;\})$ and so we can apply (\emph{counit}) reduction rule.

\vspace{0.25em}\noindent\hangindent=0.6cm 
Case (\emph{cobind}): $e = \ident{cobind}_{\cclrd{r}, \cclrd{s}}~e_1~e_2$. If $e_1$ is not a value,
  reduce using (\emph{ctx}) with context ${\ident{cobind}_{\cclrd{r},\cclrd{s}}}~\_~e$. If $e_2$ is
  not a value reduce using (\emph{ctx}) with context ${\ident{cobind}_{\cclrd{r},\cclrd{s}}}~v~\_$.
  If both are values, then from Lemma~\ref{thm:semantics-impl-canon}, we have 
  $e_2 = \ident{Impl}(v, \{ \ident{?p}_i \mapsto v_i \;|\; \ident{?p}_i \in \cclrd{r}\cup\cclrd{s} \})$ 
  and we apply the (\emph{cobind}) reduction.

\vspace{0.25em}\noindent\hangindent=0.6cm 
Case (\emph{merge}): $e = \ident{merge}_{\cclrd{r}, \cclrd{s}} e_1$. If $e_1$ is not a value, 
  reduce using (\emph{ctx}) with context $e = \ident{merge}_{\cclrd{r}, \cclrd{s}}~\_$. If $e_1$ is
  a value, it must be a pair of values $(   
    \ident{Impl}(v, \{ \ident{?p}_i \mapsto v_i \;|\; \ident{?p}_i \in\cclrd{r} \}),
    \ident{Impl}(v', \{ \ident{?p}_i \mapsto v_i \;|\; \ident{?p}_i \in\cclrd{s} \}))$
  using Lemma~\ref{thm:semantics-df-canon} and it can reduce using (\emph{merge}) reduction.

\vspace{0.25em}\noindent\hangindent=0.6cm 
Case (\emph{impl}): $e = \ident{Impl}(e', \{ \ident{?p}_1 \mapsto e_1, \ldots, \ident{?p}_n \mapsto e_n\})$. 
  If $e$ is not a value, reduce using (\emph{ctx}) with context
  $\ident{Impl}(\_, \{ \ident{?p}_1 \mapsto e_1, \ldots, \ident{?p}_n \mapsto e_n\})$. 
  If $e_i$ is not a value, reduce using (\emph{ctx}) with context 
  $\ident{Impl}(v, \{ \ident{?p}_1 \mapsto v_1, \ldots, \ident{?p}_{i-1} \mapsto v_{i-1}, \ident{?p}_i \mapsto \_, \ident{?p}_{i+1} \mapsto v_{i+1}, \ldots \ident{?p}_n \mapsto e_n\})$.
  Otherwise, $e, e_0, \ldots, e_n$ are values and so 
  $\ident{Impl}(e', \{ \ident{?p}_1 \mapsto e_1, \ldots, \ident{?p}_n \mapsto e_n\})$
  is also a value.

\vspace{0.25em}\noindent\hangindent=0.6cm 
Case (\emph{split}, \emph{letimpl}): Similar. Either sub-expression is not a value, 
  or the type guarantees that it is a comonadic value with implicit parameter bindings
  that enable the (\emph{split}) or (\emph{letimpl}) reduction, respectively.
\end{proof}

\begin{theorem}[Safety]
  If $\Gamma \vdash e : \tau$ and $e \rightarrow^* e'$ then either $e'$ is a value or there
  exists $e''$ such that $e' \rightarrow e''$
\end{theorem}
\begin{proof}
  Rule induction over $\rightarrow^*$ using Theorem~\ref{thm:semantics-impl-pres} and Theorem~\ref{thm:semantics-impl-prog}.
\end{proof}

\newpage

% --------------------------------------------------------------------------------------------------

\section{Generalised safety of comonadic embedding}
\label{sec:semantics-generalising}

In Section~\ref{sec:semantics-df} and Section~\ref{sec:semantics-proofs-impl}, we proved the
safety property of two concrete context-aware programming langauges based on the coeffect 
language framework. The proofs for the two systems were very similar and they relied on the
same key principle.

The principle is that the coeffect annotation $\cclrd{r}$ on the type modelling 
the indexed comonad structure $\ctyp{\cclrd{r}}\tau$ in the target language guarantees
that the comonadic value will provide the necessary context. As a result the reductions for
operations accessing the context do not get stuck. In case of dataflow, $\ident{prev}_\cclrd{n}$ 
can always access the tail of the stream and and $\ident{counit}_\cunit$ can always access the 
head (because the stream has a sufficient number of elements). In case of implicit parameters, 
the context passed to $\ident{lookup}_{\cclrd{\ident{?p}}}$ will always contain a binding for
$\ident{?p}$.

Our core functional target language is not expressive enough to capture the relationship between
the coeffect annotation and the structure of the \ident{Df} or \ident{Impl} value and so we
resorted to adding those as ad-hoc extensions. However, given a target language with a sufficiently 
expressive type system, the properties proved in Section~\ref{sec:semantics-proofs} would be
guaranteed directly by the target language. This includes dependently-typed languages such as
Idris or Agda \cite{other-idris,other-agda}, but type-level numerals and sets can also be encoded
in the Haskell type system \cite{effects-embedding}.

In other words, the flat coeffect type system, together with the translation for introduced in this 
chapter, can be embedded into a Haskell-like languages and it can provide a succinct and safe way 
of implementing context-aware domain specific languages.

\paragraph{Coeffects for liveness.}
As an example, we consider the third instance of coeffect calculus that was discussed in 
Chapter~\ref{ch:flat}. If we wanted to follow the development in the previous section for liveness, 
we would extend the target language with two kinds of expressions, \ident{Dead} representing a dead
context with no value and \ident{Live} representing a context with a value:
%
\begin{equation*}
\begin{array}{rcl}
e &=& \ldots \sep \ident{Dead} \sep \ident{Live}~e \\
\end{array}  
\end{equation*}
%
The typing rules promote the information about whether a value is available into the type-level 
and so a context carrying a live value is marked as $\ctyp{\cclrd{\ident{L}}}{\tau}$ while a
dead context has a type $\ctyp{\cclrd{\ident{D}}}{\tau}$.
%
\begin{equation*}
\tyrule{live}
  { \Gamma \vdash e:\tau }
  { \Gamma \vdash \ident{Live}~e : \ctyp{\cclrd{\ident{L}}}{\tau} }
\qquad
\tyrule{dead}
  { ~ }
  { \Gamma \vdash \ident{Dead} : \ctyp{\cclrd{\ident{D}}}{\tau} }
\end{equation*}
%
Finally, we need to add reduction rules that define the meaning of the comonadically-inspired
operations for liveness. Those follow the definitions given in Example~\ref{thm:semantics-indexed-opt}
when discussing the categorical semantics:
%
\begin{equation*}
\begin{array}{rl}
{\footnotesize(\emph{counit})} &
  \ident{counit}_{\cclrd{ \ident{L} }}~(\ident{Just}~v)   \rightarrow v
\\[0.5em]
{\footnotesize(\emph{cobind-1})} &
  \ident{cobind}_{\cclrd{\ident{L}}, \cclrd{\ident{L}}}~f~(\ident{Live}~v) \rightarrow \ident{Live}~(f~v)
\\
{\footnotesize(\emph{cobind-2})} &
  \ident{cobind}_{\cclrd{\ident{D}}, \cclrd{\ident{L}}}~f~(\ident{Live}~v) \rightarrow \ident{Live}~(f~\ident{Dead})
\\
{\footnotesize(\emph{cobind-3})} &
  \ident{cobind}_{\cclrd{\ident{L}}, \cclrd{\ident{D}}}~f~v \rightarrow \ident{Dead}
\\
{\footnotesize(\emph{cobind-4})} &
  \ident{cobind}_{\cclrd{\ident{D}}, \cclrd{\ident{D}}}~f~v \rightarrow \ident{Dead}
\end{array}
\end{equation*}
%
This language extension is safe because the reductions respect the typing of the
comonadically-inspired operations. The $\ident{counit}_\cunit$ reduction does not get stuck
for well-typed terms because $\cunit = \cclrd{\ident{L}}$ in the coeffect algebra and thus its 
argument is of type $\ctyp{\cclrd{\ident{L}}}{\tau}$ and will always be a value $\ident{Live}~v$.

Similarly, the when reducing $\ident{cobind}_{\cclrd{r}, \cclrd{s}}$, the typing ensures that
the value passed as the second argument is of type $\ctyp{\cclrd{r}\cseq\cclrd{s}}{\tau_1}$.
In case of liveness, $\cclrd{r}\,\cseq\,\cclrd{s}=\cclrd{\ident{L}}$ if either
$\cclrd{r}=\cclrd{\ident{L}}$ or $\cclrd{s}=\cclrd{\ident{L}}$. This means that reductions
(\emph{cobind-1}) and (\emph{cobind-2}) will not get stuck because the value will be
$\ident{Live}~v$ and not \ident{Dead}. The reduction rules also preserve typing - the resulting
value is of type $\ctyp{\cclrd{s}}\tau_2$, that is $\ident{Live}~v$ for (\emph{cobind-1}),
(\emph{cobind-2}) and \ident{Dead} for (\emph{cobind-3}) and (\emph{cobind-4}).

\paragraph{Encoding liveness in Haskell.}
The liveness example can be encoded in Haskell using type-level features such as generalized
algebraic data types (GADTs) and type families \cite{haskell-faking,haskell-promotion,haskell-families}.
We do not aim to give a complete implementation, but to show that such encoding is possible and
would provide the necessary safety guarantees.

We first define types \ident{D} and \ident{L} to capture the coeffect annotations. 
Then we define a comonadically-inspired type $\ident{C}~r~a$ as a GADT with cases for
\ident{Live} and \ident{Dead} contexts. The type parameter $r$ represents a coeffect 
annotation:
%
\begin{equation*}
\begin{array}{l}
\kvd{data}~\ident{L} \\[-0.2em]
\kvd{data}~\ident{D} \\[0.5em]
\kvd{data}~\ident{C}~r~a~\kvd{where} \\[-0.2em]
\quad \ident{Live} \hspace{0.45em}:: a \rightarrow \ident{C}~\ident{L}~a\\[-0.2em]
\quad \ident{Dead} :: \ident{C}~\ident{D}~a
\end{array}
\end{equation*}
%
The definition matches with the typing rules (\emph{live}) and (\emph{dead}). The coeffect 
annotation for a live value is \ident{L} and the annotation for a dead value is \ident{D}.
To give the type of \ident{cobind}, we need a type-level function that encodes the operations
of the flat coeffect algebra. We model $\cseq$ as $\ident{Seq}~a~b$. The operation is defined
on types $\ident{L}$ and $\ident{D}$ and returns a type $\ident{D}$ if and only if both its
arguments are $\ident{D}$:
%
\begin{equation*}
\begin{array}{l}
\kvd{type}~\kvd{family}~\ident{Seq}~r~s :: *\\[-0.2em]
\quad \kvd{type}~\kvd{instance}~\ident{Seq}~\ident{D}~\ident{D} = \ident{D}\\[-0.2em]
\quad \kvd{type}~\kvd{instance}~\ident{Seq}~\ident{L}~s = \ident{L}\\[-0.2em]
\quad \kvd{type}~\kvd{instance}~\ident{Seq}~r~\ident{L} = \ident{L}
\end{array}
\end{equation*}
%
The \ident{counit} and \ident{cobind} operations can then be defined as Haskell functions
that have types corresponding to the typing rules (\emph{counit}) and (\emph{cobind}) given
in Figure~\ref{fig:semantics-ext}:
%
\begin{equation*}
\begin{array}{rcl}
\ident{counit} &::& \ident{C}~\ident{L}~a \rightarrow a \\[-0.2em]
\ident{cobind} &::& (\ident{C}~\ident{r}~a \rightarrow b) \rightarrow \ident{C}~(\ident{Seq}~r~s)~a \rightarrow \ident{C}~\ident{s}~b
\end{array}
\end{equation*}
%
Here, the additional type annotation ensures that \ident{counit} can only be called on a context
that contains a value and so calling the operation is not going to fail. Similarly, the type of the
\ident{cobind} operation now guarantees that if a function (used as the first argument) or the result
require a live context, it will be called with a value $\ident{C}~\ident{L}~a$ that is guaranteed
to contain a value.

\paragraph{Coeffects in dependently-typed languages.}
If we use the above encoding, type preservation is guaranteed by the type 
system of the target language. The equivalent of the progress property is guaranteed by the 
fact that the the implementation of the operations is well-defined.

It is worth noting that this is where coeffects need a target language with a more expressive
type system than monads. For monadic computations, it is sufficient to use a type $\ident{M}~a$
which represents that \emph{some} effects may happen. The the type does not specify which effects
and, indeed, this means that \emph{all possible effects} may happen.

With coeffects, we need indexed comonads $\ident{C}~r~a$ where the annotation $r$ specifies what
context may be required. Without the annotation, a type $\ident{C}~a$ would represent a comonadic
context that has \emph{all possible context} available, which is rarely useful in practice.

% ==================================================================================================
%                                                 
%      ####           ##            #                # 
%      #   #           #            #                # 
%      #   #   ###     #     ###   ####    ###    ## # 
%      ####   #   #    #        #   #     #   #  #  ## 
%      # #    #####    #     ####   #     #####  #   # 
%      #  #   #        #    #   #   #  #  #      #  ## 
%      #   #   ###    ###    ####    ##    ###    ## # 
%
% ==================================================================================================

\section{Related categorical structures}
\label{sec:semantics-related}

Related work leading to coeffects has already been discussed in Chapter~\ref{ch:pathways} 
and we covered work related to individual concepts throughout the thesis. However, there is a
number of related categorical structures that are related to our \emph{indexed comonads}
(Section~\ref{sec:semantics-flat-idx}) that deserve additional discussion.

In Section~\ref{sec:semantics-related-related}, we discuss related approaches to adding 
indices to categorical structures (mostly monads). In Section~\ref{sec:semantics-related-monad},
we discuss a question that often arises when discussing coeffects and that is 
\emph{when is a coeffect (not) an effect?}

\subsection{Indexed categorical structures}
\label{sec:semantics-related-related}

Ordinary comonads have the \emph{shape preservation} property \cite{comonads-codo}. Intuitively, 
this means that the core comonad structure does not provide a way of modeling computations where
the additional context changes during the computation. For example, in the \ident{NEList} comonad, 
the length of the list stays the same after applying \ident{cobind}.

Indexed comonads are not restricted by this property of comonads. For example, given the indexed
product comonad, in the computation $\ident{cobind}_{\cclrd{r}, \cclrd{s}} f$, the shape of
the context changes from providing implicit parameters $\cclrd{r} \cup \cclrd{s}$  to providing
just implicit parameters $\cclrd{s}$. 
Thus \emph{indexed comonads} are a generalization of \emph{comonads} that captures structures that
do not form a (non-indexed) comonad. However, indexing has been also discussed in the case of
\emph{monads.}

\paragraph{Families of monads.}
When linking effect systems and monads, Wadler and Thiemann \cite{monad-notions} propose a
\emph{family of monads} as the categorical structure. The dual structure, \emph{family of
comonads}, is defined as follows.

\begin{definition}
\label{def:flat-family}
A \emph{family of comonads} is formed by triples $(\ctyp{\cclrd{r}}{}, \ident{cobind}_{\cclrd{r}}, 
  \ident{counit}_{\cclrd{r}})$ for all $\cclrd{r}$ such that each triple forms a comonad. Given 
$\cclrd{r}, \cclrd{r'}$ such that $\cclrd{r} \leq \cclrd{r'}$, there is also a mapping 
$\iota_{\cclrd{r'}, \cclrd{r}} : \ctyp{\cclrd{r'}}{} \rightarrow \ctyp{\cclrd{r}}{}$ satisfying
certain coherence conditions.
\end{definition}

\noindent
A \emph{family of comonads} is more restrictive than an \emph{indexed comonad}, because each of the data 
types needs to form a comonad separately. For example, our indexed option does not form a family of 
comonads (again, because \ident{counit} is not defined on $\ctyp{\ident{D}}{\alpha}=1$). However, given 
a family of comonads and indices such that $\cclrd{r} \leq \cclrd{r}\cseq\cclrd{s}$, we can define 
an indexed comonad. Briefly, to define $\ident{cobind}_{\cclrd{r},\cclrd{s}}$ of an indexed comonad, 
we use $\ident{cobind}_{\cclrd{r}\cseq\cclrd{s}}$ from the family, together with two lifting operations:
$\iota_{\cclrd{r}\cseq\cclrd{s}, \cclrd{r}}$ and $\iota_{\cclrd{r}\cseq\cclrd{s}, \cclrd{s}}$.

\paragraph{Parameteric effect monads.}
Parametric effect monads introduced by Katsumata \cite{monads-parametric} (independently to our 
indexed comonads) are closely related to our definition.  Although presented in a more general 
categorical framework (and using monads), the model defines the \ident{unit} operation only on the 
unit of a monoid and the \ident{bind} operation composes effect annotations using the provided monoidal 
structure.

\subsection{When is coeffect not a monad}
\label{sec:semantics-related-monad}

Coeffect systems differ from effect systems in three important ways:

\begin{itemize}
\item Semantically, coeffects capture different notions of computation. As demonstrated in 
  Chapter~\ref{ch:pathways}, coeffects track additional contextual properties required by a
  computation, many of which cannot be captured by a monad (\eg~liveness or data-flow).

\item Syntactically, coeffect calculi use a richer algebraic structure with pointwise composition,
  sequential composition and context merging ($\cpar, \cseq$, and $\czip$) while most effect systems
  only use a single operation for sequential composition (used by monadic bind).

\item Syntactically, the second difference is in the lambda abstraction (\emph{abs}). In 
  coeffect systems, the context demands of the body can be split between (or duplicated
  at) declaration site and call site, while monadic effect systems always defer all effects.
\end{itemize}

\noindent
Despite the differences, our implicit parameters example can be also represented by a monad.
Semantically, the \emph{reader} monad is equivalent to the \emph{product} comonad. Syntactically,
we use the $\cup$ operation for all three operations of the coeffect algebra. However, to enable
splitting of implicit parameter demands using the reader monad, we need to extend the 
monad structure and change the translation of monadic lambda abstraction.

% -------------------------------------------------------------------------------------------------

\subsection{When is coeffect a monad}
\label{sec:semantics-related-monads}

Implicit parameters can be captured by a monad, but \emph{just} a monad is not enough.
Lambda abstraction in effect systems does not provide a way of splitting the context
demands between declaration site and call site (or, semantically, combining the implicit 
parameters available in the scope where the function is defined and those specified by the caller).

\paragraph{Categorical relationship.}
Before looking at the necessary extensions, consider the two ways of modelling implicit 
parameters. We assume that the function $\cclrd{r} \rightarrow \sigma$ is a lookup function
for reading implicit parameter values that is defined on a set $\cclrd{r}$. The two definitions
are:
%
\begin{equation*}
\begin{array}{lll}
 \ctyp{\cclrd{r}}\tau = \tau \times (\cclrd{r} \rightarrow \sigma) &~\hspace{10em}~& (\emph{product comonad}) \\
 \mtyp{\cclrd{r}}\tau = (\cclrd{r} \rightarrow \sigma) \rightarrow \tau && (\emph{reader monad})
\end{array} 
\end{equation*}
%
The \emph{product comonad} simply pairs the value $\tau$ with the lookup function, while
the \emph{reader monad} is a function that, given a lookup function, produces a $\tau$ value.
As noted by Orchard \cite{comonads-vs-monads}, when used to model computation semantics, the 
two representations are equivalent:
%
\begin{remark}
Computations modelled as $\ctyp{\cclrd{r}}{\tau_1}\rightarrow\tau_2$ using the product comonad
are isomorphic to computations modelled as $\tau_1\rightarrow\mtyp{\cclrd{r}}{\tau_2}$ using the
reader monad via currying/uncurrying isomorphism.
\end{remark}
\begin{proof}
The isomorphism is demonstrated by the following equation:
\begin{equation*}
\begin{array}{lll}
 \ctyp{\cclrd{r}}{\tau_1} \rightarrow \tau_2 &\narrow{=}&
 (\tau_1 \times (\cclrd{r} \rightarrow \sigma)) \rightarrow \tau_2 \\
 &\narrow{=}& \tau_1 \rightarrow ((\cclrd{r} \rightarrow \sigma) \rightarrow \tau_2) =
 \tau_1 \rightarrow \mtyp{\cclrd{r}}{\tau_2} \hspace{23em}\qedhere \\
\end{array}
\end{equation*}
\end{proof}

\noindent
This equivalence holds for monads and comonads (as well as \emph{indexed} monads
and comonads), but it does not extend to \emph{flat} indexed comonads which also provide
the $\ident{merge}_{\cclrd{r}, \cclrd{s}}$ operation to model context merging.

\paragraph{Delaying effects in monads.} 
In the syntax of the language, the above difference is manifested by the (\emph{abs}) rules for 
monadic effect systems and comonadic coeffect systems. The following listing shows the two rules 
side-by-side, using the effect system notation for both of them:
%
\begin{equation*}
\tyrule{cabs}
  { \Gamma, x\!:\!\tau_1 \vdash e : \tau_2 \;\&\; \cclrd{r}\,\cclrd{\cup}\,\cclrd{s} }
  { \Gamma \vdash \lambda x.e : \tau_1 \xrightarrow{\cclrd{s}} \tau_2 \;\&\; \cclrd{r}}
\quad
\tyrule{mabs}
  { \Gamma, x\!:\!\tau_1 \vdash e : \tau_2 \;\&\; \cclrd{r}\,\cclrd{\cup}\,\cclrd{s}}
  { \Gamma \vdash \lambda x.e : \tau_1 \xrightarrow{\cclrd{r}\,\cclrd{\cup}\,\cclrd{s}} \tau_2 \;\&\; \cclrd{\emptyset} }
\qquad
\end{equation*}
%
In the comonadic (\emph{cabs}) rule, the implicit parameters of the body are split. However,
the monadic rule (\emph{mabs}) places all demands on the call site. This follows from the
fact that monadic semantics uses the \ident{unit} operation in the interpretation of lambda abstraction:
%
\begin{equation*}
\sem{\lambda x.e} \;=\; \ident{unit}~(\lambda x. \sem{e})
\end{equation*}
%
The type of $\ident{unit}$ is $\alpha\rightarrow\mtyp{\alpha}{\cclrd{\emptyset}}$, but in this specific
case, the $\alpha$ is instantiated to be $\tau_1\rightarrow\mtyp{\cclrd{r \cup s}}{\tau_2}$ and so this
use of \ident{unit} has a type:
%
\begin{equation*}
\ident{unit}~:~(\tau_1\rightarrow\mtyp{\cclrd{r \cup s}}{\tau_2}) \rightarrow \mtyp{\cclrd{\emptyset}}(\tau_1\rightarrow\mtyp{\cclrd{r \cup s}}{\tau_2})
\end{equation*}
%
In order to split the implicit parameters of the body ($\cclrd{r \cup s}$ on the left-hand side) between 
the declaration site ($\cclrd{\emptyset}$ on the outer $\mtyp{}$ on the right-hand side) and the 
call site ($\cclrd{r \cup s}$ on the inner $\mtyp{}$ on the right-hand side), we need an operation
(which we call \ident{delay}) with the following signature:
%
\begin{equation*}
\ident{delay}_{\cclrd{r}, \cclrd{s}}~:~(\tau_1\rightarrow\mtyp{\cclrd{r \cup s}}{\tau_2}) \rightarrow \mtyp{\cclrd{r}}(\tau_1\rightarrow\mtyp{\cclrd{s}}{\tau_2})
\end{equation*}
%
The operation reveals the difference between effects and coeffects -- intuitively, given a function
with effects $\cclrd{r \cup s}$, it should execute the effects $\cclrd{r}$ when wrapping the 
function, \emph{before} the function actually performs the effectful operation with the effects.
The remaining effects $\cclrd{s}$ are delayed as usual, while effects $\cclrd{r}$ are removed
from the effect annotation of the body.

Another important aspect of the signature is that the function needs to be indexed by the coeffect
annotations $\cclrd{r}, \cclrd{s}$. The indices determine how the input context demands
$\cclrd{r \cup s}$ are split -- and thus guarantee determinism of the function at run-time.

The operation cannot be implemented in a useful way for most standard monads, but the 
reader monad is, indeed, an exception. It is not difficult to see how it can be implemented
when we expand the definitions of $\mtyp{\cclrd{r}}\tau$:
%
\begin{equation*}
\ident{delay}_{\cclrd{r}, \cclrd{s}}~:~
(\tau_1\rightarrow(\cclrd{r \cup s} \rightarrow \sigma) \rightarrow \tau_2) \rightarrow 
((\cclrd{r} \rightarrow \sigma) \rightarrow \tau_1\rightarrow(\cclrd{s} \rightarrow \sigma) \rightarrow \tau_2)
\end{equation*}

\paragraph{Restricting coeffects in comonads.} 
As just demonstrated, we can extend monads so that the reader monad is capable of capturing
the semantics of implicit parameters, including the splitting of implicit parameter demands
in lambda abstraction. Can we also go the other way round and \emph{restrict} the comonadic
semantics so that all demands are delayed as in the (\emph{mabs}) rule, thus modelling 
fully dynamically scoped parameters?

This is, indeed, possible. Recall that the semantics of lambda abstraction in the flat
coeffect calculus is modelled using $\ident{merge}_{\cclrd{r}, \cclrd{s}}$. The operation takes
two contexts (wrapped in an indexed comonad $\ctyp{\cclrd{r}}\alpha$), combines their carried 
values and additional contextual information (implicit parameters). To obtain the (\emph{mabs}) 
rule, we can restrict the first parameter, which corresponds to the declaration site context:

\begin{equation*}
\begin{array}{rcll}
 \ident{merge}_{\cclrd{r}, \cclrd{s}} &\narrow{:}& \ctyp{\cclrd{r}}\alpha \times \ctyp{\cclrd{s}}\beta \rightarrow \ctyp{\cclrd{r \cup s}}(\alpha \times \beta)
 &\hspace{4em}(\textit{normal}) \\
 \ident{merge}_{\cclrd{r}, \cclrd{s}} &\narrow{:}& \ctyp{\cclrd{\emptyset}}\alpha \times \ctyp{\cclrd{s}}\beta \rightarrow \ctyp{\cclrd{s}}(\alpha \times \beta)
 &\hspace{4em}(\textit{restricted})
\end{array} 
\end{equation*}
%
In the (\emph{restricted}) version of the operation, the declaration site context requires
no implicit parameters and so all implicit parameters have to be satisfied by the call site.
The semantics using the restricted version corresponds to the (\emph{mabs}) rule shown above.

The idea of restricting the operations of the coeffect calculus semantics could be used more
generally. We could allow any of the coeffect algebra operations $\cseq, \czip, \cpar$ to be
\emph{partial} and thus the restricted (fully dynamically-scoped) version of implicit parameters
could be obtained just by changing the definition of $\czip$. Similarly, we could obtain \eg~a
fully lexically-scoped version of the system. The ability to restrict operations to partial 
functions has been used in the semantics of effectful computations by Tate~\cite{effects-producer-semantics}.

% ==================================================================================================
%
%     #####
%    #     # #    # #    # #    #   ##   #####  #   #
%    #       #    # ##  ## ##  ##  #  #  #    #  # #
%     #####  #    # # ## # # ## # #    # #    #   #
%          # #    # #    # #    # ###### #####    #
%    #     # #    # #    # #    # #    # #   #    #
%     #####   ####  #    # #    # #    # #    #   #
%
% ==================================================================================================

\section{SUmmary}

In the previous chapter, we defined \emph{type system} for flat coeffect calculi that uniformly
captures the shared structure of context-aware computations. In this chapter, we completed
the unification by providing \emph{semantics} for flat coeffect calculi and proving the
\emph{safety} of coeffect languages for dataflow and implicit parameters. The semantics shown
here also guides the implementation that is discussed later in Chapter~\ref{ch:impl}.

The development presented in this chapter follows the well-known example of effects and monads.
We introduced the notion of \emph{indexed comonad}, which generalizes comonads
and adds additional operations needed to provide categorical semantics of the flat coeffect 
calculus and we demonstrated how implicit parameters, liveness and dataflow computations form
indexed comonads.

We then used the comonadic semantics to define a \emph{comonadically-inspired translation}
that turns programs written in a domain-specific coeffect language into a functional target 
language. This is akin to the Haskell do notation for monads. Finally, we extended the target
language with concrete implementation of comonadic operations for dataflow and implicit parameters
and we presented syntactic safety proof. In summary, the proof states that well-typed 
context-aware programs written in a coeffect language \emph{do not go wrong} (when translated
to a simple functional language and evaluated).

The proof relies on the fact that coeffect annotations (provided by the coeffect type system)
guarantee that the required context is available in the comonadic value that represents the 
context and we also discussed how this would guarantee safety in languages with sufficiently
expressive type system such as Haskell.

In the following chapter, we move from \emph{flat} coeffect calculi, tracking whole-context
properties to \emph{structural} coeffect calculi, tracking per-variable information, thus
covering systems from the second half of Chapter~\ref{ch:applications}.