### 2016-04-09 text/Chapter08.tex
Atom whitespace, working on implementation chapter

% ==================================================================================================

\chapter{Implementation}
\label{ch:impl}

In the previous three chapters, we presented the theory of coeffects consisting of type system
and comonadically inspired semantics of two parameterized coeffect calculi. The theory provides
a framework that simplifies the implementation of safe context-aware programming languages. To
support this claim, this chapter presents a prototype implementation of three coeffect languages --
language with implicit parameters and both flat and structural versions of a data-flow language.

The implementation directly follows the thoery presented in the previous three chapters. It
consists of a common framework that provides type checking and translation to a simple functional
target language with comonadically-inspired primitives. Each concrete context-aware language then
adds a domain-specific rule for choosing unique typing derivation (as discussed in
Section~\ref{sec:flat-unique}) together with a domain-specific definition of the
comonadically-inspired primitives that define the runtime semantics (see Section~\ref{sec:semantics-proofs}).

The main goal of the implementation is to show that the theory is practically useful and to present it
in a more practical way. However, we do not intend to build a complete real-world programming language.
For this reason, the implementation is available primarilly as an interactive web-based essay, though
it can be also downloaded and run locally.

\paragraph{Chapter structure and contributions}

\begin{itemize}
\item We discuss how the implementation follows from the theory presented earlier (Section~\ref{sec:impl-theory}).
  This applies to the implementation of the \emph{type checker} and the implementation of the \emph{translation}
  to a simple target langauge that is then iterpereted. We also discuss how the common framework makes it
  easy to implement additional context-aware languages (Section~\ref{sec:impl-theory-ext}).

\item We consider a number of case studies (Section~\ref{sec:impl-case}) that illustrate ointeresting
  aspects of the theories discussed earlier. This includes the typing of lambda abstraction and the
  difference between flat and structural systems (Section~\ref{sec:impl-case-typing}) and the
  comonadically-inspired translation (Section~\ref{sec:impl-case-transl}).

\item We briefly outline some of the interesting technical details of the implementation
  (Section~\ref{sec:impl-tech}). To make the implementation easily accessible, we make it
  available via web and we also implement a limited support for type inference.

\item The implementation is available not just as downloadable code, but also in the format of interactive essay
  (Section~\ref{sec:impl-essay}), which aims to make coeffects accessible to a broader audience.
  We conclude this chapter by discussing the most interesting aspects of the presentation.
\end{itemize}

% ==================================================================================================
%
%    #######
%       #    #    # ######  ####  #####  #   #
%       #    #    # #      #    # #    #  # #
%       #    ###### #####  #    # #    #   #
%       #    #    # #      #    # #####    #
%       #    #    # #      #    # #   #    #
%       #    #    # ######  ####  #    #   #
%
% ==================================================================================================

\section{From theory to implementation}
\label{sec:impl-theory}

The theory discussed so far provides the two key components of the implementation. In
Chapter~\ref{ch:flat}, we discussed the type checking of context-aware programs and
Chapter~\ref{ch:semantics} models the execution of context-aware programs (in terms of
translation and operational semantics). For structural coeffects, the same components are discussed
in Chapter~\ref{ch:structural}. In this section, we discuss how those provide foundation for the
implementation.

\subsection{Type checking}
\label{sec:impl-theory-typing}

(Section~\ref{sec:flat-unique})

\paragraph{Implementation.}
From the presentation in this section, it might appear that resolving the ambiguity related to
lambda abstraction for implicit parameters requires a type system that is quite different from
the core flat coeffect type system shown earlier in Figure~\ref{fig:flat-types}. This is not the
case. As discussed in Chapter~\ref{ch:impl}, the required changes in the implementation are
simpler.

Briefly, the implementation collects constraints on the coeffects and then finds the smallest sets
of implicit parameters to satisfy the constraints. We still need to track implicit parameters in
scope $\cclrd{\Delta}$, but the rest of the (\emph{abs}) rule from the implementation is close
to the one from Figure~\ref{fig:flat-types}:
%
\begin{equation*}
\tyrule{abs}
  {\coctx{\Gamma, x\!:\!\tau_1;\cclrd{\Delta}}{\cclrd{t}} \vdash e : \tau_2 ~|~ C}
  {\coctx{\Gamma;\cclrd{\Delta}}{\cclrd{r}} \vdash \lambda x\!:\!\tau_1.e : \tau_1 \xrightarrow{\cclrd{s}} \tau_2 ~|~
    C\cup\{\cclrd{t}=\cclrd{r} \,\czip\, \cclrd{s}, \cclrd{r}=\cclrd{\Delta} \}}
\end{equation*}
%
Given a typing derivation for the body that produced constraints $C$, we generate an additional
constraint that restricts $\cclrd{r}$ (declaration-site demands) to those available in the
current static scope $\cclrd{\Delta}$. The constraint satisfaction algorithm then finds
the minimal set $\cclrd{s}$ which is $\cclrd{t}\cclrd{\setminus \Delta}$.

\paragraph{Implementation.}
As with implicit parameters, the implementation (discussed in Chapter~\ref{ch:impl}) does not
require changing the typing (\emph{abs}) rule of the flat coeffect system. The typing specified
by (\emph{idabs}) can be easily obtained by generating additional constraints:
%
\begin{equation*}
\tyrule{abs}
  {\coctx{\Gamma, x\!:\!\tau_1}{\cclrd{t}} \vdash e : \tau_2 ~|~ C}
  {\coctx{\Gamma}{\cclrd{s}} \vdash \lambda x\!:\!\tau_1.e : \tau_1 \xrightarrow{\cclrd{t}} \tau_2 ~|~
    C\cup\{\cclrd{t}=\cclrd{r} \,\czip\, \cclrd{s}, \cclrd{r}=\cclrd{t}, \cclrd{s}=\cclrd{t} \}}
\end{equation*}
%
Here, the two additional constraints restrict both $\cclrd{r}$ and $\cclrd{s}$ to be equal to the
coeffect of the body $\cclrd{t}$ and so the only possible resolution is the one specified by
(\emph{idabs}).

Type checking
 * DSL specific constraint resolution
 * easier for structural

\subsection{Translation}
\label{sec:impl-theory-transl}

Translation ~ Semantics

\subsection{Extensions}
\label{sec:impl-theory-ext}

It is really easy to add another DSL

\section{Case studies}
\label{sec:impl-case}

\subsection{Typing}
\label{sec:impl-case-typing}

let both =
  let ?fst = 100 in
  fun trd -> ?fst + ?snd + trd in
let ?fst = 200 in
both 1

num -{ ?snd:num }-> num

let struct x y = x + prev y in
struct

num -{ 1 }-> num -{ 1 }-> num

let flat x y = x + prev y in
flat

num -{ 0 }-> num -{ 1 }-> num

\subsection{Translation}
\label{sec:impl-case-transl}

let ?param = 10 in
fun x -> ?param + ?other

let (ctx2, ctx3) = split (duplicate finput) in
let ctx1 = letimpl ?param (ctx2, 10) in
fun x ->
  let ctx4 = merge (x, ctx1) in
  let (ctx5, ctx6) = split (duplicate ctx4) in
  lookup ?param ctx5 + lookup ?other ctx6

fun x -> (fun v -> prev v) (prev x)

fun x ->
  let ctx1 = merge (x, sinput) in
  (fun v ->
    let ctx3 = merge (v, choose_0 ctx1) in
    counit (prev ctx3)) (cobind (fun ctx2 ->
      counit (prev ctx2)) (choose_1 ctx1))

\section{Technical details}
\label{sec:impl-tech}

1. F# + JavaScript
2. Single AST and parsing
3. Type checking and constraint generation
4. Constraint resolution

\section{Interactive essay}
\label{sec:impl-essay}