### 2016-05-30 text/Chapter02.tex
Edits and tweaks (Chapter 8 and 9 mostly)

\chapter{Pathways to coeffects}
\label{ch:pathways}

There are many different directions from which the concept of \emph{coeffects} can be approached
and, indeed, discovered. In the previous chapter, we motivated it by practical applications, but
coeffects also naturally arise as an extension to a number of programming language theories.
Thanks to the Curry-Howard-Lambek correspondence, we can approach coeffects from the perspective of
type theory, logic and also category theory. This chapter gives an overview of the most
important directions.

We start (Section~\ref{sec:path-binding}) by discussing how coeffects arise from the most common
notion of context-dependence -- variable binding. Next, we look at coeffects as the dual of effect
systems (Section~\ref{sec:path-eff}) and we extend the duality to category theory, looking at
\emph{comonads} (Section~\ref{sec:path-sem}). We also consider type systems inspired by linear
and bunched logic, which are closely related to our structural coeffects (Section~\ref{sec:path-logic}).
Finally, we also consider practical motivations for context-aware programming (Section~\ref{sec:path-cop}).

This chapter serves two purposes. Firstly, it provides a high-level overview of the  related work,
although technical details are often postponed until later. Secondly it recasts existing ideas in
a way that naturally leads to the coeffect systems developed later in the thesis. For this reason,
we are not always faithful to the referenced work. We present the work through the coeffect view
and so we sometimes focus on aspects that authors consider unimportant or we present the work
differently than originally intended. When we do so, this is explicitly stated in the text.

%===================================================================================================
%
%    ######
%    #     # # #    # #####  # #    #  ####
%    #     # # ##   # #    # # ##   # #    #
%    ######  # # #  # #    # # # #  # #
%    #     # # #  # # #    # # #  # # #  ###
%    #     # # #   ## #    # # #   ## #    #
%    ######  # #    # #####  # #    #  ####
%
%===================================================================================================

\section{Coeffects via static and dynamic binding}
\label{sec:path-binding}

Accessing a variable is arguably the simplest form of context-dependence, to the extent that we
do not normally think of variables as a notion of context. However, variables fit well with our
earlier description of context in programming: a block of code that accesses a variable can only
be executed in an environment where the variable is available.

In this section, we look at variable binding through the perspective of context-requirements.
We discuss ordinary variable binding and Haskell's implicit parameters \cite{app-implicit-parameters},
which provide an interesting point in the design space. Implicit parameters give an example of
an ambiguity that arises more generally in context-aware programming, as well as one way of
resolving it through \emph{type-directed semantics}.

For a more general context-aware programming example, consider a program
running on a laptop that accesses an implicit parameter representing a printer. When printing, the
text may appear on my home printer (corresponding to static binding), or it may appear on the
nearest printer based on my physical location (corresponding to dynamic binding).

%---------------------------------------------------------------------------------------------------

\subsection{Variable binding}
\label{sec:path-binding-var}

Variable access represents a form of context-dependence. For example, an expression $x+y$ can be
only evaluated if the environment provides values for variables $x$ and $y$. A variable
\emph{requirements} can be satisfied in two standard ways that are characterized as \emph{dynamic}
and \emph{static} (or lexical) variable binding. Consider the following simple program:
%
\begin{equation*}
\begin{array}{l}
\kvd{let}~\ident{f}=\\[-0.25em]
\qquad\kvd{let}~\ident{x}=10~\kvd{in}\\[-0.25em]
\qquad\lambda \ident{y}\rightarrow \ident{x}+\ident{y}~\kvd{in}\\[-0.25em]
\kvd{let}~\ident{x}=5~\kvd{in}\\
\ident{f}~0
\end{array}
\end{equation*}
%
The program can be evaluated in two ways, depending on the variable binding mechanism:
%
\begin{itemize}
\item \textsc{Static (lexical) binding.} In a language with static binding (such as ML or Java),
  the variable \ident{x} inside the body of the lambda function is statically bound to the
  declaration in the lexical scope -- that is, the variable on the second line -- and the
  expression evaluates to 10.

\item \textsc{Dynamic binding.} In a language with dynamic binding (some variants of LISP), the
  variable value is dynamically bound to the last value that has been added to the environment
  during program execution. Thus, the \ident{x} variable inside the lambda function refers to
  \ident{x} defined on line 5 of the sample and the expression evaluates to 5.
\end{itemize}

\noindent
When we view variable access as a context requirement, we can see that the body of the function
($\ident{x}+\ident{y}$) requires a context that provides values for variables \ident{x} and
\ident{y}. In static binding, the context demands of the body can be placed on the scope in
which the function is defined (declaration-site). In dynamic binding, the requirements are
\emph{delayed} and are placed on the scope in which the function is called (call site).

In static and dynamic scoping, all variable requirements are always placed on one site. However,
those are not the only two options. It is conceivable that a language would use a mechanism that
splits variable requirements differently and combines aspects of dynamic and static binding. For
example, the language could use static binding by default, but resort to dynamic binding if a
variable is not available in the lexical scope. One such system is Haskell's implicit parameters
discussed in the next section.

Languages with static scoping resolve variable bindings at compile-time. In contrast,
languages with dynamic variable binding typically do not perform compile-time checks -- if a
program attempts to access a variable that is not available in the environment, a runtime error
occurs. This is possibly an artifact of their implementation -- implementing static binding
\emph{without} checking would be cumbersome and implementing static checking for dynamic binding is
requires a more sophisticated type system than traditional (Section~\ref{sec:path-effects-coeff}).

%---------------------------------------------------------------------------------------------------

\subsection{Implicit parameter binding}
\label{sec:path-binding-impl}

Haskell uses static binding for ordinary variables, but it additionally provides a feature named
\emph{implicit parameters} \cite{app-implicit-parameters} that adds a special kind of variables,
written as \ident{?param}, which use a particular combination of static and dynamic binding.

The following two examples are variations on the one discussed in Section~\ref{sec:path-binding-var},
obtained by replacing a variable \ident{x} with an implicit parameter \ident{?x}. On the left, the
implicit parameter is declared both in the static scope and in the dynamic scope. On the right, the
implicit parameter is available only in the dynamic scope:

\begin{equation*}
\begin{array}{l}
\kvd{let}~\ident{f}=\\[-0.25em]
\qquad\kvd{let}~\ident{?x}=10~\kvd{in}\\[-0.25em]
\qquad\lambda \ident{y}\rightarrow \ident{?x}+\ident{y}~\kvd{in}\\[-0.25em]
\kvd{let}~\ident{?x}=5~\kvd{in}\\
\ident{f}~0
\end{array}\hspace{5em}\begin{array}{l}
\kvd{let}~\ident{f}=\\[-0.25em]
\qquad\lambda \ident{y}\rightarrow \ident{?x}+\ident{y}~\kvd{in}\\[-0.25em]
\kvd{let}~\ident{?x}=5~\kvd{in}\\
\ident{f}~0
\end{array}
\end{equation*}
%
The binding rules for Haskell's implicit parameters can be summarized as ``\emph{static binding when possible,
dynamic binding when needed}''. If an implicit parameter is available in static scope, then the value
is statically bound and the context requirement is satisfied using the declaration-site context.
Otherwise, the context requirement is delayed and has to be satisfied at the call site.
In the example on the left, \ident{?x} is bound to $10$ and so the function \ident{f} has no
delayed context demands and thus the expression evaluates to $10$. On the right,
the context demand for \ident{?x} is \emph{delayed} and is satisfied via dynamic binding when calling
the function. The expression evaluates to $5$.

In Haskell, the type system checks that bindings for all required implicit parameters are
available and so no runtime errors can occur. The type of the function \ident{f} on the left is
$\kvd{int}\rightarrow\kvd{int}$, while the type of the \ident{f} function on the right is
$\{\ident{?x}:\kvd{int}\}\Rightarrow\kvd{int}\rightarrow\kvd{int}$.
The part before $\Rightarrow$ specifies the required implicit parameters that need to be available
in the environment when calling the function. It is worth noting that the syntax is similar to the
one used by type-class constraints. Those can be viewed as context demands too
(Section~\ref{sec:applications-flat-impl}).

\paragraph{Thesis perspective.}
The three different binding mechanisms discussed so far can be seen as different ways of splitting
context demands of a particular kind into static and dynamic parts. Dynamic and static binding
represent the opposite ends of the design spectrum and Haskell's implicit parameters are an
interesting point inside the wider spectrum.

In this thesis, we consider various notions of context, using implicit parameters as one
of several motivating examples. Implicit parameters are a particularly valuable example, because they
clearly illustrate the ambiguity inherent in context-aware programs -- the context demands of
a function can be satisfied using the context available at declaration-site or using the context
available at the call site. Recall our earlier printer example -- a language that
provides access to resources in the context needs to provide enough flexibility in handling such
ambiguities, be it implicit parameter values available in scope or printers available in the
physical environment.

We aim to find a description of context-aware languages that does not make ad-hoc decisions
about how context demands are split between the declaration-site and the call site. While
Haskell's solution for implicit parameters might be the most reasonable one for that particular
case, this thesis argues that other notions of context require different domain-specific choices
and the general framework of context-aware programming should make that possible.

%---------------------------------------------------------------------------------------------------

\subsection{Resolving ambiguity}
\label{sec:path-binding-amb}

In many practical programming languages, the value and semantics of an expression depends on its
type derivation. In order to assign unique semantics to an expression, the choice is typically
hidden behind a mechanism that selects one preferred type derivation.

This mechanism serves as an inspiration for our approach to resolving ambiguity inherent in
context-aware programs. This section discusses a brief example using the F\# language \cite{app-fsharp},
before revisiting the implicit parameters example.

Consider an F\# lambda expression $(\lambda\ident{x}\rightarrow\ident{x.Length})$, which takes
an object $\ident{x}$ and returns the value of its \ident{Length} property. F\# is a nominally-typed
language meaning that, in isolation, the function has an ambiguous meaning\footnote{In contrast,
in a structurally-typed language, the function would have a unique typing in isolation. In OCaml,
the type would be $\langle \ident{Length}:\ident{'a}\rangle \rightarrow \ident{'a}$.}. It can be
a function taking an array, it can be a function taking a string, or it can be a function taking
one of the other .NET types that are equipped with the \ident{Length} property.

The semantics of the function depends on the typing derivation. For example, for arrays, it is
compiled using the \ident{ldlen} intermediate language (IL) instruction, while or strings, it is
compiled using \ident{call} instruction (calling the property getter). In F\#, the compiler chooses
an appropriate typing derivation. For example:
%
\begin{equation*}
\begin{array}{l}
  \lbrack \str{hello};\str{world} \rbrack ~|\hspace{-0.25em}>\;\ident{List.map}~(\lambda\ident{s}\rightarrow\ident{s.Length})\\
  \lbrack \ident{Array.empty};~\ident{Array.create}~\num{100}~\num{0}\; \rbrack ~|\hspace{-0.25em}>\;\ident{List.map}~(\lambda\ident{s}\rightarrow\ident{s.Length})\\
\end{array}
\end{equation*}
%
The $|\hspace{-0.25em}>$ operator passes the value on the left to the function on the right. In
the first case, the compiler infers that the type of the input is a list of strings and so the
type of the lambda function becomes $\ident{string}\rightarrow\ident{int}$. In the second case,
the list contains two arrays (an empty array and an array containing one hundred zero values) and so
the type of the lambda function is $\ident{int}\lbrack\,\rbrack\rightarrow\ident{int}$.
The important points about the example are:
%
\begin{enumerate}
\item The semantics of the function $(\lambda\ident{x}\rightarrow\ident{x.Length})$
  depends on its type. For arrays, it is compiled using a special IL instruction, while for
  strings, it calls a property getter.

\item The compiler chooses an appropriate typing derivation. In the above case, this is done
  based on the context in which the expression appears, but other options are possible
  (in some cases, there is a \emph{default} resolution; often, the compiler requires an
  explicit typing annotation).

\item An expression without a type derivation does not have semantics. For example, given
  $\ident{List.map}~(\lambda\ident{s}\rightarrow\ident{s.Length})$, the F\# compiler fails to
  infer a type; the expression is not well-typed and does not have a semantics.\footnote{Alternatively,
  the compiler could choose default typing among multiple options. The F\# compiler does this
  for the $+$ operator, which can be used on \ident{float} and \ident{int} types, but the compiler
  chooses \ident{int} as the default.}
\end{enumerate}
%
The function $\lambda\ident{y}\rightarrow\ident{?x}+\ident{y}$ in Haskell also has multiple
possible typing derivations and its semantics varies depending on its type. If the lexical
scope contains a binding for $\ident{?x}$, the function type is $\kvd{int}\rightarrow\kvd{int}$
and it captures the value from the lexical scope. Otherwise, the type of the function is
$\{\ident{?x}:\kvd{int}\}\Rightarrow\kvd{int}\rightarrow\kvd{int}$ and it reads the parameter
value from a hidden dictionary that is passed together with the input from the call site.

\paragraph{Thesis perspective.}
Just like the F\# function in the above example, certain expressions in context-aware languages
developed in this thesis have multiple valid typing derivations and their semantics depends on the
type. In F\#, the compiler determines a unique typing derivation based on other parts of the program
(if type is not uniquely determined, it either chooses a default or fails). In our languages, we
also determine a unique typing derivation. However, rather than relying on type information from
other parts of the program, we explicitly define an algorithm that chooses the preferred unique
derivation (Section~Section~\ref{sec:flat-unique} and Section~\ref{sec:struct-unique}).

This approach decouples two important aspects of context-aware programming and lets us study them
independently -- the semantics of context-aware programs and the domain-specific way of resolving
ambiguities in how context demands are satisfied. In our treatment of implicit parameters, we consider
multiple typing derivations (representing a range with static and dynamic scoping at opposite ends),
but we uniquely choose one preferred typing (in case of implicit parameters, mimicking the
behaviour of Haskell).

%===================================================================================================
%
%    #######
%    #       ###### ###### ######  ####  #####     ####  #   #  ####  ##### ###### #    #  ####
%    #       #      #      #      #    #   #      #       # #  #        #   #      ##  ## #
%    #####   #####  #####  #####  #        #       ####    #    ####    #   #####  # ## #  ####
%    #       #      #      #      #        #           #   #        #   #   #      #    #      #
%    #       #      #      #      #    #   #      #    #   #   #    #   #   #      #    # #    #
%    ####### #      #      ######  ####    #       ####    #    ####    #   ###### #    #  ####
%
%===================================================================================================

\section{Coeffects via type and effect systems}
\label{sec:path-eff}

Introduced by Gifford and Lucassen \cite{effects-gifford,effects-polymorphic}, type and effect
systems have been designed to track effectful operations performed by computations. Examples
include tracking of reading and writing from and to memory locations \cite{effects-talpin-et-al},
communication in message-passing systems \cite{effects-messagepassing} and atomicity in concurrent
applications \cite{effects-atomicity}.

Type and effect systems are usually specified as judgements of the form $\Gamma \vdash e : \tau, \cclrd{r}$,
meaning that the expression $e$ has a type $\tau$ in a (free-variable) context $\Gamma$ and
additionally may have effects described by $\cclrd{r}$. Effect systems are typically added to a
language that already supports effectful operations as a way of increasing the safety -- the type
and effect system provides stronger guarantees than a plain type system. Filinsky
\cite{effects-comprehensive} refers to this approach as \emph{descriptive}\footnote{In contrast
to \emph{prescriptive} effect systems that implement computational effects in a pure language
-- such as monads in Haskell.}.

%---------------------------------------------------------------------------------------------------

\begin{figure}[t]
\begin{equation*}
\tyrule{var}
  {x\!:\!\tau \in \Gamma }
  {\Gamma \vdash x : \tau, \cclrd{\emptyset} }
\end{equation*}
\begin{equation*}
\tyrule{write}
  {\Gamma \vdash e : \tau, \cclrd{r} & l:\ident{ref}_\rho~\tau\in \Gamma}
  {\Gamma \vdash l \leftarrow e : \ident{unit}, \cclrd{r} \cup \cclrd{\{\ident{write}(\rho)\}} }
\end{equation*}
\begin{equation*}
\tyrule{abs}
  {\Gamma, x\!:\!\tau_1 \vdash e : \tau_2, \cclrd{r} }
  {\Gamma \vdash \lambda x.e : \tau_1 \xrightarrow{\cclrd{r}} \tau_2, \cclrd{\emptyset} }
\end{equation*}
\begin{equation*}
\tyrule{app}
  {\Gamma \vdash e_1 : \tau_1 \xrightarrow{\cclrd{r}} \tau_2, \cclrd{s} &
   \Gamma \vdash e_2 : \tau_1, \cclrd{t} }
  {\Gamma \vdash e_1~e_2 : \tau_2, \cclrd{r} \cup \cclrd{s} \cup \cclrd{t} }
\end{equation*}

\figcaption{Simple effect system}
\label{fig:path-eff}
\end{figure}

%---------------------------------------------------------------------------------------------------

\subsection{Simple effect system.}
The structure of a simple effect system\footnote{Most work on effect systems uses $\sigma$ or $F$ for
effect annotations. We use letters $\cclrd{r}, \cclrd{s}, \cclrd{t}$ and also distinguish effect
or coeffect annotations by colour.} is demonstrated in Figure~\ref{fig:path-eff}. The example
shows typing rules for a simply typed lambda calculus with an additional (effectful) operation
$l \leftarrow e$ that writes the value of $e$ to a mutable location $l$. The type of locations
($\ident{ref}_\rho~\tau$) is annotated with a \emph{memory region} $\rho$ of the location $l$.
The effects tracked by the type and effect system over-approximate the actual effects and memory
regions provide a convenient way to build such over-approximation. The effects are
represented as a set of effectful actions that an expression may perform and the effectful action
(\emph{write}) adds a primitive effect $\ident{write}(\rho)$.

The remaining rules are shared by a majority of effect systems. Variable access (\emph{var})
has no effects, application (\emph{app}) combines the effects of both expressions, together with
the latent effects of the function to be applied. Finally, lambda abstraction (\emph{abs}) is a
pure computation that turns the \emph{actual} effects of the body into \emph{latent} effects of
the created function.

%---------------------------------------------------------------------------------------------------

\subsection{Simple coeffect system.}
\label{sec:path-effects-coeff}

When writing the judgements of coeffect systems, we want to emphasize the fact that coeffect
systems talk about \emph{context} rather than \emph{results}. For this reason, we write the
judgements in the form $\coctx{\Gamma}{\cclrd{r}} \vdash e : \tau$, associating the additional
information with the context (left-hand side) of the judgement rather than with the result
(right-hand side) as in $\Gamma \vdash e : \tau, \cclrd{r}$. This change alone would not be
very interesting -- we simply used different syntax to write a predicate with four arguments.
The more interesting difference is how the lambda abstraction rule looks.

The language in Figure~\ref{fig:path-coeff} extends simple lambda calculus with resources and
with a construct $\kvd{access}~e$ that obtains the resource specified by the expression $e$.
Most of the typing rules correspond to those of effect systems. Variable access (\emph{var})
has no context demands, application (\emph{app}) combines context demands of the two
sub-expressions and latent context-requirements of the function.
The (\emph{abs}) rule is different than the corresponding rule for effect systems -- the
resource requirements of the body $\cclrd{r} \cup \cclrd{s}$ are split between the \emph{immediate
context-requirements} associated with the current context $\coctx{\Gamma}{\cclrd{r}}$ and the
\emph{latent context-requirements} of the function.

This is where context-aware languages permit multiple valid typing derivations as discussed in
Section~\ref{sec:path-binding-amb}. In the example here, a resource can be captured
when a function is declared (e.g.~when it is constructed on the server-side where database access
is available), or when a function is called (\eg~when a function created on server-side requires
access to current time-zone, it can use the resource available on the client-side). In other words,
resources in this example support both static (lexical) and dynamic scoping. Out of the multiple
valid typing derivation, we would choose one -- for example, capturing only those server-side
resources that are not available on the client-side\footnote{This can be characterized as
``\emph{dynamic binding when possible, static binding when needed}'' and it is, quite curiously, the
opposite choice than the one used by Haskell's implicit parameters.}. We discuss this
system in detail in Section~\ref{sec:applications-flat-impl}.

%---------------------------------------------------------------------------------------------------

\begin{figure}[t]
\begin{equation*}
\tyrule{var}
  {x\!:\!\tau \in \Gamma }
  {\coctx{\Gamma}{\cclrd{\emptyset}} \vdash x : \tau }
\end{equation*}
\begin{equation*}
\tyrule{access}
  {\coctx{\Gamma}{\cclrd{r}} \vdash e : \ident{res}_\rho~\tau }
  {\coctx{\Gamma}{\cclrd{r} \cup \cclrd{ \{\ident{access}(\rho)\} }} \vdash \kvd{access}~e : \tau }
\end{equation*}
\begin{equation*}
\tyrule{abs}
  {\coctx{(\Gamma, x\!:\!\tau_1)}{\cclrd{r} \cup \cclrd{s}} \vdash e : \tau_2 }
  {\coctx{\Gamma}{\cclrd{r}} \vdash \lambda x.e : \tau_1 \xrightarrow{\cclrd{s}} \tau_2}
\end{equation*}
\begin{equation*}
\tyrule{app}
  {\Gamma \vdash e_1 : \tau_1 \xrightarrow{\cclrd{r}} \tau_2, \cclrd{s} \\
   \Gamma \vdash e_2 : \tau_1, \cclrd{t} }
  {\Gamma \vdash e_1~e_2 : \tau_2, \cclrd{r}\cup\cclrd{s}\cup\cclrd{t}  }
\end{equation*}

\figcaption{Simple coeffect system}
\label{fig:path-coeff}
\end{figure}

%===================================================================================================
%
%     #####
%    #     # ###### #    #   ##   #    # ##### #  ####   ####
%    #       #      ##  ##  #  #  ##   #   #   # #    # #
%     #####  #####  # ## # #    # # #  #   #   # #       ####
%          # #      #    # ###### #  # #   #   # #           #
%    #     # #      #    # #    # #   ##   #   # #    # #    #
%     #####  ###### #    # #    # #    #   #   #  ####   ####
%
%===================================================================================================

\section{Coeffects via language semantics}
\label{sec:path-sem}

Another pathway to coeffects leads through the semantics of effectful and context-dependent
computations. In a pioneering work, Moggi \cite{monad-notions} showed that effects (including
partiality, exceptions, non-determinism and I/O) can be modelled using the category theoretic
notion of \emph{monad}.

When using monads, we distinguish effect-free values $\tau$ from programs, or
computations $\mtyp{}{\tau}$. The \emph{monad} $\mtyp{}{}$ abstracts the \emph{notion of
computation} and provides a way of constructing and composing effectful computations:

~

\begin{definition}
A \emph{monad} over a category $\catc$ is a triple $(M, \ident{unit}, \ident{bind})$ where:
\begin{compactitem}
\item $M$ is a mapping on objects (types) $M : \catc \rightarrow \catc$
\item $\ident{unit}$ is a mapping $\alpha \rightarrow \mtyp{}{\alpha}$
\item $\ident{bind}$ is a mapping $(\alpha \rightarrow \mtyp{}{\beta})
  \rightarrow (\mtyp{}{\alpha} \rightarrow \mtyp{}{\beta})$
\end{compactitem}
such that, for all $f:\alpha \rightarrow \mtyp{}{\beta}$ and $g:\beta \rightarrow \mtyp{}{\gamma}$:
\begin{align}
\tag{\emph{left identity}}
  \ident{bind}~\ident{unit} &= \idf{}
  \\
\tag{\emph{right identity}}
  \ident{bind}~f \circ \ident{unit} &= f
  \\
\tag{\emph{associativity}}
  \ident{bind}~(\ident{bind}~g \circ f) &= (\ident{bind}~f) \circ (\ident{bind}~g)
\end{align}
\end{definition}

\noindent
Without providing much details, we note that well known examples of monads include the partiality
monad ($\mtyp{}{\alpha} = \alpha + {\bot}$) also corresponding to the \ident{Maybe} type in
Haskell, list monad ($\mtyp{}{\tau} = 1 + (\tau \times \mtyp{}{\tau})$) and other.
In programming language semantics, monads can be used in two distinct ways.

%---------------------------------------------------------------------------------------------------

\subsection{Effectful languages and meta-languages}
\label{sec:path-sem-langs}

Moggi uses monads to define two formal systems. In the first formal system, a monad is used to model
the \emph{language} itself. This means that the semantics of a language is given in terms of a
one specific monad and the semantics can be used to reason about programs in that language. To quote
\emph{``When reasoning about programs one has only one monad, because the programming language is
fixed, and the main aim is to prove properties of programs''} \cite[p. 5]{monad-notions}.

In the second formal system, monads are added to the programming language as type constructors,
together with additional constructs corresponding to monadic \ident{bind} and \ident{unit}.
A single program can use multiple monads, but the key benefit is the ability to reason
about multiple languages. To quote \emph{``When reasoning about programming languages one has different
monads, one for each programming language, and the main aim is to study how they relate to each
other''} \cite[p. 5]{monad-notions}.

In this thesis, we generally follow the first approach -- this means that we work with an existing
programming language without needing to add additional constructs corresponding to the primitives
of our semantics (the alternative is discussed in Section~\ref{sec:further-meta}). To clarify the
difference, the following two sections show a minimal example of both formal systems. We follow
Moggi and start with language where judgements have the form $x\!:\!\tau_1 \vdash e : \tau_2$ with
exactly one variable\footnote{This simplifies the examples as we do not need \emph{strong} monad,
but that is an orthogonal issue to the distinction between language semantics and meta-language.}.

%---------------------------------------------------------------------------------------------------

\paragraph{Language semantics.} When using monads to provide semantics of a language, we do not
need to extend the language in any way -- we assume that the language already contains the
effectful primitives (such as the assignment operator $x \leftarrow e$ or other). A judgement
of the form $x\!:\!\tau_1 \vdash e : \tau_2$ is interpreted as a morphism $\tau_1 \rightarrow \mtyp{}{\tau_2}$,
meaning that any expression is interpreted as an effectful computation. The semantics of variable
access and the application of a primitive function $f$ is interpreted as follows:
%
\begin{equation*}
\begin{array}{rcl}
 \sem{x\!:\!\tau_1 \vdash x : \tau_1} &=& \ident{unit}_\mtyp{}{}\\
 \sem{x\!:\!\tau_1 \vdash f~e : \tau_3} &=& (\ident{bind}_\mtyp{}{}~f) \circ \sem{e}\\
\end{array}
\end{equation*}
%
Variable access is an effect-free computation, that returns the value of the variable, wrapped
using $\ident{unit}_\mtyp{}{}$. In the second rule, we assume that $e$ is an expression using
the variable $x$ and producing a value of type $\tau_2$ and that $f$ is a (primitive) function
$\tau_2 \rightarrow \mtyp{}{\tau_3}$. The semantics lifts the function $f$ using $\ident{bind}_\mtyp{}{}$
to a function $\mtyp{}{\tau_2} \rightarrow \mtyp{}{\tau_3}$ which is compatible with the
interpretation of the expression $e$.

%---------------------------------------------------------------------------------------------------

\paragraph{Meta-language interpretation.} ~When designing meta-language based on monads, we need to
extend the lambda calculus with additional type(s) and expressions that correspond to monadic
primitives:
%
\begin{align*}
\tau &:= \ident{num} \sep \tau_1 \rightarrow \tau_2 \sep \mtyp{}{\tau} \\
   e &:= x \sep f~e \sep \kvd{return}_\mtyp{}{}~e \sep \kvd{let}_\mtyp{}{}~x \Leftarrow e_1~\kvd{in}~e_2
\end{align*}
%
The types consist of the primitive type, function type and a type constructor that
represents monadic computations. Thus the expressions in the language can create both
effect-free values, such as $\tau$ and computations $\mtyp{}{\tau}$. The additional expression
$\kvd{return}_\mtyp{}{}$ is used to create a monadic computation (with no effects) from a
value and $\kvd{let}_\mtyp{}{}$ is sequences effectful computations. In the semantics,
monads are not needed to interpret variable access and application, they are only used in the
semantics of additional (monadic) constructs:
%
\begin{equation*}
\begin{array}{rcl}
\sem{x\!:\!\tau \vdash x : \tau} &=& \idf{}\\
\sem{x\!:\!\tau_1 \vdash f~e : \tau_3} &=& f \circ \sem{e}\\
\sem{x\!:\!\tau_1 \vdash \kvd{return}_\mtyp{}{}~e : \mtyp{}{\tau_2}} &=& \ident{unit}_\mtyp{}{} \circ \sem{e}\\
\sem{x\!:\!\tau_1 \vdash \kvd{let}_\mtyp{}{}~y \Leftarrow e_1~\kvd{in}~e_2 : \mtyp{}{\tau_3}} &=&
  \ident{bind}_\mtyp{}{}~\sem{e_2} \circ \sem{e_1}
\end{array}
\end{equation*}

\noindent
In this system, the interpretation of variable access becomes a simple identity function and
application is just composition. Monadic computations are constructed explicitly using
$\kvd{return}_\mtyp{}{}$ (interpreted as $\ident{unit}_\mtyp{}{}$) and they are also sequenced
explicitly using the $\kvd{let}_\mtyp{}{}$ construct. As noted by Moggi, the first formal system
can be easily translated to the latter by inserting appropriate monadic constructs.

Moggi regards the meta-language system as more fundamental, because \emph{``its models are more
general''}. This is a valid and reasonable perspective. Yet, we follow the first style,
precisely because it is \emph{less general}. Our aim is to develop concrete context-aware
programming languages (together with their type systems and semantics) rather than to build a
general framework for reasoning about languages with contextual properties.

%---------------------------------------------------------------------------------------------------

\subsection{Marriage of effects and monads}
\label{sec:path-sem-effects}

The work on effect systems and monads both tackle the same problem -- representing and tracking of
computational effects. The two lines of research have been joined by Wadler and Thiemann
\cite{monads-effects-marriage}. This requires extending the categorical structure. A monadic
computation $\tau_1 \rightarrow \mtyp{}{\tau_2}$ means that the computation has \emph{some}
effects while the judgement $x\!:\!\tau_1 \vdash e : \tau_2, \cclrd{r}$ specifies \emph{what} effects
the computation has.

To solve this mismatch, Wadler and Thiemann use a \emph{family} of monads $\mtyp{\cclrd{r}}{\tau}$
with an annotation that specifies the effects that may be performed by the computation. In their
system, an effectful function $\tau_1 \xrightarrow{\cclrd{r}} \tau_2$ is modelled as a pure
function returning monadic computation $\tau_1 \rightarrow \mtyp{\cclrd{r}}{\tau_2}$. Similarly, the
semantics of a judgement $x\!:\!\tau_1 \vdash e : \tau_2, \cclrd{r}$ can be given as a function
$\tau_1 \rightarrow \mtyp{\cclrd{r}}{\tau_2}$.
The precise nature of the family of monads has been later called \emph{indexed monads} by Tate
\cite{effects-producer-semantics} and further developed by Atkey \cite{monads-parameterised-notions}
in his work on \emph{parameterized monads} and Katsumata \cite{monads-parametric}.

\paragraph{Thesis perspective.}
The key takeaway for this thesis from the outlined line of research is that, if we want to develop a
language with type system that captures context-dependent properties of programs more precisely,
the semantics of the language also needs to be a more fine-grained structure (akin to indexed
monads). While monads have been used to model effects, an existing research links context-dependence
with \emph{comonads} -- the categorical dual of monads.

%---------------------------------------------------------------------------------------------------

\subsection{Context-dependent languages and meta-languages}
\label{sec:path-sem-contextdep}

The theoretical parts of this thesis extend the work of Uustalu and Vene who use comonads
to give the semantics of dataflow computations \cite{comonads-dataflow} and more generally,
notions of \emph{context-dependent computations} \cite{comonads-notions}. The computations discussed
in the latter work include streams, arrays and containers. This is a more diverse set of examples,
but they all mostly represent forms of collections. Ahman et al. \cite{comonads-containers} discuss
the relation between comonads and \emph{containers} \cite{types-containers} in more details.

The utility of comonads has been explored by a number of authors before. Brookes and Geva
\cite{comonads-computational} use \emph{computational} comonads for intensional semantics\footnote{The
structure of a computational comonad has been also used by the author of this thesis to abstract
evaluation order of monadic computations \cite{comonads-malias}.}. In functional programming,
Kieburtz \cite{comonads-and-codata} proposed to use comonads for stream programming, but also
handling of I/O and interoperability.

Biermann and de Paiva used comonads to model the necessity modality $\square$ in intuitionistic
modal S4 \cite{logic-intuitionistic-modal}, linking programming languages derived from modal
logics to comonads. One such language has been reconstructed by Pfenning and Davies
\cite{logic-modal-reconstruction}. Nanevski et al. extend this work to Contextual Modal Type
Theory (CMTT) \cite{logic-cmtt}, which again shows the importance of comonads for
\emph{context-dependent} computations.

While Uustalu and Vene use comonads to define the \emph{language semantics} (the first style
of Moggi), Nanevski, Pfenning and Davies use comonads as part of meta-language, in the form
of $\square$ modality, to reason about context-dependent computations (the second style of
Moggi). Before looking at the details, we use the following definition of comonad:

~

\begin{definition}
A \emph{comonad} over a category $\catc$ is a triple $(C, \ident{counit}, \ident{cobind})$ where:
\begin{compactitem}
\item $C$ is a mapping on objects (types) $C : \catc \rightarrow \catc$
\item $\ident{counit}$ is a mapping $\ctyp{}{\alpha} \rightarrow \alpha$
\item $\ident{cobind}$ is a mapping $(\ctyp{}{\alpha} \rightarrow \beta)
  \rightarrow (\ctyp{}{\alpha} \rightarrow \ctyp{}{\beta})$
\end{compactitem}
%
such that, for all $f:\ctyp{}{\alpha} \rightarrow \beta$ and $g:\ctyp{}{\beta} \rightarrow \gamma$:

\begin{align}
\tag{\emph{left identity}}
  \ident{cobind}~\ident{counit} &= \idf{}
  \\
\tag{\emph{right identity}}
  \ident{counit} \circ \ident{cobind}~f &= f
  \\
\tag{\emph{associativity}}
  \ident{cobind}~(g \circ \ident{cobind}~f) &= (\ident{cobind}~g) \circ (\ident{cobind}~f)
\end{align}
\end{definition}

\noindent
The definition is similar to a monad with ``reversed arrows''. Intuitively, the $\ident{counit}$
operation extracts a value $\alpha$ from a value that carries additional context $\ctyp{}{\alpha}$.
The $\ident{cobind}$ operation turns a context-dependent function
$\ctyp{}{\alpha} \rightarrow \beta$ into a function that takes a value with context, applies
the context-dependent function to value(s) in the context and then propagates the context. The
next section makes this intuitive definition more concrete. More detailed discussion about
comonads can be found in Orchard's PhD thesis \cite{comonads-dom-thesis}.

%---------------------------------------------------------------------------------------------------

\paragraph{Language semantics.}
To demonstrate the approach of Uustalu and Vene, we consider the non-empty list comonad
$\ctyp{}{\tau} = \tau + (\tau \times \ctyp{}{\tau})$. A value of the type is either
the last element $\tau$ or an element followed by another non-empty list $\tau \times \ctyp{}{\tau}$
(consisting of the head $\tau$ and the tail $\ctyp{}{\tau}$). Note that the list must be non-empty,
otherwise \ident{counit} would not be a complete function (it would be undefined on empty list). In
the following, we write $(l_1, \ldots, l_n)$ for a list of $n$ elements:
%
\begin{equation*}
\begin{array}{rcl}
\ident{counit}~(l_1, \ldots, l_n) &=& l_1\\
\ident{cobind}~f~(l_1, \ldots, l_n) &=& (f (l_1, \ldots, l_n), f (l_2, \ldots, l_n), \ldots, f (l_n))
\end{array}
\end{equation*}
%
The \ident{counit} operation returns the current (first) element of the (non-empty) list.
The \ident{cobind} operation creates a new list by applying the context-dependent function $f$
to the entire list, to the suffix of the list, to the suffix of the suffix and so on. Interestingly,
it preserves the \emph{shape} of the list as it turns a list of $n$ elements into another list
of $n$ elements.

In causal dataflow, we can interpret the list as a list consisting of past values, with the
current value in the head. Then, the $\ident{cobind}$ operation calculates the current value
of the output based on the current and all past values of the input; the second element is
calculated based on all past values and the last element is calculated based just on the initial
input $(l_n)$. In addition to the operations of comonad, the model also uses some operations that
are specific to causal dataflow:
%
\begin{equation*}
\begin{array}{rcl}
\ident{prev}~(l_1, \ldots, l_n) &=& (l_2, \ldots, l_n)\\
\end{array}
\end{equation*}
%
The operation drops the first element from the list. In the dataflow interpretation, this means
that it returns the previous state of a value.

Now, consider a simple dataflow language with single-variable contexts, variables,
primitive built-in functions and a construct $\kvd{prev}~e$ that returns the previous
value of the computation $e$. We omit the typing rules, but they are simple -- assuming $e$
has a type $\tau$, the expression $\kvd{prev}~e$ has also type $\tau$. The fact that
the language models dataflow and values are lists (of past values) is a matter of semantics,
which is defined as follows:

\begin{equation*}
\begin{array}{rcl}
\sem{x\!:\!\tau \vdash x : \tau} &=& \ident{counit}_\ctyp{}{}\\
\sem{x\!:\!\tau_1 \vdash f~e : \tau_3} &=& f \circ (\ident{cobind}_\ctyp{}{} ~\sem{e})\\
\sem{x\!:\!\tau_1 \vdash \kvd{prev}~e : \tau_2} &=& \ident{prev} \circ (\ident{cobind}_\ctyp{}{} ~\sem{e})\\
\end{array}
\end{equation*}
%
The semantics follows that of effectful computations using monads. A variable access is interpreted
using $\ident{counit}_\ctyp{}{}$ (extract the variable value); composition
uses $\ident{cobind}_\ctyp{}{}$ to propagate the context to the function $f$ and $\kvd{prev}$
is interpreted using the primitive $\ident{prev}$ (which takes a list and returns a list).
For example, the judgement $x\!:\!\tau \vdash \kvd{prev}~(\kvd{prev}~x) : \tau$ represents an
expression that expects context with variable $x$ and returns a stream of values before the
previous one. The semantics of the term expresses this behaviour:
$(\ident{prev} \circ \ident{prev} \circ (\ident{cobind}_\ctyp{}{}~\ident{counit}_\ctyp{}{}))$.
Note that the first operation is simply an identity function thanks to the comonad laws discussed
earlier.

In the outline presented here, we ignored lambda abstraction. Similarly to monadic semantics,
where lambda abstraction requires \emph{strong} monad, the comonadic semantics also requires
additional structure called \emph{symmetric (semi)monoidal} comonads. This structure is
responsible for the splitting of context-requirements in lambda abstraction. Note that this is
what happens in the unusual (\emph{abs}) rule in Figure~\ref{fig:path-coeff}, which distinguishes
coeffect systems from effect systems.

We return to this topic when discussing lambda abstraction in Section~\ref{sec:applications-structure-lam}
and semantics of flat coeffect systems in Section~\ref{sec:semantics-theory}.

%---------------------------------------------------------------------------------------------------

\paragraph{Meta-language interpretation.} To demonstrate the approach that employs comonads
as part of a meta-language, we look at an example inspired by the work of Pfenning et al.
\cite{logic-modal-reconstruction,logic-cmtt}. We do not attempt to provide a precise overview of
their work. The main purpose of the following discussion is to provide a different intuition behind
comonads, and to present an example of a language that includes comonad as a type constructor,
together with language primitives corresponding to comonadic operations\footnote{In fact,
Pfenning et al. never mention comonads explicitly. This is done in later work by Gabbay et al.
\cite{logic-cmtt-semantics},  but the connection between the language and comonads
is not as direct as in case of monadic or comonadic semantics covered in the previous section.}.

In languages inspired by modal logics, types can have the form $\square \tau$. In the work of
Pfenning and Davies, this is the type of a term that is provable with no assumptions. In distributed
programming language ML5 by Murphy et al. \cite{app-distributed-ml5,logic-distributed-calculus}, the
$\square \tau$ type means \emph{mobile code}, that is code that can be evaluated at any node of a
distributed system (the evaluation corresponds to the axiom $\square \tau \rightarrow \tau$).
Finally, Davies and Pfenning \cite{logic-modal-staged} consider staged computations and interpret
$\square \tau$ as a type of unevaluated expressions of type $\tau$ (with no free variables).

In Contextual Modal Type Theory, the modality $\square$ is further annotated with the free variables
of the (unevaluated) expression. We write $\square^{\cclrd{\Psi}} \tau$ for a type of expressions
that requires a context $\Psi$. The type is a comonadic counterpart to \emph{indexed monads} used by
Wadler and Thiemann when linking monads and effect systems and, indeed, it gives rise to a language
that tracks context-dependence of computations in a type system.

In staged computation, the type $\ctyp{\cclrd{\Psi}}{\tau}$ represents an expression
that requires the context $\Psi$ (i.e.~the expression is an open term that requires variables $\Psi$).
The Figure~\ref{fig:modal-meta} shows two typing rules for such language. The rules directly
correspond to the two operations of a comonad and can be interpreted as follows:

\begin{itemize}
\item (\emph{eval}) corresponds to $\ident{counit} : \ctyp{\cclrd{\emptyset}}{\alpha} \rightarrow \alpha$.
  It indicates that we can evaluate a closed (unevaluated) term and obtain a value. Interestingly, the
  rule requires a specific context annotation (empty set of free variables).
  It is not possible to evaluate an open term.

\item (\emph{letbox}) corresponds to $\ident{cobind} : (\ctyp{\cclrd{\Psi}}{\alpha} \rightarrow \beta)
  \rightarrow \ctyp{\cclrd{\Psi}, \cclrd{\Phi}}{\alpha} \rightarrow \ctyp{\cclrd{\Phi}}{\beta}$.
  Given a term which requires variable context $\cclrd{\Psi}, \cclrd{\Phi}$
  (expression $e_1$) and a function that turns a term needing $\cclrd{\Psi}$ into an evaluated
  value (expression $e_2$), we can construct a term that requires just $\cclrd{\Phi}$.
\end{itemize}

\noindent
The fact that the (\emph{eval}) rule requires a specific context is an interesting relaxation
from ordinary comonads where \ident{counit} needs to be defined for all values. Here, the indexed
\ident{counit} operation needs to be defined \emph{only} on values annotated with $\emptyset$.

The annotated \ident{cobind} operation that corresponds to (\emph{letbox}). An interesting aspect
is that it propagates the context-requirements ``backwards''. The input expression (second parameter)
requires a combination of contexts that are required by the two components -- those required by the
input of the function (first argument) and those required by the resulting expression (result).
This is another key aspect that distinguishes coeffects from effect systems. We return back to
the meta-language approach of embedding comonads in Section~\ref{sec:further-meta}.

%---------------------------------------------------------------------------------------------------

\begin{figure}
\begin{equation*}
\inference[(eval)]
  {\Gamma \vdash e : \square^{\cclrd{\emptyset}}{\tau}}
  {\Gamma \vdash !e : \tau}
\end{equation*}
\begin{equation*}
\inference[(letbox)]
  { \Gamma \vdash e_1 : \square^{\cclrd{\Phi}, \cclrd{\Psi}}{\tau_1} &
    \Gamma, x\!:\!\square^{\cclrd{\Phi}}{\tau_1} \vdash e_2 : \tau_2 }
  { \Gamma \vdash \kvd{let~box}~x=e_1~\kvd{in}~e_2 : \square^{\cclrd{\Psi}}{\tau_2}}
\end{equation*}

\figcaption{Typing for a comonadic language with contextual staged computations}
\label{fig:modal-meta}
\end{figure}

%---------------------------------------------------------------------------------------------------

\paragraph{Thesis perspective.}
As mentioned earlier, we are interested in designing context-dependent languages and so we
use comonads for \emph{language semantics}. Uustalu and Vene present a semantics of
context-dependent computations in terms of comonads. We provide the rest of the story known
from the marriage of monads and effects. We develop coeffect calculus with an type system that
tracks the context demands more precisely (by annotating the types) and we add indexing
to comonads and link the two by giving a formal semantics. The indexing allows us to capture
applications that do not fit into the model provided by plain comonads.

The \emph{meta-language} approach of Pfenning et al. is closely related to
our work. Most importantly, Contextual Modal Type Theory (CMTT) uses indexed $\square$ modality
which corresponds to indexed comonads (in a similar way in which effect systems correspond to
indexed monads). The relation between CMTT and comonads has been suggested by
Gabbay et al. \cite{logic-cmtt-semantics}, but the meta-language employed by CMTT does not
directly correspond to comonadic operations. For example, our (\emph{letbox}) typing rule from
Figure~\ref{fig:modal-meta} is not a primitive of CMTT and would correspond to
$\ident{box}(\cclrd{\Psi}, \ident{letbox}(e_1, x, e_2))$. Nevertheless, the indexing in
CMTT provides a useful hint for adding indexing to the work of Uustalu and Vene.

%===================================================================================================
%
%     #####
%    #     # #    # #####   ####  ##### #####  #    #  ####  ##### #    # #####    ##   #
%    #       #    # #    # #        #   #    # #    # #    #   #   #    # #    #  #  #  #
%     #####  #    # #####   ####    #   #    # #    # #        #   #    # #    # #    # #
%          # #    # #    #      #   #   #####  #    # #        #   #    # #####  ###### #
%    #     # #    # #    # #    #   #   #   #  #    # #    #   #   #    # #   #  #    # #
%     #####   ####  #####   ####    #   #    #  ####   ####    #    ####  #    # #    # ######
%
%===================================================================================================

\section{Coeffects via substructural and bunched logics}
\label{sec:path-logic}

In the coeffect system for tracking resource usage outlined earlier, we associated additional
contextual information (set of available resources) with the variable context of the typing
judgement: $\coctx{\Gamma}{\cclrd{r}} \vdash e : \tau$. In other words, our work focuses on
what is happening on the left hand side of $\vdash$.

In the case of resources, the additional information about the context are simply added to the
variable context (as a products), but we will later look at contextual properties that affect
how variables are represented. More importantly, \emph{structural coeffects} link additional
information to individual variables in the context, rather than the context as a whole.

In this section, we look at type systems that reconsider $\Gamma$ in a number of ways.
First of all, substructural type systems \cite{substruct-attpl-intro} restrict the use of variables
in the language. Most famously linear type systems introduced by Wadler \cite{substruct-linear-change}
can guarantee that a variable is used exactly once. This has interesting implications for memory
management and I/O.

In bunched typing developed by O'Hearn \cite{substruct-bunched}, the variable context is a tree
formed by multiple different constructors (e.g.~one that allows sharing and one that does not).
Most famously, bunched typing has contributed to the development of separation logic
\cite{substruct-separation-logic} (starting a fruitful line of research in software verification),
but it is also interesting on its own.

%---------------------------------------------------------------------------------------------------

\begin{figure}
\begin{equation*}
\tyrule{exchange}
  {\Gamma, x\!:\!\tau_1, y\!:\!\tau_2 \vdash e : \gamma}
  {\Gamma, y\!:\!\tau_2, x\!:\!\tau_1 \vdash e : \gamma}
\end{equation*}
\begin{equation*}
\tyrule{weakening}
  {\Gamma, \Delta \vdash e : \gamma}
  {\Gamma, x\!:\!\tau, \Delta \vdash e : \gamma}
\end{equation*}
\begin{equation*}
\tyrule{contraction}
  {\Gamma, x\!:\!\tau_1, y\!:\!\tau_1, \Delta \vdash e : \tau_2}
  {\Gamma, x\!:\!\tau_1, \Delta \vdash \subst{e}{y}{x} : \tau_2}
\end{equation*}

\figcaption{Exchange, weakening and contraction typing rules}
\label{fig:substructural-rules}
\end{figure}

%---------------------------------------------------------------------------------------------------

\subsection{Substructural type systems.}

Traditionally, $\Gamma$ is viewed as a set of assumptions and typing rules admit (or explicitly
include) three transformations that manipulate the variable contexts which are shown in
Figure~\ref{fig:substructural-rules}. The (\emph{exchange}) rule allows reordering of variables
(which is implicit when assumptions are treated as set); (\emph{weakening}) makes it possible
to discard an assumption -- this has the implication that a variable may be declared but never
used. Finally, (\emph{contraction}) makes it possible to use a single variable multiple times
(in the rule, this is done explicitly by joining multiple variables into a single one using
substitution).

In substructural type systems, the assumptions are typically treated as a list. As a result,
they have to be manipulated explicitly. Different systems allow different subsets of the rules.
For example, \emph{affine} systems allows exchange and weakening, leading to a system where
variable may be used at most once; in \emph{linear} systems, only exchange is permitted and so
every variable has to be used exactly once.

When tracking context-dependent properties associated with individual variables, we need to
be more explicit in how variables are used. Substructural type systems provide a way to do this.
Even if we allow all three operations, we can use a variation on the three rules (exchange,
weakening and contraction) to track which variables are used and how (and to track additional
contextual information about variables).

%---------------------------------------------------------------------------------------------------

\subsection{Bunched type systems.}
Bunched typing makes one more refinement to how $\Gamma$ is treated. Rather than having a list
of assumptions, the context becomes a tree that contains variable typings (or special identity
values) in the leaves and has multiple different types of nodes. The context can be defined,
for example, as follows:
%
\begin{equation*}
\Gamma, \Delta, \Sigma := x\!:\!\alpha \sep I \sep \Gamma, \Gamma \sep 1 \sep \Gamma; \Gamma
\end{equation*}
%
The values $I$ and $1$ represent two kinds of ``empty'' contexts. More interestingly, non-empty
variable contexts may be constructed using two distinct constructors -- $\Gamma, \Gamma$ and
$\Gamma; \Gamma$ -- that have different properties. In particular, weakening and contraction is
only allowed for the $;$ constructor, while exchange is allowed for both.

The structural rules for bunched typing are shown in Figure~\ref{fig:substructural-bunched}.
The syntax $\Gamma(\Delta)$ is used to mean an assumption tree that contains $\Delta$ as a
sub-tree and so, for example, (\emph{exchange1}) can switch the order of contexts anywhere in the
tree. The remaining rules are similar to the rules of linear logic.

One important note about bunched typing is that it requires a different interpretation. The omission
of weakening and contraction in linear logic means that variable can be used exactly once.
In bunched typing, variables may still be duplicated, but only using the ``;'' separator.
The type system can be interpreted as specifying whether a variable may be shared between the
body of a function and the context where a function is declared.

The system introduces two
distinct function types $\tau_1 \rightarrow \tau_2$ and $\tau_1~ \textendash\!\!\!\ast \tau_2$
(corresponding to ``;'' and ``,'' respectively). The key property is that only the first kind
of functions can share variables with the context where a function is declared, while the second
restricts such sharing. We do not attempt to give a detailed description here as it is not
immediately related to coeffects -- for more information, refer to O'Hearn's introduction
\cite{substruct-bunched}.

\begin{figure}
\begin{equation*}
\tyrule{exchange1}
  {\Gamma(\Delta, \Sigma) \vdash e : \alpha}
  {\Gamma(\Sigma, \Delta) \vdash e : \alpha}
\end{equation*}
\begin{equation*}
\tyrule{exchange2}
  {\Gamma(\Delta; \Sigma) \vdash e : \alpha}
  {\Gamma(\Sigma; \Delta) \vdash e : \alpha}
\end{equation*}
\begin{equation*}
\tyrule{weakening}
  {\Gamma(\Delta) \vdash e : \alpha}
  {\Gamma(\Delta; \Sigma) \vdash e : \alpha}
\end{equation*}
\begin{equation*}
\tyrule{contraction}
  {\Gamma(\Delta; \Sigma) \vdash e : \alpha}
  {\Gamma(\Delta) \vdash \subst{e}{\Sigma}{\Delta} : \alpha}
\end{equation*}
\figcaption{Exchange, weakening and contraction rules for bunched typing}
\label{fig:substructural-bunched}
\end{figure}

%---------------------------------------------------------------------------------------------------

\paragraph{Thesis perspective.}

From the perspective of substructural and bunched types, our work can be viewed as annotating
bunches. Such annotations then specify additional information about the context -- or, more
specifically, about the sub-tree of the context. Although this is not the exact definition used in
Chapter~\ref{ch:structural}, we could define contexts as follows:
%
\begin{equation*}
\Gamma, \Delta, \Sigma := x\!:\!\alpha \sep 1 \sep \Gamma, \Gamma \sep \coctx{\Gamma}{\cclrd{r}}
\end{equation*}
%
Now we can not only annotate an entire context with some information (as in the simple coeffect
system for tracking resources that used judgements of a form $\cclrd{\Gamma}{\cclrd{r}} \vdash e : \tau$).
We can also annotate individual components. For example, a context containing variables $x,y,z$
where only $x$ is used could be written as $\coctx{(x\!:\!\tau_1)}{\ident{\cclrd{used}}}, \coctx{(y\!:\!\tau_2, z\!:\!\tau_3)}{\ident{\cclrd{unused}}}$.

For the purpose of this introduction, we ignore important aspects such as how are nested annotations
interpreted. The main goal is to show that coeffects can be easily viewed as an extension to the
work on bunched logic. Aside from this principal connection, \emph{structural coeffects} also
use some of the proof techniques from the work on bunched logics.

%===================================================================================================
%
%     #####  ####### ######
%    #     # #     # #     #
%    #       #     # #     #
%    #       #     # ######
%    #       #     # #
%    #     # #     # #
%     #####  ####### #
%
%===================================================================================================

\section{Context oriented programming}
\label{sec:path-cop}

The importance of context-aware computations is perhaps most obvious when considering mobile
application, client/server web applications or even the internet of things. A pioneering work
in the area using functional languages has been done by Serrano \cite{app-hop-diffuse,app-hop-lang}
(which also inspired the motivating example presented in Chapter~\ref{ch:intro}). His HOP language supports
cross-compilation and programs execute in different contexts. However, HOP is not statically
type checked.

In the software engineering community, a number of authors have addressed the
problem of context-aware computations. Hirschfeld et al. propose \emph{Context-Oriented Programming}
(COP) as a methodology \cite{app-cop-method}. The COP paradigm has been later implemented by
programming language features. Costanza \cite{app-cop-contextl} develops a domain-specific LISP-like
language ContextL and Bardram \cite{app-cop-javafwk} proposes a Java framework for COP.

Finally, the subject of context-awareness has also been addressed in work focusing on the development
of mobile applications \cite{app-cop-mobile,app-cop-mobile2}. Here, the \emph{context} focuses more
on concrete physical context (obtained from the device sensors) than context as an abstract
language feature.

We approach the problem from a different perspective, building on the tradition of
statically-typed functional programming languages, focusing on type systems as the primary way
of capturing contextual properties.

%===================================================================================================
%
%     #####
%    #     # #    # #    # #    #   ##   #####  #   #
%    #       #    # ##  ## ##  ##  #  #  #    #  # #
%     #####  #    # # ## # # ## # #    # #    #   #
%          # #    # #    # #    # ###### #####    #
%    #     # #    # #    # #    # #    # #   #    #
%     #####   ####  #    # #    # #    # #    #   #
%
%===================================================================================================

\section{Summary}

This chapter presented four different pathways leading to the idea of coeffects. We also
introduced the most important related work, although presenting related work was not the
primary goal of the chapter. The primary goal was to present the idea of coeffects as a logical
follow up to a number of research directions. For this reason, we highlighted only certain aspects
of the discussed related work -- the remaining aspects as well as important technical details are
covered throughout the thesis.

The first pathway follows as a generalization of static and dynamic variable binding. Variable
binding can be seen as the most primitive form of context-dependence and coeffects provide a
generalization that can capture different binding mechanisms in a unified way. In the second
pathway, we looked at the dual of well-known work on effect systems. However, this is not simply
a syntactic transformation. As we further discuss in the next chapter, coeffect systems treat lambda
abstraction differently. The third pathway follows by extending comonadic semantics of context-dependent
computations with indexing and building a type system analogous to effect system from the ``marriage of
effects and monads''. Finally, the fourth pathway starts with substructural type systems. Coeffect
systems naturally arise by annotating bunches in bunched logics with additional information. In this
thesis, we mostly follow the first two approaches.